### 2022-07-26 Reports/Week 2/weekly_report2.tex
Upload New File

\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{makecell}
\usepackage[margin=1in]{geometry}

\author{Lukas Evers}
\title{Weekly Report 2}

\renewcommand\theadalign{bl}
\renewcommand{\cellalign}{tc}
\renewcommand\theadfont{\bfseries}
\renewcommand\theadgape{\Gape[4pt]}
\renewcommand\cellgape{\Gape[4pt]}

\begin{document}

\maketitle
\newpage

\section{Collection of Scenarios for Autonomy and Safety Tests}

\subsection{Scenarios Derived From Intel RSS Guidelines}

\begin{table}[!h]

	\caption{RSS Guideline Scenarios}
	\label{tab:table1}
	\begin{tabular}{|c|l|l|}
		\hline
      	\thead{Name of Scenario} & \thead{RSS Rule} & \thead{Description} \\
      	\hline
      	SafeDistance &  \makecell{ Safe linear and \\ lateral distance}  &  \makecell{ An obstacle gets too close to the \\ robot. 				Obstacle gets spawned near \\ the current location of the robot}  \\
     	 \hline
      	RightOfWay &  \makecell{Right of way \\ is given} & \makecell{Another robot drives into \\ the path of the robot.} \\
      	\hline
      	Occlusions & \makecell{Limited Visibility} & \makecell{The robot drives through a densely \\ occupied area with many occlusions \\ in the lidar data ahead.} \\
      	\hline
      	CollisionAvoidance & \makecell{Avoid collisions} & \makecell{Robot performs a dodge maneuver \\ if anything gets too close in the 			zone \\ what is considered to be a safe distance.} \\
      	\hline 
	\end{tabular}
\end{table}

\subsection{Scenarios Derived Fom Sense - Plan - Act - Principle}

\begin{table}[!h]

	\caption{Sense-Plan-Act Scenarios}
	\label{tab:table2}
	\begin{tabular}{|c|l|l|}
		\hline
      	\thead{Name of Scenario} & \thead{Activity} & \thead{Description} \\
      	\hline
      	LidarCrash &  \makecell{Sense} & \makecell{Crash of the Lidar node.} \\
     	\hline
      	OdomCrash &  \makecell{Sense} & \makecell{Crash of the Odometry/Localization node.} \\
      	\hline
      	NoPathFound & \makecell{Plan}  & \makecell{No valid path to goal is found.} \\
      	\hline
      	PlannerCrash & \makecell{Plan}  & \makecell{Path Planning Node crashed.} \\
      	\hline
      	ForcedCollision & \makecell{Act}  & \makecell{A collision with an obstacle is forced.} \\
      	\hline 
      	BatteryLow & \makecell{Act} & \makecell{The battery is too low to at the destination.} \\
      	\hline
      	MotorFailure & \makecell{Act} & \makecell{The motor has a hardware failure.} \\
      	\hline
      	NetworkFailure & \makecell{Act} & \makecell{The method to transmit the motor commands crashes.} \\
      	\hline

	\end{tabular}
\end{table}

\newpage

\section{Current Behaviour vs. Planned Behaviour during Scenarios}

\begin{table}[!h]

	\caption{Current Behaviour vs. Behaviour Planning Extension}
	\label{tab:table3}
	\begin{tabular}{|c|l|l|}
		\hline
      	\thead{Scenario} & \thead{Current Behaviour} & \thead{With Behaviour Planning} \\
      	\hline
      	SafeDistance &  \makecell{Robot replans to avoid the\\ obstacles, but does not factor in the \\speed into what is considered safe.}  &  \makecell{The robot adapts the allowed safe distance \\to the current speed.}  \\
     	 \hline
      	RightOfWay &  \makecell{The robot would replan the \\ path around the around but \\ not predict where things will \\ be moving.} &  \makecell{The robot would calculate the point of \\collision and search for strategies to \\avoid the crash in the future. Slowing down, \\ slowing down hard, or planning around. } \\
      	\hline
      	Occlusions &  \makecell{Robot is unaware of occlusions \\ and sudden appearances of \\dynamic obstacles} 						& \makecell{The robot slows down in areas with \\ occlusions.} \\
      	\hline
      	CollisionAvoidance & \makecell{The robot would replan the \\ path around the around but \\ not predict where things will \\ be moving.} 	& \makecell{The robot would calculate the point of \\collision and search for strategies to \\avoid the crash in the future. Slowing down, \\ slowing down hard, or planning around.} \\
      	\hline 
      	LidarCrash &  \makecell{The robot would continue to drive, \\ but it has no information about the \\ environment anymore.} 															& \makecell{The robot would continue to drive with \\ lowered speed and try to restart the \\ Lidar. Taking into consideration the motion \\ prediction of obstacles. After not receiving\\ information for a longer period the robot\\ would come to a stop.} \\
     	\hline
      	OdomCrash &  \makecell{Robot would drive indefinitely and \\ never reach the goal} 												& \makecell{The robot tries to change the method of \\localization and to meanwhile restart or \\reset the odometry.} \\
      	\hline
      	NoPathFound &  \makecell{The robot would never move.} & \makecell{The robot tries to restart the planner and \\ replan. Else, it tries to change the path \\planning algorithm. If planning is still not \\ possible the robot tries to move to a \\ different starting position. } \\
      	\hline
      	PlannerCrash &  \makecell{The robot would never move.}  & \makecell{The robot will reuse the last path and \\drive slower and restart the path planning \\ node.} \\
      	\hline
      	ForcedCollision &  \makecell{The robot would not be able\\ to move away from the crash.}  											& \makecell{The robot slowly reverses out\\ of the crash and resets \\navigation and odometry at the \\last known safe location. } \\
      	\hline 
      	BatteryLow &  \makecell{The robot would take goals and \\ execute the commands even though\\ the battery would  fail during \\the driving.} & \makecell{The robot would calculate if the battery \\ would be enough to make it to the goal\\ and only then would start to drive. } \\
      	\hline 
      	MotorFailure &  \makecell{The robot would not be able\\ to move.}  											& \makecell{The robot would not be able to move. } \\
      	\hline 
	\end{tabular}
\end{table}

\newpage

\section{Robot Definitions}
\begin{itemize}
\item Turtlebot 3 Waffle
\item ROS2 Foxy
\item Ubuntu 20.04
\item Lidar
\item Wheel Odometry
\item Raspberry Pi Camera
\item Add an additional IMU Sensor in the Simulation

\end{itemize}

\section{Next Week}

\begin{itemize}
\item Work on creating the first BT nodes with ROS Interfaces
\item Make concrete Sub-Scenarios
\item Create a template for scenario creation (Launch file(s), Gazebo Setup, etc.)
\end{itemize}

\end{document}

### 2022-07-26 Code/bt_workspace/src/bt/src/bt_test.cpp
Expand BT further

@@ -36,7 +36,7 @@ int main()
 
     factory.registerNodeType<LidarExecutionCheck>("LidarExecutionCheck");
 
-    auto tree = factory.createTreeFromFile("../../resources/bt.xml");
+    auto tree = factory.createTreeFromFile("/home/levers/behavioural-planning/Code/bt_workspace/src/bt/resources/bt.xml");
 
     tree.tickRoot();
 

### 2022-08-30 Code/bt_workspace/src/bt/include/accel_to_normal_speed.hpp
Map Updates working

@@ -3,33 +3,33 @@
 #include "behaviortree_cpp_v3/action_node.h"
 #include "std_srvs/srv/set_bool.hpp"
 
-class DecelToMinDrivingSpeed : public BT::SyncActionNode
+class AccelToNormalSpeed : public BT::SyncActionNode
 {   
     public:
 
-    DecelToMinDrivingSpeed(const std::string& name) : BT::SyncActionNode(name, {})
+    AccelToNormalSpeed(const std::string& name) : BT::SyncActionNode(name, {})
     {
-        node_= rclcpp::Node::make_shared("decel_to_min_driving_speed");
+        node_= rclcpp::Node::make_shared("accel_to_normal_speed");
         // cmd_vel_publisher_ = node_->create_publisher<geometry_msgs::msg::Twist>("cmd_vel_bt", 1);
         if(debug)
-            RCLCPP_INFO(node_->get_logger(), "DecelMinDrivingSpeed Constructor");
+            RCLCPP_INFO(node_->get_logger(), "AccelToNormalSpeed Constructor");
 
         // std::cout << "DecelMinDrivingSpeed Constructor" << std::endl;
 
         service_client_ = node_->create_client<std_srvs::srv::SetBool>("decel_min_speed_service");
         
         if(debug)
-            RCLCPP_INFO(node_->get_logger(), "DecelMinDrivingSpeed Service Client created.");
+            RCLCPP_INFO(node_->get_logger(), "AccelToNormalSpeed Service Client created.");
 
         request_ = std::make_shared<std_srvs::srv::SetBool_Request>();
 
-        request_->data = true;
+        request_->data = false;
 
         if(debug)
             RCLCPP_INFO(node_->get_logger(), "Exiting Constructor.");
     }
 
-    virtual ~DecelToMinDrivingSpeed()
+    virtual ~AccelToNormalSpeed()
     {
         if(debug)
             std::cout << "DecelToMinDrivingSpeed destroyed" << std::endl;
@@ -47,7 +47,7 @@ class DecelToMinDrivingSpeed : public BT::SyncActionNode
                 RCLCPP_ERROR(rclcpp::get_logger("rclcpp"), "Interrupted while waiting for the service. Exiting.");
                 return BT::NodeStatus::FAILURE;
             }
-            RCLCPP_INFO(rclcpp::get_logger("rclcpp"), "PubCmdVel Service not available, waiting again...");
[…]

### 2022-08-30 Code/bt_workspace/src/bt/include/bt_lidar.hpp

@@ -3,6 +3,7 @@
 #include "bt_lidar.hpp"
 #include "complete_stop.hpp"
 #include "decel_to_min_driving_speed.hpp"
+#include "accel_to_normal_speed.hpp"
 #include "lidar_execution_check.hpp"
 #include "restart_lidar.hpp"
 #include "execution_checker.hpp"
@@ -21,8 +22,22 @@
 #include "disable_cmd_vel_override.hpp"
 #include "reset_occupancy_map.hpp"
 #include "slam_execution_check.hpp"
+#include "update_map_after_collision.hpp"
+#include "save_updated_map.hpp"
+#include "load_updated_map.hpp"
 
-#include "battery_suffient_check.hpp"
+#include "battery_sufficient_check.hpp"
+
+#include "cancel_nav_goal.hpp"
+#include "republish_last_goal.hpp"
+
+#include "global_planner_execution_check.hpp"
+#include "restart_global_planner.hpp"
+#include "publish_closer_goal.hpp"
+#include "path_possible_check.hpp"
+
+
+
 
 #include "behaviortree_cpp_v3/loggers/bt_zmq_publisher.h"
 #include "behaviortree_cpp_v3/bt_factory.h"

### 2022-08-30 Code/bt_workspace/src/bt/include/cancel_nav_goal.hpp

@@ -1,0 +1,37 @@
+#include "rclcpp/rclcpp.hpp"
+#include "behaviortree_cpp_v3/action_node.h"
+#include "nav2_msgs/action/navigate_to_pose.hpp"
+#include "rclcpp_action/rclcpp_action.hpp"
+
+
+class CancelNavGoal : public BT::SyncActionNode
+{   
+    public:
+
+        CancelNavGoal(const std::string& name) : BT::SyncActionNode(name, {})
+        {
+            node_= rclcpp::Node::make_shared("complete_stop");
+
+            action_client_ = rclcpp_action::create_client<nav2_msgs::action::NavigateToPose>(node_, "navigate_to_pose");
+        }
+
+        // void cancel_callback()
+        // {
+
+        // }
+
+        BT::NodeStatus tick() override
+        {
+            auto response = action_client_->async_cancel_all_goals();
+            //auto return_code = response.get()->return_code;
+            return BT::NodeStatus::SUCCESS;
+        }
+
+    private:
+        std::shared_ptr<rclcpp::Node> node_;
+        rclcpp_action::Client<nav2_msgs::action::NavigateToPose>::SharedPtr action_client_;
+        // auto goal_handle_;
+
+};
+
+

### 2022-08-30 Code/bt_workspace/src/bt/include/execution_checker_service.hpp

@@ -2,6 +2,7 @@
 #include <math.h>
 #include "rclcpp/rclcpp.hpp"
 #include "std_srvs/srv/set_bool.hpp"
+#include "std_srvs/srv/empty.hpp"
 #include "sensor_msgs/msg/laser_scan.hpp"
 #include "sensor_msgs/msg/imu.hpp"
 #include "nav_msgs/msg/odometry.hpp"
@@ -11,9 +12,11 @@
 #include "tf2/LinearMath/Quaternion.h"
 #include "tf2/LinearMath/Matrix3x3.h"
 
+#include "nav_msgs/msg/path.hpp"
 #include "geometry_msgs/msg/pose_stamped.hpp"
 #include "geometry_msgs/msg/pose_with_covariance_stamped.hpp"
 #include "bt_msgs/srv/get_distance.hpp"
+#include "action_msgs/msg/goal_status_array.hpp"
 // #include "tf2_conversion_helpers.hpp"
 // #include "tf2_helpers/tf2_conversion_helpers.hpp"
 
@@ -77,6 +80,7 @@ class ExecutionCheckerService : public rclcpp::Node
             const std_srvs::srv::SetBool_Response::SharedPtr response);
 
         rclcpp::Service<std_srvs::srv::SetBool>::SharedPtr collision_service_;
+        rclcpp::Client<std_srvs::srv::Empty>::SharedPtr save_collision_pose_client;
         rclcpp::Subscription<gazebo_msgs::msg::ContactsState>::SharedPtr sub_collision_;
         rclcpp::Time last_msg_received_collision_;
         bool collision_detected_;
@@ -96,13 +100,13 @@ class ExecutionCheckerService : public rclcpp::Node
             const bt_msgs::srv::GetDistance_Request::SharedPtr request,
             const bt_msgs::srv::GetDistance_Response::SharedPtr response);
 
-        void goal_callback(const geometry_msgs::msg::PoseStamped::SharedPtr msg);
+        void goal_callback(const nav_msgs::msg::Path::SharedPtr msg);
         void pose_update_callback(const geometry_msgs::msg::PoseWithCovarianceStamped::SharedPtr msg);
 
         void calc_distance();
 
         rclcpp::Service<bt_msgs::srv::GetDistance>::SharedPtr distance_service_;
-        rclcpp::Subscription<geometry_msgs::msg::PoseStamped>::SharedPtr sub_goal_;
+        rclcpp::Subscription<nav_msgs::msg::Path>::SharedPtr sub_path_;
[…]

### 2022-08-30 Code/bt_workspace/src/bt/include/global_planner_execution_check.hpp

@@ -1,0 +1,83 @@
+#include "behaviortree_cpp_v3/condition_node.h"
+
+#include <chrono>
+
+#include "rclcpp/rclcpp.hpp"
+
+#include "std_srvs/srv/set_bool.hpp"
+#include "lifecycle_msgs/srv/get_state.hpp"
+
+
+class GlobalPlannerExecutionCheck : public BT::ConditionNode
+{
+public:
+
+    GlobalPlannerExecutionCheck(const std::string& name) : BT::ConditionNode(name, {})
+    {   
+        is_planner_running_ = true;
+        debug = true;
+        node_= rclcpp::Node::make_shared("global_planner_execution_check");
+        service_client_ = node_->create_client<lifecycle_msgs::srv::GetState>("planner_server/get_state");
+        // request_ = std::make_shared<lifecycle_msgs::srv::GetState::Request>();
+    }
+
+    BT::NodeStatus tick() override
+    {
+        while (!service_client_->wait_for_service(std::chrono::seconds(1)))
+        {
+            if (!rclcpp::ok()) 
+            {
+                RCLCPP_ERROR(rclcpp::get_logger("rclcpp"), "Interrupted while waiting for the service. Exiting.");
+                return BT::NodeStatus::FAILURE;
+            }
+            RCLCPP_INFO(rclcpp::get_logger("rclcpp"), "Lidar Execution Service not available, waiting again...");
+        }
+
+        auto request = std::make_shared<lifecycle_msgs::srv::GetState::Request>();
+
+        auto result = service_client_->async_send_request(request);
+
+        lifecycle_msgs::msg::State state;
+
+        if(rclcpp::spin_until_future_complete(node_, result) == rclcpp::FutureReturnCode::SUCCESS)
+        {
+            state = result.get()->current_state;
+        }
+        else
+        {
+            return BT::NodeStatus::FAILURE;
+        }
+
+        if (state.id == lifecycle_msgs::msg::State::PRIMARY_STATE_ACTIVE)
+        {
+            is_planner_running_ = true;
+        }
+        else
+        {
+            is_planner_running_ = false;
+        }
+
+        if(debug)
+        {
+            RCLCPP_INFO(node_->get_logger(), "Is Global Planner running: ");
+            RCLCPP_INFO(node_->get_logger(), (is_planner_running_) ? "true" : "false");
+        }
+
+        if (is_planner_running_)
+        {
[…]

### 2022-08-30 Code/bt_workspace/src/bt/include/load_updated_map.hpp

@@ -1,0 +1,71 @@
+#include "rclcpp/rclcpp.hpp"
+#include "behaviortree_cpp_v3/action_node.h"
+#include "nav2_msgs/srv/load_map.hpp"
+
+class LoadUpdatedMap : public BT::SyncActionNode
+{   
+    public:
+
+    LoadUpdatedMap(const std::string& name) : BT::SyncActionNode(name, {})
+    {
+        node_= rclcpp::Node::make_shared("load_updated_map");
+        map_service_client_ = node_->create_client<nav2_msgs::srv::LoadMap>("map_server/load_map");
+
+        //clear_costmap_service_client_ = node_->create_client<nav2_msgs::srv::ClearEntireCostmap>("global_costmap/clear_entirely_global_costmap");
+
+        debug = true;
+    }
+
+    bool load_map_service_call()
+    {
+        while (!map_service_client_->wait_for_service(std::chrono::seconds(1)))
+        {
+            if (!rclcpp::ok()) 
+            {
+                RCLCPP_ERROR(rclcpp::get_logger("load_updated_map"), "Interrupted while waiting for the service. Exiting.");
+                return false;
+            }
+            RCLCPP_INFO(rclcpp::get_logger("load_updated_map"), "LoadMap Service not available, waiting again...");
+        }
+
+        auto request = std::make_shared<nav2_msgs::srv::LoadMap::Request>();
+        request->map_url = "/home/luke/map_updated.yaml";
+
+        auto result = map_service_client_->async_send_request(request);
+
+        if (debug)
+            RCLCPP_INFO(node_->get_logger(), "Sending Service Request.");
+
+        if(rclcpp::spin_until_future_complete(node_, result) == rclcpp::FutureReturnCode::SUCCESS)
+        {
+            // result.get()->
+            return true;
+        }
+        else
+        {
+            return false;
+        }
+
+        return false;
+    }
+
+    BT::NodeStatus tick() override
+    {   
+        if(!load_map_service_call())
+            return BT::NodeStatus::FAILURE;
+
+        return BT::NodeStatus::SUCCESS;
+    }
+
+    private:
+    bool debug;
+    std::shared_ptr<rclcpp::Node> node_;
+    rclcpp::Client<nav2_msgs::srv::LoadMap>::SharedPtr map_service_client_;
[…]

### 2022-08-30 Code/bt_workspace/src/bt/include/path_possible_check.hpp

@@ -1,0 +1,80 @@
+#include "behaviortree_cpp_v3/condition_node.h"
+
+#include "rclcpp/rclcpp.hpp"
+
+#include "std_srvs/srv/set_bool.hpp"
+
+/**
+ * @brief Returns Success when all 4 wheels are still on the ground 
+ * Determined by the robots imu orientation,
+ * if the terrain contains slopes this method does not work 
+ * 
+ */
+class PathPossibleCheck : public BT::ConditionNode
+{
+public:
+
+    PathPossibleCheck(const std::string& name) : BT::ConditionNode(name, {})
+    {   
+        node_= rclcpp::Node::make_shared("path_possible_check");
+        if (debug)
+            RCLCPP_INFO(node_->get_logger(), "Constructor PathPossibleCheck");
+        is_path_possible = true;
+        service_client_ = node_->create_client<std_srvs::srv::SetBool>("path_possible_check_service");
+        request_ = std::make_shared<std_srvs::srv::SetBool::Request>();
+        request_->data = true;
+        if (debug)
+            RCLCPP_INFO(node_->get_logger(), "Constructor PathPossibleCheck finished.");
+        // debug = true;
+        debug = true;
+    }
+
+    BT::NodeStatus tick() override
+    {
+        while (!service_client_->wait_for_service(std::chrono::seconds(1)))
+        {
+            if (!rclcpp::ok()) 
+            {
+                RCLCPP_ERROR(rclcpp::get_logger("rclcpp"), "Interrupted while waiting for the service. Exiting.");
+                return BT::NodeStatus::FAILURE;
+            }
+            RCLCPP_INFO(rclcpp::get_logger("rclcpp"), "PathPossibleCheck Service not available, waiting again...");
+        }
+        
+        auto result = service_client_->async_send_request(request_);
+
+        if(rclcpp::spin_until_future_complete(node_, result) == rclcpp::FutureReturnCode::SUCCESS)
+        {
+            is_path_possible = result.get()->success;
+        }
+        else
+        {
+            return BT::NodeStatus::FAILURE;
+        }
+
+        // getInput("is_lidar_running", is_lidar_running_);
+        // (is_lidar_running_) ? is_lidar_running_ = true : is_lidar_running_ = false;
+
+        if(debug)
+        {
+            RCLCPP_INFO(node_->get_logger(), "Calculated Path can be followed: ");
[…]

### 2022-08-30 Code/bt_workspace/src/bt/include/publish_closer_goal.hpp

@@ -1,0 +1,186 @@
+#include "behaviortree_cpp_v3/action_node.h"
+
+#include "rclcpp/rclcpp.hpp"
+
+#include "std_srvs/srv/empty.hpp"
+#include "bt_msgs/srv/get_last_goal.hpp"
+#include "bt_msgs/srv/get_pose.hpp"
+#include "geometry_msgs/msg/pose_stamped.hpp"
+#include "geometry_msgs/msg/vector3.hpp"
+
+/**
+ * @brief Publish a new /goal_pose which is on the vector between robot pose and old goal pose closer to the robot.
+ * 
+ */
+class PublishCloserGoal : public BT::SyncActionNode
+{
+public:
+
+    PublishCloserGoal(const std::string& name) : BT::SyncActionNode(name, {})
+    {   
+        node_= rclcpp::Node::make_shared("publish_closer_goal");
+        debug = false;
+
+        if (debug)
+            RCLCPP_INFO(node_->get_logger(), "Constructor PublishCloserGoal");
+
+        goal_service_client_ = node_->create_client<bt_msgs::srv::GetLastGoal>("get_last_goal_service");
+        pose_service_client_ = node_->create_client<bt_msgs::srv::GetPose>("get_last_pose_service");
+
+        goal_publisher_ = node_->create_publisher<geometry_msgs::msg::PoseStamped>("goal_pose", 1);
+
+
+        if (debug)
+            RCLCPP_INFO(node_->get_logger(), "Constructor PublishCloserGoal finished.");
+    }
+
+
+    bool last_goal_service_call(geometry_msgs::msg::Pose& pose)
+    {
+        while (!goal_service_client_->wait_for_service(std::chrono::seconds(1)))
+        {
+            if (!rclcpp::ok()) 
+            {
+                RCLCPP_ERROR(rclcpp::get_logger("rclcpp"), "Interrupted while waiting for the service. Exiting.");
+                return false;
+            }
+            RCLCPP_INFO(rclcpp::get_logger("rclcpp"), "Get LastGoal Service not available, waiting again...");
+        }
+
+        auto request = std::make_shared<bt_msgs::srv::GetLastGoal::Request>();
+
+        auto result = goal_service_client_->async_send_request(request);
+
+        if (debug)
+            RCLCPP_INFO(node_->get_logger(), "Sending Service Request.");
+
+        if(rclcpp::spin_until_future_complete(node_, result) == rclcpp::FutureReturnCode::SUCCESS)
+        {
[…]

### 2022-08-30 Code/bt_workspace/src/bt/include/republish_last_goal.hpp

@@ -1,0 +1,71 @@
+#include "behaviortree_cpp_v3/action_node.h"
+
+#include "rclcpp/rclcpp.hpp"
+
+#include "std_srvs/srv/empty.hpp"
+
+/**
+ * @brief Republish the last known goal
+ * 
+ */
+class RepublishLastGoal : public BT::SyncActionNode
+{
+public:
+
+    RepublishLastGoal(const std::string& name) : BT::SyncActionNode(name, {})
+    {   
+        node_= rclcpp::Node::make_shared("republish_last_goal");
+        if (debug)
+            RCLCPP_INFO(node_->get_logger(), "Constructor RepublishLastGoal");
+
+        service_client_ = node_->create_client<std_srvs::srv::Empty>("pub_last_goal_service");
+        request_ = std::make_shared<std_srvs::srv::Empty::Request>();
+
+        if (debug)
+            RCLCPP_INFO(node_->get_logger(), "Constructor RepublishLastGoal finished.");
+        // debug = true;
+        debug = false;
+    }
+
+    BT::NodeStatus tick() override
+    {
+
+        if (debug)
+            RCLCPP_INFO(node_->get_logger(), "RepublishLastGoal ticked.");
+
+        while (!service_client_->wait_for_service(std::chrono::seconds(1)))
+        {
+            if (!rclcpp::ok()) 
+            {
+                RCLCPP_ERROR(rclcpp::get_logger("rclcpp"), "Interrupted while waiting for the service. Exiting.");
+                return BT::NodeStatus::FAILURE;
+            }
+            RCLCPP_INFO(rclcpp::get_logger("rclcpp"), "RepublishLastGoal Service not available, waiting again...");
+        }
+        
+        auto result = service_client_->async_send_request(request_);
+
+        if (debug)
+            RCLCPP_INFO(node_->get_logger(), "Sending Service Request.");
+
+        if(rclcpp::spin_until_future_complete(node_, result) == rclcpp::FutureReturnCode::SUCCESS)
+        {
+            return BT::NodeStatus::SUCCESS;
+        }
+        else
+        {
+            return BT::NodeStatus::FAILURE;
+        }
+
+
+        return BT::NodeStatus::FAILURE;
+    }
+
+
+private:
+    rclcpp::Node::SharedPtr node_;
+    rclcpp::Client<std_srvs::srv::Empty>::SharedPtr service_client_;
+    std::shared_ptr<std_srvs::srv::Empty::Request> request_;
+    bool debug;
+
+};

### 2022-08-30 Code/bt_workspace/src/bt/include/restart_imu.hpp

@@ -26,7 +26,7 @@ class RestartImu : public BT::SyncActionNode
         try
         {
             system("gnome-terminal -e 'sh -c \"ros2 run gazebo_sensor_drivers imu_driver; exec bash\"'");
-            std::this_thread::sleep_for(std::chrono::milliseconds(3000));
+            std::this_thread::sleep_for(std::chrono::milliseconds(2000));
 
             if(debug)
                 RCLCPP_INFO(rclcpp::get_logger("restart_imu"), "Trying to restart IMU");

### 2022-08-30 Code/bt_workspace/src/bt/include/restart_lidar.hpp

@@ -54,7 +54,7 @@ class RestartLidar : public BT::SyncActionNode
         try
         {
             system("gnome-terminal -e 'sh -c \"ros2 run gazebo_sensor_drivers lidar_driver; exec bash\"'");
-            std::this_thread::sleep_for(std::chrono::milliseconds(3000));
+            std::this_thread::sleep_for(std::chrono::milliseconds(2000));
 
             if(debug)
                 RCLCPP_INFO(rclcpp::get_logger("restart_lidar"), "Trying to restart Lidar");

### 2022-08-30 Code/bt_workspace/src/bt/include/reverse_cmd_vel.hpp

@@ -27,8 +27,10 @@ class ReverseCmdVel : public BT::SyncActionNode
         if (debug)
             RCLCPP_INFO(node_->get_logger(), "Constructor ReverseCmdVel");
 
-        service_client_pub_cmd_vel = node_->create_client<bt_msgs::srv::PubCmdVel>("pub_cmd_vel_service");
-        request_pub_cmd_vel = std::make_shared<bt_msgs::srv::PubCmdVel_Request>();
+        cmd_vel_publisher_ = node_->create_publisher<geometry_msgs::msg::Twist>("cmd_vel_bt", 10);
+
+        // service_client_pub_cmd_vel = node_->create_client<bt_msgs::srv::PubCmdVel>("pub_cmd_vel_service");
+        // request_pub_cmd_vel = std::make_shared<bt_msgs::srv::PubCmdVel_Request>();
 
         service_client_get_twist_array = node_->create_client<bt_msgs::srv::GetTwistArray>("get_last_cmd_velocities");
 
@@ -36,32 +38,32 @@ class ReverseCmdVel : public BT::SyncActionNode
             RCLCPP_INFO(node_->get_logger(), "Finished Constructor ReverseCmdVel");
     }
 
-    bool pub_cmd_vel_service_call()
-    {   
-        while (!service_client_get_twist_array->wait_for_service(std::chrono::seconds(1)))
-        {
-            if (!rclcpp::ok()) 
-            {
-                RCLCPP_ERROR(rclcpp::get_logger("rclcpp"), "Interrupted while waiting for the service. Exiting.");
-                return false;
-            }
-            RCLCPP_INFO(rclcpp::get_logger("rclcpp"), "Get Twist Array Service not available, waiting again...");
-        }
+    // bool pub_cmd_vel_service_call()
+    // {   
+    //     while (!service_client_get_twist_array->wait_for_service(std::chrono::seconds(1)))
+    //     {
+    //         if (!rclcpp::ok()) 
+    //         {
+    //             RCLCPP_ERROR(rclcpp::get_logger("rclcpp"), "Interrupted while waiting for the service. Exiting.");
+    //             return false;
+    //         }
+    //         RCLCPP_INFO(rclcpp::get_logger("rclcpp"), "Get Twist Array Service not available, waiting again...");
+    //     }
 
-        auto result = service_client_pub_cmd_vel->async_send_request(request_pub_cmd_vel);
[…]

### 2022-08-30 Code/bt_workspace/src/bt/include/save_updated_map.hpp

@@ -1,0 +1,71 @@
+#include "rclcpp/rclcpp.hpp"
+#include "behaviortree_cpp_v3/action_node.h"
+#include "nav2_msgs/srv/save_map.hpp"
+
+class SaveUpdatedMap : public BT::SyncActionNode
+{   
+    public:
+
+    SaveUpdatedMap(const std::string& name) : BT::SyncActionNode(name, {})
+    {
+        node_= rclcpp::Node::make_shared("save_updated_map");
+        map_service_client_ = node_->create_client<nav2_msgs::srv::SaveMap>("map_saver/save_map");
+
+        //clear_costmap_service_client_ = node_->create_client<nav2_msgs::srv::ClearEntireCostmap>("global_costmap/clear_entirely_global_costmap");
+
+        debug = true;
+    }
+
+    bool save_map_service_call()
+    {
+        while (!map_service_client_->wait_for_service(std::chrono::seconds(1)))
+        {
+            if (!rclcpp::ok()) 
+            {
+                RCLCPP_ERROR(rclcpp::get_logger("save_updated_map"), "Interrupted while waiting for the service. Exiting.");
+                return false;
+            }
+            RCLCPP_INFO(rclcpp::get_logger("save_updated_map"), "SaveMap Service not available, waiting again...");
+        }
+
+        auto request = std::make_shared<nav2_msgs::srv::SaveMap::Request>();
+        request->map_topic = "map_updated";
+        request->map_url = "/home/luke/map_updated";
+
+        auto result = map_service_client_->async_send_request(request);
+
+        if (debug)
+            RCLCPP_INFO(node_->get_logger(), "Sending Service Request.");
+
+        if(rclcpp::spin_until_future_complete(node_, result) == rclcpp::FutureReturnCode::SUCCESS)
+        {
+            return true;
+        }
+        else
+        {
+            return false;
+        }
+
+        return false;
+    }
+
+    BT::NodeStatus tick() override
+    {   
+        if(!save_map_service_call())
+            return BT::NodeStatus::FAILURE;
+
+        return BT::NodeStatus::SUCCESS;
+    }
+
+    private:
+    bool debug;
+    std::shared_ptr<rclcpp::Node> node_;
+    rclcpp::Client<nav2_msgs::srv::SaveMap>::SharedPtr map_service_client_;
[…]

### 2022-08-30 Code/bt_workspace/src/bt/include/update_map_after_collision.hpp

@@ -1,0 +1,310 @@
+#include "rclcpp/rclcpp.hpp"
+#include "behaviortree_cpp_v3/action_node.h"
+#include "nav_msgs/msg/occupancy_grid.hpp"
+#include "bt_msgs/srv/get_last_map.hpp"
+#include "bt_msgs/srv/get_pose.hpp"
+#include <vector>
+#include "visualization_msgs/msg/marker.hpp"
+#include "tf2/convert.h"
+#include "tf2/LinearMath/Quaternion.h"
+#include "tf2/LinearMath/Matrix3x3.h"
+#include "nav2_msgs/srv/clear_entire_costmap.hpp"
+#include "bt_msgs/srv/send_updated_map.hpp"
+
+#define ROBOT_RADIUS 0.45
+
+class UpdateMapAfterCollision : public BT::SyncActionNode
+{   
+    public:
+
+    UpdateMapAfterCollision(const std::string& name) : BT::SyncActionNode(name, {})
+    {
+        node_= rclcpp::Node::make_shared("update_map_after_collision");
+        map_service_client_ = node_->create_client<bt_msgs::srv::GetLastMap>("get_last_map_service");
+        pose_service_client_ = node_->create_client<bt_msgs::srv::GetPose>("get_collision_pose_service");
+        // map_publisher_ = node_->create_publisher<nav_msgs::msg::OccupancyGrid>("map_updated", rclcpp::QoS(rclcpp::KeepLast(1)).transient_local().reliable());
+        // node_->create_publisher<nav_msgs::msg::OccupancyGrid>("map", )
+        marker_publisher_ = node_->create_publisher<visualization_msgs::msg::Marker>("marker", 1);
+        send_map_client_ = node_->create_client<bt_msgs::srv::SendUpdatedMap>("send_updated_map_service");
+        clear_costmap_service_client_ = node_->create_client<nav2_msgs::srv::ClearEntireCostmap>("global_costmap/clear_entirely_global_costmap");
+
+        debug = true;
+    }
+
+    bool get_map_service_call(nav_msgs::msg::OccupancyGrid& map)
+    {
+        while (!map_service_client_->wait_for_service(std::chrono::seconds(1)))
+        {
+            if (!rclcpp::ok()) 
+            {
+                RCLCPP_ERROR(rclcpp::get_logger("rclcpp"), "Interrupted while waiting for the service. Exiting.");
+                return false;
+            }
[…]

### 2022-08-30 Code/bt_workspace/src/bt/launch/bt_utils.launch.py

@@ -1,0 +1,57 @@
+import os
+import launch
+
+from launch_ros.actions import Node
+from launch.actions import IncludeLaunchDescription
+from ament_index_python.packages import get_package_share_directory
+
+def generate_launch_description():
+    ld = launch.LaunchDescription()
+
+    bt_directory = get_package_share_directory("bt")
+
+    param_file = os.path.join(bt_directory, "config", "util_params.yaml")
+
+    decision_gate_cmd = Node(
+        package="cmd_vel_decision_gate",
+        executable="cmd_vel_decision_gate",
+        output="screen",
+        parameters=[param_file]
+    )
+
+    # robot_battery_cmd = Node(
+    #     package="robot_battery",
+    #     executable="robot_battery",
+    #     output="screen",
+    #     parameters=[{"charge": 100.0},
+	# 	            {"idle_decrease": 0.01},
+	# 	            {"drive_decrease": 0.02}]
+    # )
+
+    robot_battery_cmd = Node(
+        package="robot_battery",
+        executable="robot_battery",
+        output="screen",
+        parameters=[param_file]
+    )
+
+    execution_checker_cmd = Node(
+        package="bt",
+        executable="execution_checker_service",
+        output="screen",
+        parameters=[param_file]
+    )
+
+    sensor_data_cmd = Node(
+        package="sensor_data_backup",
+        executable="sensor_data_backup",
+        output="screen",
+        parameters=[param_file]
+    )
+
+    ld.add_action(decision_gate_cmd)
+    ld.add_action(robot_battery_cmd)
+    ld.add_action(execution_checker_cmd)
+    ld.add_action(sensor_data_cmd)
+
+    return ld

### 2022-08-30 Code/bt_workspace/src/bt/src/bt_lidar.cpp

@@ -17,6 +17,7 @@ int main(int argc, const char* argv[])
     RCLCPP_INFO(rclcpp::get_logger("root"), "Creating Nodes"); 
     // BT::PortsList lidar_exec_check_ports = {BT::InputPort<bool>("is_lidar_running")};
     factory.registerNodeType<LidarExecutionCheck>("LidarExecutionCheck");
+    factory.registerNodeType<AccelToNormalSpeed>("AccelToNormalSpeed");
     factory.registerNodeType<DecelToMinDrivingSpeed>("DecelToMinDrivingSpeed");
     factory.registerNodeType<CompleteStop>("CompleteStop");
     factory.registerNodeType<RestartLidar>("RestartLidar");
@@ -35,9 +36,22 @@ int main(int argc, const char* argv[])
     factory.registerNodeType<DisableCmdVelOverride>("DisableCmdVelOverride");
     factory.registerNodeType<SlamExecutionCheck>("SlamExecutionCheck");
     factory.registerNodeType<ResetOccupancyMap>("ResetOccupancyMap");
+    factory.registerNodeType<UpdateMapAfterCollision>("UpdateMapAfterCollision");
+    factory.registerNodeType<SaveUpdatedMap>("SaveUpdatedMap");
+    factory.registerNodeType<LoadUpdatedMap>("LoadUpdatedMap");
+
+    factory.registerNodeType<BatterySufficientCheck>("BatterySufficientCheck");
+
+    factory.registerNodeType<CancelNavGoal>("CancelNavGoal");
+    factory.registerNodeType<RepublishLastGoal>("RepublishLastGoal");
+
+    factory.registerNodeType<GlobalPlannerExecutionCheck>("GlobalPlannerExecutionCheck");
+    factory.registerNodeType<RestartGlobalPlanner>("RestartGlobalPlanner");
+    factory.registerNodeType<PathPossibleCheck>("PathPossibleCheck");
+    factory.registerNodeType<PublishCloserGoal>("PublishCloserGoal");
 
     RCLCPP_INFO(rclcpp::get_logger("root"), "Creating BT"); 
-    auto tree = factory.createTreeFromFile("/home/luke/behavioural-planning/Code/bt_workspace/src/bt/resources/map_bt.xml");
+    auto tree = factory.createTreeFromFile("/home/luke/behavioural-planning/Code/bt_workspace/src/bt/resources/save_new_map.xml");
[…]

### 2022-08-30 Code/bt_workspace/src/cmd_vel_decision_gate/include/cmd_vel_decision_gate/cmd_vel_decision_gate.hpp

@@ -2,8 +2,11 @@
 #include "rclcpp/rclcpp.hpp"
 #include "geometry_msgs/msg/twist.hpp"
 #include "bt_msgs/srv/pub_cmd_vel.hpp"
+#include "std_srvs/srv/set_bool.hpp"
 #include "rcl_interfaces/msg/parameter.hpp"
 #include "rcl_interfaces/msg/set_parameters_result.hpp"
+
+#define MIN_DRIVING_SPEED 0.05
 
 class CmdVelDecisionGate : public rclcpp::Node 
 {
@@ -16,9 +19,15 @@ class CmdVelDecisionGate : public rclcpp::Node
                     const bt_msgs::srv::PubCmdVel_Request::SharedPtr request,
                     const bt_msgs::srv::PubCmdVel_Response::SharedPtr response);
 
+        void decel_service_callback(
+            const std_srvs::srv::SetBool_Request::SharedPtr request,
+            const std_srvs::srv::SetBool_Response::SharedPtr response);
+
         void pub_complete_stop();
 
         void cmd_vel_nav_callback(const geometry_msgs::msg::Twist::SharedPtr msg);
+
+        void cmd_vel_bt_callback(const geometry_msgs::msg::Twist::SharedPtr msg);
 
         void add_node_to_executor(rclcpp::Node::SharedPtr node_ptr);
 
@@ -31,22 +40,29 @@ class CmdVelDecisionGate : public rclcpp::Node
         OnSetParametersCallbackHandle::SharedPtr parameter_callback_handle_;
 
         rclcpp::Subscription<geometry_msgs::msg::Twist>::SharedPtr sub_nav_;
+        rclcpp::Subscription<geometry_msgs::msg::Twist>::SharedPtr sub_bt_;
 
         rclcpp::Publisher<geometry_msgs::msg::Twist>::SharedPtr pub_cmd_vel_;
 
-        rclcpp::Service<bt_msgs::srv::PubCmdVel>::SharedPtr service_pub_cmd_vel_;
-        rclcpp::executors::SingleThreadedExecutor exec_;
-        double publish_duration;
-        bool service_called; 
+        // rclcpp::Service<bt_msgs::srv::PubCmdVel>::SharedPtr service_pub_cmd_vel_;
 
-        bool bt_override_flag;
+
+        rclcpp::Service<std_srvs::srv::SetBool>::SharedPtr decel_service;
+
[…]

### 2022-08-30 Code/bt_workspace/src/gazebo_sensor_drivers/src/odom_driver.cpp

@@ -37,7 +37,8 @@ class OdomDriver : public rclcpp::Node
 
     void callback(const nav_msgs::msg::Odometry::SharedPtr msg)
     {
-      RCLCPP_INFO(this->get_logger(), "%d", msg->header.stamp.sec);
+      if(debug)
+        RCLCPP_INFO(this->get_logger(), "%d", msg->header.stamp.sec);
 
       // tf = buffer.lookupTransform()
 
@@ -46,7 +47,7 @@ class OdomDriver : public rclcpp::Node
 
   private:
 
-    bool debug = true;
+    bool debug = false;
     rclcpp::Publisher<nav_msgs::msg::Odometry>::SharedPtr pub_;
     rclcpp::Subscription<nav_msgs::msg::Odometry>::SharedPtr sub_;
     //tf2_ros::TransformListener tf_listener;

### 2022-08-30 Code/bt_workspace/src/robot_battery/include/robot_battery/robot_battery.hpp

@@ -2,9 +2,11 @@
 #include "rclcpp/rclcpp.hpp"
 #include "geometry_msgs/msg/twist.hpp"
 #include "bt_msgs/srv/get_charge.hpp"
+#include "rcl_interfaces/msg/parameter.hpp"
+#include "rcl_interfaces/msg/set_parameters_result.hpp"
 
-#define IDLE_CHARGE_DECREASE_PER_SEC 0.1
-#define DRIVING_CHARGE_DECREASE_PER_MSG 0.02 // 20 msg per second -> 0.4% / sec driving
+// #define IDLE_CHARGE_DECREASE_PER_SEC 0.1
+// #define DRIVING_CHARGE_DECREASE_PER_MSG 0.02 // 20 msg per second -> 0.4% / sec driving
 
 class RobotBattery : public rclcpp::Node
 {
@@ -13,7 +15,12 @@ class RobotBattery : public rclcpp::Node
 
         ~RobotBattery() = default;
 
+        rcl_interfaces::msg::SetParametersResult parameters_callback(const std::vector<rclcpp::Parameter> &parameters);
+    
     private:
+
+
+        OnSetParametersCallbackHandle::SharedPtr parameter_callback_handle_;
 
         rclcpp::TimerBase::SharedPtr timer_;
         void timer_callback();
@@ -27,7 +34,7 @@ class RobotBattery : public rclcpp::Node
             const bt_msgs::srv::GetCharge_Response::SharedPtr response);
 
         float charge_;
-        float capacity_;
+        // float capacity_;
         float idle_decrease_;
         float drive_decrease_;
              

### 2022-08-30 Code/bt_workspace/src/sensor_data_backup/src/sensor_data_backup.cpp

@@ -2,23 +2,95 @@
 
 SensorDataBackup::SensorDataBackup(const std::string & node_name) : Node(node_name)
 {
-    sub_cmd_vel = this->create_subscription<geometry_msgs::msg::Twist>("cmd_vel", 10, std::bind(&SensorDataBackup::cmd_vel_callback, this, std::placeholders::_1));
-    service_cmd_vel = this->create_service<bt_msgs::srv::GetTwistArray>("get_last_cmd_velocities", std::bind(&SensorDataBackup::cmd_vel_service_callback, this, std::placeholders::_1, std::placeholders::_2));
-    bool debug = true;
+    sub_cmd_vel_ = this->create_subscription<geometry_msgs::msg::Twist>("cmd_vel_nav", 10, std::bind(&SensorDataBackup::cmd_vel_callback, this, std::placeholders::_1));
+    service_cmd_vel_ = this->create_service<bt_msgs::srv::GetTwistArray>("get_last_cmd_velocities", std::bind(&SensorDataBackup::cmd_vel_service_callback, this, std::placeholders::_1, std::placeholders::_2));
+
+    pub_map_ = this->create_publisher<nav_msgs::msg::OccupancyGrid>("map_updated", rclcpp::QoS(rclcpp::KeepLast(1)).transient_local().reliable());
+    pub_goal_ = this->create_publisher<geometry_msgs::msg::PoseStamped>("goal_pose", 1);
+    sub_goal_ = this->create_subscription<geometry_msgs::msg::PoseStamped>("goal_pose", 1, std::bind(&SensorDataBackup::goal_callback, this, std::placeholders::_1));
+    service_pub_last_goal_ = this->create_service<std_srvs::srv::Empty>("pub_last_goal_service", std::bind(&SensorDataBackup::pub_last_goal_service_callback, this, std::placeholders::_1, std::placeholders::_2));
+    
+    sub_map_ = this->create_subscription<nav_msgs::msg::OccupancyGrid>("map", rclcpp::QoS(rclcpp::KeepLast(1)).transient_local().reliable(), std::bind(&SensorDataBackup::map_callback, this, std::placeholders::_1));
+    sub_pose = this->create_subscription<geometry_msgs::msg::PoseWithCovarianceStamped>("amcl_pose", 1, std::bind(&SensorDataBackup::pose_update_callback, this, std::placeholders::_1));
+
[…]

### 2022-08-30 Code/bt_workspace/src/bt/include/battery_sufficient_check.hpp
Map Updates working

@@ -38,6 +38,8 @@ public:
 
     bool distance_service_call(float &distance)
     {
+        if (debug)
+            RCLCPP_INFO(node_->get_logger(), "Distance Service Call");
         while (!distance_service_client_->wait_for_service(std::chrono::seconds(1)))
         {
             if (!rclcpp::ok()) 
@@ -65,7 +67,10 @@ public:
     }
 
     bool charge_service_call(float &charge, float &idle_decrease, float &drive_decrease)
-    {
+    {   
+        if (debug)
+            RCLCPP_INFO(node_->get_logger(), "Charge Service Call");
+
         while (!charge_service_client_->wait_for_service(std::chrono::seconds(1)))
         {
             if (!rclcpp::ok()) 
@@ -82,7 +87,7 @@ public:
             RCLCPP_INFO(node_->get_logger(), "Sent Service Request");
 
         if(rclcpp::spin_until_future_complete(node_, charge_result) == rclcpp::FutureReturnCode::SUCCESS)
-        {
+        {   
             charge = charge_result.get()->charge;
             idle_decrease = charge_result.get()->idle_decrease_per_sec;
             drive_decrease = charge_result.get()->drive_decrease_per_sec;
@@ -101,6 +106,9 @@ public:
         if(!distance_service_call(distance))
             return BT::NodeStatus::FAILURE;
 
+        if (debug)
+            RCLCPP_INFO(node_->get_logger(), "Distance to Goal: %f", distance);
+
         if(distance == 0.0)
             return BT::NodeStatus::SUCCESS;
         
@@ -109,9 +117,15 @@ public:
         if(!charge_service_call(charge, idle_decrease, drive_decrease))
             return BT::NodeStatus::FAILURE;
         
+        if (debug)
+            RCLCPP_INFO(node_->get_logger(), "Charge: %f", charge);
+
         float time = distance / avrg_speed_;
 
         float consumption = time * (idle_decrease + drive_decrease);
+
+        if (debug)
+            RCLCPP_INFO(node_->get_logger(), "Consumption: %f", consumption);
 
         if (charge >= (consumption * safety_factor_))
         {

### 2022-08-30 Code/bt_workspace/src/bt/include/complete_stop.hpp

@@ -1,7 +1,6 @@
 #include "rclcpp/rclcpp.hpp"
 #include "geometry_msgs/msg/twist.hpp"
 #include "behaviortree_cpp_v3/action_node.h"
-#include "bt_msgs/srv/pub_cmd_vel.hpp"
 
 class CompleteStop : public BT::SyncActionNode
 {   
@@ -11,62 +10,34 @@ class CompleteStop : public BT::SyncActionNode
     {
         node_= rclcpp::Node::make_shared("complete_stop");
         //cmd_vel_publisher_ = node_->create_publisher<geometry_msgs::msg::Twist>("cmd_vel_bt", 10);
-        service_client_ = node_->create_client<bt_msgs::srv::PubCmdVel>("pub_cmd_vel_service");
-        request_ = std::make_shared<bt_msgs::srv::PubCmdVel_Request>();
+        // service_client_ = node_->create_client<bt_msgs::srv::PubCmdVel>("pub_cmd_vel_service");
+        // request_ = std::make_shared<bt_msgs::srv::PubCmdVel_Request>();
+        cmd_vel_publisher_ = node_->create_publisher<geometry_msgs::msg::Twist>("cmd_vel_bt", 1);
 
-        request_->cmd_vel.linear.x = 0.00;
-        request_->cmd_vel.linear.y = 0.0;
-        request_->cmd_vel.linear.z = 0.0;
-        request_->cmd_vel.angular.z = 0.0;
-        request_->cmd_vel.angular.x = 0.0;
-        request_->cmd_vel.angular.y = 0.0;
-        request_->time_in_seconds = 0.001;
+        stop_msg.linear.x = 0.0;
+        stop_msg.linear.y = 0.0;
+        stop_msg.linear.z = 0.0;
+        stop_msg.angular.z = 0.0;
+        stop_msg.angular.x = 0.0;
+        stop_msg.angular.y = 0.0;
+       
     }
 
     BT::NodeStatus tick() override
     {
-        // geometry_msgs::msg::Twist message = geometry_msgs::msg::Twist();
-
-        // message.linear.x = 0.0;
-        // message.linear.y = 0.0;
-        // message.linear.z = 0.0;
-        // message.angular.z = 0.0;
-        // message.angular.x = 0.0;
-        // message.angular.y = 0.0;
-
-        // cmd_vel_publisher_->publish(message);
         if (debug)
             RCLCPP_INFO(node_->get_logger(), "CompleteStop ticked");
[…]

### 2022-08-30 Code/bt_workspace/src/bt/include/restart_global_planner.hpp

@@ -12,11 +12,11 @@
 
 #include "behaviortree_cpp_v3/action_node.h"
 
-class RestartOdom : public BT::SyncActionNode
+class RestartGlobalPlanner : public BT::SyncActionNode
 {   
     public:
 
-    RestartOdom(const std::string& name) : BT::SyncActionNode(name, {})
+    RestartGlobalPlanner(const std::string& name) : BT::SyncActionNode(name, {})
     {
         
     }
@@ -25,18 +25,18 @@ class RestartOdom : public BT::SyncActionNode
     {
         try
         {
-            system("gnome-terminal -e 'sh -c \"ros2 run gazebo_sensor_drivers odom_driver; exec bash\"'");
-            std::this_thread::sleep_for(std::chrono::milliseconds(2000));
+            system("gnome-terminal -e 'sh -c \"ros2 launch bt launch_global_planner.launch.py ; exec bash\"'");
+            std::this_thread::sleep_for(std::chrono::milliseconds(3000));
 
             if(debug)
-                RCLCPP_INFO(rclcpp::get_logger("restart_Odom"), "Trying to restart Odom");
+                RCLCPP_INFO(rclcpp::get_logger("restart_global_planner"), "Trying to restart Global Planner");
             return BT::NodeStatus::SUCCESS;
         }
         catch(const std::exception& e)
         {   
             //std::cerr << e.what() << '\n';
             if(debug)
-                RCLCPP_ERROR(rclcpp::get_logger("restart_lidar"), e.what());
+                RCLCPP_ERROR(rclcpp::get_logger("restart_global_planner"), e.what());
             return BT::NodeStatus::FAILURE;
         }
     }

### 2022-08-30 Code/bt_workspace/src/bt/launch/launch_global_planner.launch.py

@@ -1,0 +1,41 @@
+import os
+
+import launch.actions
+import launch_ros.actions
+from ament_index_python.packages import get_package_share_directory
+from launch import LaunchDescription
+from launch_ros.actions import Node
+
+def generate_launch_description():
+    
+    # rviz_config_dir = os.path.join(get_package_share_directory('nav2_course'), 'config', 'nav2_default_view.rviz')
+    # map_file = os.path.join(get_package_share_directory('nav2_course'), 'config', 'turtlebot3_world.yaml')
+    params = os.path.join(
+        get_package_share_directory('turtlebot3_navigation2'),
+        'param',
+        'waffle.yaml'
+    )
+
+    return LaunchDescription([
+        Node(
+            package='nav2_planner',
+            executable='planner_server',
+            name='planner_server',
+            output='screen',
+            parameters=[{params} 
+                       ]
+            ),
+        Node
+        (
+            package='nav2_lifecycle_manager',
+            executable='lifecycle_manager',
+            name='lifecycle_manager_localization',
+            output='screen',
+            parameters=
+            [
+                {'use_sim_time': True},
+                {'autostart': True},
+                {'node_names': ['planner_server']}
+            ]
+        )            
+    ])

### 2022-08-30 Code/bt_workspace/src/bt/src/execution_checker_service.cpp

@@ -43,6 +43,13 @@ ExecutionCheckerService::ExecutionCheckerService (const std::string & node_name)
             this,
             std::placeholders::_1,
             std::placeholders::_2));
+
+    plan_possible_service_ = this->create_service<std_srvs::srv::SetBool>(
+        "path_possible_check_service",
+        std::bind(&ExecutionCheckerService::plan_possible_service_callback,
+        this,
+        std::placeholders::_1,
+        std::placeholders::_2));
     
 
     // topics_qos = rclcpp::QoS(rclcpp::KeepLast(1)).transient_local().reliable(); // Transient Local Sub Settings
@@ -50,13 +57,16 @@ ExecutionCheckerService::ExecutionCheckerService (const std::string & node_name)
     sub_imu_ = this->create_subscription<sensor_msgs::msg::Imu>("imu", 10, std::bind(&ExecutionCheckerService::imu_callback, this, std::placeholders::_1));
     sub_odom_ = this->create_subscription<nav_msgs::msg::Odometry>("odom", 10, std::bind(&ExecutionCheckerService::odom_callback, this, std::placeholders::_1));
     sub_collision_ = this->create_subscription<gazebo_msgs::msg::ContactsState>("bumper_states", 10, std::bind(&ExecutionCheckerService::collision_callback, this, std::placeholders::_1));
-    sub_goal_ = this->create_subscription<geometry_msgs::msg::PoseStamped>("goal_pose", 1, std::bind(&ExecutionCheckerService::goal_callback, this, std::placeholders::_1));
+    sub_path_ = this->create_subscription<nav_msgs::msg::Path>("plan", 1, std::bind(&ExecutionCheckerService::goal_callback, this, std::placeholders::_1));
     //sub_pose_ = this->create_subscription<geometry_msgs::msg::PoseWithCovarianceStamped>("amcl_pose", 1, std::bind(&ExecutionCheckerService::pose_update_callback, this, std::placeholders::_1));
     sub_pose_ = this->create_subscription<geometry_msgs::msg::PoseWithCovarianceStamped>("amcl_pose", 10, std::bind(&ExecutionCheckerService::pose_update_callback, this, std::placeholders::_1));
[…]

### 2022-08-30 Code/bt_workspace/src/cmd_vel_decision_gate/src/cmd_vel_decision_gate.cpp

@@ -5,97 +5,96 @@ CmdVelDecisionGate::CmdVelDecisionGate(
   const rclcpp::NodeOptions & options = rclcpp::NodeOptions().allow_undeclared_parameters(true)) : Node(node_name, options)
 {
     sub_nav_ = this->create_subscription<geometry_msgs::msg::Twist>("cmd_vel_nav", 1, std::bind(&CmdVelDecisionGate::cmd_vel_nav_callback, this, std::placeholders::_1));
+    sub_bt_ = this->create_subscription<geometry_msgs::msg::Twist>("cmd_vel_bt", 1, std::bind(&CmdVelDecisionGate::cmd_vel_bt_callback, this, std::placeholders::_1));
+    
     pub_cmd_vel_ = this->create_publisher<geometry_msgs::msg::Twist>("cmd_vel", 10);
 
-    service_pub_cmd_vel_ = this->create_service<bt_msgs::srv::PubCmdVel>("pub_cmd_vel_service", std::bind(&CmdVelDecisionGate::pub_cmd_vel_service_callback, this, std::placeholders::_1, std::placeholders::_2));
+    // service_pub_cmd_vel_ = this->create_service<bt_msgs::srv::PubCmdVel>("pub_cmd_vel_service", std::bind(&CmdVelDecisionGate::pub_cmd_vel_service_callback, this, std::placeholders::_1, std::placeholders::_2));
 
+    decel_service = this->create_service<std_srvs::srv::SetBool>("decel_min_speed_service", std::bind(&CmdVelDecisionGate::decel_service_callback, this, std::placeholders::_1, std::placeholders::_2));
+    
+    // service_called = false;
 
-    service_called = false;
     debug = true;
 
-    this->declare_parameter("bt_override", false);
+    // this->declare_parameter("bt_override", false);
+
+    this->declare_parameter("decel_flag", false);
+
+    decel_override_flag = this->get_parameter("decel_flag").as_bool();
     
     // What is the diff between these two lines?
-    bt_override_flag = this->get_parameter("bt_override").as_bool();
+    // bt_override_flag = this->get_parameter("bt_override").as_bool();
 
-    this->get_parameter("bt_override", bt_override_flag);
+    // this->get_parameter("bt_override", bt_override_flag);
 
[…]

### 2022-08-30 Code/bt_workspace/src/gazebo_sensor_drivers/launch/sensor_drivers.launch.py

@@ -1,0 +1,41 @@
+import os
+import launch
+
+from launch_ros.actions import Node
+from launch.actions import IncludeLaunchDescription
+from ament_index_python.packages import get_package_share_directory
+
+def generate_launch_description():
+    ld = launch.LaunchDescription()
+
+    # bt_directory = get_package_share_directory("bt")
+
+    # param_file = os.path.join(bt_directory, "config", "util_params.yaml")
+
+    lidar_cmd = Node(
+        package="gazebo_sensor_drivers",
+        executable="lidar_driver",
+        output="screen",
+
+    )
+
+    imu_cmd = Node(
+        package="gazebo_sensor_drivers",
+        executable="imu_driver",
+        output="screen",
+
+    )
+
+    odom_cmd = Node(
+        package="gazebo_sensor_drivers",
+        executable="odom_driver",
+        output="screen",
+
+    )
+
+
+    ld.add_action(lidar_cmd)
+    ld.add_action(imu_cmd)
+    ld.add_action(odom_cmd)
+
+    return ld

### 2022-08-30 Code/bt_workspace/src/robot_battery/src/robot_battery.cpp

@@ -2,6 +2,8 @@
 
 RobotBattery::RobotBattery(const std::string & node_name) : Node(node_name)
 {
+    debug = false;
+
     sub_cmd_vel_ = this->create_subscription<geometry_msgs::msg::Twist>(
         "cmd_vel", 
         10, 
@@ -20,17 +22,63 @@ RobotBattery::RobotBattery(const std::string & node_name) : Node(node_name)
         this, std::placeholders::_1, 
         std::placeholders::_2));
     
-    this->declare_parameter("capacity", 100.0);
+    // this->declare_parameter("capacity", 100.0);
     this->declare_parameter("charge", 100.0);
     this->declare_parameter("idle_decrease", 0.01);
     this->declare_parameter("drive_decrease", 0.02);
 
-    this->get_parameter("capacity", capacity_);
+    // this->get_parameter("capacity", capacity_);
     this->get_parameter("charge", charge_);
     this->get_parameter("idle_decrease", idle_decrease_);
     this->get_parameter("drive_decrease", drive_decrease_);
 
-    debug = false;
+    parameter_callback_handle_ = this->add_on_set_parameters_callback(std::bind(&RobotBattery::parameters_callback, this, std::placeholders::_1));
+
+}
+
+rcl_interfaces::msg::SetParametersResult RobotBattery::parameters_callback(const std::vector<rclcpp::Parameter> &parameters)
+{
+    if(debug)
+        RCLCPP_INFO(this->get_logger(), "Parameter Callback called.");
+
+    for (auto &param : parameters)
+    {
+
+        RCLCPP_INFO(this->get_logger(), "%s", param.get_name().c_str());
+        RCLCPP_INFO(this->get_logger(), "%s", param.get_type_name().c_str());
+        RCLCPP_INFO(this->get_logger(), "%s", param.value_to_string().c_str());
+
+        if(param.get_name() == "charge")
+        {
+            if (param.get_type() == rclcpp::ParameterType::PARAMETER_DOUBLE)
+            {
+                charge_ = param.as_double();
+
+                if(debug)
+                    RCLCPP_INFO(this->get_logger(), "Changed charge.");
+            }
+        }
+        else if(param.get_name() == "idle_decrease")
+        {
+            if (param.get_type() == rclcpp::ParameterType::PARAMETER_DOUBLE)
+            {
+                idle_decrease_ = param.as_double();
+            }
+        }
+        else if(param.get_name() == "drive_decrease")
+        {
[…]

### 2022-08-30 Code/bt_workspace/src/sensor_data_backup/include/sensor_data_backup/sensor_data_backup.hpp

@@ -9,6 +9,12 @@
 #include "geometry_msgs/msg/twist.hpp"
 // #include "bt_msgs/srv/pub_cmd_vel.hpp"
 #include "bt_msgs/srv/get_twist_array.hpp"
+#include "std_srvs/srv/empty.hpp"
+#include "geometry_msgs/msg/pose_with_covariance_stamped.hpp"
+#include "bt_msgs/srv/get_last_goal.hpp"
+#include "bt_msgs/srv/get_pose.hpp"
+#include "bt_msgs/srv/get_last_map.hpp"
+#include "bt_msgs/srv/send_updated_map.hpp"
 
 class SensorDataBackup : public rclcpp::Node
 {
@@ -25,26 +31,77 @@ class SensorDataBackup : public rclcpp::Node
                 const bt_msgs::srv::GetTwistArray_Request::SharedPtr request,
                 const bt_msgs::srv::GetTwistArray_Response::SharedPtr response);
 
+        void pub_last_goal_service_callback(
+            const std_srvs::srv::Empty_Request::SharedPtr request,
+            const std_srvs::srv::Empty_Response::SharedPtr response);
+
+        void get_last_goal_service_callback(
+            const bt_msgs::srv::GetLastGoal_Request::SharedPtr request,
+            const bt_msgs::srv::GetLastGoal_Response::SharedPtr response);
+        
+        void get_last_pose_service_callback(
+            const bt_msgs::srv::GetPose_Request::SharedPtr request,
+            const bt_msgs::srv::GetPose_Response::SharedPtr response);        
+
+        void get_collision_pose_service_callback(
+            const bt_msgs::srv::GetPose_Request::SharedPtr request,
+            const bt_msgs::srv::GetPose_Response::SharedPtr response);
+
+        void save_collision_pose_service_callback(
+            const std_srvs::srv::Empty_Request::SharedPtr request,
+            const std_srvs::srv::Empty_Response::SharedPtr response);
+
+        void get_last_map_service_callback(
+            const bt_msgs::srv::GetLastMap_Request::SharedPtr request,
+            const bt_msgs::srv::GetLastMap_Response::SharedPtr response);
+
+        void send_updated_map_callback(
[…]

### 2022-08-30 Reports/Week5/weekly_report5.tex
weekly report 5

\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{makecell}
\usepackage[margin=1in]{geometry}

\author{Lukas Evers}
\title{Weekly Report 3}

\renewcommand\theadalign{bl}
\renewcommand{\cellalign}{tc}
\renewcommand\theadfont{\bfseries}
\renewcommand\theadgape{\Gape[4pt]}
\renewcommand\cellgape{\Gape[4pt]}

\begin{document}

\maketitle
\newpage

\section{Scenarios}

\subsection{Scenarios with Subscenarios}

\begin{table}[!h]

	\caption{Sense-Plan-Act Scenarios}
	\label{tab:table2}
	\begin{tabular}{|c|l|l|}
		\hline
      	\thead{Scenario Name} & \thead{Sub Scenario} & \thead{Description} \\
      	\hline
      	LidarCrash_1 &  \makecell{Lidar Node dies while stationary} & \makecell{Crash of the Lidar node.} \\
      	\hline
      	LidarCrash_2 &  \makecell{Lidar Node dies while driving straight} & \makecell{Crash of the Lidar node.} \\
      	\hline
      	LidarCrash_3 &  \makecell{Lidar Node dies while driving around a corner} & \makecell{Crash of the Lidar node.} \\
     	\hline
      	OdomCrash_1 &  \makecell{Odometry Node dies while standing} & \makecell{Crash of the Odometry.} \\
      	\hline
      	OdomCrash_2 &  \makecell{Odometry Node dies while driving} & \makecell{Crash of the Odometry.} \\
      	\hline
      	IMUCrash_1 &  \makecell{Odometry Node dies while standing} & \makecell{Crash of the IMU Driver.} \\
      	\hline
      	IMUCrash_2 &  \makecell{Odometry Node dies while driving} & \makecell{Crash of the IMU Driver.} \\
      	\hline
      	NoPathFound_1 & \makecell{No valid global path can be found}  & \makecell{Path can not be calculated.} \\
		\hline      	
      	NoPathFound_2 & \makecell{Global Planner Node crashed.}  & \makecell{No response from Path Planner.} \\
      	\hline
      	ForcedCollision_1 & \makecell{Obstacle crashes into robot}  & \makecell{The robot is following its path and something else crashes into the robot.} \\
      	\hline
      	ForcedCollision_2 & \makecell{The robot crashes into undetected obstacle}  & \makecell{The robot is following its path crashes into an .} \\
      	\hline 
      	BatteryLow_1 & \makecell{Battery Voltage Low} & \makecell{The remaining battery is too low to reach the destination.} \\
      	\hline
      	MotorFailure_1 & \makecell{The motor driver fails.} & \makecell{The motor has a hardware failure.} \\
      	\hline
      	NetworkFailure_1 & \makecell{The robot can not receive any commands.} & \makecell{The method to transmit the motor commands crashes.} \\
      	\hline

	\end{tabular}
\end{table}

\newpage

\section{Next Week}

\end{document}

### 2022-09-06 Thesis/text/diskussion.tex
Upload New File

\chapter{Summary}
\label{cha:diskussion}

\section{Conclusion}

\section{Prospect}

### 2022-09-06 Thesis/text/implementierung.tex
Upload New File

\chapter{Implementation}
\label{cha:implementierung}

\section{System Architecture}

\section{Behavior Tree Structure}

### 2022-09-06 Thesis/text/verwandt.tex
Upload New File

\chapter{State of the Art}
\label{cha:state of the art}

\section{Automated and Autonomous Vehicle}

\section{Autonomous Driving Architectures}

\section{Autonomous Robot Navigation}

\section{Behavior Planning Approaches}

FSM, POMDP, BT

\section{Behavior Hierarchy}

Reactive Sequential and Deliberative Behaviors

### 2022-09-06 Code/bt_workspace/src/bt/include/update_map_after_collision.hpp
map for testing change

@@ -53,6 +53,8 @@ class UpdateMapAfterCollision : public BT::SyncActionNode
         if(rclcpp::spin_until_future_complete(node_, result) == rclcpp::FutureReturnCode::SUCCESS)
         {
             map = result.get()->map;
+            if(debug)
+                RCLCPP_INFO(node_->get_logger(), "Received map with dimensions: width: %d and height: %d", map.info.width, map.info.height);
             return true;
         }
         else
@@ -131,15 +133,15 @@ class UpdateMapAfterCollision : public BT::SyncActionNode
     {
         if(debug)
             RCLCPP_INFO(node_->get_logger(), "Converting to vector 2d.");
-        for (size_t i = 0; i < map.info.height -1; i++)
+        for (size_t i = 0; i < map.info.width-1; i++)
         {   
             // if(debug)
                 // RCLCPP_INFO(node_->get_logger(), "i: %d", i);
-            for (size_t j = 0; j < map.info.width -1; j++)
+            for (size_t j = 0; j < map.info.height -1; j++)
             {   
                 // if(debug)
                     // RCLCPP_INFO(node_->get_logger(), "j: %d", j);
-                vec_2d.at(i).at(j) = map.data.at(i * map.info.width + j);
+                vec_2d.at(i).at(j) = map.data.at(i * map.info.height + j);
             }            
         }       
     }
@@ -196,6 +198,9 @@ class UpdateMapAfterCollision : public BT::SyncActionNode
         goal_point.y = (sin(yaw)*ROBOT_RADIUS) + pose.position.y;
         goal_point.z = 0.0;
 
+        if(debug)
+            RCLCPP_INFO(node_->get_logger(), "Calculated Collision Point: %f | %f ", goal_point.x, goal_point.y); 
+
         return goal_point;
     }
 
@@ -209,7 +214,7 @@ class UpdateMapAfterCollision : public BT::SyncActionNode
         {
             for (size_t j = y_px -2; j < y_px +2; j++)
             {
-                costmap_2d.at(j).at(i) = 100; 
+                costmap_2d.at(i).at(j) = 100; 
             }
             
         }
@@ -225,12 +230,37 @@ class UpdateMapAfterCollision : public BT::SyncActionNode
             RCLCPP_INFO(node_->get_logger(), "Converting to one dimensional data array and putting it into the map msg.");
         std::vector<int8_t> data((map.info.height * map.info.width), -1);
 
[…]

### 2022-09-06 Thesis/text/einleitung.tex

\chapter{Introduction}
\label{cha:introduction}

Core Problem - too little autonomy in a standard autonomous mobile robot running ROS navigation(AMR)

Description of the Problem

Robots regurlarly needs operator in case smth unexpected happens\\

this may include: unexpected collisions, a system restart is needed, battery is empty\\

The system is not able to react and handle unforeseen situations\\

This can lead to safety issues during the operation of a robot. These safety issues can lead to uncontrolled driving maneuvers due to wrong sensor data or a wrong environment represantation.\\

Due to the need of the robot to require an operator, the robot is closer to an automated instead of an autonomous vehicle. 
Difference is defined by the SAE Levels for automonous driving in cars.\\

It is important for the robot to possess the ability to navigate in an uncertain environment to guarantee the safety of the robot itself and all other actors in its environment, these can be other robots and humans.\\

In theory the robot can perform fully autonomous navigation including mapping and localization but as soon as something out of the ordinary happens the robots autonomy is not guaranteed to be reliable anymore.
Default robot is not able to represent all possible fail states and does not detect failures on a systematic level, but only inside its navigation related subsystem. And even when failures inside this subsystems are discovered, the options to handle these problems are very limited and often do not deal with the problems in an autonomous way. The reliance on human-needed problem solving decreases the robots usefulness when tasked with real goals.

Having a robust and safe robot behaviour opens the door to many applications in a more diverse set of challenging environments. Usually a reinforcement learning approach with broad training data set is a great way to handle a multitude of unknown environments and situations, but one of the biggest problems is the lacking determinism in such systems.

In the robotic world exists a need for a deterministic system which can enhance and override the decision making process and navigation capabilities of the robot.

This thesis proposes a comprehensive approach for behavior planning for an autonomous mobile robot.

\textbf{How can behavior planning increase the safety and autonomy of an autonomous mobile robot running ROS?}

### 2022-09-06 Thesis/text/ergebnisse.tex

\chapter{Evaluation}
\label{cha:ergebnisse}

\section{Scenarios}

### 2022-09-06 Thesis/text/methoden.tex

\chapter{Methodology}
\label{cha:methoden}

\section{Requirements}

\section{Libraries}

### 2022-09-09 Thesis/text/verwandt.tex
state of the art

@@
 
-\section{Automated and Autonomous Vehicle}
+\section{Automated and Autonomous Vehicles}
 
-\section{Autonomous Driving Architectures}
+This thesis looks for a way to improve the autonomy of mobile robots.
+In order to define what qualifies as an improvement in the autonomous behaviour of a mobile robot, one needs to look at the defintion of automated and autonomy.
+The EFI defines the term autonomy in the context of robotics as a system which can act without human instructions and still solve complex tasks, make decisions, learn independently aswell as react to unforeseen circumstances \cite{efi2018}.
+The definition specifies the needed requirements to make a robot fully autonomous. 
+The automotive industry defines vehicle autonomy in levels based on the human driver's need for supervision and possible intervention during the execution of complex driving tasks \cite{J3016_202104}.
+Figure xx depicts the different levels of autonomy of passenger cars on a scale between no automation to completely autonomous.
+\\
 
-\section{Autonomous Robot Navigation}
+\begin{figure}
+	\includegraphics[width=1.0\textwidth]{images/j3016graphic_2021.png}
+	\caption{SAE Levels Of Driving Automation \cite{J3016_202104}  }
+\end{figure}
+
+Both definitions entail the reliance on human supervision and guidance as the central part of what hinders a system or vehicle to become autonomous.
+By decreasing the instances a human intervention during the operation of a robot, one can increase the automation level towards full autonomy.
+But this has to be done by equipping the robot with robust and context-driven decision-making and planning capabilities, otherwise a robot which would never need a human to operate could easily be created.
+But this robot could not be considered autonomous as it would not able to solve complex tasks and make intelligent decisions.
+
+Different levels of autonomy require different methodologies to achieve them.
+
+SAE Levels
+EFI Gutachten
+Fachforum
+
+\section{Autonomous Driving Navigation Architectures}
+
+To gain a better understanding how behavior planning influences the autonomy of a robot, one can take a look at the latest and most used software architectures for autonomous driving on a functional level.
+Most of the current autonomous driving architectures are implementing a hierarchical structure that follows the "Sense - Think - Act" paradigm \cite{murphy2000}.
+As shown in figure xx the sensory inputs, usually from multiple sensors, get processed to create an environment representation.
+Then the system calculates how to get from its current position to the destination and creates a path.
+A second planning calculation is triggered to compute the motion commands with the constraints of the system (e.g.
+size, turning radius, etc.).
+These motion commands then get converted to control the motors.
+The planning sequence can be further divided into global planning, behavior planning and local planning.
+Global planning,also named route planning, is responsible for outputting a path from start to goal.
+This is comparable to a person using Google Maps to find the shortest or fastest path to a far away city.
+In this analogy, the behavior planner is responsible to follow the traffic rules on the trip to the destination, e.g.
+stopping at red lights, giving right of way to other vehicles, following the speed limit.
+The local planning, also named motion planning, module takes the input from the behavior layer and has the task to let the vehicle drive in accordance to the executed behavior, so that it stays in its lane and comes to a stop at a red light.
+\\
+\begin{figure}
+	\includegraphics[width=1.0\textwidth]{images/autonomous_driving_architecture.png}
+	\caption{Common Autonomous Driving Architecture \cite{1087032} \cite{9266268}  }
+\end{figure}
+ \\
+
+Due to safety considerations, many autonomous driving architectures implement an additional system supervision layer which monitors the execution of the other system components.
+This supervisor checks the health of all software components.
+In case one or more components fail to pass the health check the supervisor is responsible for ensuring that the system does not continue the normal driving routine.
+The supervisor triggers measures to restore the normal functionality or if that is not possible bring the robot into a safe state (zimmermann2020).
+
+Considering that with these architectures, vehicles can achieve an SAE level of autonomy up to level 4 (bacha 2008),  a good starting point to increase autonomy in a robotic system would be to ensure that all of the functional modules in figure xx are implemented and available.
+On this basis higher levels of autonomy are achievable through the expansion of the behavior planning module inside of the planning layer.
+
+bacha2008
+reke 2020
+wei2014
+brooks1986
+dhillon 2002
+gonzales2016
+
+\section{Behavior Types}
+
+The behavior planning module contains a set of behaviors to operate in a environment.
+The task of the module is then to choose the best behavior to execute.
+In this sense a behavior is defined as a mapping of sensory inputs to a pattern of actions that in turn carry out a specific task \cite{murphy2000}.
+A good behavior planning approach provides the robot with adequate behaviors for every possible scenario the robot is operating in.
+Different scenarios pose different challenges for the behavior planning module which has to decide on the best behavior option to execute in order to achieve a higher level goal.
+
+\subsection{Reactive}
+During the emergence of behavior based robots, the first type of behavior that was implemented in many robots was a simple reactive behavior.
+This behavior maps a sensory input directly to motor commands.
+In human behavior this type of behavior resembles reflexes, like the tapping on the kneecap which unwillingly results in motion in the knee joint.
+This behavior pattern is not following the typical "Sense, Think, Act" loop, but shortcuts directly from sensing to acting \cite{desilva2008}.
+Reactive behaviors can be chained together to result in more advanced and goal-driven behaviors.
+Despite the possibility of more complex robot behaviors, sequential behaviors remain on the level of reactive behaviors due to the lack of a planning and decision-making cycle during their execution.
+The computational load of reactive behaviours is low and thereby fast as no decision and planning process is taking place, which makes them suited in scenarios where real-time safety is of concern.
+This property makes these kind of behaviours useful for system supervisor, where sometimes immediate reactions with low latency are required to ensure vehicle safety.
+But a system with a reactive behavior planning approach will always fail to meet the requirements for higher levels of autonomy due to the fact.
+This does not mean that reactive behaviors can not be part in highly autonomous systems as their quick response time to sensor inputs makes them valuable in improving vehicle safety and reliability.
+
+\subsection{Deliberative}
+Reactive behaviors suffer the drawback that they are not suitable enable complex behaviors.
+In addition, the creation of sequences of reactive behaviors that produce the target behavior is a complicated task \cite{murphy2000}.
+Behaviors that involve a planning and decision making aspect are defined as deliberative.
+Unlike reactive behaviors, deliberative behaviors are following the Sense, Think, Act paradigm.
+They are not mapping sensory inputs directly into motor commands.
+Instead deliberative-type behaviors are deciding the best course of action before acting and only then proceed with the exution.
+The use of deliberative behaviors in the planning module is what enables a robot to meet the defintion of autonomy, e.g.
+making decisions and react to unforeseen circumstances.
+So in order to improve the autonomy of a robot, the focus should be on the creation of deliberative behaviors.
+A behaviour planner with a deliberative approach makes decision based on current sensor data, as well as previously processed data.
+By incorporating older sensor information into the decision making process, the deliberative behavior planner can make predictions about the changes that will happen in the environment.
+This allows the planner to proactively change the motion planning commands to better adapt to the environment.
+For example, a motion prediction of obstacles in highly dynamic environments would improve the motion planning considerably.
+This is due to a better understanding of how obstacles are interacting with the planned path of the robot in relation to the time at which the robot and obstacle are predicted to intersect paths.
+A purely reactive planner could never determine if an obstacle is destined to intersect the robots path in the future and would reroute constantly in order to accomplish the goal.
+A deliberative planner could determine if the best course of action is to stay on the current path as the dynamic obstacle has already moved away by the time the robot reaches to intersection point.
+Other actions in scenario are possibly slowing down or speeding up briefly to avoid an obstacle and staying on the current path as this is calculated to be quicker than rerouting and taking a longer path.
+Generally, these cognitively more complex behaviours mimic human skills and thus make a system more autonomous.
+
+\textit{Talk about machine learning as deliberative behavior -> true autonomy bc learning and unknown scenarios, programmers can not cover every scenario, so a well designed/trained ai is needed to make a system really autonomous, but -> missing determinism and unaccountability of machines, hence we need behaviors that are still hard coded and deterministic, these can be highly automated.
+traditional behaviour  and ai generated behaviours can and must coexist in the system (adler2019)}
+
+silva2008
+ingrand 2014
+vasiolopolous 2022
+
+\subsection{Hybrid}
+To create reliable aswell as intelligent systems one can combine reactive and deliberative behaviours in a hybrid model.
+This behavior planning module is then able to quickly react to incoming sensor data and still exhibit high levels of automation.
+\\
+
+\begin{figure}
+	\includegraphics[width=1.0\textwidth]{images/Deliberative_hierarchical_planning.png} 
+	\caption{Hybrid Planning Architecture \cite{arkin1998}  }
+\end{figure}
+
+Figure xx shows the areas where reactive and deliberative behaviours have their advantages and limitations.
+Deliberative planning does possess limitations during the execution of generated plans as the time horizon is long-term and not able to react to fast changes in the environment .
+When the system is not adressing these points, deliberative robots are very focussed on a small problem domain and are not robust \cite{arkin1998}.
+The incorporation of reactive behaviors into the behavior planning combats this problem.
+In this multi-layer approach the higher level, more deliberative planners are able to override the reactive planners in order to allow the system to be more flexible and adaptible but still maintain fast reaction times.
+This architecture leads to more robust robots that are able to be deployed in a wider variety of environments, thus leading to higher levels of autonomy.
 
@@
 
-FSM, POMDP, BT
+Robots that use a hybrid, hierarchical behavior planning approach need a system which decides on a high level which behaviors are to be executed.
+This allows overriding simple reactive behaviors with more complex, deliberative behaviors when the system benefits from it.
 
-\section{Behavior Hierarchy}
+The different solution for such behavior execution system all share the fact that they make use of states a robot can be in.
+Based upon a robots state and additional information about the environment, different strategies and behaviors are executed during the runtime of the system.
 
-Reactive Sequential and Deliberative Behaviors
+\subsection{Finite State Machines}
+Finite State Machines are commonly used in the environment of autonomous driving.
+A finite state machine is defined by a list of possible states S, a starting state s0, a set of state transitions delta, a set of final states F and the input alphabet sigma.
+\\
+
+\begin{figure}
+	\includegraphics[width=1.0\textwidth]{images/fsm.png} 
+	\caption{Example of a finite state machine transition diagram \cite{wagner2006} }
+\end{figure}
+
+allgeuer 2013
+connor 2017
+ziegler 2014
+
+\subsection{POMDPs}
+
+\subsection{Behavior Trees}
+marzinotto 2014
+colledanchise 2018
+
+\section{Deterministic Behavior}

### 2022-09-09 Thesis/text/einleitung.tex
state of the art

@@
 
+Research question:
+\textbf{How can behavior planning increase the robustness and autonomy of a robot running ROS2?}
+
 Core Problem - too little autonomy in a standard autonomous mobile robot running ROS navigation(AMR)
 
-Description of the Problem
+Description of the Problem:
 
-Robots regurlarly needs operator in case smth unexpected happens\\
+Despite the advancements in autonomous driving the robot regurlarly needs an operator in case smth unexpected happens.
 
-this may include: unexpected collisions, a system restart is needed, battery is empty\\
+This may include: unexpected collisions, a system restart is needed, battery is empty, the environment is too crowded and the robot gets stuck
 
-The system is not able to react and handle unforeseen situations\\
+The system is not able to react and handle unforeseen situations.
 
-This can lead to safety issues during the operation of a robot. These safety issues can lead to uncontrolled driving maneuvers due to wrong sensor data or a wrong environment represantation.\\
+This can lead to safety issues during the operation of a robot. These safety issues can lead to uncontrolled driving maneuvers due to wrong sensor data or a wrong environment represantation.
 
-Due to the need of the robot to require an operator, the robot is closer to an automated instead of an autonomous vehicle. 
-Difference is defined by the SAE Levels for automonous driving in cars.\\
+Due to the need of the robot to require an operator for safe, the robot is not really autonomous despite it being labeled as such.
 
+(Difference is defined by the SAE Levels for automonous driving in cars.) \\
+
+Relevance, need for research:\\
 It is important for the robot to possess the ability to navigate in an uncertain environment to guarantee the safety of the robot itself and all other actors in its environment, these can be other robots and humans.\\
@@
 
-This thesis proposes a comprehensive approach for behavior planning for an autonomous mobile robot.
+Planned solution:\\
 
-\textbf{How can behavior planning increase the safety and autonomy of an autonomous mobile robot running ROS?}
+This thesis will explore an approach how an extension of the current software architecture and behavior planning capabilities for an autonomous mobile robot that is running ROS2 can improve the autonomy of the robot.
+
+Expected results:\\
+
+The desired outcomes of these implemented extension will be an increase of robustness and decrease of required human actions during a number of scenarios.
+In these scenarios, failures and and problems are artificially induced to test the robots abilities to behave autonomously.
+
+\textit{Research gap is not named so far}

### 2022-09-09 Code/bt_workspace/src/bt_tests/src/scenario1.cpp

@@ -1,0 +1,10 @@
+#include <cstdio>
+
+int main(int argc, char ** argv)
+{
+  (void) argc;
+  (void) argv;
+
+  printf("hello world bt_tests package\n");
+  return 0;
+}

### 2022-09-09 Thesis/text/verwandt.tex
state of the art thesis, finished but behavior tree subsection

@@
 stopping at red lights, giving right of way to other vehicles, following the speed limit.
-The local planning, also named motion planning, module takes the input from the behavior layer and has the task to let the vehicle drive in accordance to the executed behavior, so that it stays in its lane and comes to a stop at a red light.
+The local planning, also named motion planning, module takes the input from the behavior layer and has the task to let the vehicle drive in accordance to the executed behavior, so that it stays in its lane and comes to a stop at a red light \cite{reke2020}.
 \\
@@
 	\includegraphics[width=1.0\textwidth]{images/autonomous_driving_architecture.png}
-	\caption{Common Autonomous Driving Architecture \cite{1087032} \cite{9266268}  }
+	\caption{Common Autonomous Driving Architecture \cite{brooks1986} \cite{velasco2020}  }
 \end{figure}
@@
 \subsection{Finite State Machines}
+
 Finite State Machines are commonly used in the environment of autonomous driving.
+They are used to describe behavior in a form of states, which in turn execute action.
 A finite state machine is defined by a list of possible states S, a starting state s0, a set of state transitions delta, a set of final states F and the input alphabet sigma.
+At any given time only a single state is selected and its containing actions are executed.
 \\
@@
 
+Figure xx depicts an example state machine with different states and examplatory transitions between them.
+Inside of the states the letters indicate the existence of actions.
+E stands for entry, I for Input and X for Exit.
+The arrows between the states illustrate the possible transition and transition conditions for every state.
+A state transition diagram is a good way to understand the system behavior but can not show the details of the modedeled behavior, like the actions \cite{wagner2006}.
+\\
+
+Creating small state machines can easily and quickly be accomplished through the use of various feature-rich and performant libraries \cite{fourakis2014}. 
+Finite state machines are often being used for comprehensive modeling of reactive behavior and sequential behaviors in autonomous driving functions \cite{ziegler2014}.
+These finite state machines make use of nested state machines inside of another state machine.
+This reduces the complexity of the state machine as the transition do not need to be modeled, because the start and final state of a nested state machine is treated as just one state inside of the larger state machine.
+Despite this possibility to simplify state machines, as a modeled behavior system grows larger to accomodate more complex behaviors into the state machine the effort to expand and maintain the state machine grows rapidly.
+This is due to the modeling effort of the transitions, transition conditions and transition events growing with each new state that gets introduced into the state machine as existing states need to be checked and updated accordingly \cite{conner2017}.
+\\
+
+The execution time of finite state machines can be fast enough that they are used to control the motors of a bipedal robot to enable the robot to balance and walk despite unexpected height variations between steps \cite{park2013}.
+This property makes state machines suited very well suited for fast, reactive type behaviors and still allows a hierarchical approach to behavior planning
+
 allgeuer 2013
@@
 ziegler 2014
-
-\subsection{POMDPs}
 
@@
 
-\section{Deterministic Behavior}
+% \section{Deterministic Behavior}

### 2022-09-16 Thesis/text/methoden.tex
thesis update_ chapter 3

@@
-\chapter{Methodology}
+\chapter{Concept}
 \label{cha:methoden}
+
+\section{Current State}
+This chapter briefly presents the current state of mobile navigation with ROS and its limitations. 
+The goal of this chapter is to then identify the key areas in which standard mobile robots running ROS lack robustness and autonomy during autonomous mobile navigation.
+According to these key areas requirements are derived from them and prioritized with a risk analysis.
+Furthermore this chapter proposes a way how behavior planning can improve the current systems to meet the requirements.
+
+\subsection{Turtlebot3}
+A standard mobile research plattform that has an available ROS interface to control the robot is the Turtlebot3.
+The robot is well integrated in the ROS ecosystem and is established in the literature as a system to develop and integrate new methods on.
+The robot has two motors with attached wheel encoders to drive and steer, so it classifies as an differential drive robot. A third omnidirectional wheel is added for stability.
+
+The manufacturer provides an open-source model for simulating the robot in physics-based simulators like Gazebo, as depicted in figure xx, which enables faster development and testing for the robot. 
+The equipped sensors and easy simulation capabilities make the robot very well suited for the use of the Nav2 stack and further development for behavior planning.
+
+The robots specifications are specified in table xx.
+
+\begin{figure}[h!]
+	\includegraphics[width=1.0\textwidth]{images/turtlebot_sim.png}
+	\caption{Turtlebot3 model in Gazebo simulator}
+\end{figure}
+
+The turtlebot will be used for the analysis of the current state of mobile navigation.
+Furthermore, the simulated turtlebot will be the plattform on which the behavior planning will be implemented and tested.
+
+\subsection{ROS?!}
+\textit{write sth about that?}
+
+\subsection{Navigation2}
+The ROS Navigation2 Stack (Nav2) is a combination of different packages that allows mobile robots to navigate from point A to point B.
+Navigation2 is the de-facto standard to achieve mobile navigation with a wide-range of supported robots.
+Supported robot types are holonomic, differential-drive, legged and ackerman (car-like).
+For the mobile robot to make use of the Nav2 stack, it has to be set up in a certain way to be able to generate plans and execute commands in the right way.
+Nav2 requires the mobile robot to possess a laserscan or pointcloud sensor, an odometry source (such as an IMU, wheel encoders) and a set of transformations to be available for planning and navigation.
+The stack contains tools that save and load occupancy maps, localize the robot, plan paths, execute the path, provide costmaps, build behaviors and execute recoveries \cite{macenski2020}.
+
+\begin{figure}[h!]
+	\includegraphics[width=1.0\textwidth]{images/nav2_architecture.png}
+	\caption{Navigation2 Architecture \cite{macenski2020}}
+\end{figure}
+These packages are often combined with a SLAM method to build or enlarge maps.
+The path planning and path execution capabilities, which correspond to the global and local planner as described in section 2.2, are further supported by ready-to-use planning plugins which use A* and Dijkstra algorithms for path planning and a dynamic window approach for path execution.
+The system sequencing is done in a hierarchical structure as there is central behavior tree which calls asynchronous actions from the respective planners after another as seen in figure xx.
+
+\begin{figure} [h!]
+	\includegraphics[width=1.0\textwidth]{images/nav_bt-modified.png}
+	\caption{Navigation2 Behavior Tree}
+\end{figure}
+
+The behavior tree depicted in figure xx is the one that is the default Nav2 behvaior and has a set of recovery behaviors already implemented (spin, wait, backup, clear map).
+These behaviors are executed if the robot can not make progess towards the set goal.
+Nav2 makes use of a node lifecycle management system which is used to start, activate, deactivate and shutdown nodes in a controlled way.
+This allows the system degree to monitor the system's nodes startups, executions and failures.
+
+With this quantity of functions and hierarchical architecture the ROS Navigation2 stack is a great basis to achieve high levels of autonomy in robots.
+Due to the expandability of the software with plugins the software can be modified and expanded to fit the users needs.
+When comparing the functionalities of the software stack to the proposed autonomous system architecture in figure xx (2.2), the things that Nav2 are missing are a system-wide supervision layer and and a more deliberative approach to behavior planning as mentioned in 2.3.
+The absence of these components leads to a limited set of circumstances in which the robot can perform navigation reliably.
+One of the core problems that behavior-based robots face is to generate a dependable environment representation as this is the basis for all of the latter steps in the planning part of the "Sense - Think - Act" cycle.
+Unforeseen events can severely limit the reliability of the environment representation without the robot being aware of it.
+Examples for such types of events that are currently not handled appropriatly are slipping wheels, orientation changes by external influences or undetected, more specifically undetectable obstacles.
+\textit{Add statistics about random sensor failures every ~10000h} This can lead for example to erratic movement of the robot, as reactive behaviors and more deliberative behaviors compete for authority over the robots movement.
+By design, in order to allow more intelligent robot behaviors, the deliberative behavior can override commands from reactive behaviors.
+But the deliberative behavior's planning is based upon a wrong representation of reality which then leads to unsafe commands which in turn triggers reactive-type behaviors again to counteract the commands from the deliberative behavior.
+Possibly the reactive behaviors do not get activated at all and the robot is continuing to drive despite being unaware of the surroundings.
+This would be a situation in which a human operator would need to step in and restore the robot to full functionality by hand.
+
+\textit{Add a closing statement about that suitability of the Nav stack for achieving autonomy}
+
+\section{Risk Analysis For Requirement Prioritization}
+
+\textit{Do that?}
 
@@
 
-\section{Libraries}
+\textbf{How do I justify these requirements? Atm they are made arbitrarily with no real substance or reasoning.}
+
+Functional Requirement List
+
+number .. Name .. Description .. Success Criteria -.. Priority
+freq1 Recovery The system can recover from crashes Ensure that the system can successfully reach goals despite previous crashes
+freq2 Sensor Failure	The system detects sensor failures	Ensure that the system can restart sensors and decrease the speed during time sensor delivers limited information
+freq3	Maintain operability The robot executes commands as long as it is safe	Ensure that the robot keeps driving if it is safe even when system functions are not working correctly
+freq5	Emergency Detection	The system detects emergency. Ensure that the system can detect when the continuation on the calculated path is no longer safe (sensor failures, blockage)
+freq6	Emergency Stop The system can initiate emergeny stops	Ensure that the system can override all commands and stop in case an emergency is detected
+freq7	Override Navigation2 The system can override navigation2	Ensure that the system's commands can always override the commands coming from navigation2.
+freq8 Reset Goals The system can reset and override goals set by the user so that the goal is reachable by planners.
+freq9 Determinism The outcome for a given set of inputs has to be deterministic, meaning that the behavior is always executed in the same way.
+
+\begin{table}[h!]
+\caption{Functional Requirements}
+	\begin{tabular}{ | m{0.05\textwidth} | m{0.13\textwidth}| m{0.07\textwidth} | m{0.65\textwidth} |} 
+  	\hline
+  	Nr. & Name & Priority & Description \\ 
+  	\hline
+  	freq1 & Recovery & High &  The system can recover from crashes. Ensure that the system can successfully reach goals despite previous crashes\\ 
+  	\hline
+  	freq2 & Sensor Failure & High & The system detects sensor failure.Ensure that the system can restart sensors and decrease the speed during time sensor delivers limited information \\ 
+  	\hline
+  	freq3 & Maintain operability & High & The robot executes commands as long as it is safe	Ensure that the robot keeps driving if it is safe even when system functions are not working correctly \\
+  	\hline
+  	freq4 & Emergency Detection & High & The system detects emergency. Ensure that the system can detect when the continuation on the calculated path is no longer safe (sensor failures, blockage) \\
+  	\hline
+  	freq5 & Emergency Stop & High & The system can initiate emergeny stops. Ensure that the system can override all commands and stop in case an emergency is detected \\
+  	\hline
+  	freq6 & Override Navigation2 & High & The system can override navigation2. Ensure that the system's commands can always override the commands coming from navigation2. \\
+  	\hline
+  	freq7 & Reset Goals & Medium & The system can reset and override goals set by the user so that the goal is reachable by planners. \\
+  	\hline
+  	freq8 & Determinism & High & The outcome for a given set of inputs has to be deterministic, meaning that the behavior is always executed in the same way. \\
+  	\hline
+
+	\end{tabular}
+\end{table}
+Non-functional requirement list
+
+number .. 	Name .. 				Description ..										Success Criteria
+nfreq1 Single Point of Failure	The system does not have a single point of failure	When parts of the system fail the system maintains operability to a degree
+nfreq2 Performance	The system's control loop guarantees fast reactions	The average frequence with which the system checks the sensors/navigation2 is higher than 100Hz (10ms)
+nfreq3
+
+\begin{table}[h!]
+\caption{Non-Functional Requirements}
+	\begin{tabular}{ | m{0.05\textwidth} | m{0.13\textwidth}| m{0.07\textwidth} | m{0.65\textwidth} |} 
+  	\hline
+  	Nr. & Name & Priority & Description \\ 
+  	\hline
+  	nfreq1 & Single Point of Failure & High &  The system does not have a single point of failure	When parts of the system fail the system maintains operability to a degree \\ 
+  	\hline
+  	nfreq2 & Performance & High & The system's control loop guarantees fast reactions. The average frequence with which the system checks the sensors/navigation2 is higher than 100Hz (10ms) \\ 
+  	\hline
+  	nfreq3 & Determinism & High & The outcome for a given set of inputs has to be deterministic, meaning that the behavior is always executed in the same way. \\
+  	\hline
+  	nfreq4 & Deliberate & High & The robot is able to execute deliberative behaviors, meaning that behaviors new implemented behaviors go beyond reacting to sensor input and have a planning aspect. \\
+	\end{tabular}
+\end{table}
+
+\section{System Solution}
+
+In order for a Turtlebot3 running Nav2 to appropriatly deal with the listed requirements, the system needs to be extended with new external modules.
+To eliminate a single point of failure for the system one needs to create an independent second system outside of the existing one.
+A suitable name for the second system is autonomy layer as it aims to increase the robustness and autonomy of the whole system.
+The autonomy layer provides a holistic system supervisor which monitors the health of all components (Sensors, Navigation, Robot Controller).
+This equips the system with the ability to deal with the event of node failures and component malfunctions.
+Another required addition to the autonomy layer is a data store of relevant system information.
+This not only includes sensor data, but also data like generated maps, costmaps, positions and speed commands.
+Making use of past datapoints allows behaviors to become more deliberative in their approach and opens up many possibilities for smart behaviors like movement predictions of obstacles.
+A third component to add to the autonomy layer is a behavior planner that is largely independent from the Nav2 execution and planning. This behavior planner
+
+\section{Software}
+
+\subsection{Behaviortree.CPP}
+The Navigation2 stack uses a behavior tree to coordinate the planning layer.
+The behavior tree that is used is an expansion of the Behaviortree.CPP (BT.CPP) library.
+This library offers a fast way to create, execute, monitor and edit behavior trees.
+Additionally to the types of control nodes that are mentioned in chapter 2.4.2 the library adds the concept of reactivity into the catalogue of control nodes.
+Reactive sequences and reactive fallbacks differ from normal ones in the handling of nodes that return the state "running".
+Instead of ticking the node again, the whole sequence restarts, which is very useful for continously checking a condition and executing an asynchronous action node, that gets halted when a condition is not met anymore.
+
+Also, it offers Decorator Nodes which allow more control over the child node and its output.
+These Decorator Nodes can only have one child node each.
+A comprehensive list of the available nodes and descriptions is in table xx.
+
+\textit{TABLE HERE}
+
+As a way to allow the tree nodes to communicate with each other, the library provides the developer with two possibilities.
+Either node can make use of the blackboard which is a dictionary (key/value) that all nodes of the tree can read and write.
+Or two nodes can be connected through ports which allows the direct communication between two nodes via a key/value.

### 2022-09-16 Thesis/text/verwandt.tex
thesis update_ chapter 3

@@
 
-Finite State Machines are commonly used in the environment of autonomous driving.
+Finite State Machines (FSM) are commonly used in the environment of autonomous driving.
 They are used to describe behavior in a form of states, which in turn execute action.
@@
 The execution time of finite state machines can be fast enough that they are used to control the motors of a bipedal robot to enable the robot to balance and walk despite unexpected height variations between steps \cite{park2013}.
-This property makes state machines suited very well suited for fast, reactive type behaviors and still allows a hierarchical approach to behavior planning
+This property makes state machines suited very well suited for fast, reactive type behaviors and still allows a hierarchical approach to behavior planning.
 
@@
 \subsection{Behavior Trees}
+Behavior Trees (BT) are another way to model and control the behavior of autonomous systems.
+Behavior Trees first found big acceptance in the computer game industry where they are mainly used to model artificial intelligence for non-player characters \cite{florez2009}.
+Every tree consists of one root and many child, parent and leaf nodes.
+Leaf nodes are also called execution nodes, while non-leaf nodes are called control or control flow nodes.
+The execution of a behavior tree is done by ticking the root node of tree.
+This signal then travels down to the child node of the root node, where either control nodes are being ticked or action nodes are executed.
+Nodes return either "Success", "Running" or "Failure" which influences how the tick signal gets processed by the rest of the behavior tree.
+Control nodes can be of the type "Sequence", "Fallback" or "Condition".
+The condition node can not have child nodes and can only return success or failure.
+A sequence node can be compared to a logical "and" condition.
+This means that once a sequence gets ticked from the parent node, it will send the tick signal to every child node as long as they return "success" of "running", and only return "success" itself when every child node was successfully ticked.
+If any of the child nodes return "failure", the sequence node will stop ticking the remaining children and return "failure".
+On the other hand the fallback node is an equivalent of a logical "or" condition.
+The fallback node will tick its child nodes as long as they return "failure" or "running" and will return "success" if one of the ticked node returned "success".
+It will only return "failure" if all of the child nodes "returned success".
+Figure xx depicts an example of a BT with multiple levels and action nodes.
+The nodes with the arrow (->) symbol are sequence nodes, the question mark (?) symbol denotes fallback nodes.
+The condition node are depicted as white ovals.
+The action nodes are represented as gray rectangles in the BT example.
+
+\begin{figure}
+	\includegraphics[width=1.0\textwidth]{images/bt_example.jpg} 
+	\caption{Example of a behavior tree \cite{iovino2022} }
+\end{figure}
+
+One of the core differences to Finite State Machine is that transitions between states are not distributed across all the states but they are organized in a hierarchical tree, where the leaves are representing the states \cite{iovino2022}.
+While both FSM and BT are capable of producing the same behavior in robots, the fundamental shift in how the two system are created lead to significant advantages in the modularity, synthesis and analysis of the systems at hand.
+These effects are more significant with an increase in size of the system.
+BT offer a lot more flexibility when creating an advanced behavior layer for an autonomous system.
+
+iovino2022
 marzinotto 2014

### 2022-09-18 Thesis/text/verwandt.tex
chapter 4.4 ready

@@
 It will only return "failure" if all of the child nodes "returned success".
+
+\begin{table}[h!]
+	\caption{The five types of nodes \cite{iovino2022}}
+	\begin{tabular}{ | m{0.13\textwidth} | m{0.13\textwidth}| m{0.2\textwidth} | m{0.2\textwidth} | m{0.2\textwidth} |} 
+  	\hline
+  	Node type & Symbol & Succeeds & Fails & Running \\ 
+  	\hline
+  	Sequence & - > & If all children succeed &  If one child fails & If one child returns running \\ 
+  	\hline
+  	Fallback & ? & If one child succeeds & If all children fail & If one child returns running \\ 
+  	\hline
+  	Parallel & - > - > & If >= M children succeed & If > N-M children fail & else \\
+  	\hline
+  	Action & shaded box & Upon completion & When impossible to complete & During completion \\
+  	\hline
+  	Condition & white oval & If true & If false & Never \\
+  	\hline
+	\end{tabular}
+\end{table}
+
 Figure xx depicts an example of a BT with multiple levels and action nodes.
@@
 
-\begin{figure}
+\begin{figure}[h!]
 	\includegraphics[width=1.0\textwidth]{images/bt_example.jpg} 
@@
 
+\subsection{POMPDs}
+
 % \section{Deterministic Behavior}

### 2022-09-18 Thesis/text/implementierung.tex
state of the art
thesis update_ chapter 3
chapter 4.4 ready

@@
 \section{System Architecture}
+\begin{figure}
+	\includegraphics[width=1.0\textwidth]{images/component_diagram_bt_architecture.png}
+	\caption{Component Diagram of System Architecture}
+\end{figure}
+
+\section{Data Backup}
+The data backup component functions as a way to efficiently gather data from all components of the system and bereitstellen necessary data to the behavior tree when the planning process requires the use of past data points.
+This component subscribes is a to all sensor data and saves the messages in a queue data structure.
+The most recent data gets pushed into the array and if the array exceeds a bestimmte length, the oldest data this data gets deleted from the array.
+The limited array size is twofold.
+On the on side, the goal is to decrease the size of the message that gets sent to the BT to limit the network auslastung and on the other side, the BT often does not need to look back further than a few seconds to make a decision.
+This way the computational load and network load is decreased.
+The BT can access the saved data by sending a service call to one of the provided services by this component.
+Service calls in ROS2 are executed with hihg priority and the execution is done quickly but it adds an additional step than just saving the data directly in the behavior tree.
+The reason for the outsourcing of the data backup in another component is that ROS node need an executor which spins the node constantly.
+The spinning allows the node to check the network if there is any work to do for the node.
+This means that if one of the subscribed topic received a message, the node has to execute one of its callback functions that is associated with that topic.
+This spinning has to be done in regular intervals, otherwise the node might miss incoming messages from subscribed topics.
+The way to do the spinning is to have the node spun constantly in one thread.
+Have a constant spinning node would block the execution of the BT completely.
+If one would try to spin a node once when a node gets ticked in the BT, the problem of missing messages would become relevant because the BT might need longer than exepected to return to a node to tick it.
+That way the BT node that would be responsible for saving all the relevant data can not guarantee that it has picked up on all of the messages between to spins.
+The synchronous nature of the ROS executor spinning the nodes is the reason why the data backup is not stored within the BT.
+
+The saved data inside of this component with the corresponding lookback time is listed in table xx.
+
+\begin{table}[h!]
+	\caption{Types of saved data}
+	\begin{tabular}{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
+  	\hline
+  	Name & Type & Lookback Time \\ 
+  	\hline
+  	Lidar & Laserscan & 3 seconds \\
+  	\hline
+  	Poses & Pose with Covariance & 2 seconds\\ 
+  	\hline
+  	Map & Occupancy Grid & Only saved last one \\ 
+  	\hline
+  	Collision Pose & Pose with Covariance & Only last one \\
+  	\hline
+  	Command Velocities & Twist & 2 seconds\\
+  	\hline  	
+  	Global Costmap & Occupancy Grid & Only last one \\
+  	\hline
+	\end{tabular}
+\end{table}
 
 \section{Behavior Tree Structure}
+
+A simplified behavior tree only depicting the top level behavior control nodes is shown in figure xx.
+The tree receives the inital tick signal from the root in an infinite loop.
+The root control node is a sequence node.
+We made the root a simple sequence because we want the whole system to function with increasing complexity in the later parts (on the right side) of the tree.
+The complex behaviors require that all of the previous components and conditions are working as they should.
+The root sequence will not allow the execution of higher level behaviors if the system is detecting problems on more fundamental levels.
+The root sequence is the first step for increasing the robustness of the system to combat competing reactive and deliberative behaviors which could endanger the safety of the robot and other actors in its environments.
+
+\subsection{BT System Supervision}
+
+The system supervisor component is partly outsourced into a component outside of the thread of the BT, because of the same reasons as the sensor backup component discussed in the previous section.
+The synchronous executor would stop the execution of the whole behavior tree, which is why the component needs to be moved inside of a new thread.
+The system supervisor, nachfolgend also called execution checker, is receiving messages from nodes that are not implemented as ROS2 lifecycle nodes.
+If a node stops sending messages or stops responding completely the external execution checker can inform the BT via a provided service of a problem with the node.
+ROS2 lifecycle nodes provide information about their health via an implementation of a FSM inside of the node.
+The current node state can be found out via a service call.
+However, to ensure the node is working correctly the output of the node needs to be checked too, because it might be that the FSM inside of the node did not pick up on a error.
+The system supervision is constantly checking the system's health. The BT is using service calls to get the health information about the compononts which are checked inside of Condition Nodes. 
+The components get checked by the BT inside of a Fallback.
+If the condition for the execution is met, the Condition Node returns "success" and the fallback would exit to move on to the next fallback and execution condition check (see fig.
+xx).
+In case the execution checker detects a node failure, a sequence to react to the sensor failure is executed in which the robot slows down to a minimum and tries to restart the sensor.
+Should the restart fail the robot will not be able to navigate in a safe manner anymore and will come to a complete stop.
+This behavior pattern is realized for the lidar, the IMU and the wheel odometry.
+
+\begin{figure}[h!]
+	\includegraphics[width=1.0\textwidth]{images/sensor_fallback.png}
+	\caption{Lidar Fallback}
+\end{figure}
+
+The timeout period for the execution checker to declare a node failed after no message was received is set at one second. 
+After that period the goal is to enable to robot to continue to make progress towards the goal but if the system can not restart the sensor in a short time after failure, the robot has to come to a stop as a sensor failure is a severe function loss for the robot.
+
+\subsection{BT Advanced Behaviors}
+After the BT has checked for sensor failures and can guarantee that the environment representation is accurate, the system can check for more complicated scenarios to improve the autonomy.
+
+\subsubsection{Collision Behavior}
+One of the biggest problems mentioned in chapter 3.1 is the inability to react to collisions.
+The collision can happen due to inadequate sensory coverage for detecting obstacles or general sensor inadequacy which can not detect obstacles because of their shape or surface material.
+With a multitude of sensors covering all angles and heights and computer vision capabilities to detect obstacles collisions would be very unlikely.
+But even if the robot could perrfectly detect obstacles, it would still be vulnerable to outside pertubations like other robots driving into it, humans accidently touching the robot etc.
+.
+To combat this the robot must be able to first detect that a collision occured, get out of the collision state, update the map and reset all costmaps to generate a new path to the original goal.
+
+Figure for the Collision Behavior from Groot.
+
+The collision checker uses a simulated sensor to detect collisions.
+On a real robot the collision checker can use a IMU-based collision detection, but this has not been implemented in this behavior tree due to time constraints of this thesis.
+The collision checker itself is integrated into the system supervisor and the BT is making service calls to the execution checker to get the collision state of the robot periodically.
+The child nodes of the collision sequence make use of the sensor data backup. 
+To get out of the collision state, the BT requests the last saved command velocities from the data backup component and modifies them to reverse with minimum speed the same way the robot got into the collision.
+The reversing action takes longer than the two seconds as the commands get published with longer time intervals between them to adjust the driven distance to be the same as the navigation commands.
+An additional check is implemented before the reversing to make sure the robot only reverses once, otherwise the robot would be driving back and forth as the data backup service would provide the BT with the commands from the last reversing action.
+
+Regarding the map reset, we have to determine if we are creating a new map with a SLAM method or if the map server provided a previously saved map. 
+In case of the running SLAM, the map needs to be reset to before the collision occurred.
+
+\subsubsection{Battery Behavior}
+
+\subsubsection{Path Planning Behavior}
+
+\section{Command Velocity Decision Gate}

### 2022-09-18 Thesis/text/methoden.tex
chapter 4.4 ready

--- 2022-09-06 Thesis/text/methoden.tex
@@
-\chapter{Methodology}
+\chapter{Concept}
 \label{cha:methoden}
+
+\section{Current State}
+This chapter briefly presents the current state of mobile navigation with ROS and its limitations. 
+This chapter aims to identify the critical areas in which standard mobile robots running ROS lack robustness and autonomy during autonomous mobile navigation.
+According to these critical areas, requirements are derived from them and prioritized with a risk analysis.
+This chapter proposes how behavior planning can improve the current default systems to meet these requirements.
+
+\subsection{ROS}
+The Robot Operating System (ROS) is an open-source middleware for creating robot applications. It is not an operating system (OS) since it needs an underlying OS, mainly Linux-based, to work. 
+ROS can handle the communication between different programs (called nodes) with standardized interfaces and network protocols.
+Nodes can communicate via topics, services, and actions, which differ in their application domain (compare table xx).
+
+\begin{table}[h!]
+	\caption{ROS Communication Options}
+	\begin{tabular}{ | m{0.15\textwidth} | m{0.3\textwidth}| m{0.15\textwidth} | m{0.4\textwidth}|} 
+  	\hline
+  	Name & Communication Pattern & Area of Use & Cardinality \\ 
+  	\hline
+  	Topics & Publisher/Subscriber & Continuous stream of data e.g. sensor data &  n:m \\ 
+  	\hline
+  	Services & Request/Response & Get specific data only once & One service, many clients \\ 
+  	\hline
+  	Actions & Request/Response and Publish/Subscribe & Trigger executing of asynchronous goal-driven processes and receive updates & One action, many clients \\
+  	\hline
+	\end{tabular}
+\end{table}
+
+This design choice makes robotic applications built with ROS highly reusable and interchangeable, allowing developers to integrate foreign libraries more easily into their applications. 
+Due to the availability of many high-quality open-source libraries for many different use cases, ROS has become the leading methodology for creating robotic applications in the research environment \cite{quigley2009}.
+
+With ROS2, the framework received fundamental changes and updates to its architecture and design to gain more acceptance and use in the industry.
+These changes allow for real-time safety, simpler certification, and security when building industrial applications and products using ROS2 \cite{ros2022}.
+
+\subsection{Navigation2}
+The ROS Navigation2 Stack (Nav2) combines different packages that allow mobile robots to navigate from point A to point B.
+Navigation2 is the de-facto standard for mobile navigation with a wide range of supported robots.
+Supported robot types are holonomic, differential-drive, legged, and ackerman (car-like).
+For the mobile robot to make use of the Nav2 stack, it has to be set up in a certain way to be able to generate plans and execute commands in the right way.
+Nav2 requires the mobile robot to possess a laser scan or point cloud sensor, an odometry source (such as an IMU or wheel encoders), and a set of transformations for planning and navigation.
+The stack contains tools that save and load occupancy maps, localize the robot, plan paths, execute the path, provide cost maps, build behaviors and execute recoveries \cite{macenski2020}.
+
+\begin{figure}[h!]
+	\includegraphics[width=1.0\textwidth]{images/nav2_architecture.png}
+	\caption{Navigation2 Architecture \cite{macenski2020}}
+\end{figure}
+These packages are often integrating a SLAM method to build or enlarge maps.
+Path planning and execution capabilities, which correspond to the global and local planner described in section 2.2, are further supported by ready-to-use planning plugins that use A* and Dijkstra algorithms for path planning and a dynamic window approach for path execution.
+The system sequencing is done in a hierarchical structure as a central behavior tree calls asynchronous actions from the respective planners after another, as seen in figure xx.
+
+\begin{figure} [h!]
+	\includegraphics[width=1.0\textwidth]{images/nav_bt-modified.png}
+	\caption{Navigation2 Behavior Tree}
+\end{figure}
+
+The behavior tree depicted in figure xx is the one that is the default Nav2 behavior and has a set of recovery behaviors already implemented (spin, wait, back up, clear map).
+These behaviors are executed if the robot can not make progress towards the set goal.
+Nav2 uses a node lifecycle management system to start, activate, deactivate, and shut down nodes in a controlled way.
+The lifecycle approach allows the system degree to monitor the system node's startups, executions, and failures.
+
+With this quantity of functions and hierarchical architecture, the ROS Navigation2 stack is an excellent basis for achieving high levels of robot autonomy.
+The software can be modified and expanded with plugins to fit the users' needs.
+When comparing the functionalities of the software stack to the proposed autonomous system architecture in figure xx (2.2), the things that Nav2 is missing are a system-wide supervision layer and a more deliberative approach to behavior planning, as mentioned in 2.3.
+The absence of these components leads to a limited set of circumstances in which the robot can perform navigation reliably.
+One of the core problems behavior-based robots face is generating a dependable environment representation.
+An accurate representation is the basis for the last steps in the planning part of the "Sense - Think - Act" cycle.
+Unforeseen events can severely limit the reliability of the environment representation without the robot being aware of it.
+Examples of such events currently not handled appropriately are slipping wheels, orientation changes by external influences, or undetected, more specifically, undetectable obstacles.
+\textit{Add statistics about random sensor failures every ~10000h} This can lead to erratic movement of the robot, as reactive behaviors and more deliberative behaviors compete for authority over the robot's movement.
+By design, to allow more intelligent robot behaviors, the deliberative behavior can override commands from reactive behaviors.
+However, the deliberative behavior's planning is based upon a false representation of reality, leading to unsafe commands.
+The unsafe planning triggers reactive-type behaviors again to counteract the commands from the deliberative behavior.
+Even worse would be that the reactive behaviors may not get activated, and the robot continues to drive despite being unaware of the surroundings.
+The unsafe behavior requires a human operator would need to step in and restore the robot to full functionality by hand.
+
+\textit{Add a closing statement about that suitability of the Nav stack for achieving autonomy}
+
+\subsection{Turtlebot3}
+The Turtlebot3 is a standard mobile research platform with an available ROS interface to control the robot.
+The robot is well integrated into the ROS ecosystem and is established in the literature as a system to develop and integrate new methods.
+The robot has two motors with attached wheel encoders to drive and steer, classifying it as a differential drive robot. A third omnidirectional wheel stabilizes the robot.
+
+The manufacturer provides an open-source model for simulating the robot in physics-based simulators like Gazebo, as depicted in figure xx, which enables faster development and testing for the robot. 
+The equipped sensors and easy simulation capabilities make the robot well suited for using the Nav2 stack and further development for behavior planning.
+
+The robot specifications are specified in table xx.  \\
+
+\begin{figure}[h!]
+	\includegraphics[width=1.0\textwidth]{images/turtlebot_sim.png}
+	\caption{Turtlebot3 model in Gazebo simulator}
+\end{figure}
+
+The turtlebot will be used for the analysis of the current state of mobile navigation.
+Furthermore, the simulated turtlebot will be the platform on which the behavior planning will be implemented and tested.
+
+\textit{Do that?}
 
@@
 
-\section{Libraries}
+The software requirements that improve the robot's behavior are derived from the current limitations of the standard ROS and Navigation2 setup for mobile robots described in the previous chapter.
+
+\begin{table}[h!]
+\caption{Functional Requirements}
+	\begin{tabular}{| m{0.08\textwidth} | m{0.13\textwidth}| m{0.07\textwidth} | m{0.62\textwidth}|} 
+  	\hline
+  	Nr. & Name & Priority & Description/Acceptance Criteria \\ 
+  	\hline
+  	fn$\_$req1 & Recovery & High &  The system can recover from crashes. Ensure that the system can successfully reach goals despite previous crashes\\ 
+  	\hline
+  	fn$\_$req2 & Sensor Failure & High & The system detects sensor failure. Ensure that the system can restart sensors and decrease the speed during the time the sensor delivers limited information \\ 
+  	\hline
+  	fn$\_$req3 & Maintain operability & High & The robot executes commands as long as it is safe.
+Ensure that the robot keeps driving if it is safe even when system functions are not working correctly.
+\\
+  	\hline
+  	fn$\_$req4 & Emergency Detection & High & The system detects emergency.
+Ensure that the system can detect when the continuation on the calculated path is no longer safe (sensor failures, blockage) \\
+  	\hline
+  	fn$\_$req5 & Emergency Stop & High & The system can initiate emergency stops. Ensure that the system can override all commands and stop in case an emergency is detected \\
+  	\hline
+  	fn$\_$req6 & Override Navigation2 & High & The system can override navigation2. Ensure that the system's commands can always override the commands coming from navigation2. \\
+  	\hline
+  	fn$\_$req7 & Reset Goals & Medium & The system can reset and override goals set by the user so that the goal is reachable by planners. \\
+  	\hline
+  	fn$\_$req8 & Robot Range & Medium & Ensure that the robot will not run out of battery during navigation to a goal.  \\	
+  	\hline
+	\end{tabular}
+\end{table}
+
+Non-functional requirement list
+
+number .. 	Name .. 				Description ..										Success Criteria
+nfreq1 Single Point of Failure	The system does not have a single point of failure	When parts of the system fail, the system maintains operability to a degree
+nfreq2 Performance	The system's control loop guarantees fast reactions	The average frequency with which the system checks the sensors/navigation2 is higher than 100Hz (10ms)
+nfreq3
+
+\begin{table}[h!]
+	\caption{Non-Functional Requirements}
+	\begin{tabular}{ | m{0.05\textwidth} | m{0.13\textwidth}| m{0.07\textwidth} | m{0.65\textwidth} |} 
+  	\hline
+  	Nr. & Name & Priority & Description \\ 
+  	\hline
+  	nfreq1 & Single Point of Failure & High &  The system does not have a single point of failure	When parts of the system fail, the system maintains operability to a degree \\ 
+  	\hline
+  	nfreq2 & Performance & High & The system's control loop guarantees fast reactions. The average frequency with which the system checks the sensors/navigation2 is higher than 100Hz (10ms) \\ 
+  	\hline
+  	nfreq3 & Determinism & High & The outcome for a given set of inputs must be deterministic, meaning that the behavior is always executed similarly. \\
+  	\hline
+  	nfreq4 & Deliberate & High & The robot can execute deliberative behaviors, meaning that behaviors new implemented behaviors go beyond reacting to sensor input and have a planning aspect. \\
+  	\hline
+	\end{tabular}
+\end{table}
+
+\newpage
+
+\section{Risk Analysis For Requirement Prioritization}
+
+\section{System Solution}
+
+For a Turtlebot3 running Nav2 to appropriately deal with the listed requirements, the system needs to be extended with new external modules.
+To eliminate a single point of failure for the system, one needs to create an independent second system outside the existing one.
+This is to create a robust fallback behavior that does not rely on Nav2 to move.
+Also, an external system can more reliably monitor other systems when its execution is detached from the performance and execution of other systems.
+
+A suitable name for the second system is the "autonomy layer", as it aims to increase the robustness and autonomy of the whole system. 
+The autonomy layer has to provide a broad system supervisor that monitors all components' health (sensors, navigation, robot controller).
+The constant execution monitoring equips the system with the ability to deal with the event of node failures and component malfunctions.
+Another required addition to the autonomy layer is a data storage of relevant system information.
+The stored data includes sensor data and generated maps, cost maps, positions, and speed commands.
+Using past data points allows behaviors to become more deliberative in their approach and opens up many possibilities for intelligent behaviors like movement predictions of obstacles.
+A third component to the autonomy layer is a behavior planner largely independent of Nav2 execution and planning.
+This behavior planner must be capable of overriding the Nav2 behaviors if needed.
+The component processes information from the execution checker, Nav2, and past and present sensor data to decide if and how to override the default behaviors.
+Based on this information, the behavior planner can make a deliberative, strategic decision to increase robustness and autonomy, thus decreasing reliance on human supervision.
+Additionally, a component to switch intelligently between speed commands from Nav2 and the autonomy layer is needed.
+This component could be realized as a decision gate that receives speed commands from Nav2 and the new behavior planning component and can forward, block or modify commands sent to the controller.
+This modification aspect allows high-quality local planning capabilities but enables more deliberative behaviors to happen upon this planning.
+
+\section{Software and Simulation}
+A comparison and assessment of the behavior planning approach were made in chapter 2.4.
+The decision on which approach to take fell on the behavior tree method.
+The reason for this is mainly the increased flexibility when creating and expanding the behavior planning compared to FSMs.
+Also, behavior trees allow complete determinism and introspection, which is why POMPDs are not suitable as the behavior planning approach for the specified requirements.
+
+\subsection{Behaviortree.CPP}
+The Navigation2 stack uses a behavior tree to coordinate the planning layer.
+The behavior tree that is used is an expansion of the Behaviortree.CPP (BT.CPP) library.
+This library offers a fast way to create, execute, monitor, and edit behavior trees.
+Additionally to the types of control nodes that are mentioned in chapter 2.4.2, the library adds the concept of reactivity into the catalog of control nodes.
+Reactive sequences and reactive fallbacks differ from normal ones in handling nodes that return the state "running".
+Instead of ticking the node again, the whole sequence restarts, which is very useful for continuously checking a condition and executing an asynchronous action node that gets halted when a condition is not met anymore (compare table xx).
+
+\begin{table}[h!]
+	\caption{Control Nodes in BT.CPP}
+	\begin{tabular}{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
+  	\hline
+  	Name & Child returns Failure & Child Returns Running \\ 
+  	\hline
+  	Sequence & Restart & Tick Again\\ 
+  	\hline
+  	Reactive Sequence & Restart & Restart \\ 
+  	\hline
+  	Sequence Star & Tick again & Tick again \\
+  	\hline
+  	Fallback & Tick next & Tick again\\
+  	\hline  	
+  	Reactive Fallback & Tick next & Restart \\
+  	\hline
+	\end{tabular}
+\end{table}
+
+Also, it offers Decorator Nodes which allow more control over the child node and its output.
+These Decorator Nodes can only have one child node each.
+A comprehensive list of the available nodes and descriptions is in table xx.
+
+\begin{table}[h!]
+	\caption{Decorator Nodes in BT.CPP}
+	\begin{tabular}{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
+  	\hline
+  	Name & Succeeds & Fails & Running \\ 
+  	\hline
+  	Inverter Node & If the child returns false & If child return success &  If the child returns running \\ 
+  	\hline
+  	ForceSuccess Node & Always & Never & If child returns running \\ 
+  	\hline
+  	ForceFailure Node & Never & Always & If child returns running\\
+  	\hline
+  	Repeat Node & Ticks child as long as it returns success. The maximum number of repeats can be defined & If the child returns failure & If the child returns running \\
+  	\hline  	
+  	Retry Node & If the child returns success & Ticks child as long as it returns failure. The maximum number of ticks can be defined & If the child returns running. \\
+  	\hline
+	\end{tabular}
+\end{table}
+
+To allow the tree nodes to communicate, the library provides the developer with two possibilities.
+Either node can use the blackboard, a dictionary (key/value) that all nodes of the tree can read and write.
+Alternatively, two nodes can be connected through ports which allows direct communication between two nodes via a key/value.
+
+\subsection{Groot}
+
+Groot is a software to create, edit, monitor, and debug behavior trees with a graphical user interface.
+The software allows the creation of behavior trees in XML files which can be directly loaded into and used by the BT.CPP library.
+Groot can monitor the live execution of behavior trees and allows the introspection of how the tree reacts to different scenarios.
+
+\textit{Screenshot of live monitoring}
+
+\subsection{Gazebo Simulation Environment}
+The development and testing of behaviors will be done with the Gazebo simulator and the Turtlebot3 model.
+Gazebo is well-integrated into ROS and ROS2 and offers many ROS interfaces to control the simulation.
+Gazebo accurately models the physics of the robot and can simulate and publish the wheel odometry directly on a ROS topic.
+Furthermore, the simulator can simulate the readings for the laser scanner and IMU data via plugins.
+The use of the simulator allows easier repeatability for test cases as various obstacles can be spawned at any time, and sensor failures can be easily induced, resulting in quicker overall development.
+Using a simulator allows slowing down or speeding up the environment time.
+The time control allows more intricate observation of fast reactive type behaviors in slow motion while also enabling the observation of the long-time robustness of the system during extended tests.
+The ROS community offers many different simulation environments to test the robot in. The environments range from simple geometric structurs, to office spaces up to complete race tracks.

### 2022-09-19 Thesis/text/implementierung.tex
implementation chapter almost finished, validation chapter tomorrow

@@
 
+\section{Command Velocity Decision Gate}
+
 \section{Behavior Tree Structure}
@@
 This behavior pattern is realized for the lidar, the IMU and the wheel odometry.
+\\
 
@@
 
-Figure for the Collision Behavior from Groot.
+\begin{figure}
+	\includegraphics[width=1.0\textwidth]{images/collision_fallback.png}
+	\caption{Collision Fallback Behavior}
+\end{figure}
 
@@
 
+The modification is done in the following manner
+
+\textit{Formula for determining execution time}
+
 Regarding the map reset, we have to determine if we are creating a new map with a SLAM method or if the map server provided a previously saved map. 
-In case of the running SLAM, the map needs to be reset to before the collision occurred.
+In case of the running SLAM, the recent updates to the map must be verwerfen to a map prior to when the collision occurred.
+When a static map is provided by the map server component from Nav2 the map needs to be updated with a new obstacle at the point of collision and republished by the map server.
+
+The activity of the "MapUpdateAfterCollision" action node is depicted in figure xx.
+The collision point is estimated by the robot's pose which refers to the robot base link and the robot footprint.
+We assumed that the collision point is right in front of the robot, because the local planner favours forward drive kinematics.
+The simulated bumper sensor can provide more nuanced information about the point of contact, but this functionality can not be achieved on a real robot with an IMU-based collsion detection, hence we simplified the obstacle position to always be in front of the robot.
+The collision position is calculated with
+
+\textit{Formula here}
+
+A 25 by 25 cm square with full occupation of that space is added to the current map and then saved to be reloaded by the map server as depicted in figure xx.
+
+The original goal is then republished to replan the global path on the updated map.
+
+In this manner obstacles lower than the lidar scan height can be detected.
+With this behavior the robot should also be able to eventually navigate around transparent obstacles which do not reflect the laser.
+Glass doors are known to cause issues for robots based only on lidars.
 
@@
 
+The battery behavior has the goal to monitor the robot's battery and has to make to the decision if the robot should navigate to a new goal location or not.
+The decision is based on the current battery charge and the estimated consumption of energy to reach the new goal.
+The robot should not begin the navigation process if it is probable that the remaining charge will not suffice for the driving.
+
+\begin{figure}
+	\includegraphics[width=0.5\textwidth]{images/battery_fallback.png}
+	\caption{Battery Fallback Behavior}
+\end{figure}
+The robot has to take into account the idle consumption, the energy to drive the motors.
+With this information we created a simple linear function for the energy expenditure to predict the required power depending on the length of the path.
+To make the algorithm fail safe, we added an additional safety factor into the function, to berücksichtigen the scenario where the robot has to reroute and drive a longer route or has to stop for a moment.
+The resulting function can be described as: \\
+
+\textit{Formel here}
+
+On a real robot the average values for the idle and driving consumption could be experimentally determined.
+For the simulated turtlebot we created a simulated battery package which updates the battery charge every second.
+The battery package subscribes to the comman velocity topic and decreases the battery charge for every message that is received.
+We created the battery package because the behavior for this scenario needs to be tested.
+This is way the consumption rates are not close to reality.
+The battery package provides a service to get the battery charge and the values for the angenommenen consumption rates.
+The BT node "IsBatterySufficientCheck" calls the service to compare the current battery to the calculated consumption.
+The battery behavior fallback will cancel the navigation goal if the robot can not reach the given goal with the remaining charge.
+\\
+
 \subsubsection{Path Planning Behavior}
 
-\section{Command Velocity Decision Gate}
+With this behavior, we implement a way to make the global planner more flexible. 
+The execution checker component monitors the health of the Nav2 global planner.
+Additionally, the BT analyses the global planners output.
+If one of the condition fails, the given navigation goal is not reachable by the planning algorithm.
+We combat this by shifting the goal to a nearby alternative to get the robot as close to the original goal as possible.
+Is the location still unreachable, the goal gets moved further away and closer to the robot again.
+This routine gets repeated for up to ten times, which leads to a distance to the original goal of up to one meter.
+The goal gets moved on a straight vector towards the robot. Other strategies for finding alternative goals were not implemented in this behavior.
+
+\begin{figure}
+	\includegraphics[width=0.75\textwidth]{images/global_planning_fallback.png}
+	\caption{Path Planning Fallback}
+\end{figure}

### 2022-09-23 Thesis/text/diskussion.tex
almost done, summary not finished, introduction is not polished

@@
 
-\section{Conclusion}
+What went well? 
+This thesis demonstrated a way on how to incorporate advanced behavior planning in ROS2 mobile robots.
+The goal was to achieve higher levels of autonomy and robustness.
+The implemented behaviors and the system supervision component have shown to improve the behavior in the simulated scenarios.
 
-\section{Prospect}
+However, the results will need to be reproduced on a real mobile robot.
+Additionaly, the implemented behaviors are a first step to increase the autonomy, but many more behaviors and fallbacks need to be created for the behavior planner.
+
+The used system architecture is a good base upon which more development can take place without running into problems or limitations in later stages of the development.
+
+More work is needed to define what is considered to be a safe state or position for the robot to be in after an emergency stop is triggered.
+Reaching these minimum risk states after emergencies is then the duty of the behavior planner.
+
+Some behaviors that incorporate rules for autonomous driving, like keeping appropriate distances, right of way, etc., into the planning need to be implemented in future works.
+
+Right now, the robot has no special behaviors for multi-robot system which is a field for many more deliberative behaviors like sharing sensor data with other robots to overcome complete sensor failures.
+
+What is missing? Further development?
+
+New research questions?

### 2022-09-23 Thesis/text/implementierung.tex
almost done, summary not finished, introduction is not polished

@@
 \section{System Architecture}
-\begin{figure}
-	\includegraphics[width=1.0\textwidth]{images/component_diagram_bt_architecture.png}
+
+The system architecture, depicted in figure xx, integrates the behavior tree, responsible for the behavior planning, in a position to override Navigation2.
+This is done through the help of the Decision Gate, that can selectively forward speed commands to the robot controller.
+The behavior tree can access the sensor data when needed through the Data Backup component.
+The System Supervision component controls the health of all components, but the connections are not depicted in the component diagram for Übersichtlichkeit.
+
+\begin{figure}[h!]
+	\includegraphics[width=1.05\textwidth]{images/component_diagram_bt.png}
 	\caption{Component Diagram of System Architecture}
@@
 
+This component is responsible for modyfying or blocking the Velocity Commands from Nav2.
+The component receives messages from Navigation2 on the "/cmd$\_$vel$\_$nav" topic and on the "/cmd$\_$vel$\_$bt" topic from the behavior tree.
+This component publishes the received messages on the actual topic "/cmd$\_$vel" which is subscribed by the robot controller.
+The BT can modify parameters of the Decision Gate, to either disregard incoming messages of Nav2 completely or to modify them.
+The modification is useful when the situation demands the robot to slow down.
+If the Decision Gate receives messages from both Nav2 and the BT, it will prioritize the BT.
+This situation should theoretically never occur as the BT will set the parameters for the Decision Gate accordingly and/or cancel the current Navigation Goal before publishing commands on its own.
+It is implemented as an additional safety measure, in case the cancellation of the goal is not possible if Navigation2 is not responsive to the cancellation command.
+
 \section{Behavior Tree Structure}
@@
 The root sequence is the first step for increasing the robustness of the system to combat competing reactive and deliberative behaviors which could endanger the safety of the robot and other actors in its environments.
+
+\begin{figure}[h!]
+	\includegraphics[width=1.0\textwidth]{images/simplified_bt.png}
+	\caption{Top Level Behavior Tree Structure}
+\end{figure}
 
@@
 
-The original goal is then republished to replan the global path on the updated map.
+The original goal is then republished to replan the global path on the updated map. A more detailed view of the whole behavior is depicted in figure xx.
+
+\begin{figure}
+	\includegraphics[width=0.9\textwidth]{images/activity_diagram_collision.png}
+	\caption{Activity Diagram for the Collision Behavior}
+\end{figure}
 
@@
 
+\begin{figure}[h!]
+	\includegraphics[width=0.7\textwidth]{images/global_planning_fallback.png}
+	\caption{Path Planning Fallback Behavior}
+\end{figure}
+
+\section{System Support Tools}
+\subsection{Gazebo Sensor Drivers}
+For testing purposes, the simulated sensor data is not published directly on to the actual topic by gazebo.
+To control the flow of information in the system, an intermediary node is used, which subscribes to the Gazebo topic and republishes the information on the original topic.
+With this methodology, the failure of a sensor can be simulated as the simulation does not offer this functionality.
+Figure xx shows the RQT graph, which depicts the ROS nodes and their subscribed/published topics.
+
 \begin{figure}
-	\includegraphics[width=0.75\textwidth]{images/global_planning_fallback.png}
-	\caption{Path Planning Fallback}
+	\includegraphics[width=0.75\textwidth]{images/rqt_graph_sensor_drivers.png}
+	\caption{Graph for the Sensor Driver nodes and topics}
 \end{figure}
+
+\subsection{Custom ROS Interfaces}
+
+Because of the neccessity to have ROS nodes spinning outside of the BT thread as discussed in 4.4.1, additional interfaces needed to be created in order to receive informations through service calls.
+The created service types are listed in table xx.
+All of the service definitions are created in a dedicated package named "bt$\_$msgs".
+
+\begin{table}[h!]
+	\caption{Implemented Custom Services}
+	\begin{tabular}{ | m{0.2\textwidth} | m{0.33\textwidth}| m{0.33\textwidth} | m{0.2\textwidth} |} 
+  	\hline
+  	Name & Request & Response \\ 
+  	\hline
+  	GetCharge & Empty & float charge, float idle$\_$decrease$\_$per$\_$sec, float drive$\_$decrease$\_$per$\_$sec \\
+  	\hline
+  	GetDistance & Empty & float distance$\_$in$\_$meter\\ 
+  	\hline
+  	GetLastGoal & Empty & geometry$\_$msgs/Pose last$\_$goal$\_$pose \\ 
+  	\hline
+  	GetLastMap & Empty & nav$\_$msgs/OccupancyGrid map \\
+  	\hline
+  	GetTwistArray & Empty & geometry$\_$msgs/Twist[] cmd$\_$vel$\_$array\\
+  	\hline  	
+  	PubCmdVel & geometry$\_$msgs/Twist cmd$\_$vel, float time$\_$in$\_$seconds & bool success \\
+  	\hline
+  	SendUpdatedMap & nav$\_$msgs/OccupancyGrid updated$\_$map & Empty \\
+  	\hline
+	\end{tabular}
+\end{table}

### 2022-09-23 Thesis/text/ergebnisse.tex
almost done, summary not finished, introduction is not polished

@@
 
+To test the efficacy of the behavior planning approach in relation to the requirements from chapter 3.xx, the robot is put into specific scenarios that should trigger the behavior responses.
+
+The scenarios are executed in a simulated apartment environment as seen in figure xx and in a simulated world with small round obstacles in an enclosed space as depicted in figure xx.
+
+\begin{figure}
+	\includegraphics[width=0.5\textwidth]{images/house_env.png}
+	\caption{Apartment Environment in Gazebo}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=0.5\textwidth]{images/world_env.png}
+	\caption{Enclosed Environment in Gazebo}
+\end{figure}
+
+The environments was mapped beforehand with the ROS2 "Slam Toolbox" package and provided via the Nav2 map server.
+
+In order to achieve comparability between test runs of each scenario the robot location and navigation goals are defined in another ROS Node which programmatically spawns the robot into the environment and sets the goal this way.
+This way, the exact same scenario can be run with and without the behavior planning active.
+For testing the behavior of the sensor failure fallbacks, the respective sensor driver, mentioned in chapter 4.5.1, was shutdown to emulate the failure of the sensor.
+
 \section{Scenarios}
+
+The scenarios are derived from the functional requirements and can be used to test multiple requirements.
+The acceptance criteria often contains multiple binary conditions that can only be met or failed. All acceptance criteria for a given scenario must be met, for the test to be counted as success.
+
+For the Lidar, IMU and Odometry scenarios the sensor driver was shut down when reaching the desired speed as listed in the scenario. 
+Scenario NoPathFound1 is a scenario in which the lifecycle state gets transitioned into "inactive" and can not execute the planning action anymore when triggered.
+The BT has to recover from the inability to plan in this scenario.
+NoPathFound$\_$2 is simulating the scenario when the planner is active but fails to find a collision-free path to the desired goal.
+This scenario tests the ability to find alternative goals and navigate to a nearby goal.
+In the Collision$\_$1 scenario, an obstacle gets moved so close to the robot that a collision is registered.
+This tests the ability to get out of a collision and resume the navigation.
+After the BT exits the Collision Fallback Routine a new goal is published for the robot to navigate to.
+Collision$\_$2 scenario is testing the same behavior but during an active navigation.
+A flat obstacle is spawned in the path of the robot.
+The obstacle is too low to be detected by the laserscanner and a collision is happening.
+The robot must be able to recover from the crash and keep on navigating towards the goal with an updated global path in this scenario.
+The Battery$\_$1 scenario tests if the robot tries to navigate towards a far away goal with a low battery charge.
+For this, the simulated battery is emptied to a low level and a goal is set.
+If the battery charge runs below zero percent charge during the navigation the test is failed.
+
+\begin{table}[h!]
+\caption{Scenarios}
+	\begin{tabular}{| m{0.1\textwidth} | m{0.13\textwidth}| m{0.33\textwidth} | m{0.34\textwidth}|} 
+  	\hline
+  	Name & Related Requirements & Description & Success Criteria\\ 
+  	\hline
+  	Lidar$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Lidar node crashes & Reset the system\\ 
+  	\hline
+  	Lidar$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
+  	\hline
+  	Lidar$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
+  	\hline
+  	IMU$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, IMU node crashes & Reset system \\
+  	\hline
+  	IMU$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
+  	\hline
+  	IMU$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
+  	\hline
+  	Odom$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Odom node crashes & Reset the system\\ 
+  	\hline
+  	Odom$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
+  	\hline
+  	Odom$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
+  	\hline
+ 	\end{tabular}
+\end{table}
+
+\begin{table}[h!]
+	\caption{Scenarios}
+	\begin{tabular}{| m{0.1\textwidth} | m{0.13\textwidth}| m{0.33\textwidth} | m{0.34\textwidth}|} 
+  	\hline	
+  	NoPath Found$\_$1 & fn$\_$req7 & Path to goal can not be calculated, robot is standing still &  
+Reset system, calculate path to goal, reach goal \\
+	\hline
+	NoPath Found$\_$2 & fn$\_$req7, fn$\_$req8 & Goal is unreachable, robot is standing still & Alternative goals, close to the original are tested to be reachable \\
+	\hline
+	Collision$\_$1 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & Robot is standing still and a collision with the robot is caused & Robot can get out of collision state, navigation to goals still working \\
+	\hline
+	Collision$\_$2 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & The robot is driving and collides with an undetected obstacle (0.25m/s) &  
+Robot can get out of collision state, navigation to goals still working, undetected obstacle gets added to map \\
+	\hline
+	Battery$\_$1 & fn$\_$req9 & The robot battery runs low & The robot will not drive to a goal which is outside of its reachable range \\
+	\hline
+	Motor$\_$1 & fn$\_$req2, fn$\_$req5 & Hardware failure on the motors & n.a. \\
+	\hline
+
+	\end{tabular}
+\end{table}
+
+\section{Results}
+
+The results in table xx show an increase in the successful handling of the test scenarios for most test cases.
+The system supervision component is a very reliable component and enables the system to recover from an unexpected sensor failure.
+The more advanced behaviors that were implemented also show good improvements in the autonomous handling of the scenarios.
+The Battery Scenario was successful every time it was induced.
+However, the path planning behavior for finding alternative goals was not successful in about 50 percent of the time.
+The carried out test showed that alternative methods for finding new goals would probably have resulted in better performance.
+The collision behavior during driving worked really well and the map updates are a important feature for the success of the scenarios.
+Nevertheless, there is a lack of intelligence and autonomy when the robot is standing still and a collision is forced as it relies on the same mechanism as during the driving scenario.
+This led to a case in which the robot reversed into a wall because the collision turned the robot by about 90 degrees.
+This led to a failed test in which a operator would be needed to correct the robot.
+This apparent lack of robustness was not apparent in the driving collision scenario because the force of the robot was not enough to cause a major shift in orientation.
+
+\begin{table}[h!]
+\caption{Results}
+	\begin{tabular}{| m{0.11\textwidth} | m{0.13\textwidth}| m{0.33\textwidth} | m{0.33\textwidth}|} 
+  	\hline
+  	Name & Number of Runs &  Percentage Succesful Without Behavior Planning & Percentage Succesful With Behavior Planning\\ 
+  	\hline
+  	Lidar$\_$1 & 5 & 0 & 100 \\ 
+  	\hline
+  	Lidar$\_$2 & 5 & 0 & 100 \\ 
+  	\hline
+  	Lidar$\_$3 & 5 & 0 & 100 \\ 
+  	\hline
+  	IMU$\_$1 & 5 & 0 & 100 \\ 
+  	\hline
+  	IMU$\_$2 & 5 & 0 & 100 \\ 
+  	\hline
+  	IMU$\_$3 & 5 & 0 & 100 \\ 
+  	\hline
+  	Odom$\_$1 & 5 & 0 & 100 \\ 
+  	\hline
+  	Odom$\_$2 & 5 & 0 & 100 \\ 
+  	\hline
+  	Odom$\_$3 & 5 & 0 & 100 \\ 
+  	\hline
+  	NoPath Found$\_$1 & 20 & 0 & 100 \\ 
+  	\hline
+  	NoPath Found$\_$2 & 20 & 0 & 55 \\ 
+  	\hline
+  	Collision$\_$1 & 20 & 0 & 95 \\ 
+  	\hline
+  	Collision$\_$2 & 20 & 0 & 100 \\ 
+  	\hline
+  	Battery$\_$1 & 10 & 0 & 100 \\ 
+  	\hline
+  	Motor$\_$1 & 10 & 0 & 0 \\ 
+  	\hline
+	\end{tabular}
+\end{table}
+
+The non-functional requirements are mostly met by the architectural design choices. 
+The implementation of the whole behavior planning and system supervision system is eliminating a single point of failure for the system.
+Even when the whole Navigation2 system collapses is the robot able to move to a degree and more importantly stop completely (compare table 3.3, non$\_$fn$\_$req).
+The behavior tree itself and algorithms that are implemented in the behavior tree are all deterministic (non$\_$fn$\_$req3).
+The robots behavior goes beyond pure reactivity when navigating through environments, as the advanced behaviors are making use of a dedicated planning phase before the actions are carried out (non$\_$fn$\_$req4).
+And finally, the performance of the system when checking the conditions of the BT is below the desired threshold of 10ms, as the whole system runs in multiple threads which allows the behavior tree to complete one complete cycle in about 5ms with current system (non$\_$fn$\_$req2).

### 2022-09-23 Thesis/text/methoden.tex
almost done, summary not finished, introduction is not polished

@@
 	\caption{ROS Communication Options}
-	\begin{tabular}{ | m{0.15\textwidth} | m{0.3\textwidth}| m{0.15\textwidth} | m{0.4\textwidth}|} 
+	\begin{tabular}{ | m{0.12\textwidth} | m{0.2\textwidth}| m{0.35\textwidth} | m{0.2\textwidth}|} 
   	\hline
@@
 \begin{figure}[h!]
-	\includegraphics[width=1.0\textwidth]{images/turtlebot_sim.png}
+	\includegraphics[width=0.7\textwidth]{images/turtlebot_sim.png}
 	\caption{Turtlebot3 model in Gazebo simulator}
@@
 The turtlebot will be used for the analysis of the current state of mobile navigation.
-Furthermore, the simulated turtlebot will be the platform on which the behavior planning will be implemented and tested.
-
-\textit{Do that?}
+Furthermore, the simulated turtlebot will be the platform on which the behavior planning will be implemented and tested.\\
 
@@
   	\hline
-  	fn$\_$req1 & Recovery & High &  The system can recover from crashes. Ensure that the system can successfully reach goals despite previous crashes\\ 
+  	fn$\_$req1 & Sensor Failure & High & The system detects sensor failure. Ensure that the system can restart sensors and decrease the speed during the time the sensor delivers limited information \\ 
   	\hline
-  	fn$\_$req2 & Sensor Failure & High & The system detects sensor failure. Ensure that the system can restart sensors and decrease the speed during the time the sensor delivers limited information \\ 
+  	fn$\_$req2 & Emergency Detection & High & The system detects emergency.
+Ensure that the system can detect when the continuation on the calculated path is no longer safe (sensor failures, blockage) \\
   	\hline
-  	fn$\_$req3 & Maintain operability & High & The robot executes commands as long as it is safe.
+  	fn$\_$req3 & Emergency Stop & High & The system can initiate emergency stops. Ensure that the system can override all commands and stop in case an emergency is detected \\
+  	\hline  	
+  	fn$\_$req4 & Override Navigation2 & High & The system can override navigation2. Ensure that the system's commands can always override the commands coming from navigation2. \\
+  	\hline
+  	fn$\_$req5 & Maintain operability & High & The robot executes commands as long as it is safe.
 Ensure that the robot keeps driving if it is safe even when system functions are not working correctly.
@@
   	\hline
-  	fn$\_$req4 & Emergency Detection & High & The system detects emergency.
-Ensure that the system can detect when the continuation on the calculated path is no longer safe (sensor failures, blockage) \\
+  	fn$\_$req6 & Recovery & High &  The system can recover from crashes. Ensure that the system can successfully reach goals despite previous crashes\\ 
   	\hline  
-  	fn$\_$req5 & Emergency Stop & High & The system can initiate emergency stops. Ensure that the system can override all commands and stop in case an emergency is detected \\
+  	fn$\_$req7 & Control Path Planning & Medium & The system controls and rates the quality of the planned paths. \\
   	\hline	
-  	fn$\_$req6 & Override Navigation2 & High & The system can override navigation2. Ensure that the system's commands can always override the commands coming from navigation2. \\
+  	fn$\_$req8 & Reset Goals & Medium & The system can reset and override goals set by the user so that the goal is reachable by planners. \\
   	\hline
-  	fn$\_$req7 & Reset Goals & Medium & The system can reset and override goals set by the user so that the goal is reachable by planners. \\
-  	\hline
-  	fn$\_$req8 & Robot Range & Medium & Ensure that the robot will not run out of battery during navigation to a goal.  \\	
+  	fn$\_$req9 & Robot Range & Medium & Ensure that the robot will not run out of battery during navigation to a goal.  \\	
   	\hline
@@
 \end{table}
-
-Non-functional requirement list
-
-number .. 	Name .. 				Description ..										Success Criteria
-nfreq1 Single Point of Failure	The system does not have a single point of failure	When parts of the system fail, the system maintains operability to a degree
-nfreq2 Performance	The system's control loop guarantees fast reactions	The average frequency with which the system checks the sensors/navigation2 is higher than 100Hz (10ms)
-nfreq3
 
@@
 \end{table}
-
-\newpage
-
-\section{Risk Analysis For Requirement Prioritization}
 

### 2022-09-23 Code/bt_workspace/src/bt_tests/include/bt_tests/scenario_1.hpp

@@ -1,0 +1,27 @@
+#include <cstdio>
+#include "rclcpp/rclcpp.hpp"
+#include "gazebo_msgs/srv/spawn_entity.hpp"
+#include "geometry_msgs/msg/pose.hpp"
+#include "tf2/convert.h"
+#include "tf2/LinearMath/Quaternion.h"
+#include "tf2/LinearMath/Matrix3x3.h"
+
+class BehaviorTester : rclcpp::Node
+{
+  public:
+    BehaviorTester(std::string &name);
+
+    void spawn_entity();
+
+  private:
+
+  rclcpp::Client<gazebo_msgs::srv::SpawnEntity>::SharedPtr spawn_client_handle_;
+  geometry_msgs::msg::Pose robot_spawn_pose_;
+  float robot_spawn_yaw_;
+
+  geometry_msgs::msg::Pose goal_pose_;
+  float goal_yaw_;
+  std::shared_ptr<gazebo_msgs::srv::SpawnEntity::Request> request_;
+
+
+};

### 2022-09-23 Code/bt_workspace/src/bt_tests/src/scenario1.cpp

@@ -1,4 +1,61 @@
-#include <cstdio>
+#include "scenario_1.hpp"
+
+BehaviorTester::BehaviorTester(std::string &name) : rclcpp::Node(name)
+    {
+        // sdf_file_path = (os.path.join(get_package_share_directory(package_name), 'urdf', 'radu_gazebo_compiled.urdf')),
+
+        this->declare_parameter("spawn_pose_x", 0.0);
+        this->declare_parameter("spawn_pose_y", 0.0);
+        this->declare_parameter("spawn_pose_yaw", 0.0);
+
+        this->declare_parameter("goal_pose_x", 1.0);
+        this->declare_parameter("goal_pose_y", 1.0);
+        this->declare_parameter("goal_pose_yaw", 0.0);
+
+        this->get_parameter("spawn_pose_x", robot_spawn_pose_.position.x);
+        this->get_parameter("spawn_pose_y", robot_spawn_pose_.position.y);
+        this->get_parameter("spawn_pose_yaw", robot_spawn_yaw_);
+        tf2::Quaternion quat_robot;
+        quat_robot.setRPY(0.0, 0.0, robot_spawn_yaw_);
+        robot_spawn_pose_.orientation.w = quat_robot.getW();
+        robot_spawn_pose_.orientation.x = quat_robot.getX();
+        robot_spawn_pose_.orientation.y = quat_robot.getY();
+        robot_spawn_pose_.orientation.z = quat_robot.getZ();
+
+        this->get_parameter("goal_pose_x", goal_pose_.position.x);
+        this->get_parameter("goal_pose_y", goal_pose_.position.y);
+        this->get_parameter("goal_pose_yaw", goal_yaw_);
+        tf2::Quaternion quat_goal;
+        quat_goal.setRPY(0.0, 0.0, goal_yaw_);
+        goal_pose_.orientation.w = quat_goal.getW();
+        goal_pose_.orientation.x = quat_goal.getX();
+        goal_pose_.orientation.y = quat_goal.getY();
+        goal_pose_.orientation.z = quat_goal.getZ();
+
+        spawn_client_handle_ = this->create_client<gazebo_msgs::srv::SpawnEntity>("spawn_entity");
+
+        request_->name = "turtlebot";
+        request_->robot_namespace = "";
+        request_->reference_frame = "world";
+        request_->initial_pose = robot_spawn_pose_;
+        // request_->xml = 
+    }
+
[…]

### 2022-09-23 Thesis/text/diskussion.tex
little commit summary 1st draft done

--- 2022-09-06 Thesis/text/diskussion.tex
@@
 \label{cha:diskussion}
+\paragraph*{}
 
-\section{Conclusion}
+This thesis demonstrated a way on how to incorporate advanced behavior planning in ROS2 mobile robots.
+The goal was to achieve higher levels of autonomy and robustness.
+The implemented behaviors and the system supervision component have shown to improve the behavior in the simulated scenarios.
+With the additional components, the system is able to react and recover from multiple scenarios.
+\paragraph*{}
 
-\section{Prospect}
+The used system architecture is a good base upon which more development can take place without running into problems or limitations in later stages of the development.
+Nevertheless, the design of the behavior tree could be modified to make more use of blackboard values to increase the information flow between action node in the tree.
+Also, the tree does not utilize the reactive control flow nodes and no action node in the tree is designed to allow multiple threads right now.
+The incorporation of so-called stateful nodes into the structure of the tree to allow co-routines would increase the possibility to create better and more deliberative behaviors.
+The implemented behaviors are a first step to increase the autonomy, but many more behaviors and fallbacks need to be created for the behavior planner.
+Some behaviors which were not implemented in this thesis are the incorporation of the Intel RSS rules for autonomous driving, like keeping appropriate distances, right of way rules, and adaptable speed based on occlusions.
+Another are of future work are special behaviors for multi-robot system like sharing sensor data with other robots to overcome complete sensor failures or intelligent path planning to avoid collisions when robots can access the planned paths of other robots.
+Also, more work is needed to define what is considered to be a safe state or position for the robot to be in after an emergency stop is triggered.
+Reaching these minimum risk states after emergencies is then the duty of the behavior planner.
+
+\paragraph*{}
+
+To achieve more independence of human operators, the behaviors need to be smarter and more robust in general.
+One of the key question in autonomous driving is the responsibility of a vehicle in a crash, because with higher levels of automation the human is not needed, but a machine can not be made responsible for its action, because it is programmed to do so.
+But smart, robust and independent decision making almost always involves machine learning which makes decisions based on a training with dedicated datasets and not programmed lines of code.
+Because a system in which the outcome for a high-risk scenario is unclear, machine learning behaviors may not be suited to be used on their own.
+A way to allow a system to make use of machine learning behaviors for smarter decisions and still maintain a deterministic character would be to have a behavior planner, like a behavior tree, to execute and monitor the behaviors and equip the planner with highly automated, but still deterministic, behaviors to ensure safety in case the learned behavior can not guarantee it to be deterministic.
+Research is needed on how behavior trees can manage a system with autonomous machine learning behaviors inside of the tree.
+
+\paragraph*{}
+
+Lastly, the results must to be reproduced on a real mobile robot.
+The layout of the behavior tree will stay the same, but further effort is needed to adapt the action nodes of the tree for a real system.

### 2022-09-25 Thesis/text/einleitung.tex
cleanup of tjhe doc

--- 2022-09-06 Thesis/text/einleitung.tex
@@
 
-Core Problem - too little autonomy in a standard autonomous mobile robot running ROS navigation(AMR)
-
-Description of the Problem
-
-Robots regurlarly needs operator in case smth unexpected happens\\
-
-this may include: unexpected collisions, a system restart is needed, battery is empty\\
-
-The system is not able to react and handle unforeseen situations\\
-
-This can lead to safety issues during the operation of a robot. These safety issues can lead to uncontrolled driving maneuvers due to wrong sensor data or a wrong environment represantation.\\
-
-Due to the need of the robot to require an operator, the robot is closer to an automated instead of an autonomous vehicle. 
-Difference is defined by the SAE Levels for automonous driving in cars.\\
-
-It is important for the robot to possess the ability to navigate in an uncertain environment to guarantee the safety of the robot itself and all other actors in its environment, these can be other robots and humans.\\
-
+Mobile robots are becoming more widespread in many domains.
+They are marketed as autonomous mobile robots, but often the autonomy of the robot is limited to very specific environments and require frequent human help to function correctly.
+This hardly qualifies the robot to be labeled autonomous.
+One of the quickest and most popular ways to build and program a robot is to use the Robot Operating System (ROS).
+ROS offers many tools and functionalities to create and program robots, but even then a normal robot needs human supervision during the operation to ensure the proper function and safety.
+There are ROS packages to safely navigate from point A to point B, but currently they are lacking the ability to react to unforeseen events.
+It is important for the robot to possess the ability to navigate in an uncertain environment to guarantee the safety of the robot itself and all other actors in its environment, these can be other robots and humans.
 In theory the robot can perform fully autonomous navigation including mapping and localization but as soon as something out of the ordinary happens the robots autonomy is not guaranteed to be reliable anymore.
@@
 The reliance on human-needed problem solving decreases the robots usefulness when tasked with real goals.
+Therefore, this thesis centers around the research question how behavior planning can increase the robustness and autonomy of a robot running ROS2. 
+Behavior planning enables the robot to react to failures and problems of the system and can decide alternative courses of action to mitigate risks and finish the given tasks.
 
-Having a robust and safe robot behaviour opens the door to many applications in a more diverse set of challenging environments.
+%Core Problem - too little autonomy in a standard autonomous mobile robot running ROS navigation(AMR)
+%
+%Description of the Problem:
+%
+%Despite the advancements in autonomous driving the robot regurlarly needs an operator in case smth unexpected happens.
+%
+%This may include: unexpected collisions, a system restart is needed, battery is empty, the environment is too crowded and the robot gets stuck
+%
+%The system is not able to react and handle unforeseen situations.
+%
+%This can lead to safety issues during the operation of a robot. These safety issues can lead to uncontrolled driving maneuvers due to wrong sensor data or a wrong environment represantation.
+%
+%Due to the need of the robot to require an operator for safe, the robot is not really autonomous despite it being labeled as such. 
+%
+%(Difference is defined by the SAE Levels for automonous driving in cars.) \\
+%
+%Relevance, need for research:\\
+%\\
+
+%
+%Having a robust and safe robot behaviour opens the door to many applications in a more diverse set of challenging environments.
 Usually a reinforcement learning approach with broad training data set is a great way to handle a multitude of unknown environments and situations, but one of the biggest problems is the lacking determinism in such systems.
+%
+%In the robotic world exists a need for a deterministic system which can enhance and override the decision making process and navigation capabilities of the robot. 
+%
+%Planned solution:\\
 
-In the robotic world exists a need for a deterministic system which can enhance and override the decision making process and navigation capabilities of the robot.
+This thesis will explore possibilities how current systems can be improved through the addition of a dedicated behavior planning component.
+Furthermore, the thesis will provide an implementation of examplatory behaviors and provide a system architecture to incorporate behavior planning in other robots, too
 
-This thesis proposes a comprehensive approach for behavior planning for an autonomous mobile robot.
+%The thesis will explore how an extension of the current software architecture and behavior planning capabilities for an autonomous mobile robot that is running ROS2 can improve the autonomy of the robot.
+%
+%Expected results:\\
 
-\textbf{How can behavior planning increase the safety and autonomy of an autonomous mobile robot running ROS?}
+The desired outcomes of these implemented software will be an increase of robustness and decrease of required human actions during a number of scenarios.
+In these scenarios, failures and problems are artificially induced to test the robots abilities to behave autonomously.
+
+%\textit{Research gap is not named so far}

### 2022-09-25 Thesis/text/ergebnisse.tex
cleanup of tjhe doc

@@
 
-To test the efficacy of the behavior planning approach in relation to the requirements from chapter 3.xx, the robot is put into specific scenarios that should trigger the behavior responses.
+To test the efficacy of the behavior planning approach in relation to the requirements from chapter 3.2, the robot is put into specific scenarios that should trigger the behavior responses.
 
-The scenarios are executed in a simulated apartment environment as seen in figure xx and in a simulated world with small round obstacles in an enclosed space as depicted in figure xx.
-
-\begin{figure}
-	\includegraphics[width=0.5\textwidth]{images/house_env.png}
+The scenarios are executed in a simulated apartment environment as seen in figure \ref{fig:house_gazebo} and in a simulated world with small round obstacles in an enclosed space as depicted in figure \ref{fig:world_gazebo}.
+\begin{center}
+\begin{figure}[ht]
+	\label{fig:house_gazebo}
+	\includegraphics[width=0.75\textwidth]{images/house_env.png}
 	\caption{Apartment Environment in Gazebo}
 \end{figure}
+\end{center}
 
+\begin{center}
 \begin{figure}
-	\includegraphics[width=0.5\textwidth]{images/world_env.png}
+	\label{fig:world_gazebo}
+	\includegraphics[width=0.75\textwidth]{images/world_env.png}
 	\caption{Enclosed Environment in Gazebo}
 \end{figure}
+\end{center}
 
@@
 
-The scenarios are derived from the functional requirements and can be used to test multiple requirements.
+The scenarios are derived from the functional requirements and can be used to test multiple requirements and listed in the tables \ref{tab:sensor_scenarios} and \ref{tab:behavior_scenarios}.
 The acceptance criteria often contains multiple binary conditions that can only be met or failed. All acceptance criteria for a given scenario must be met, for the test to be counted as success.
@@
 
-\begin{table}[h!]
+\begin{table}[ht]
+\label{tab:sensor_scenarios}
 \caption{Scenarios}
@@
 
-\begin{table}[h!]
+\begin{table}[ht]
+	\label{tab:behavior_scenarios}
 	\caption{Scenarios}
 	\begin{tabular}{| m{0.1\textwidth} | m{0.13\textwidth}| m{0.33\textwidth} | m{0.34\textwidth}|} 
+	\hline
+	Name & Related Requirements & Description & Success Criteria\\ 
   	\hline	
@@
 
-The results in table xx show an increase in the successful handling of the test scenarios for most test cases.
+The results in table \ref{tab:results} show an increase in the successful handling of the test scenarios for most test cases.
 The system supervision component is a very reliable component and enables the system to recover from an unexpected sensor failure.
@@
 
-\begin{table}[h!]
+\begin{table}[ht]
+\label{tab:results}
 \caption{Results}
@@
 The implementation of the whole behavior planning and system supervision system is eliminating a single point of failure for the system.
-Even when the whole Navigation2 system collapses is the robot able to move to a degree and more importantly stop completely (compare table 3.3, non$\_$fn$\_$req).
+Even when the whole Navigation2 system collapses is the robot able to move to a degree and more importantly stop completely (compare table \ref{tab:tab:nonfn_req}, non$\_$fn$\_$req).
 The behavior tree itself and algorithms that are implemented in the behavior tree are all deterministic (non$\_$fn$\_$req3).

### 2022-09-25 Thesis/text/verwandt.tex
cleanup of tjhe doc

@@
 In order to define what qualifies as an improvement in the autonomous behaviour of a mobile robot, one needs to look at the defintion of automated and autonomy.
-The EFI defines the term autonomy in the context of robotics as a system which can act without human instructions and still solve complex tasks, make decisions, learn independently aswell as react to unforeseen circumstances \cite{efi2018}.
+The "Expertenkommission Forschung und Innovation" (EFI) defines the term autonomy in the context of robotics as a system which can act without human instructions and still solve complex tasks, make decisions, learn independently aswell as react to unforeseen circumstances \cite{efi2018}.
 The definition specifies the needed requirements to make a robot fully autonomous. 
 The automotive industry defines vehicle autonomy in levels based on the human driver's need for supervision and possible intervention during the execution of complex driving tasks \cite{J3016_202104}.
-Figure xx depicts the different levels of autonomy of passenger cars on a scale between no automation to completely autonomous.
+Figure \ref{fig:sae_levels} depicts the different levels of autonomy of passenger cars on a scale between no automation to completely autonomous.
 \\
 
-\begin{figure}
+\begin{figure}[ht]
+	\label{fig:sae_levels}
 	\includegraphics[width=1.0\textwidth]{images/j3016graphic_2021.png}
@@
 
-SAE Levels
-EFI Gutachten
-Fachforum
+%SAE Levels
+%EFI Gutachten
+%Fachforum
 
@@
 Most of the current autonomous driving architectures are implementing a hierarchical structure that follows the "Sense - Think - Act" paradigm \cite{murphy2000}.
-As shown in figure xx the sensory inputs, usually from multiple sensors, get processed to create an environment representation.
+As shown in figure \ref{fig:autonomous_driving_architecture} the sensory inputs, usually from multiple sensors, get processed to create an environment representation.
 Then the system calculates how to get from its current position to the destination and creates a path.
@@
 These motion commands then get converted to control the motors.
+\begin{figure}[ht]
+	\label{fig:autonomous_driving_architecture}
+	\includegraphics[width=1.0\textwidth]{images/autonomous_driving_architecture.png}
+	\caption{Common Autonomous Driving Architecture \cite{brooks1986} \cite{velasco2020}}
+\end{figure}
+\subparagraph*{}
 The planning sequence can be further divided into global planning, behavior planning and local planning.
@@
 The local planning, also named motion planning, module takes the input from the behavior layer and has the task to let the vehicle drive in accordance to the executed behavior, so that it stays in its lane and comes to a stop at a red light \cite{reke2020}.
-\\
-\begin{figure}
-	\includegraphics[width=1.0\textwidth]{images/autonomous_driving_architecture.png}
-	\caption{Common Autonomous Driving Architecture \cite{brooks1986} \cite{velasco2020}  }
-\end{figure}
- \\
 
+\subparagraph*{}
 Due to safety considerations, many autonomous driving architectures implement an additional system supervision layer which monitors the execution of the other system components.
@@
 In case one or more components fail to pass the health check the supervisor is responsible for ensuring that the system does not continue the normal driving routine.
-The supervisor triggers measures to restore the normal functionality or if that is not possible bring the robot into a safe state (zimmermann2020).
-
-Considering that with these architectures, vehicles can achieve an SAE level of autonomy up to level 4 (bacha 2008),  a good starting point to increase autonomy in a robotic system would be to ensure that all of the functional modules in figure xx are implemented and available.
+The supervisor triggers measures to restore the normal functionality or if that is not possible bring the robot into a safe state \cite{zimmermann2020adaptive}.
+Considering that with these architectures, vehicles can achieve an SAE level of autonomy up to level 4 \cite{bacha2008odin},  a good starting point to increase autonomy in a robotic system would be to ensure that all of the functional modules in figure xx are implemented and available.
 On this basis higher levels of autonomy are achievable through the expansion of the behavior planning module inside of the planning layer.
 
-bacha2008
-reke 2020
-wei2014
-brooks1986
-dhillon 2002
-gonzales2016
+%bacha2008
+%reke 2020
+%wei2014
+%brooks1986
+%dhillon 2002
+%gonzales2016
 
@@
 Reactive behaviors can be chained together to result in more advanced and goal-driven behaviors.
+\subparagraph*{}
 Despite the possibility of more complex robot behaviors, sequential behaviors remain on the level of reactive behaviors due to the lack of a planning and decision-making cycle during their execution.
@@
 making decisions and react to unforeseen circumstances.
+\subparagraph*{}
 So in order to improve the autonomy of a robot, the focus should be on the creation of deliberative behaviors.
@@
 This is due to a better understanding of how obstacles are interacting with the planned path of the robot in relation to the time at which the robot and obstacle are predicted to intersect paths.
+\subparagraph*{}
 A purely reactive planner could never determine if an obstacle is destined to intersect the robots path in the future and would reroute constantly in order to accomplish the goal.
@@
 
-\textit{Talk about machine learning as deliberative behavior -> true autonomy bc learning and unknown scenarios, programmers can not cover every scenario, so a well designed/trained ai is needed to make a system really autonomous, but -> missing determinism and unaccountability of machines, hence we need behaviors that are still hard coded and deterministic, these can be highly automated.
+%\textit{Talk about machine learning as deliberative behavior -> true autonomy bc learning and unknown scenarios, programmers can not cover every scenario, so a well designed/trained ai is needed to make a system really autonomous, but -> missing determinism and unaccountability of machines, hence we need behaviors that are still hard coded and deterministic, these can be highly automated.
 traditional behaviour  and ai generated behaviours can and must coexist in the system (adler2019)}
-
-silva2008
-ingrand 2014
-vasiolopolous 2022
+%
+%silva2008
+%ingrand 2014
+%vasiolopolous 2022
 
@@
 This behavior planning module is then able to quickly react to incoming sensor data and still exhibit high levels of automation.
-\\
 
-\begin{figure}
+\begin{figure}[ht]
+	\label{fig:hybrid_planning}
 	\includegraphics[width=1.0\textwidth]{images/Deliberative_hierarchical_planning.png} 
@@
 
-Figure xx shows the areas where reactive and deliberative behaviours have their advantages and limitations.
+Figure \ref{fig:hybrid_planning} shows the areas where reactive and deliberative behaviours have their advantages and limitations.
 Deliberative planning does possess limitations during the execution of generated plans as the time horizon is long-term and not able to react to fast changes in the environment .
@@
 At any given time only a single state is selected and its containing actions are executed.
-\\
 
-\begin{figure}
+\begin{figure}[ht]
+	\label{fig:fsm}
 	\includegraphics[width=1.0\textwidth]{images/fsm.png} 
@@
 
-Figure xx depicts an example state machine with different states and examplatory transitions between them.
+\subparagraph*{}
+Figure \ref{fig:fsm} depicts an example state machine with different states and examplatory transitions between them.
 Inside of the states the letters indicate the existence of actions.
@@
 A state transition diagram is a good way to understand the system behavior but can not show the details of the modedeled behavior, like the actions \cite{wagner2006}.
-\\
-
+\subparagraph*{}
 Creating small state machines can easily and quickly be accomplished through the use of various feature-rich and performant libraries \cite{fourakis2014}. 
@@
 This is due to the modeling effort of the transitions, transition conditions and transition events growing with each new state that gets introduced into the state machine as existing states need to be checked and updated accordingly \cite{conner2017}.
-\\
-
+\subparagraph*{}
 The execution time of finite state machines can be fast enough that they are used to control the motors of a bipedal robot to enable the robot to balance and walk despite unexpected height variations between steps \cite{park2013}.
@@
 
-allgeuer 2013
-connor 2017
-ziegler 2014
+%
+%allgeuer 2013
+%connor 2017
+%ziegler 2014
 
@@
 Leaf nodes are also called execution nodes, while non-leaf nodes are called control or control flow nodes.
+\subparagraph*{}
 The execution of a behavior tree is done by ticking the root node of tree.
@@
 It will only return "failure" if all of the child nodes "returned success".
+All of the basic node types and their control flow modification are listed in table \ref{tab:node_types}.
 
-\begin{table}[h!]
+\begin{table}[ht]
+	\label{tab:node_types}
 	\caption{The five types of nodes \cite{iovino2022}}
@@
   	\hline
-  	Sequence & - > & If all children succeed &  If one child fails & If one child returns running \\ 
+  	Sequence & -$>$ & If all children succeed &  If one child fails & If one child returns running \\ 
   	\hline
@@
   	\hline
-  	Parallel & - > - > & If >= M children succeed & If > N-M children fail & else \\
+  	Parallel & -$>$ -$>$ & If $>$= M children succeed & If $>$ N-M children fail & else \\
   	\hline
@@
 
-Figure xx depicts an example of a BT with multiple levels and action nodes.
-The nodes with the arrow (->) symbol are sequence nodes, the question mark (?) symbol denotes fallback nodes.
+Figure \ref{fig:bt_example} depicts an example of a BT with multiple levels and action nodes.
+The nodes with the arrow (-$>$) symbol are sequence nodes, the question mark (?) symbol denotes fallback nodes.
 The condition node are depicted as white ovals.
@@
 \begin{figure}[h!]
+	\label{fig:bt_example}
 	\includegraphics[width=1.0\textwidth]{images/bt_example.jpg} 
@@
 
-iovino2022
-marzinotto 2014
-colledanchise 2018
+%iovino2022
+%marzinotto 2014
+%colledanchise 2018
 
-\subsection{POMPDs}
+%\subsection{POMPDs}
 

### 2022-09-25 Thesis/text/implementierung.tex
cleanup of tjhe doc

@@
 
-The system architecture, depicted in figure xx, integrates the behavior tree, responsible for the behavior planning, in a position to override Navigation2.
-This is done through the help of the Decision Gate, that can selectively forward speed commands to the robot controller.
-The behavior tree can access the sensor data when needed through the Data Backup component.
-The System Supervision component controls the health of all components, but the connections are not depicted in the component diagram for Übersichtlichkeit.
+The system architecture, depicted in figure \ref{fig:component_diagram}, integrates the behavior tree into the existing architecture. 
+The already existing components are coloured in gray and the newly added components are coloured in blue.
+The new components and their functions are described in the following sections of this chapter.  
+%This is done with the addition of the Decision Gate, that can selectively forward speed commands to the robot controller.
+The behavior tree's commands have a higher priority and disable the forwarding of the Navigation2 commands.
+%The behavior tree can access the sensor data when needed through the Data Backup component.
+The System Supervision component controls the health of all components, but the connections are not depicted in the component diagram for reasons of clarity and comprehensibility.
 
-\begin{figure}[h!]
-	\includegraphics[width=1.05\textwidth]{images/component_diagram_bt.png}
+\begin{figure}[ht]
+	\label{fig:component_diagram}
+	\includegraphics[width=1.00\textwidth]{images/component_diagram_bt.png}
 	\caption{Component Diagram of System Architecture}
@@
 This way the computational load and network load is decreased.
+\paragraph*{}
+
 The BT can access the saved data by sending a service call to one of the provided services by this component.
@@
 The synchronous nature of the ROS executor spinning the nodes is the reason why the data backup is not stored within the BT.
+The saved data inside of this component with the corresponding lookback time is listed in table \ref{tab:data_backup_types}.
 
-The saved data inside of this component with the corresponding lookback time is listed in table xx.
-
-\begin{table}[h!]
+\begin{center}
+\begin{table}[ht]
+	\label{tab:data_backup_types}
 	\caption{Types of saved data}
@@
   	\hline
-  	Name & Type & Lookback Time \\ 
+  	\textbf{Name} & \textbf{Type} & \textbf{Lookback Time} \\ 
   	\hline
@@
 \end{table}
+\end{center}
 
@@
 
-A simplified behavior tree only depicting the top level behavior control nodes is shown in figure xx.
-The tree receives the inital tick signal from the root in an infinite loop.
+A simplified behavior tree only depicting the top level behavior control nodes is shown in figure \ref{fig:top_level_bt}. The tree receives the inital tick signal from the root in an infinite loop. 
+\begin{figure}[ht]
+	\label{fig:top_level_bt}
+	\includegraphics[width=1.0\textwidth]{images/simplified_bt.png}
+	\caption{Top Level Behavior Tree Structure}
+\end{figure}
 The root control node is a sequence node.
@@
 The root sequence is the first step for increasing the robustness of the system to combat competing reactive and deliberative behaviors which could endanger the safety of the robot and other actors in its environments.
-
-\begin{figure}[h!]
-	\includegraphics[width=1.0\textwidth]{images/simplified_bt.png}
-	\caption{Top Level Behavior Tree Structure}
-\end{figure}
 
@@
 This behavior pattern is realized for the lidar, the IMU and the wheel odometry.
-\\
 
-\begin{figure}[h!]
+\begin{figure}[ht]
 	\includegraphics[width=1.0\textwidth]{images/sensor_fallback.png}
@@
 After the BT has checked for sensor failures and can guarantee that the environment representation is accurate, the system can check for more complicated scenarios to improve the autonomy.
+The implemented behaviors are outlined in this section.
 
 \subsubsection{Collision Behavior}
-One of the biggest problems mentioned in chapter 3.1 is the inability to react to collisions.
+
+One of the biggest problems, mentioned in chapter 3.1, is the inability to react to collisions.
 The collision can happen due to inadequate sensory coverage for detecting obstacles or general sensor inadequacy which can not detect obstacles because of their shape or surface material.
@@
 
-\begin{figure}
+\begin{figure}[ht]
+	\label{fig:collision_fallback}
 	\includegraphics[width=1.0\textwidth]{images/collision_fallback.png}
@@
 
+\subparagraph*{}
 The collision checker uses a simulated sensor to detect collisions.
@@
 
-The modification is done in the following manner
-
-\textit{Formula for determining execution time}
-
+%The modification is done in the following manner
+%
+%\textit{Formula for determining execution time}
+\subparagraph*{}
 Regarding the map reset, we have to determine if we are creating a new map with a SLAM method or if the map server provided a previously saved map.
-In case of the running SLAM, the recent updates to the map must be verwerfen to a map prior to when the collision occurred.
-When a static map is provided by the map server component from Nav2 the map needs to be updated with a new obstacle at the point of collision and republished by the map server.
-
-The activity of the "MapUpdateAfterCollision" action node is depicted in figure xx.
+In case of the running SLAM, the recent updates to the map must be verwerfen to a map prior to when the collision occurred.When a static map is provided by the map server component from Nav2 the map needs to be updated with a new obstacle at the point of collision and republished by the map server.
+The activity of the "MapUpdateAfterCollision" action node is depicted in figure \ref{fig:activity_diagram_collision}.
 The collision point is estimated by the robot's pose which refers to the robot base link and the robot footprint.
@@
 The simulated bumper sensor can provide more nuanced information about the point of contact, but this functionality can not be achieved on a real robot with an IMU-based collsion detection, hence we simplified the obstacle position to always be in front of the robot.
-The collision position is calculated with
+%The collision position is calculated with 
+%
+%\textit{Formula here}
 
-\textit{Formula here}
+A 25 by 25 cm square with full occupation of that space is added to the current map and then saved to be reloaded by the map server. 
+%as depicted in figure xx. 
+% add the
 
-A 25 by 25 cm square with full occupation of that space is added to the current map and then saved to be reloaded by the map server as depicted in figure xx.
+\subparagraph*{}
+The original goal is then republished to replan the global path on the updated map. A more detailed view of the whole behavior is depicted in figure \ref{fig:activity_diagram_collision}.
 
-The original goal is then republished to replan the global path on the updated map. A more detailed view of the whole behavior is depicted in figure xx.
-
-\begin{figure}
-	\includegraphics[width=0.9\textwidth]{images/activity_diagram_collision.png}
+\begin{figure}[ht]
+	\label{fig:activity_diagram_collision}
+	\includegraphics[width=0.92\textwidth]{images/activity_diagram_collision.png}
 	\caption{Activity Diagram for the Collision Behavior}
@@
 The robot should not begin the navigation process if it is probable that the remaining charge will not suffice for the driving.
+The behavior tree fallback is depicted in figure \ref{fig:battery_fallback}.
 
-\begin{figure}
+\begin{center}
+\begin{figure}[ht]
+	\label{fig:battery_fallback}
 	\includegraphics[width=0.5\textwidth]{images/battery_fallback.png}
@@
 \end{figure}
+\end{center}
+
 The robot has to take into account the idle consumption, the energy to drive the motors.
@@
 To make the algorithm fail safe, we added an additional safety factor into the function, to berücksichtigen the scenario where the robot has to reroute and drive a longer route or has to stop for a moment.
-The resulting function can be described as: \\
-
-\textit{Formel here}
+%The resulting function can be described as: 
+%
+%\textit{Formel here}
+\subparagraph*{}
 
@@
 The battery behavior fallback will cancel the navigation goal if the robot can not reach the given goal with the remaining charge.
-\\
 
@@
 
-With this behavior, we implement a way to make the global planner more flexible. 
+With this behavior, a way to make the global planner more flexible was implemented. 
 The execution checker component monitors the health of the Nav2 global planner.
@@
 If one of the condition fails, the given navigation goal is not reachable by the planning algorithm.
-We combat this by shifting the goal to a nearby alternative to get the robot as close to the original goal as possible.
+This gets combatted by shifting the goal to a nearby alternative to get the robot as close to the original goal as possible.
 Is the location still unreachable, the goal gets moved further away and closer to the robot again.
 This routine gets repeated for up to ten times, which leads to a distance to the original goal of up to one meter.
-The goal gets moved on a straight vector towards the robot. Other strategies for finding alternative goals were not implemented in this behavior.
+The goal gets moved on a straight vector towards the robot.
+Other strategies for finding alternative goals were not implemented in this behavior.
+The behavior tree for the behavior is pictured in figure \ref{fig:global_planning_fallback}.
 
-\begin{figure}[h!]
+\begin{center}
+	\begin{figure}[ht]
+	\label{fig:global_planning_fallback}
 	\includegraphics[width=0.7\textwidth]{images/global_planning_fallback.png}
@@
 	\end{figure}
+\end{center}
 
@@
 With this methodology, the failure of a sensor can be simulated as the simulation does not offer this functionality.
-Figure xx shows the RQT graph, which depicts the ROS nodes and their subscribed/published topics.
+Figure \ref{fig:rqt_graph} shows the RQT graph, which depicts the ROS nodes and their subscribed/published topics.
 
-\begin{figure}
+\begin{center}
+\begin{figure}[ht]
+	\label{fig:rqt_graph}
 	\includegraphics[width=0.75\textwidth]{images/rqt_graph_sensor_drivers.png}
@@
 \end{figure}
+\end{center}
 
@@
 Because of the neccessity to have ROS nodes spinning outside of the BT thread as discussed in 4.4.1, additional interfaces needed to be created in order to receive informations through service calls.
-The created service types are listed in table xx.
+The created service types are listed in table \ref{tab:custom_interfaces}.
 All of the service definitions are created in a dedicated package named "bt$\_$msgs".
 
-\begin{table}[h!]
+\begin{table}[ht]
+	\label{tab:custom_interfaces}
 	\caption{Implemented Custom Services}

### 2022-09-25 Thesis/text/methoden.tex
cleanup of tjhe doc

@@
 
-\section{Current State}
 This chapter briefly presents the current state of mobile navigation with ROS and its limitations. 
@@
 
+\section{Current State}
 \subsection{ROS}
@@
 ROS can handle the communication between different programs (called nodes) with standardized interfaces and network protocols.
-Nodes can communicate via topics, services, and actions, which differ in their application domain (compare table xx).
+Nodes can communicate via topics, services, and actions, which differ in their application domain (compare table \ref{tab:ros_interfaces}).
 
-\begin{table}[h!]
+\begin{table}[ht]
+	\label{tab:ros_interfaces}
 	\caption{ROS Communication Options}
-	\begin{tabular}{ | m{0.12\textwidth} | m{0.2\textwidth}| m{0.35\textwidth} | m{0.2\textwidth}|} 
+	\begin{tabular}{ | m{0.15\textwidth} | m{0.2\textwidth}| m{0.35\textwidth} | m{0.2\textwidth}|} 
   	\hline
@@
   	\hline
-  	Topics & Publisher/Subscriber & Continuous stream of data e.g. sensor data &  n:m \\ 
+  	Topics & Publisher/ Subscriber & Continuous stream of data e.g. sensor data &  n:m \\ 
   	\hline
-  	Services & Request/Response & Get specific data only once & One service, many clients \\ 
+  	Services & Request/ Response & Get specific data only once & One service, many clients \\ 
   	\hline
-  	Actions & Request/Response and Publish/Subscribe & Trigger executing of asynchronous goal-driven processes and receive updates & One action, many clients \\
+  	Actions & Request/ Response and Publish/ Subscribe & Trigger executing of asynchronous goal-driven processes and receive updates & One action, many clients \\
   	\hline
@@
 
-\begin{figure}[h!]
+\begin{figure}[ht]
+	\label{fig:nav_architecture}
 	\includegraphics[width=1.0\textwidth]{images/nav2_architecture.png}
@@
 \end{figure}
+\subparagraph*{}
 These packages are often integrating a SLAM method to build or enlarge maps.
 Path planning and execution capabilities, which correspond to the global and local planner described in section 2.2, are further supported by ready-to-use planning plugins that use A* and Dijkstra algorithms for path planning and a dynamic window approach for path execution.
-The system sequencing is done in a hierarchical structure as a central behavior tree calls asynchronous actions from the respective planners after another, as seen in figure xx.
+The system sequencing is done in a hierarchical structure as a central behavior tree calls asynchronous actions from the respective planners after another, as seen in figure \ref{fig:nav_architecture}.
 
-\begin{figure} [h!]
+\begin{figure}[ht]
+	\label{fig:bt_nav}
 	\includegraphics[width=1.0\textwidth]{images/nav_bt-modified.png}
@@
 \end{figure}
-
-The behavior tree depicted in figure xx is the one that is the default Nav2 behavior and has a set of recovery behaviors already implemented (spin, wait, back up, clear map).
+\subparagraph*{}
+The behavior tree depicted in figure \ref{fig:bt_nav} is the one that is the default Nav2 behavior and has a set of recovery behaviors already implemented (spin, wait, back up, clear map).
 These behaviors are executed if the robot can not make progress towards the set goal.
@@
 The lifecycle approach allows the system degree to monitor the system node's startups, executions, and failures.
-
+\subparagraph*{}
 With this quantity of functions and hierarchical architecture, the ROS Navigation2 stack is an excellent basis for achieving high levels of robot autonomy.
 The software can be modified and expanded with plugins to fit the users' needs.
-When comparing the functionalities of the software stack to the proposed autonomous system architecture in figure xx (2.2), the things that Nav2 is missing are a system-wide supervision layer and a more deliberative approach to behavior planning, as mentioned in 2.3.
+When comparing the functionalities of the software stack to the proposed autonomous system architecture in figure \ref{fig:autonomous_driving_architectures}, the things that Nav2 is missing are a system-wide supervision layer and a more deliberative approach to behavior planning, as mentioned in chapter 2.3.
+\subparagraph*{}
 The absence of these components leads to a limited set of circumstances in which the robot can perform navigation reliably.
@@
 Examples of such events currently not handled appropriately are slipping wheels, orientation changes by external influences, or undetected, more specifically, undetectable obstacles.
-\textit{Add statistics about random sensor failures every ~10000h} This can lead to erratic movement of the robot, as reactive behaviors and more deliberative behaviors compete for authority over the robot's movement.
+%\textit{Add statistics about random sensor failures every ~10000h} 
+This can lead to erratic movement of the robot, as reactive behaviors and more deliberative behaviors compete for authority over the robot's movement.
 By design, to allow more intelligent robot behaviors, the deliberative behavior can override commands from reactive behaviors.
@@
 
-\textit{Add a closing statement about that suitability of the Nav stack for achieving autonomy}
+%\textit{Add a closing statement about that suitability of the Nav stack for achieving autonomy}
 
@@
 The robot has two motors with attached wheel encoders to drive and steer, classifying it as a differential drive robot. A third omnidirectional wheel stabilizes the robot. 
-
-The manufacturer provides an open-source model for simulating the robot in physics-based simulators like Gazebo, as depicted in figure xx, which enables faster development and testing for the robot. 
+\subparagraph*{}
+The manufacturer provides an open-source model for simulating the robot in physics-based simulators like Gazebo, as depicted in figure \ref{fig:turtlebot}, which enables faster development and testing for the robot.
 The equipped sensors and easy simulation capabilities make the robot well suited for using the Nav2 stack and further development for behavior planning.
 
-The robot specifications are specified in table xx.  \\
+%The robot specifications are specified in table xx.  \\
 
-\begin{figure}[h!]
+\begin{figure}[ht]
+    \label{fig:turtlebot}
 	\includegraphics[width=0.7\textwidth]{images/turtlebot_sim.png}
@@
 
-\begin{table}[h!]
+\begin{table}[ht]
+\label{tab:fn_req}
 \caption{Functional Requirements}
@@
 
-\begin{table}[h!]
+\begin{table}[ht]
 	\caption{Non-Functional Requirements}
-	\begin{tabular}{ | m{0.05\textwidth} | m{0.13\textwidth}| m{0.07\textwidth} | m{0.65\textwidth} |} 
+	\label{tab:nonfn_req}
+	\begin{tabular}{ | m{0.13\textwidth} | m{0.13\textwidth}| m{0.08\textwidth} | m{0.56\textwidth} |} 
   	\hline
-  	Nr. & Name & Priority & Description \\ 
+  	\textbf{Nr.} & \textbf{Name} & \textbf{Priority} & \textbf{Description} \\ 
   	\hline
-  	nfreq1 & Single Point of Failure & High &  The system does not have a single point of failure	When parts of the system fail, the system maintains operability to a degree \\ 
+  	non$\_$freq1 & Single Point of Failure & High &  The system does not have a single point of failure	When parts of the system fail, the system maintains operability to a degree \\ 
   	\hline
-  	nfreq2 & Performance & High & The system's control loop guarantees fast reactions. The average frequency with which the system checks the sensors/navigation2 is higher than 100Hz (10ms) \\ 
+  	non$\_$freq2 & Performance & High & The system's control loop guarantees fast reactions. The average frequency with which the system checks the sensors/navigation2 is higher than 100Hz (10ms) \\ 
   	\hline
-  	nfreq3 & Determinism & High & The outcome for a given set of inputs must be deterministic, meaning that the behavior is always executed similarly. \\
+  	non$\_$freq3 & Determinism & High & The outcome for a given set of inputs must be deterministic, meaning that the behavior is always executed similarly. \\
   	\hline
-  	nfreq4 & Deliberate & High & The robot can execute deliberative behaviors, meaning that behaviors new implemented behaviors go beyond reacting to sensor input and have a planning aspect. \\
+  	non$\_$freq4 & Deliberate & High & The robot can execute deliberative behaviors, meaning that behaviors new implemented behaviors go beyond reacting to sensor input and have a planning aspect. \\
   	\hline
@@
 Also, an external system can more reliably monitor other systems when its execution is detached from the performance and execution of other systems.
-
 A suitable name for the second system is the "autonomy layer", as it aims to increase the robustness and autonomy of the whole system. 
@@
 The constant execution monitoring equips the system with the ability to deal with the event of node failures and component malfunctions.
+\subparagraph*{}
 Another required addition to the autonomy layer is a data storage of relevant system information.
@@
 Using past data points allows behaviors to become more deliberative in their approach and opens up many possibilities for intelligent behaviors like movement predictions of obstacles.
+\subparagraph*{}
 A third component to the autonomy layer is a behavior planner largely independent of Nav2 execution and planning.
@@
 Based on this information, the behavior planner can make a deliberative, strategic decision to increase robustness and autonomy, thus decreasing reliance on human supervision.
+\subparagraph*{}
 Additionally, a component to switch intelligently between speed commands from Nav2 and the autonomy layer is needed.
@@
 The reason for this is mainly the increased flexibility when creating and expanding the behavior planning compared to FSMs.
-Also, behavior trees allow complete determinism and introspection, which is why POMPDs are not suitable as the behavior planning approach for the specified requirements.
+Also, behavior trees allow complete determinism and introspection.
+%which is why POMPDs are not suitable as the behavior planning approach for the specified requirements.
 
@@
 Reactive sequences and reactive fallbacks differ from normal ones in handling nodes that return the state "running".
-Instead of ticking the node again, the whole sequence restarts, which is very useful for continuously checking a condition and executing an asynchronous action node that gets halted when a condition is not met anymore (compare table xx).
+Instead of ticking the node again, the whole sequence restarts, which is very useful for continuously checking a condition and executing an asynchronous action node that gets halted when a condition is not met anymore (compare table \ref{tab:control_nodes_bt}).
 
-\begin{table}[h!]
+\begin{table}[ht]
 	\caption{Control Nodes in BT.CPP}
+	\label{tab:control_nodes_bt}
 	\begin{tabular}{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
@@
 These Decorator Nodes can only have one child node each.
-A comprehensive list of the available nodes and descriptions is in table xx.
+A comprehensive list of the available nodes and descriptions is in table \ref{tab:decorators_bt}.
 
-\begin{table}[h!]
+\begin{center}
+\begin{table}[ht]
 	\caption{Decorator Nodes in BT.CPP}
+	\label{tab:decorators_bt}
 	\begin{tabular}{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
@@
 \end{table}
+\end{center}
 
+\subparagraph*{}
 To allow the tree nodes to communicate, the library provides the developer with two possibilities.
@@
 
-\textit{Screenshot of live monitoring}
+%\textit{Screenshot of live monitoring}
 

### 2022-09-26 Thesis/text/einleitung.tex
grammarly

--- 2022-09-06 Thesis/text/einleitung.tex
@@
 
-Core Problem - too little autonomy in a standard autonomous mobile robot running ROS navigation(AMR)
+Mobile robots are becoming more widespread in many domains.
+They are marketed as autonomous mobile robots, but their autonomy is often limited to particular environments and requires frequent human help to function correctly.
+This reliance on a human operator hardly qualifies the robot to be labeled autonomous.
+\subparagraph*{}
+One of the quickest and most popular ways to build and program a robot is to use the Robot Operating System (ROS).
+ROS offers many tools and functionalities to create and program robots, but a standard robot needs human supervision during the operation to ensure proper function and safety.
+There are ROS packages to safely navigate from point A to point B, but currently, typical robots cannot react to unforeseen events.
+The robot needs to be able to navigate in an uncertain environment to guarantee the safety of the robot itself and all other actors in its environment.
+These can be either other robots or humans.
+\subparagraph*{}
+In theory, the robot can perform fully autonomous navigation, including mapping and localization. However, when something unexpected happens, the robot's autonomy can not be guaranteed any more. 
+A default robot cannot represent all possible fail states and does not detect failures on a systematic level but only inside its navigation-related subsystems.
+Moreover, when failures inside these subsystems are discovered, the options to handle these problems are minimal and often do not deal with the problems in an autonomous way.
+The reliance on external problem-solving decreases the robot's usefulness when tasked with tangible goals.
+Therefore, this thesis centers around the research question of how behavior planning can increase the robustness and autonomy of a robot running ROS2. 
+Behavior planning enables the robot to react to failures and problems of the system and can decide alternative courses of action to mitigate risks and finish the given tasks.
 
-Description of the Problem
+\subparagraph*{}
+This thesis will explore possibilities for improving current systems by adding a dedicated behavior planning component.
+Furthermore, the thesis will provide an implementation of exemplary behaviors and provide a system architecture to incorporate behavior planning in other robots, too.
 
-Robots regurlarly needs operator in case smth unexpected happens\\
+\subparagraph*{}
+The desired outcomes of this implemented software will be an increase in robustness and a decrease in required human actions during several scenarios.
+In these scenarios, failures and problems are artificially induced to test the robot's abilities to behave autonomously.
 
-this may include: unexpected collisions, a system restart is needed, battery is empty\\
+% BEFORE GRAMMARLY BELOW
 
-The system is not able to react and handle unforeseen situations\\
-
-This can lead to safety issues during the operation of a robot. These safety issues can lead to uncontrolled driving maneuvers due to wrong sensor data or a wrong environment represantation.\\
-
-Due to the need of the robot to require an operator, the robot is closer to an automated instead of an autonomous vehicle. 
-Difference is defined by the SAE Levels for automonous driving in cars.\\
-
-It is important for the robot to possess the ability to navigate in an uncertain environment to guarantee the safety of the robot itself and all other actors in its environment, these can be other robots and humans.\\
-
-In theory the robot can perform fully autonomous navigation including mapping and localization but as soon as something out of the ordinary happens the robots autonomy is not guaranteed to be reliable anymore.
-Default robot is not able to represent all possible fail states and does not detect failures on a systematic level, but only inside its navigation related subsystem.
+%\chapter{Introduction}
+%\label{cha:introduction}
+%
+%Mobile robots are becoming more widespread in many domains.
+They are marketed as autonomous mobile robots, but often the autonomy of the robot is limited to very specific environments and require frequent human help to function correctly.
+This hardly qualifies the robot to be labeled autonomous.
+%\subparagraph*{}
+%One of the quickest and most popular ways to build and program a robot is to use the Robot Operating System (ROS).
+ROS offers many tools and functionalities to create and program robots, but even then a normal robot needs human supervision during the operation to ensure the proper function and safety.
+%There are ROS packages to safely navigate from point A to point B, but currently they are lacking the ability to react to unforeseen events.
+It is important for the robot to possess the ability to navigate in an uncertain environment to guarantee the safety of the robot itself and all other actors in its environment, these can be other robots and humans.
+%\subparagraph*{}
+%In theory, the robot can perform fully autonomous navigation including mapping and localization but as soon as something out of the ordinary happens the robots autonomy is not guaranteed to be reliable anymore.
+%Default robot is not able to represent all possible fail states and does not detect failures on a systematic level, but only inside its navigation related subsystem.
 And even when failures inside this subsystems are discovered, the options to handle these problems are very limited and often do not deal with the problems in an autonomous way.
 The reliance on human-needed problem solving decreases the robots usefulness when tasked with real goals.
-
-Having a robust and safe robot behaviour opens the door to many applications in a more diverse set of challenging environments.
+%Therefore, this thesis centers around the research question how behavior planning can increase the robustness and autonomy of a robot running ROS2. 
+%Behavior planning enables the robot to react to failures and problems of the system and can decide alternative courses of action to mitigate risks and finish the given tasks.
+%
+%%Core Problem - too little autonomy in a standard autonomous mobile robot running ROS navigation(AMR)
+%%
+%%Description of the Problem:
+%%
+%%Despite the advancements in autonomous driving the robot regurlarly needs an operator in case smth unexpected happens.
+%%
+%%This may include: unexpected collisions, a system restart is needed, battery is empty, the environment is too crowded and the robot gets stuck
+%%
+%%The system is not able to react and handle unforeseen situations.
+%%
+%%This can lead to safety issues during the operation of a robot. These safety issues can lead to uncontrolled driving maneuvers due to wrong sensor data or a wrong environment represantation.
+%%
+%%Due to the need of the robot to require an operator for safe, the robot is not really autonomous despite it being labeled as such. 
+%%
+%%(Difference is defined by the SAE Levels for automonous driving in cars.) \\
+%%
+%%Relevance, need for research:\\
+%%\\
+%
+%%
+%%Having a robust and safe robot behaviour opens the door to many applications in a more diverse set of challenging environments.
 Usually a reinforcement learning approach with broad training data set is a great way to handle a multitude of unknown environments and situations, but one of the biggest problems is the lacking determinism in such systems.
-
-In the robotic world exists a need for a deterministic system which can enhance and override the decision making process and navigation capabilities of the robot.
-
-This thesis proposes a comprehensive approach for behavior planning for an autonomous mobile robot.
-
-\textbf{How can behavior planning increase the safety and autonomy of an autonomous mobile robot running ROS?}
+%%
+%%In the robotic world exists a need for a deterministic system which can enhance and override the decision making process and navigation capabilities of the robot. 
+%%
+%%Planned solution:\\
+%\subparagraph*{}
+%This thesis will explore possibilities how current systems can be improved through the addition of a dedicated behavior planning component.
+Furthermore, the thesis will provide an implementation of examplatory behaviors and provide a system architecture to incorporate behavior planning in other robots, too
+%
+%%The thesis will explore how an extension of the current software architecture and behavior planning capabilities for an autonomous mobile robot that is running ROS2 can improve the autonomy of the robot.
+%%
+%%Expected results:\\
+%\subparagraph*{}
+%The desired outcomes of these implemented software will be an increase of robustness and decrease of required human actions during a number of scenarios.
+In these scenarios, failures and problems are artificially induced to test the robots abilities to behave autonomously.
+%
+%
+%%\textit{Research gap is not named so far}

### 2022-09-26 Thesis/text/ergebnisse.tex
grammarly

--- 2022-09-06 Thesis/text/ergebnisse.tex
@@
 
+The robot is put into specifically designed scenarios that should trigger behavior responses from the behavior tree component.
+The scenario testing is done to evaluate the efficacy of the implemented behaviors in relation to the functional requirements listed in chapter 3.2.
+
+The scenarios are executed in a simulated apartment environment as seen in figure \ref{fig:house_gazebo} and a simulated world with small round obstacles in an enclosed space as depicted in figure \ref{fig:world_gazebo}.
+\begin{center}
+\begin{figure}[ht]
+	\includegraphics[width=0.75\textwidth]{images/house_env.png}
+	\caption{Apartment Environment in Gazebo}
+	\label{fig:house_gazebo}
+\end{figure}
+\end{center}
+
+\begin{center}
+\begin{figure}
+	\includegraphics[width=0.75\textwidth]{images/world_env.png}
+	\caption{Enclosed Environment in Gazebo}
+	\label{fig:world_gazebo}
+\end{figure}
+\end{center}
+
+The environments were mapped beforehand with the ROS2 "Slam Toolbox" package and provided via the Nav2 map server.
+
+The robot location and navigation goals are defined in another ROS Node, which programmatically spawns the robot into the environment and sets the goal this way.
+This way, the same scenario can be run with and without behavior planning.
+To test the behavior of the sensor failure fallbacks, the respective sensor driver, mentioned in chapter 4.5.1, was shut down to emulate the sensor's failure.
+
 \section{Scenarios}
+
+The scenarios are derived from the functional requirements and can be used to test multiple requirements as listed in the tables \ref{tab:sensor_scenarios} and \ref{tab:behavior_scenarios}.
+The acceptance criteria often contain multiple binary conditions that can only be met or failed. All acceptance criteria for a given scenario must be met for the test to be counted as a success.
+
+For the Lidar, IMU, and Odometry scenarios, the sensor driver was shut down when reaching the desired speed listed in the scenario. 
+Scenario NoPathFound1 is a scenario in which the lifecycle state gets transitioned into "inactive" and can not execute the planning action anymore when triggered.
+The BT has to recover from the inability to plan in this scenario.
+NoPathFound$\_$2 is simulating the scenario when the planner is active but fails to find a collision-free path to the desired goal.
+This scenario tests the ability to find alternative goals and navigate to a nearby goal.
+In the Collision$\_$1 scenario, an obstacle gets moved so close to the robot that a collision is registered.
+The forced collision tests the ability to escape a collision and resume navigation afterward.
+A new navigation goal is published after the behavior tree exits the Collision Fallback Routine.
+Collision$\_$2 scenario is testing the same behavior but during active navigation.
+A flat obstacle is spawned in the path of the robot.
+The obstacle is too low to be detected by the laserscanner, and a collision occurs.
+The robot must be able to recover from the crash and keep on navigating towards the goal with an updated global path in this scenario.
+The Battery$\_$1 scenario tests if the robot tries to navigate toward a distant goal with a low battery charge.
+The simulated battery is emptied to a low level, and a new navigation goal is set.
+If the battery charge runs below zero percent during the navigation, the test fails.
+
+\begin{table}[ht]
+	\caption{Sensor Scenarios}
+	\label{tab:sensor_scenarios}
+	\begin{tabular}{| m{0.1\textwidth} | m{0.17\textwidth}| m{0.3\textwidth} | m{0.3\textwidth}|} 
+  	\hline
+  	\textbf{Name} & \textbf{Related Requirements }& \textbf{Description} & \textbf{Success Criteria}\\ 
+  	\hline
+  	Lidar$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Lidar node crashes & Reset the system\\ 
+  	\hline
+  	Lidar$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
+  	\hline
+  	Lidar$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
+  	\hline
+  	IMU$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, IMU node crashes & Reset system \\
+  	\hline
+  	IMU$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
+  	\hline
+  	IMU$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
+  	\hline
+  	Odom$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Odom node crashes & Reset the system\\ 
+  	\hline
+  	Odom$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
+  	\hline
+  	Odom$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
+  	\hline
+ 	\end{tabular}
+\end{table}
+
+\begin{table}[ht]
+	\caption{Behavior Scenarios}
+	\label{tab:behavior_scenarios}
+	\begin{tabular}{| m{0.1\textwidth} | m{0.13\textwidth}| m{0.33\textwidth} | m{0.34\textwidth}|} 
+	\hline
+	\textbf{Name} & \textbf{Related Requirements} & \textbf{Description} & \textbf{Success Criteria}\\ 
+  	\hline	
+  	NoPath Found$\_$1 & fn$\_$req7 & Path to goal can not be calculated, robot is standing still &  
+Reset system, calculate path to goal, reach goal \\
+	\hline
+	NoPath Found$\_$2 & fn$\_$req7, fn$\_$req8 & Goal is unreachable, robot is standing still & Alternative goals, close to the original are tested to be reachable \\
+	\hline
+	Collision$\_$1 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & Robot is standing still and a collision with the robot is caused & Robot can get out of collision state, navigation to goals still working \\
+	\hline
+	Collision$\_$2 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & The robot is driving and collides with an undetected obstacle (0.25m/s) &  
+Robot can get out of collision state, navigation to goals still working, undetected obstacle gets added to map \\
+	\hline
+	Battery$\_$1 & fn$\_$req9 & The robot battery runs low & The robot will not drive to a goal that is outside of its reachable range \\
+	\hline
+	Motor$\_$1 & fn$\_$req2, fn$\_$req5 & Hardware failure on the motors & n.a. \\
+	\hline
+
+	\end{tabular}
+\end{table}
+
+\section{Results}
+
+The results in table \ref{tab:results} show an increase in the successful handling of the test scenarios for most test cases.
+The system supervision component is very reliable and enables the system to recover from an unexpected sensor failure.
+The more advanced behaviors implemented can improve the autonomous handling of the scenarios.
+The Battery Scenario was successful every time it was induced.
+However, the path planning behavior for finding alternative goals was unsuccessful about 50 percent of the time.
+The test showed that alternative methods for finding new goals would probably have resulted in better performance.
+The collision behavior during driving worked well, and the map updates are an essential feature for the success of the scenarios.
+Nevertheless, there is a lack of intelligence and autonomy when the robot stands still.
+A collision is forced, but the implemented behavior relies on the same mechanism as during the driving scenario.
+The missing differentiation led to a test case where the robot reversed into a wall because it was turned 90 degrees by the collision.
+That test failed because a human operator would be needed to correct the robot.
+This lack of robustness was not apparent in the driving collision scenario because the robot's force was insufficient to cause a significant shift in orientation when colliding with obstacles.
+
+\begin{table}[ht]
+
+	\caption{Results}
+	\label{tab:results}
+	\begin{tabular}{| m{0.11\textwidth} | m{0.13\textwidth}| m{0.33\textwidth} | m{0.33\textwidth}|} 
+  	\hline
+  	\textbf{Name} & \textbf{Number of Runs} &  \textbf{Percentage Succesful Without Behavior Planning} & \textbf{Percentage Succesful With Behavior Planning}\\ 
+  	\hline
+  	Lidar$\_$1 & 5 & 0 & 100 \\ 
+  	\hline
+  	Lidar$\_$2 & 5 & 0 & 100 \\ 
+  	\hline
+  	Lidar$\_$3 & 5 & 0 & 100 \\ 
+  	\hline
+  	IMU$\_$1 & 5 & 0 & 100 \\ 
+  	\hline
+  	IMU$\_$2 & 5 & 0 & 100 \\ 
+  	\hline
+  	IMU$\_$3 & 5 & 0 & 100 \\ 
+  	\hline
+  	Odom$\_$1 & 5 & 0 & 100 \\ 
+  	\hline
+  	Odom$\_$2 & 5 & 0 & 100 \\ 
+  	\hline
+  	Odom$\_$3 & 5 & 0 & 100 \\ 
+  	\hline
+  	NoPath Found$\_$1 & 20 & 0 & 100 \\ 
+  	\hline
+  	NoPath Found$\_$2 & 20 & 0 & 55 \\ 
+  	\hline
+  	Collision$\_$1 & 20 & 0 & 95 \\ 
+  	\hline
+  	Collision$\_$2 & 20 & 0 & 100 \\ 
+  	\hline
+  	Battery$\_$1 & 10 & 0 & 100 \\ 
+  	\hline
+  	Motor$\_$1 & 10 & 0 & 0 \\ 
+  	\hline
+	\end{tabular}
+\end{table}
+
+The architectural design choices meet all the non-functional requirements. 
+Implementing the whole behavior planning and system supervision system eliminates a single point of failure for the system.
+Even when the Navigation2 system collapses, is the robot able to move to a degree and, more importantly, stop completely (compare table \ref{tab:tab:nonfn_req}, non$\_$fn$\_$req).
+The behavior tree and algorithms implemented in the behavior tree are all deterministic (non$\_$fn$\_$req3).
+The robots' behavior goes beyond pure reactivity when navigating through environments, as the advanced behaviors use a dedicated planning phase before the actions are carried out (non$\_$fn$\_$req4).
+Finally, the system's performance when checking the conditions of the BT is below the desired threshold of 10ms, as the whole system runs in multiple threads, which allows the behavior tree to complete one complete execution cycle in about 5ms with the current system.
+(non$\_$fn$\_$req2).

### 2022-09-26 Thesis/text/methoden.tex
grammarly

@@
 \begin{table}[ht]
+	\caption{ROS Communication Options}
 	\label{tab:ros_interfaces}
-	\caption{ROS Communication Options}
 	\begin{tabular}{ | m{0.15\textwidth} | m{0.2\textwidth}| m{0.35\textwidth} | m{0.2\textwidth}|} 
   	\hline
-  	Name & Communication Pattern & Area of Use & Cardinality \\ 
+  	\textbf{Name} & \textbf{Communication Pattern} & \textbf{Area of Use} & \textbf{Cardinality} \\ 
   	\hline
@@
 \begin{figure}[ht]
-	\label{fig:nav_architecture}
 	\includegraphics[width=1.0\textwidth]{images/nav2_architecture.png}
 	\caption{Navigation2 Architecture \cite{macenski2020}}
+	\label{fig:nav_architecture}
 \end{figure}
@@
 \begin{figure}[ht]
-	\label{fig:bt_nav}
 	\includegraphics[width=1.0\textwidth]{images/nav_bt-modified.png}
 	\caption{Navigation2 Behavior Tree}
+	\label{fig:bt_nav}
 \end{figure}
@@
 \begin{figure}[ht]   
-    \label{fig:turtlebot}
 	\includegraphics[width=0.7\textwidth]{images/turtlebot_sim.png}
 	\caption{Turtlebot3 model in Gazebo simulator}
+	\label{fig:turtlebot}
 \end{figure}
@@
 The software requirements that improve the robot's behavior are derived from the current limitations of the standard ROS and Navigation2 setup for mobile robots described in the previous chapter. 
+The functional requirements and their acceptance criteria are listed in \ref{tab:fn_req}. The non-functional requirements are listed seperately in \ref{tab:nonfn_req}. 
+To better contextualize the context of the systems requirements, figure \ref{fig:use_case} depicts a Use-Case diagram with the needed functions to improve the system's autonomy.
 
 \begin{table}[ht]
+\caption{Functional Requirements}
 \label{tab:fn_req}
-\caption{Functional Requirements}
-	\begin{tabular}{| m{0.08\textwidth} | m{0.13\textwidth}| m{0.07\textwidth} | m{0.62\textwidth}|} 
+	\begin{tabular}{| m{0.08\textwidth} | m{0.13\textwidth}| m{0.09\textwidth} | m{0.60\textwidth}|} 
   	\hline
-  	Nr. & Name & Priority & Description/Acceptance Criteria \\ 
+  	\textbf{Nr.} & \textbf{Name} & \textbf{Priority} & \textbf{Description/Acceptance Criteria} \\ 
   	\hline
@@
 	\label{tab:nonfn_req}
-	\begin{tabular}{ | m{0.13\textwidth} | m{0.13\textwidth}| m{0.08\textwidth} | m{0.56\textwidth} |} 
+	\begin{tabular}{ | m{0.13\textwidth} | m{0.13\textwidth}| m{0.1\textwidth} | m{0.55\textwidth} |} 
   	\hline
@@
 \end{table}
+
+\begin{figure}[ht]   
+	\includegraphics[width=0.7\textwidth]{images/use_case.png}
+	\caption{Use Case Diagram}
+	\label{fig:use_case}
+\end{figure}
 
@@
   	\hline
-  	Name & Child returns Failure & Child Returns Running \\ 
+  	\textbf{Name }& \textbf{Child returns Failure} & \textbf{Child Returns Running} \\ 
   	\hline
@@
   	\hline
-  	Name & Succeeds & Fails & Running \\ 
+  	\textbf{Name} & \textbf{Succeeds} & \textbf{Fails} & \textbf{Running} \\ 
   	\hline

### 2022-09-26 Thesis/text/verwandt.tex
grammarly

@@
 This thesis looks for a way to improve the autonomy of mobile robots.
-In order to define what qualifies as an improvement in the autonomous behaviour of a mobile robot, one needs to look at the defintion of automated and autonomy.
-The "Expertenkommission Forschung und Innovation" (EFI) defines the term autonomy in the context of robotics as a system which can act without human instructions and still solve complex tasks, make decisions, learn independently aswell as react to unforeseen circumstances \cite{efi2018}.
+In order to define an improvement in autonomous behavior, one needs to explore the definitions of the terms "automated" and "autonomy".
+The "Expertenkommission Forschung und Innovation" (EFI) defines the term autonomy in the context of robotics as a system that can act without human instructions and still solve complex tasks, make decisions, learn independently as well as react to unforeseen circumstances \cite{efi2018}.
 The definition specifies the needed requirements to make a robot fully autonomous.
-The automotive industry defines vehicle autonomy in levels based on the human driver's need for supervision and possible intervention during the execution of complex driving tasks \cite{J3016_202104}.
+\begin{figure}[ht]
+	\includegraphics[width=1.0\textwidth]{images/j3016graphic_2021.png}
+	\caption{SAE Levels Of Driving Automation \cite{J3016_202104}}
+	\label{fig:sae_levels}
+\end{figure}
+The automotive industry defines vehicle autonomy based on the human driver's need for supervision and possible intervention when executing complex driving tasks \cite{J3016_202104}.
 Figure \ref{fig:sae_levels} depicts the different levels of autonomy of passenger cars on a scale between no automation to completely autonomous.
-\\
+Both definitions rely on human supervision and guidance as the central part of what hinders a system or vehicle from becoming fully autonomous.
+By decreasing the instances of human intervention during the operation of a robot, one can increase the automation level towards full autonomy.
+However, this has to be done by equipping the robot with robust and context-driven decision-making and planning capabilities.
+Otherwise, a robot that never needs a human to operate could easily be created.
+Nevertheless, this robot could not be considered autonomous as it would not be able to solve complex tasks and make intelligent decisions.
 
-\begin{figure}[ht]
-	\label{fig:sae_levels}
-	\includegraphics[width=1.0\textwidth]{images/j3016graphic_2021.png}
-	\caption{SAE Levels Of Driving Automation \cite{J3016_202104}  }
-\end{figure}
-
-Both definitions entail the reliance on human supervision and guidance as the central part of what hinders a system or vehicle to become autonomous.
-By decreasing the instances a human intervention during the operation of a robot, one can increase the automation level towards full autonomy.
-But this has to be done by equipping the robot with robust and context-driven decision-making and planning capabilities, otherwise a robot which would never need a human to operate could easily be created.
-But this robot could not be considered autonomous as it would not able to solve complex tasks and make intelligent decisions.
-
-Different levels of autonomy require different methodologies to achieve them.
+%Different levels of autonomy require different methodologies to achieve them.
 
@@
 
-To gain a better understanding how behavior planning influences the autonomy of a robot, one can take a look at the latest and most used software architectures for autonomous driving on a functional level.
-Most of the current autonomous driving architectures are implementing a hierarchical structure that follows the "Sense - Think - Act" paradigm \cite{murphy2000}.
-As shown in figure \ref{fig:autonomous_driving_architecture} the sensory inputs, usually from multiple sensors, get processed to create an environment representation.
-Then the system calculates how to get from its current position to the destination and creates a path.
-A second planning calculation is triggered to compute the motion commands with the constraints of the system (e.g.
-size, turning radius, etc.).
-These motion commands then get converted to control the motors.
+To better understand how behavior planning influences a robot's autonomy, one can look at the latest and most used software architectures for autonomous driving on a functional level. 
+Most of the current autonomous driving architectures implement a hierarchical structure that follows the "Sense - Think - Act" paradigm \cite{murphy2000}. 
 \begin{figure}[ht]
-	\label{fig:autonomous_driving_architecture}
 	\includegraphics[width=1.0\textwidth]{images/autonomous_driving_architecture.png}
 	\caption{Common Autonomous Driving Architecture \cite{brooks1986} \cite{velasco2020}}
+	\label{fig:autonomous_driving_architecture}
 \end{figure}
-\subparagraph*{}
-The planning sequence can be further divided into global planning, behavior planning and local planning.
-Global planning,also named route planning, is responsible for outputting a path from start to goal.
-This is comparable to a person using Google Maps to find the shortest or fastest path to a far away city.
-In this analogy, the behavior planner is responsible to follow the traffic rules on the trip to the destination, e.g.
-stopping at red lights, giving right of way to other vehicles, following the speed limit.
-The local planning, also named motion planning, module takes the input from the behavior layer and has the task to let the vehicle drive in accordance to the executed behavior, so that it stays in its lane and comes to a stop at a red light \cite{reke2020}.
+As shown in figure \ref{fig:autonomous_driving_architecture}, the sensory inputs, usually from multiple sensors, get processed to create an environment representation.
+Then the system calculates how to get from its current position to the destination and creates a path.
+A second planning calculation is triggered to compute the motion commands with the system's constraints (e.g., size, turning radius).
+These motion commands then get converted to control the motors.
 
 \subparagraph*{}
-Due to safety considerations, many autonomous driving architectures implement an additional system supervision layer which monitors the execution of the other system components.
+The planning sequence can be further divided into global, behavioral, and local planning.
+Global planning, also named route planning, is responsible for outputting a path from start to goal.
+Global planning is comparable to a person using Google Maps to find the shortest or fastest path to a faraway city.
+In this analogy, the behavior planner is responsible for following the traffic rules on the trip to the destination, e.g., stopping at red lights, giving the right of way to other vehicles, and following speed limits.
+The local planning, also named motion planning, takes the input from the behavior layer and is tasked to let the vehicle drive according to the executed behavior so that it stays in its lane and stops at a red light at the correct position\cite{reke2020}.
+
+\subparagraph*{}
+Due to safety considerations, many autonomous driving architectures implement an additional system supervision layer that monitors the execution of the other system components.
 This supervisor checks the health of all software components.
-In case one or more components fail to pass the health check the supervisor is responsible for ensuring that the system does not continue the normal driving routine.
-The supervisor triggers measures to restore the normal functionality or if that is not possible bring the robot into a safe state \cite{zimmermann2020adaptive}.
+If one or more components fail to pass the health check, the supervisor is responsible for ensuring that the system does not continue the regular driving routine.
+The supervisor triggers measures to restore the standard functionality or, if that is not possible, bring the robot into a safe state \cite{zimmermann2020adaptive}.
 Considering that with these architectures, vehicles can achieve an SAE level of autonomy up to level 4 \cite{bacha2008odin},  a good starting point to increase autonomy in a robotic system would be to ensure that all of the functional modules in figure xx are implemented and available.
-On this basis higher levels of autonomy are achievable through the expansion of the behavior planning module inside of the planning layer.
+On this basis, higher autonomy levels are achievable by expanding the behavior planning module inside the planning layer.
 
@@
 
-The behavior planning module contains a set of behaviors to operate in a environment.
-The task of the module is then to choose the best behavior to execute.
-In this sense a behavior is defined as a mapping of sensory inputs to a pattern of actions that in turn carry out a specific task \cite{murphy2000}.
-A good behavior planning approach provides the robot with adequate behaviors for every possible scenario the robot is operating in.
-Different scenarios pose different challenges for the behavior planning module which has to decide on the best behavior option to execute in order to achieve a higher level goal.
+The behavior planning module contains a set of behaviors to operate in an environment.
+The module's task is then to choose the best behavior to execute.
+In this sense, behavior is defined as a mapping of sensory inputs to a pattern of actions that carry out a specific task \cite{murphy2000}.
+A good behavior planning approach provides the robot with acceptable behaviors for every possible scenario the robot is operating in.
+Different scenarios pose different challenges for the behavior planning module, which has to decide on the best behavior option to achieve a higher level goal.
 
 \subsection{Reactive}
-During the emergence of behavior based robots, the first type of behavior that was implemented in many robots was a simple reactive behavior.
-This behavior maps a sensory input directly to motor commands.
-In human behavior this type of behavior resembles reflexes, like the tapping on the kneecap which unwillingly results in motion in the knee joint.
-This behavior pattern is not following the typical "Sense, Think, Act" loop, but shortcuts directly from sensing to acting \cite{desilva2008}.
-Reactive behaviors can be chained together to result in more advanced and goal-driven behaviors.
+During the emergence of behavior-based robots in the 1960s, the first type of behaviors implemented were simple reactive behaviors.
+This behavior type maps a sensory input directly to motor commands.
+In human behavior, this behavior resembles reflexes, like tapping on the kneecap, which unwillingly results in motion in the knee joint.
+This behavior pattern is not following the typical "Sense, Think, Act" loop but shortcuts directly from sensing to acting \cite{desilva2008}.
+Reactive behaviors can be chained together to achieve more advanced and goal-driven behaviors.
 \subparagraph*{}
 Despite the possibility of more complex robot behaviors, sequential behaviors remain on the level of reactive behaviors due to the lack of a planning and decision-making cycle during their execution.
-The computational load of reactive behaviours is low and thereby fast as no decision and planning process is taking place, which makes them suited in scenarios where real-time safety is of concern.
-This property makes these kind of behaviours useful for system supervisor, where sometimes immediate reactions with low latency are required to ensure vehicle safety.
-But a system with a reactive behavior planning approach will always fail to meet the requirements for higher levels of autonomy due to the fact.
-This does not mean that reactive behaviors can not be part in highly autonomous systems as their quick response time to sensor inputs makes them valuable in improving vehicle safety and reliability.
+The computational load of reactive behaviors is low and thereby fast as no decision and planning process is taking place, which makes them suited in scenarios where real-time safety is of concern.
+This property makes this behavior useful for system supervisors, where immediate reactions with low latency are sometimes required to ensure system safety.
+However, a system with a reactive behavior planning approach will always fail to meet the requirements for higher levels of autonomy due to the reasons discussed in previous chapters.
+This systematic lack of autonomy does not mean reactive behaviors can not be part of highly autonomous systems.
+Their quick response time to sensor inputs makes them valuable in improving vehicle safety and reliability.
 
 \subsection{Deliberative}
-Reactive behaviors suffer the drawback that they are not suitable enable complex behaviors.
-In addition, the creation of sequences of reactive behaviors that produce the target behavior is a complicated task \cite{murphy2000}.
-Behaviors that involve a planning and decision making aspect are defined as deliberative.
-Unlike reactive behaviors, deliberative behaviors are following the Sense, Think, Act paradigm.
-They are not mapping sensory inputs directly into motor commands.
-Instead deliberative-type behaviors are deciding the best course of action before acting and only then proceed with the exution.
-The use of deliberative behaviors in the planning module is what enables a robot to meet the defintion of autonomy, e.g.
-making decisions and react to unforeseen circumstances.
+Reactive behaviors suffer the drawback that they are not suitable to enable complex behaviors.
+In addition, creating sequences of reactive behaviors that produce the target behavior is a complicated task \cite{murphy2000}.
+Behaviors that involve planning and decision-making aspects are defined as deliberative.
+Unlike reactive behaviors, deliberative behaviors follow the Sense, Think, Act paradigm, and they are not mapping sensory inputs directly into motor commands.
+Instead, deliberative-type behaviors can decide the best course of action before they act.
+Using deliberative behaviors in the planning module enables a robot to meet the definition of autonomy, e.g., making decisions and reacting to unforeseen circumstances.
 \subparagraph*{}
-So in order to improve the autonomy of a robot, the focus should be on the creation of deliberative behaviors.
-A behaviour planner with a deliberative approach makes decision based on current sensor data, as well as previously processed data.
-By incorporating older sensor information into the decision making process, the deliberative behavior planner can make predictions about the changes that will happen in the environment.
-This allows the planner to proactively change the motion planning commands to better adapt to the environment.
-For example, a motion prediction of obstacles in highly dynamic environments would improve the motion planning considerably.
-This is due to a better understanding of how obstacles are interacting with the planned path of the robot in relation to the time at which the robot and obstacle are predicted to intersect paths.
-\subparagraph*{}
-A purely reactive planner could never determine if an obstacle is destined to intersect the robots path in the future and would reroute constantly in order to accomplish the goal.
-A deliberative planner could determine if the best course of action is to stay on the current path as the dynamic obstacle has already moved away by the time the robot reaches to intersection point.
-Other actions in scenario are possibly slowing down or speeding up briefly to avoid an obstacle and staying on the current path as this is calculated to be quicker than rerouting and taking a longer path.
-Generally, these cognitively more complex behaviours mimic human skills and thus make a system more autonomous.
+Concluding, the focus should be on creating deliberative behaviors to improve autonomy.
+A behavior planner with a deliberative approach decides based on current sensor data and previously processed data.
+By incorporating older sensor information into the decision-making process, the deliberative behavior planner can predict the changes in the environment.
+Using older data points allows the planner to proactively change the motion planning commands to better adapt to the environment.
+For example, a motion prediction of obstacles in highly dynamic environments would improve motion planning considerably.
+The behavior planner can adjust the motion planning with improved information on how dynamic obstacles interact with the robot's planned path in the future.
+A purely reactive planner could never determine if an obstacle is destined to intersect the robot's path in the future and would constantly reroute to accomplish the given goal.
+On the other hand, a deliberative planner can determine if the best course of action is to stay on the current path as the dynamic obstacle has already moved away by the time the robot reaches the intersection point.
+Other actions in the scenario are possibly slowing down or speeding up briefly to avoid an obstacle and staying on the current path, which is calculated to be quicker than rerouting and taking a longer path.
+Generally, these cognitively more complex behaviors mimic human skills and thus make a system more autonomous.
 
@@
 \subsection{Hybrid}
-To create reliable aswell as intelligent systems one can combine reactive and deliberative behaviours in a hybrid model.
-This behavior planning module is then able to quickly react to incoming sensor data and still exhibit high levels of automation.
+One can combine reactive and deliberative behaviors in a hybrid model to create reliable and intelligent systems.
+This behavior planning module can quickly react to incoming sensor data and still exhibit high levels of automation.
 
 \begin{figure}[ht]
+	\includegraphics[width=1.0\textwidth]{images/Deliberative_hierarchical_planning.png} 
+	\caption{Hybrid Planning Architecture \cite{arkin1998}}
 	\label{fig:hybrid_planning}
-	\includegraphics[width=1.0\textwidth]{images/Deliberative_hierarchical_planning.png} 
-	\caption{Hybrid Planning Architecture \cite{arkin1998}  }
 \end{figure}
 
-Figure \ref{fig:hybrid_planning} shows the areas where reactive and deliberative behaviours have their advantages and limitations.
-Deliberative planning does possess limitations during the execution of generated plans as the time horizon is long-term and not able to react to fast changes in the environment .
-When the system is not adressing these points, deliberative robots are very focussed on a small problem domain and are not robust \cite{arkin1998}.
-The incorporation of reactive behaviors into the behavior planning combats this problem.
-In this multi-layer approach the higher level, more deliberative planners are able to override the reactive planners in order to allow the system to be more flexible and adaptible but still maintain fast reaction times.
-This architecture leads to more robust robots that are able to be deployed in a wider variety of environments, thus leading to higher levels of autonomy.
+Figure \ref{fig:hybrid_planning} shows the areas where reactive and deliberative behaviors have advantages and limitations.
+Deliberative planning does possess limitations during the execution of generated plans as the time horizon is long-term and not able to react to fast changes in the environment.
+When the system is not addressing these points, deliberative robots are very focused on a small problem domain and are not robust \cite{arkin1998}.
+The incorporation of reactive behaviors into behavior planning combats this problem.
+In this multi-layer approach, the higher level, more deliberative planners can override the reactive planners to allow the system to be more flexible and adaptable while maintaining fast reaction times.
+This architecture leads to more robust robots deployed in a wider variety of environments, thus leading to higher levels of autonomy.
 
@@
 
-Robots that use a hybrid, hierarchical behavior planning approach need a system which decides on a high level which behaviors are to be executed.
-This allows overriding simple reactive behaviors with more complex, deliberative behaviors when the system benefits from it.
+Robots that use a hybrid, hierarchical behavior planning approach need a system that decides on a high level which behaviors are to be executed.
+That system allows overriding simple reactive behaviors with more complex, deliberative behaviors when the whole system benefits from them.
 
-The different solution for such behavior execution system all share the fact that they make use of states a robot can be in.
-Based upon a robots state and additional information about the environment, different strategies and behaviors are executed during the runtime of the system.
+The different solutions for behavior execution systems all share that they use states a robot can be in.
+Different strategies and behaviors are executed during the system's runtime based on a robot's state and additional information about the environment.
+
+In the following sections, different approaches for behavior planning are presented and compared.
 
@@
 
-Finite State Machines (FSM) are commonly used in the environment of autonomous driving.
-They are used to describe behavior in a form of states, which in turn execute action.
-A finite state machine is defined by a list of possible states S, a starting state s0, a set of state transitions delta, a set of final states F and the input alphabet sigma.
-At any given time only a single state is selected and its containing actions are executed.
+Finite State Machines (FSM) are commonly used in the autonomous driving domain.
+They describe behavior in the form of states, which trigger the execution of actions.
+A finite state machine is defined by a list of possible states S, a starting state s0, a set of state transitions delta, a set of final states F, and the input alphabet sigma.
+At any given time, only a single state is selected, and its containing actions are executed.
 
 \begin{figure}[ht]
+	\includegraphics[width=1.0\textwidth]{images/fsm.png} 
+	\caption{Example of a finite state machine transition diagram \cite{wagner2006}}
 	\label{fig:fsm}
-	\includegraphics[width=1.0\textwidth]{images/fsm.png} 
-	\caption{Example of a finite state machine transition diagram \cite{wagner2006} }
 \end{figure}
@@
 \subparagraph*{}
-Figure \ref{fig:fsm} depicts an example state machine with different states and examplatory transitions between them.
+Figure \ref{fig:fsm} depicts an example state machine with different states and exemplary transitions between them.
 Inside of the states the letters indicate the existence of actions.
@@
 The arrows between the states illustrate the possible transition and transition conditions for every state.
-A state transition diagram is a good way to understand the system behavior but can not show the details of the modedeled behavior, like the actions \cite{wagner2006}.
+A state transition diagram is an excellent way to understand the system behavior but can not show the details of the modeled behavior, like the actions \cite{wagner2006}.
 \subparagraph*{}
-Creating small state machines can easily and quickly be accomplished through the use of various feature-rich and performant libraries \cite{fourakis2014}. 
-Finite state machines are often being used for comprehensive modeling of reactive behavior and sequential behaviors in autonomous driving functions \cite{ziegler2014}.
-These finite state machines make use of nested state machines inside of another state machine.
-This reduces the complexity of the state machine as the transition do not need to be modeled, because the start and final state of a nested state machine is treated as just one state inside of the larger state machine.
-Despite this possibility to simplify state machines, as a modeled behavior system grows larger to accomodate more complex behaviors into the state machine the effort to expand and maintain the state machine grows rapidly.
-This is due to the modeling effort of the transitions, transition conditions and transition events growing with each new state that gets introduced into the state machine as existing states need to be checked and updated accordingly \cite{conner2017}.
+Creating small state machines can easily and quickly be accomplished using various feature-rich, and performant libraries \cite{fourakis2014}. 
+Finite state machines are often used to model reactive behavior comprehensively, and sequential behaviors in autonomous driving functions \cite{ziegler2014}.
+These finite state machines use nested state machines inside another state machine.
+The nested state machines reduce the complexity of the state machine as the transitions do not need to be modeled because a nested state machine's start and final state are treated as just one state inside the bigger state machine.
+Despite this possibility to simplify state machines, as a modeled behavior system grows larger to accommodate more complex behaviors into the state machine, the effort to expand and maintain the state machine multiplies rapidly.
+The higher effort is due to the modeling of the transitions, transition conditions, and transition events growing with each new state that gets introduced into the state machine as existing states need to be checked and updated accordingly \cite{conner2017}.
 \subparagraph*{}
 The execution time of finite state machines can be fast enough that they are used to control the motors of a bipedal robot to enable the robot to balance and walk despite unexpected height variations between steps \cite{park2013}.
-This property makes state machines suited very well suited for fast, reactive type behaviors and still allows a hierarchical approach to behavior planning.
+This property makes state machines very well suited for fast, reactive type behaviors and still allows a hierarchical approach to behavior planning.
 
@@
 Behavior Trees (BT) are another way to model and control the behavior of autonomous systems.
-Behavior Trees first found big acceptance in the computer game industry where they are mainly used to model artificial intelligence for non-player characters \cite{florez2009}.
-Every tree consists of one root and many child, parent and leaf nodes.
-Leaf nodes are also called execution nodes, while non-leaf nodes are called control or control flow nodes.
+Behavior Trees first found considerable acceptance in the computer game industry, where they are mainly used to model artificial intelligence for non-player characters \cite{florez2009}.
+Every tree has one root and many children, parent, and leaf nodes.
+Leaf nodes are also called execution nodes, while non-leaf nodes are control or control flow nodes.
 \subparagraph*{}
-The execution of a behavior tree is done by ticking the root node of tree.
-This signal then travels down to the child node of the root node, where either control nodes are being ticked or action nodes are executed.
-Nodes return either "Success", "Running" or "Failure" which influences how the tick signal gets processed by the rest of the behavior tree.
-Control nodes can be of the type "Sequence", "Fallback" or "Condition".
-The condition node can not have child nodes and can only return success or failure.
-A sequence node can be compared to a logical "and" condition.
-This means that once a sequence gets ticked from the parent node, it will send the tick signal to every child node as long as they return "success" of "running", and only return "success" itself when every child node was successfully ticked.
-If any of the child nodes return "failure", the sequence node will stop ticking the remaining children and return "failure".
-On the other hand the fallback node is an equivalent of a logical "or" condition.
-The fallback node will tick its child nodes as long as they return "failure" or "running" and will return "success" if one of the ticked node returned "success".
-It will only return "failure" if all of the child nodes "returned success".
-All of the basic node types and their control flow modification are listed in table \ref{tab:node_types}.
-
 \begin{table}[ht]
+	\caption{The five types of nodes \cite{iovino2022}}
 	\label{tab:node_types}
-	\caption{The five types of nodes \cite{iovino2022}}
 	\begin{tabular}{ | m{0.13\textwidth} | m{0.13\textwidth}| m{0.2\textwidth} | m{0.2\textwidth} | m{0.2\textwidth} |} 
   	\hline
-  	Node type & Symbol & Succeeds & Fails & Running \\ 
+  	\textbf{Node type} & \textbf{Symbol} & \textbf{Succeeds} & \textbf{Fails} & \textbf{Running} \\ 
   	\hline
@@
 \end{table}
-
+The execution of a behavior tree is done by ticking the tree's root node.
+This signal then travels down to the child node of the root node, where either control nodes are ticked, or action nodes are executed.
+Nodes return either "Success", "Running" or "Failure" which influences how the tick signal gets processed by the rest of the behavior tree.
+Control nodes can be of the type "Sequence", "Fallback" or "Condition".
+The condition node can not have child nodes and can only return success or failure.
+A sequence node can be compared to a logical "and" condition.
+This property of the sequence node means that once a sequence gets ticked from the parent node, it will send the tick signal to every child node as long as they return "Success" of "Running" and only return "Success" itself when every child node was successfully ticked.
+If any children nodes return "Failure", the sequence node will stop ticking the remaining children and return "Failure".
+On the other hand, the fallback node is an equivalent of a logical "or" condition.
+The fallback node ticks its children nodes as long as they return "Failure" or "Running" and will return "Success" if one of the ticked nodes returned "success".
+It will only return "Failure" if all children nodes returned "Success".
+The primary node types and their control flow modification are listed in table \ref{tab:node_types}.
 Figure \ref{fig:bt_example} depicts an example of a BT with multiple levels and action nodes.
-The nodes with the arrow (-$>$) symbol are sequence nodes, the question mark (?) symbol denotes fallback nodes.
-The condition node are depicted as white ovals.
-The action nodes are represented as gray rectangles in the BT example.
+The nodes with the arrow (-$>$) symbol are sequence nodes, and the question mark (?) symbol denotes fallback nodes.
+The condition nodes are depicted as white ovals, and the action nodes are represented as gray rectangles in the behavior tree example.
 
 \begin{figure}[h!]
+	\includegraphics[width=1.0\textwidth]{images/bt_example.jpg} 
+	\caption{Example of a behavior tree \cite{iovino2022}}
 	\label{fig:bt_example}
-	\includegraphics[width=1.0\textwidth]{images/bt_example.jpg} 
-	\caption{Example of a behavior tree \cite{iovino2022} }
 \end{figure}
 
-One of the core differences to Finite State Machine is that transitions between states are not distributed across all the states but they are organized in a hierarchical tree, where the leaves are representing the states \cite{iovino2022}.
-While both FSM and BT are capable of producing the same behavior in robots, the fundamental shift in how the two system are created lead to significant advantages in the modularity, synthesis and analysis of the systems at hand.
-These effects are more significant with an increase in size of the system.
-BT offer a lot more flexibility when creating an advanced behavior layer for an autonomous system.
+One of the core differences to the Finite State Machine is that transitions between states are not distributed across all the states.
+In contrast, they are organized in a hierarchical tree, where the leaves represent the states \cite{iovino2022}.
+FSM and BT can produce the same behavior in robots.
+However, the fundamental shift in how the two systems are created leads to significant advantages in the modularity, synthesis, and analysis of the systems at hand.
+These effects are more significant with an increase in the size of the system.
+Behavior trees offer much more flexibility when creating an advanced behavior layer for an autonomous system.
 
@@
 
-%\subsection{POMPDs}
+\subsection{Partially Observable Markov Decision Processes}
+Another approach to behavior planning is the Markov Decision Process(MDP) and the further advanced Partially Observable Markov Decision Process (POMDPs).
+MDPs share structural features with Finite State Machines, but the behavior planning and state transitions are stochastic and not deterministic in contrast to FSMs.
+An MDP can be described as a tuple <S, D, A, T, R>.
+Like a finite state machine, the MDP has a set of possible states \textit{S} and a set of possible transitions \textit{T} between them.
+The transitions contain a probability of execution and must add up to 1.0 for each associated state.
+An MDP adds a set of finite actions \textit{A} and rewards \textit{R} to the system.
+Additionally, MDPs possess a set of discrete time steps, in this context often called epochs, \textit{D}, which can be infinitely long.
+The guiding principle of MDPs is to maximize the expected reward by deciding on the best action.
+This policy Pi is what guides the behavior of the MDP \cite{feyzabadi2014riskaware}.
+
+\begin{figure}[ht]
+	\includegraphics[width=1.0\textwidth]{images/mdp.png} 
+	\caption{Example of a fully observable Markov Decision Process \cite{andrew1999reinforcement}}
+	\label{fig:mdp}
+\end{figure}
+
+Figure \ref{fig:mdp} shows an exemplary MDP for determining a fisherman's best course of action.
+The states are the population sizes of available fish, and the only available actions are to either fish or not to fish.
+The actions possess different rewards depending on the current state of the system \cite{tanwar2019markov}.
+
+The difference between an MDP and a POMDP is that the system's current state can be unknown, and the system can make observations and estimate the state to a degree.
+The POMDP is defined as a tuple <S, D, A, T, R, Z, O> in which the MDP tuple gets extended by \textit{Z} and \textit{O}.
+\textit{Z} is the observation, and \textit{O} is the probability function of seeing Z in a given state S.
+\subparagraph*{}
+POMDPs are well suited for applications in unknown environments because the system can make decisions when the perception information is imperfect, or the environment is not fully observable.
+The decision process is probabilistic and is more adaptable when confronted with unknown scenarios.
+This flexibility comes at the cost of determinism.
+The system's behavior is not guaranteed to be the same whenever it is subjected to the same sensory input \cite{krishnamurthy2016partially}.
 

### 2022-09-26 Thesis/text/implementierung.tex
grammarly

@@
 The system architecture, depicted in figure \ref{fig:component_diagram}, integrates the behavior tree into the existing architecture. 
-The already existing components are coloured in gray and the newly added components are coloured in blue.
+The already existing components are colored in gray, and the newly added components are colored in blue.
 The new components and their functions are described in the following sections of this chapter.  
@@
 \begin{figure}[ht]
-	\label{fig:component_diagram}
 	\includegraphics[width=1.00\textwidth]{images/component_diagram_bt.png}
 	\caption{Component Diagram of System Architecture}
+	\label{fig:component_diagram}
 \end{figure}
@@
 \section{Data Backup}
-The data backup component functions as a way to efficiently gather data from all components of the system and bereitstellen necessary data to the behavior tree when the planning process requires the use of past data points.
-This component subscribes is a to all sensor data and saves the messages in a queue data structure.
-The most recent data gets pushed into the array and if the array exceeds a bestimmte length, the oldest data this data gets deleted from the array.
-The limited array size is twofold.
-On the on side, the goal is to decrease the size of the message that gets sent to the BT to limit the network auslastung and on the other side, the BT often does not need to look back further than a few seconds to make a decision.
-This way the computational load and network load is decreased.
+The data backup component functions as a way to efficiently gather data from all system components and provide necessary data to the behavior tree when the planning process requires past data points.
+This component subscribes to all sensor data and saves the messages in a queue data structure.
+The most recent data gets pushed into the array, and if the array exceeds a specified length, the oldest data this data gets deleted from the array.
+There are two main reasons to limit the array size in this way.
+On the one side, the goal is to decrease the size of the message that gets sent to the BT to limit the network load, and on the other side, the BT often does not need to look back further than a few seconds to make a decision.
+This way, the computational load, and the network load are decreased.
 \paragraph*{}
@@
 The BT can access the saved data by sending a service call to one of the provided services by this component.
-Service calls in ROS2 are executed with hihg priority and the execution is done quickly but it adds an additional step than just saving the data directly in the behavior tree.
-The reason for the outsourcing of the data backup in another component is that ROS node need an executor which spins the node constantly.
-The spinning allows the node to check the network if there is any work to do for the node.
-This means that if one of the subscribed topic received a message, the node has to execute one of its callback functions that is associated with that topic.
-This spinning has to be done in regular intervals, otherwise the node might miss incoming messages from subscribed topics.
-The way to do the spinning is to have the node spun constantly in one thread.
-Have a constant spinning node would block the execution of the BT completely.
-If one would try to spin a node once when a node gets ticked in the BT, the problem of missing messages would become relevant because the BT might need longer than exepected to return to a node to tick it.
-That way the BT node that would be responsible for saving all the relevant data can not guarantee that it has picked up on all of the messages between to spins.
-The synchronous nature of the ROS executor spinning the nodes is the reason why the data backup is not stored within the BT.
-The saved data inside of this component with the corresponding lookback time is listed in table \ref{tab:data_backup_types}.
+Service calls in ROS2 are executed with high priority and are done quickly.
+However, it adds steps to the algorithm instead of just saving the data directly in the behavior tree.
+The reason for outsourcing the data backup to another component is that the ROS node needs an executor which constantly spins the node.
+The spinning action checks the network to see if there is any work for the node.
+Work, in this context, means that if one of the subscribed topics receives a message, the node has to execute one of its callback functions associated with that subscription.
+This spinning has to be done in regular intervals.
+Otherwise, the node might miss incoming messages from subscribed topics.
+The way to do the spinning is to have the node constantly spun in one thread.
+Having a constant spinning node would completely block the behavior tree's execution.
+If one tries to spin a node once when a node gets ticked in the BT, the problem of missing messages will become relevant because the BT might need longer than expected to return to a node to tick it.
+That way, the BT node responsible for saving all the relevant data can not guarantee that it has picked up on all the messages between spins.
+The synchronous nature of the ROS executor spinning the nodes is why the data backup is not stored within the BT.
+The saved data inside this component with the corresponding lookback time is listed in table \ref{tab:data_backup_types}.
 
@@
 \begin{table}[ht]
+	\caption{Types of saved data}
 	\label{tab:data_backup_types}
-	\caption{Types of saved data}
 	\begin{tabular}{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
@@
 
-This component is responsible for modyfying or blocking the Velocity Commands from Nav2.
-The component receives messages from Navigation2 on the "/cmd$\_$vel$\_$nav" topic and on the "/cmd$\_$vel$\_$bt" topic from the behavior tree.
+This component is responsible for modifying or blocking the Velocity Commands from Nav2.
+The component receives messages from Navigation2 on the "/cmd$\_$vel$\_$nav" topic and the "/cmd$\_$vel$\_$bt" topic from the behavior tree.
 This component publishes the received messages on the actual topic "/cmd$\_$vel" which is subscribed by the robot controller.
-The BT can modify parameters of the Decision Gate, to either disregard incoming messages of Nav2 completely or to modify them.
-The modification is useful when the situation demands the robot to slow down.
-If the Decision Gate receives messages from both Nav2 and the BT, it will prioritize the BT.
-This situation should theoretically never occur as the BT will set the parameters for the Decision Gate accordingly and/or cancel the current Navigation Goal before publishing commands on its own.
-It is implemented as an additional safety measure, in case the cancellation of the goal is not possible if Navigation2 is not responsive to the cancellation command.
+The BT can modify the Decision Gate parameters to either disregard incoming messages of Nav2 entirely or to modify them. The modification is helpful when the situation demands the robot to slow down. 
+If the Decision Gate receives messages from Nav2 and the BT, it will prioritize the BT.
+This situation should theoretically never occur as the BT will set the parameters for the Decision Gate accordingly and cancel the current Navigation Goal before publishing commands on its own.
+It is implemented as an additional safety measure in case the cancellation of the goal is not possible if Navigation2 is not responsive to the cancellation command.
 
@@
 
-A simplified behavior tree only depicting the top level behavior control nodes is shown in figure \ref{fig:top_level_bt}. The tree receives the inital tick signal from the root in an infinite loop. 
+A simplified behavior tree only depicting the top-level behavior control fallback nodes is shown in figure \ref{fig:top_level_bt}.
+The tree receives the initial tick signal from the root in an infinite loop.
 \begin{figure}[ht]
-	\label{fig:top_level_bt}
 	\includegraphics[width=1.0\textwidth]{images/simplified_bt.png}
 	\caption{Top Level Behavior Tree Structure}
+	\label{fig:top_level_bt}
 \end{figure}
 The root control node is a sequence node.
-We made the root a simple sequence because we want the whole system to function with increasing complexity in the later parts (on the right side) of the tree.
-The complex behaviors require that all of the previous components and conditions are working as they should.
-The root sequence will not allow the execution of higher level behaviors if the system is detecting problems on more fundamental levels.
-The root sequence is the first step for increasing the robustness of the system to combat competing reactive and deliberative behaviors which could endanger the safety of the robot and other actors in its environments.
+The root node is a simple sequence because the whole system must function with increasing complexity in the tree's later parts (on the right side).
+The complex behaviors require that all previous components and conditions work as they should.
+The root sequence will not allow the execution of higher-level behaviors if the system detects problems on more fundamental levels.
+The root sequence is the first step for increasing the system's robustness to combat competing reactive and deliberative behaviors that could endanger the robot's safety and other actors in its environment.
 
@@
 
-The system supervisor component is partly outsourced into a component outside of the thread of the BT, because of the same reasons as the sensor backup component discussed in the previous section.
+The system supervisor component is partly outsourced into a component outside of the BT's thread for the same reasons as the sensor backup component discussed in the previous section.
 The synchronous executor would stop the execution of the whole behavior tree, which is why the component needs to be moved inside of a new thread.
-The system supervisor, nachfolgend also called execution checker, is receiving messages from nodes that are not implemented as ROS2 lifecycle nodes.
-If a node stops sending messages or stops responding completely the external execution checker can inform the BT via a provided service of a problem with the node.
-ROS2 lifecycle nodes provide information about their health via an implementation of a FSM inside of the node.
-The current node state can be found out via a service call.
-However, to ensure the node is working correctly the output of the node needs to be checked too, because it might be that the FSM inside of the node did not pick up on a error.
-The system supervision is constantly checking the system's health. The BT is using service calls to get the health information about the compononts which are checked inside of Condition Nodes. 
+The system supervisor, also called the execution checker, is receiving messages from nodes not implemented as ROS2 lifecycle nodes.
+If a node stops sending messages or responding completely, the external execution checker can inform the BT via a provided service of a problem with the node.
+ROS2 lifecycle nodes provide information about their health via an FSM implementation inside the node.
+The current node state can be requested via a service call.
+However, the node's output needs to be checked, too, because it might be that the FSM inside of the node did not pick up on an error and failed to change the internal lifecycle state.
+The system supervision is constantly checking the system's health. The BT is using service calls to get the health information about the components checked inside Condition Nodes. 
 The components get checked by the BT inside of a Fallback.
-If the condition for the execution is met, the Condition Node returns "success" and the fallback would exit to move on to the next fallback and execution condition check (see fig.
-xx).
-In case the execution checker detects a node failure, a sequence to react to the sensor failure is executed in which the robot slows down to a minimum and tries to restart the sensor.
-Should the restart fail the robot will not be able to navigate in a safe manner anymore and will come to a complete stop.
-This behavior pattern is realized for the lidar, the IMU and the wheel odometry.
+If the condition for the execution is met, the Condition Node returns "Success", and the fallback would exit to move on to the next fallback and execution condition check (see fig.
+\ref{lidar_fallback}).
+If the execution checker detects a node failure, a sequence to react to the sensor failure is executed in which the robot slows down to a minimum and tries to restart the sensor.
+Should the restart fail twice, the robot will not be able to navigate safely anymore and will come to a complete stop.
+This behavior pattern is realized for the lidar, the IMU, and the wheel odometry.
 
@@
 	\caption{Lidar Fallback}
+	\label{fig:lidar_fallback}
 \end{figure}
 
-The timeout period for the execution checker to declare a node failed after no message was received is set at one second. 
-After that period the goal is to enable to robot to continue to make progress towards the goal but if the system can not restart the sensor in a short time after failure, the robot has to come to a stop as a sensor failure is a severe function loss for the robot.
+The timeout period for the execution checker to declare a node failed after receiving no message is set at one second. 
+After that period, the goal is to enable the robot to continue progressing towards the goal.
+However, if the system can not restart the sensor in a short time after failure, the robot has to come to a stop as a sensor failure is a severe function loss for the robot.
 
 \subsection{BT Advanced Behaviors}
-After the BT has checked for sensor failures and can guarantee that the environment representation is accurate, the system can check for more complicated scenarios to improve the autonomy.
+After the BT has checked for sensor failures and can guarantee that the environment representation is accurate, the system can check for more complicated scenarios to improve autonomy.
 The implemented behaviors are outlined in this section.
@@
 
-One of the biggest problems, mentioned in chapter 3.1, is the inability to react to collisions.
-The collision can happen due to inadequate sensory coverage for detecting obstacles or general sensor inadequacy which can not detect obstacles because of their shape or surface material.
-With a multitude of sensors covering all angles and heights and computer vision capabilities to detect obstacles collisions would be very unlikely.
-But even if the robot could perrfectly detect obstacles, it would still be vulnerable to outside pertubations like other robots driving into it, humans accidently touching the robot etc.
-.
-To combat this the robot must be able to first detect that a collision occured, get out of the collision state, update the map and reset all costmaps to generate a new path to the original goal.
+One of the biggest problems mentioned in chapter 3.1 is the inability to react to collisions.
+The collision can happen due to inadequate sensory coverage for detecting obstacles or general sensor inadequacy, which can not detect obstacles because of their shape or surface material.
+Collisions would be improbable with many sensors covering all angles and heights and computer vision capabilities to detect obstacles.
+However, even if the robot could detect all obstacles, it would still be vulnerable to outside perturbations, e.g., other robots driving into it or humans accidentally touching the robot.
+In order to combat this weakness, the robot must be able to detect that a collision occurred, get out of the collision state, update the map and reset all costmaps to generate a new path to the original goal.
 
 \begin{figure}[ht]
-	\label{fig:collision_fallback}
 	\includegraphics[width=1.0\textwidth]{images/collision_fallback.png}
 	\caption{Collision Fallback Behavior}
+	\label{fig:collision_fallback}
 \end{figure}
@@
 The collision checker uses a simulated sensor to detect collisions.
-On a real robot the collision checker can use a IMU-based collision detection, but this has not been implemented in this behavior tree due to time constraints of this thesis.
-The collision checker itself is integrated into the system supervisor and the BT is making service calls to the execution checker to get the collision state of the robot periodically.
-The child nodes of the collision sequence make use of the sensor data backup. 
+On a real robot, the collision checker can use an IMU-based collision detection, but this has not been implemented in this behavior tree due to the time constraints of this thesis.
+The collision checker is integrated into the system supervisor, and the BT makes service calls to the execution checker to get the collision state of the robot periodically.
+The child nodes of the collision sequence use the sensor data backup. 
 To get out of the collision state, the BT requests the last saved command velocities from the data backup component and modifies them to reverse with minimum speed the same way the robot got into the collision.
-The reversing action takes longer than the two seconds as the commands get published with longer time intervals between them to adjust the driven distance to be the same as the navigation commands.
-An additional check is implemented before the reversing to make sure the robot only reverses once, otherwise the robot would be driving back and forth as the data backup service would provide the BT with the commands from the last reversing action.
+The reversing action takes longer than the two seconds as the commands get published with more extended time intervals between them to adjust the driven distance to be the same as the navigation commands.
+An additional check is implemented before the reversing to ensure the robot only reverses once.
+Otherwise, the robot would drive back and forth as the data backup service would provide the BT with the commands from the last reversing action.
 
@@
 \subparagraph*{}
-Regarding the map reset, we have to determine if we are creating a new map with a SLAM method or if the map server provided a previously saved map.
-In case of the running SLAM, the recent updates to the map must be verwerfen to a map prior to when the collision occurred.When a static map is provided by the map server component from Nav2 the map needs to be updated with a new obstacle at the point of collision and republished by the map server.
+Regarding the map reset, it has to be determined if a new map is being created with a SLAM method or if the map server provided a previously saved map.
+In the case of the running SLAM, the recent updates to the map must be discarded, and a previous map version before the collision has to be re-utilized.
+When the map server component from Nav2 provides a static map, the map needs to be updated with a new obstacle at the point of collision and republished by the map server.
 The activity of the "MapUpdateAfterCollision" action node is depicted in figure \ref{fig:activity_diagram_collision}.
-The collision point is estimated by the robot's pose which refers to the robot base link and the robot footprint.
-We assumed that the collision point is right in front of the robot, because the local planner favours forward drive kinematics.
-The simulated bumper sensor can provide more nuanced information about the point of contact, but this functionality can not be achieved on a real robot with an IMU-based collsion detection, hence we simplified the obstacle position to always be in front of the robot.
+The collision point is estimated by the robot's pose, which refers to the robot base link and the robot footprint.
+It is assumed that the collision point was right in front of the robot because the local planner favors forward drive kinematics and probably is driving forwards when a collision occurs.
+The simulated bumper sensor can provide more nuanced information about the point of contact, but this functionality can not be achieved on a real robot with IMU-based collision detection.
+Hence the obstacle position is simplified to always be in front of the robot.
 %The collision position is calculated with 
@@
 
-A 25 by 25 cm square with full occupation of that space is added to the current map and then saved to be reloaded by the map server. 
+A 25 by 25 cm square with 100 percent occupation at the collision point is added to the current map and then saved to be reloaded by the map server. 
 %as depicted in figure xx. 
@@
 
-\begin{figure}[ht]
+\begin{figure}[h!]
+	\includegraphics[width=0.8\textwidth]{images/activity_diagram_collision.png}
+	\caption{Activity Diagram for the Collision Behavior}
 	\label{fig:activity_diagram_collision}
-	\includegraphics[width=0.92\textwidth]{images/activity_diagram_collision.png}
-	\caption{Activity Diagram for the Collision Behavior}
 \end{figure}
 
-In this manner obstacles lower than the lidar scan height can be detected.
-With this behavior the robot should also be able to eventually navigate around transparent obstacles which do not reflect the laser.
+This way, obstacles lower than the lidar scan height can be detected.
+With this behavior, the robot should be able to eventually navigate around transparent obstacles that do not reflect the laser.
 Glass doors are known to cause issues for robots based only on lidars.
@@
 
-The battery behavior has the goal to monitor the robot's battery and has to make to the decision if the robot should navigate to a new goal location or not.
+The battery behavior has the goal of monitoring the robot's battery and has to decide if the robot should navigate to a new goal location or not.
 The decision is based on the current battery charge and the estimated consumption of energy to reach the new goal.
 The robot should not begin the navigation process if it is probable that the remaining charge will not suffice for the driving.
-The behavior tree fallback is depicted in figure \ref{fig:battery_fallback}.
+Figure \ref{fig:battery_fallback} depicts the behavior tree fallback.
 
@@
 \begin{figure}[ht]
-	\label{fig:battery_fallback}
 	\includegraphics[width=0.5\textwidth]{images/battery_fallback.png}
 	\caption{Battery Fallback Behavior}
+	\label{fig:battery_fallback}
 \end{figure}
@@
 
-The robot has to take into account the idle consumption, the energy to drive the motors.
-With this information we created a simple linear function for the energy expenditure to predict the required power depending on the length of the path.
-To make the algorithm fail safe, we added an additional safety factor into the function, to berücksichtigen the scenario where the robot has to reroute and drive a longer route or has to stop for a moment.
+The robot has to consider the idle consumption and the energy to drive the motors.
+With this information, a linear function for energy expenditure was created to predict battery usage depending on the path length.
+An additional safety factor was added to the energy estimation function to consider the scenario in which the robot must reroute and drive a longer route or stop for a moment.
+The safety factor was included to increase the behavior's robustness and cover more scenarios.
 %The resulting function can be described as: 
@@
 
-On a real robot the average values for the idle and driving consumption could be experimentally determined.
-For the simulated turtlebot we created a simulated battery package which updates the battery charge every second.
-The battery package subscribes to the comman velocity topic and decreases the battery charge for every message that is received.
-We created the battery package because the behavior for this scenario needs to be tested.
-This is way the consumption rates are not close to reality.
-The battery package provides a service to get the battery charge and the values for the angenommenen consumption rates.
+The average values for idle and driving consumption on a real robot could be experimentally determined.
+For the simulated turtlebot, a simulated battery package was created, which updates the battery charge every second.
+The battery package subscribes to the command velocity topic and decreases the battery charge for every message received.
+The battery package was created because the behavior for this scenario needs to be tested, and Gazebo does not offer this function.
+The consumption rates are not close to reality but are chosen to allow the rapid testing of the behavior.
+The battery package provides a service to get the battery charge and the values for the assumed consumption rates while idle and driving (compare \ref{tab:custom_interfaces}.
 The BT node "IsBatterySufficientCheck" calls the service to compare the current battery to the calculated consumption.
@@
 The execution checker component monitors the health of the Nav2 global planner.
-Additionally, the BT analyses the global planners output.
-If one of the condition fails, the given navigation goal is not reachable by the planning algorithm.
-This gets combatted by shifting the goal to a nearby alternative to get the robot as close to the original goal as possible.
-Is the location still unreachable, the goal gets moved further away and closer to the robot again.
-This routine gets repeated for up to ten times, which leads to a distance to the original goal of up to one meter.
+Additionally, the behavior tree analyses the global planner's output.
+If the latter condition fails, the given navigation goal is not reachable by the planning algorithm.
+By shifting the goal to a nearby alternative goal, the robot tries to navigate, although the original goal is not reachable.
+If the location is still unreachable, the goal gets moved further away from the original.
+The routine repeats up to ten times, leading to a maximum distance to the original goal of up to one meter.
 The goal gets moved on a straight vector towards the robot.
 Other strategies for finding alternative goals were not implemented in this behavior.
-The behavior tree for the behavior is pictured in figure \ref{fig:global_planning_fallback}.
+Figure \ref{fig:global_planning_fallback} depicts the behavior tree fallback for the behavior.
 
@@
 	\begin{figure}[ht]
-	\label{fig:global_planning_fallback}
 	\includegraphics[width=0.7\textwidth]{images/global_planning_fallback.png}
 	\caption{Path Planning Fallback Behavior}
+	\label{fig:global_planning_fallback}
 	\end{figure}
@@
 \subsection{Gazebo Sensor Drivers}
-For testing purposes, the simulated sensor data is not published directly on to the actual topic by gazebo.
-To control the flow of information in the system, an intermediary node is used, which subscribes to the Gazebo topic and republishes the information on the original topic.
-With this methodology, the failure of a sensor can be simulated as the simulation does not offer this functionality.
+For testing purposes, the simulated sensor data is not published directly on the actual topic by Gazebo.
+An intermediary node is used, which subscribes to the Gazebo topic and republishes the information on the original topic.
+The failure of a sensor can be simulated this way by stopping the execution of the intermediary node.
 Figure \ref{fig:rqt_graph} shows the RQT graph, which depicts the ROS nodes and their subscribed/published topics.
@@
 \begin{figure}[ht]
-	\label{fig:rqt_graph}
 	\includegraphics[width=0.75\textwidth]{images/rqt_graph_sensor_drivers.png}
 	\caption{Graph for the Sensor Driver nodes and topics}
+	\label{fig:rqt_graph}
 \end{figure}
@@
 
-Because of the neccessity to have ROS nodes spinning outside of the BT thread as discussed in 4.4.1, additional interfaces needed to be created in order to receive informations through service calls.
+Because of the necessity to have ROS nodes spinning outside of the BT thread, as discussed in 4.4.1, additional interfaces needed to be created in order to receive information through service calls.
 The created service types are listed in table \ref{tab:custom_interfaces}.
-All of the service definitions are created in a dedicated package named "bt$\_$msgs".
+All service definitions are created in a dedicated package named "bt$\_$msgs".
 
 \begin{table}[ht]
+	\caption{Implemented Custom Services}
 	\label{tab:custom_interfaces}
-	\caption{Implemented Custom Services}
 	\begin{tabular}{ | m{0.2\textwidth} | m{0.33\textwidth}| m{0.33\textwidth} | m{0.2\textwidth} |} 
   	\hline
-  	Name & Request & Response \\ 
+  	\textbf{Name} & \textbf{Request} & \textbf{Response} \\ 
   	\hline

### 2022-09-26 Thesis/text/abstract.tex
split summary in conclusion and outlook

%\section*{Abstract}
\chapter*{Abstract}

%Introduction
This thesis explores ways to improve the autonomy of mobile robots running the Robot Operating System (ROS) with behavioral planning approaches. Autonomous mobile robots often require close human supervision or intervention during their operation.

%Methods
A system supervision system, a sensor data storage and a behavior tree were designed and implemented. The system was tested on a simulated robot in the Gazebo simulator. Behaviors to handle sensor failures, collisions, unreachable goals, and low battery state.

%Results
The different test scenarios were artificially induced in a standardized manner to trigger the behaviors to react to the scenarios. The results show an increase in the robot's autonomy and robustness to failures.

%Discussion
The results need to be tested and verified on a real robot, but the increase in the robot's autonomy in the simulation tests are promising an equally positive result. Behavior trees to control and improve the behavior of robots offer the ability to effortlessly add new behaviors to an existing system. This thesis shows a comprehensive way how behavior trees can be implemented to an existing ROS2 system.

### 2022-09-26 Thesis/text/diskussion.tex
split summary in conclusion and outlook

--- 2022-09-06 Thesis/text/diskussion.tex
@@
-\chapter{Summary}
+\chapter{Conclusion}
 \label{cha:diskussion}
+\paragraph*{}
 
-\section{Conclusion}
+This thesis demonstrated how to incorporate advanced behavior planning in ROS2 mobile robots.
+The goal was to achieve higher levels of autonomy and robustness by reducing the number of scenarios in which a human operator is needed.
+Implementing the system supervisor component improves the system's robustness and provides a safe basis for executing more intelligent behaviors.
+The implemented behaviors to react to sensor failures and robot collisions are valuable improvements to the system and improve current ROS2 mobile robots significantly.
+The layout of the behavior tree is structured to allow more complex behaviors to be safely executed because possible scenarios that interfere with a correct environment representation are checked and dealt with before complex behaviors get executed based on that representation.
+The implemented layout choice for the tree is not mentioned in the literature. However, it has proven to be a sensible approach to model robot behavior safely and robustly. 
+%The implemented behaviors and the system supervision component have improved the behavior in the simulated scenarios.
+The system can react and recover from multiple scenarios with additional components.
+\paragraph*{}
+Furthermore, the system architecture is a good base for further development without facing problems or limitations in the later stages of developing behaviors.
+The architecture shows a novel way that works with the asynchronous ROS2 executor properties and allows fast execution of both the behavior planner and the system components.
 
-\section{Prospect}
+Nevertheless, the behavior tree design could be modified to use blackboard values more effectively to increase the information flow between action nodes in the tree.
+Also, the tree does not utilize the reactive control flow nodes, and no action node in the tree is designed to allow multiple threads at the current state.
+Incorporating so-called stateful nodes into the tree's structure to allow co-routines would increase the possibility of creating better and more deliberative behaviors.
+
+\chapter{Outlook}
+\label{cha:outlook}
+
+The implemented behaviors are a first step to increasing autonomy, but the behavior planner needs to create many more behaviors and fallbacks.
+Some behaviors not implemented in this thesis incorporate the Intel RSS rules for autonomous driving, like keeping appropriate distances, right-of-way rules, and adaptable speed based on occlusions.
+Another area of future work is the creation of dedicated behaviors for multi-robot systems.
+E.g., sharing sensor data with other robots to overcome complete sensor failures or intelligent path planning to avoid collisions when robots can access the planned paths of other robots.
+Also, more work is needed to define a robot's safe state or position after triggering an emergency stop. Reaching these minimum risk states after emergencies is then the duty of the behavior planner.
+
+\paragraph*{}
+
+The behaviors must be more deliberative and robust to achieve more independence from human operators.
+One key question in autonomous driving is the responsibility of a vehicle in a crash because, with higher levels of automation, the human is not needed anymore.
+However, a machine can not be responsible for its action because it was programmed to do so.
+Nevertheless, intelligent, robust, and independent decision-making almost always involves machine learning which makes decisions based on training with dedicated datasets and not explicitly written lines of code.
+Because a system in which the outcome for a high-risk scenario is unclear, machine learning behaviors may not be suited to be used independently.
+Therefore, an important work area is incorporating machine learning algorithms into a deterministic system, like a behavior tree.
+A behavior tree could execute and monitor the behaviors to safeguard the indeterministic behaviors and override them if necessary.
+%Further research is needed on how behavior trees can manage a system with autonomous machine learning behaviors inside the tree.
+
+\paragraph*{}
+
+Lastly, the results must be reproduced on an actual mobile robot.
+The layout of the behavior tree will stay the same, but further effort is needed to adapt the action nodes of the tree for a real system.

### 2022-09-26 Thesis/text/implementierung.tex
split summary in conclusion and outlook

@@
 The new components and their functions are described in the following sections of this chapter.
+
 %This is done with the addition of the Decision Gate, that can selectively forward speed commands to the robot controller.
@@
 The tree receives the initial tick signal from the root in an infinite loop.
+
 \begin{figure}[ht]
@@
 \end{figure}
+
 The root control node is a sequence node.
@@
 \subsection{BT System Supervision}
+
+The design pattern of the system supervisor is an "Watchdog" pattern.
+This pattern is depicted in figure \ref{fig:watchdog_pattern}.
+The pattern is used to monitor a system from the outside and directly take corrective measures when problems are detected.
+
+\begin{figure}[ht]
+	\includegraphics[width=0.4\textwidth]{images/watchdog_pattern.png}
+	\caption{Examiner pattern for the system supervisor component \cite{konrad2003defining}}
+	\label{fig:watchdog_pattern}
+\end{figure}
 
@@
 This behavior pattern is realized for the lidar, the IMU, and the wheel odometry.
-
+\begin{center}
 \begin{figure}[ht]
@@
 
-\begin{figure}[h!]
+\begin{figure}[ht]
 	\includegraphics[width=0.8\textwidth]{images/activity_diagram_collision.png}
@@
 \subsection{Gazebo Sensor Drivers}
+%\label{sec:gazebo_driver}
 For testing purposes, the simulated sensor data is not published directly on the actual topic by Gazebo.

### 2022-09-26 Thesis/text/methoden.tex
split summary in conclusion and outlook

@@
 The manufacturer provides an open-source model for simulating the robot in physics-based simulators like Gazebo, as depicted in figure \ref{fig:turtlebot}, which enables faster development and testing for the robot.
-The equipped sensors and easy simulation capabilities make the robot well suited for using the Nav2 stack and further development for behavior planning.
-
-%The robot specifications are specified in table xx.  \\
-
+\begin{center}
 \begin{figure}[ht]   
@@
 \end{figure}
+\end{center}
+The equipped sensors and easy simulation capabilities make the robot well suited for using the Nav2 stack and further development for behavior planning.
+
+The robot specifications are specified in table \ref{tab:turtlebot_spec}.
+
+\begin{table}[ht]
+\caption{Turtlebot3 Specifications}
+\label{tab:turtlebot_spec}
+	\begin{tabular}{| m{0.3\textwidth} | m{0.3\textwidth} |} 
+  	\hline
+  	\textbf{Specification} & \textbf{Value} \\ 
+  	\hline
+  	Maximum translational velocity & 0.26 m/s \\ 
+  	\hline
+  	Maximum rotational velocity & 1.82 rad/s (104.27 deg/s) \\
+  	\hline
+  	Maximum payload & 30kg \\
+  	\hline
+  	Size (L x W x H) & 281mm x 306mm x 141mm \\
+  	\hline
+  	Weight & 1.8 kg \\
+  	\hline
+  	Singe Board Computer & Raspberry Pi (3 or 4) \\
+  	\hline
+  	Laser Distance Sensor & 360 Degree Laser Distance Sensor LDS-01 \\
+  	\hline
+  	IMU & Gyroscope 3 Axis, Accelerometer 3 Axis \\
+  	\hline
+  	Actuators & XM430-W210 \\
+  	\hline
+	\end{tabular}
+\end{table}
 
 The turtlebot will be used for the analysis of the current state of mobile navigation.
-Furthermore, the simulated turtlebot will be the platform on which the behavior planning will be implemented and tested.\\
+Furthermore, the simulated turtlebot will be the platform on which the behavior planning will be implemented and tested.
 
@@
 A suitable name for the second system is the "autonomy layer", as it aims to increase the robustness and autonomy of the whole system. 
-The autonomy layer has to provide a broad system supervisor that monitors all components' health (sensors, navigation, robot controller).
+The autonomy layer has to provide a comprehensive system supervisor that monitors all components' health (sensors, navigation, robot controller).
 The constant execution monitoring equips the system with the ability to deal with the event of node failures and component malfunctions.
@@
 This modification aspect allows high-quality local planning capabilities but enables more deliberative behaviors to happen upon this planning.
+A simplified system block diagram is depicted in figure \ref{fig:block_diagram}.
+The proposed additional components, discussed in this chapter are colored in blue.
+Existing components, discussed in chapter 3.1, are colored in gray.
+
+\begin{center}
+\begin{figure}[ht]   
+	\includegraphics[width=1.0\textwidth]{images/block_diagram.png}
+	\caption{System Block Diagram}
+	\label{fig:block_diagram}
+\end{figure}
+\end{center}
 
@@
 These Decorator Nodes can only have one child node each.
-A comprehensive list of the available nodes and descriptions is in table \ref{tab:decorators_bt}.
+A comprehensive list of the available Decorators and descriptions is presented in table \ref{tab:decorators_bt}.
 
@@
 The ROS community offers many different simulation environments to test the robot in. The environments range from simple geometric structurs, to office spaces up to complete race tracks.
+
+\subsection{OS and Software Versions}
+The software versions used to implment and test the system on are listed below in table \ref{tab:versions}.
+Many of the software versions are determined by the choice of the ROS distribution (distro).
+ROS2 Foxy is not the newest ROS2 distribution but is well supported and developed, so the decision was made to use this distribution.
+The decision why the turtlebot is used is laid out in chapter 3.1.3.
+Although Python and C++ can be used to develop applications with ROS2, the behavior tree library relies on C++, which is why C++ was used for developing the whole system.
+
+\begin{center}
+\begin{table}[ht]
+	\caption{Decorator Nodes in BT.CPP}
+	\label{tab:versions}
+	\begin{tabular}{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
+  	\hline
+  	\textbf{Name} & \textbf{Version / Release} & \textbf{Comments} \\ 
+  	\hline
+  	Operating System & Ubuntu 20.04 LTS &  \\ 
+  	\hline
+  	ROS & ROS2 Foxy & Long-Term-Supported Distribution \\ 
+  	\hline
+  	Navigation2 & Foxy-devel & Newest Release for ROS2 Foxy\\
+  	\hline
+  	Simulation & Gazebo 11 & Determined by choice of ROS Distro \\
+  	\hline  	
+  	Robot & Turtlebot3 &  \\
+  	\hline
+  	Behavior Tree & BT.CPP 3.7 & \\
+  	\hline
+  	Programming Language & C++ 14 & Determined by choice of ROS Distro \\
+  	\hline
+  	Build System & CMake and Colcon & Determined by choice of ROS Distro \\
+  	\hline
+  	Development Environment & Visual Studio Code & Offers IDE debugging options for ROS \\
+  	\hline
+	\end{tabular}
+\end{table}
+\end{center}

### 2022-09-26 Thesis/text/methoden.tex
update to last chapter

@@
 The software can be modified and expanded with plugins to fit the users' needs.
-When comparing the functionalities of the software stack to the proposed autonomous system architecture in figure \ref{fig:autonomous_driving_architectures}, the things that Nav2 is missing are a system-wide supervision layer and a more deliberative approach to behavior planning, as mentioned in chapter 2.3.
+When comparing the functionalities of the software stack to the proposed autonomous system architecture in figure 2.2, the things that Navigation2 is missing are a system-wide supervision layer and a more deliberative approach to behavior planning, as mentioned in chapter 2.3.
 \subparagraph*{}
@@
 \begin{figure}[ht]   
-	\includegraphics[width=0.7\textwidth]{images/turtlebot_sim.png}
+	\includegraphics[width=0.5\textwidth]{images/turtlebot_sim.png}
 	\caption{Turtlebot3 model in Gazebo simulator}
@@
 The reason for this is mainly the increased flexibility when creating and expanding the behavior planning compared to FSMs.
-Also, behavior trees allow complete determinism and introspection.
-%which is why POMPDs are not suitable as the behavior planning approach for the specified requirements.
+Also, behavior trees allow complete determinism and introspection which is why POMPDs are not suitable as the behavior planning approach for the specified requirements.
 
@@
 The software allows the creation of behavior trees in XML files which can be directly loaded into and used by the BT.CPP library.
-Groot can monitor the live execution of behavior trees and allows the introspection of how the tree reacts to different scenarios.
+Groot can monitor the live execution of behavior trees and allows the introspection of how the tree reacts to different scenarios (compare fig.
+\ref{fig:groot}).
+
+\begin{center}
+	\begin{figure}[ht]   
+		\includegraphics[width=0.7\textwidth]{images/groot.png}
+		\caption{Live monitoring of behavior trees with Groot}
+		\label{fig:groot}
+	\end{figure}
+\end{center}
 
@@
 \begin{table}[ht]
-	\caption{Decorator Nodes in BT.CPP}
+	\caption{Used Software Versions}
 	\label{tab:versions}

### 2022-09-26 Thesis/text/diskussion.tex
ahmed's work of god

@@
 \label{cha:diskussion}
-\paragraph*{}
 
@@
 The system can react and recover from multiple scenarios with additional components.
-\paragraph*{}
+
 Furthermore, the system architecture is a good base for further development without facing problems or limitations in the later stages of developing behaviors.
@@
 
-\paragraph*{}
-
 The behaviors must be more deliberative and robust to achieve more independence from human operators.
@@
 
-\paragraph*{}
-
 Lastly, the results must be reproduced on an actual mobile robot.

### 2022-09-26 Thesis/text/einleitung.tex
ahmed's work of god

@@
 This reliance on a human operator hardly qualifies the robot to be labeled autonomous.
-\subparagraph*{}
+
 One of the quickest and most popular ways to build and program a robot is to use the Robot Operating System (ROS).
@@
 These can be either other robots or humans.
-\subparagraph*{}
-In theory, the robot can perform fully autonomous navigation, including mapping and localization. However, when something unexpected happens, the robot's autonomy can not be guaranteed any more. 
+
+In theory, the robot can perform fully autonomous navigation, including mapping and localization.
+However, when something unexpected happens, the robot's autonomy can not be guaranteed any more.
 A default robot cannot represent all possible fail states and does not detect failures on a systematic level but only inside its navigation-related subsystems.
@@
 
-\subparagraph*{}
-This thesis will explore possibilities for improving current systems by adding a dedicated behavior planning component.
-Furthermore, the thesis will provide an implementation of exemplary behaviors and provide a system architecture to incorporate behavior planning in other robots, too.
+This thesis explores possibilities for improving current systems by adding a dedicated behavior planning component.
+Furthermore, the thesis provides an implementation of exemplary behaviors and provide a system architecture to incorporate behavior planning in other robots, too.
 
-\subparagraph*{}
-The desired outcomes of this implemented software will be an increase in robustness and a decrease in required human actions during several scenarios.
+The desired outcomes of this implemented software is an increase in robustness and a decrease in required human actions during several scenarios.
 In these scenarios, failures and problems are artificially induced to test the robot's abilities to behave autonomously.
+
+The remainder of this thesis is organized as follows:
+
+\begin{itemize}
+	\item Chapter \ref{cha:state of the art} presents the state of the art and compares the different behavior planning approaches
+
+	\item Chapter \ref{cha:methoden} compares the current systems and derives requirements from the key problem areas
+
+	\item Chapter \ref{cha:implementierung} describes the implemented software approach including the overall system architecture
+
+	\item Chapter \ref{cha:ergebnisse} shows the obtained results from all selected scenarios and discusses the acceptance criteria of the requirements
+
+	\item Chapter \ref{cha:diskussion} summarizes the thesis novelty by showing how the implemented approach answers the research question
+
+	\item Chapter \ref{cha:outlook} provides an outlook of future work and recommendations
+\end{itemize}
 
@@
 This hardly qualifies the robot to be labeled autonomous.
-%\subparagraph*{}
+%
 %One of the quickest and most popular ways to build and program a robot is to use the Robot Operating System (ROS).
@@
 It is important for the robot to possess the ability to navigate in an uncertain environment to guarantee the safety of the robot itself and all other actors in its environment, these can be other robots and humans.
-%\subparagraph*{}
+%
 %In theory, the robot can perform fully autonomous navigation including mapping and localization but as soon as something out of the ordinary happens the robots autonomy is not guaranteed to be reliable anymore.
@@
 %%Planned solution:\\
-%\subparagraph*{}
+%
 %This thesis will explore possibilities how current systems can be improved through the addition of a dedicated behavior planning component.
@@
 %%Expected results:\\
-%\subparagraph*{}
+%
 %The desired outcomes of these implemented software will be an increase of robustness and decrease of required human actions during a number of scenarios.

### 2022-09-26 Thesis/text/ergebnisse.tex
update to last chapter
ahmed's work of god

@@
 The robot is put into specifically designed scenarios that should trigger behavior responses from the behavior tree component.
-The scenario testing is done to evaluate the efficacy of the implemented behaviors in relation to the functional requirements listed in chapter 3.2.
+The scenario testing is done to evaluate the efficacy of the implemented behaviors in relation to the functional requirements listed in section \ref{sec:Requirements}.
 
 The scenarios are executed in a simulated apartment environment as seen in figure \ref{fig:house_gazebo} and a simulated world with small round obstacles in an enclosed space as depicted in figure \ref{fig:world_gazebo}.
-\begin{center}
+The environments were mapped beforehand with the ROS2 "Slam Toolbox" package and provided via the Nav2 map server.
+
 \begin{figure}[ht]
-	\includegraphics[width=0.75\textwidth]{images/house_env.png}
+	\centering
+	\begin{subfigure}[b]{0.45\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{images/house_env.png}
 		\caption{Apartment Environment in Gazebo}
 		\label{fig:house_gazebo}
-\end{figure}
-\end{center}
-
-\begin{center}
-\begin{figure}
-	\includegraphics[width=0.75\textwidth]{images/world_env.png}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.45\textwidth}
+		\centering
+		\includegraphics[width=\textwidth]{images/world_env.png}
 \caption{Enclosed Environment in Gazebo}
 		\label{fig:world_gazebo}
+	\end{subfigure}
+	\caption{Gazebo World Environment}
+	\label{fig:three graphs}
 \end{figure}
-\end{center}
-
-The environments were mapped beforehand with the ROS2 "Slam Toolbox" package and provided via the Nav2 map server.
 
@@
 This way, the same scenario can be run with and without behavior planning.
-To test the behavior of the sensor failure fallbacks, the respective sensor driver, mentioned in chapter 4.5.1, was shut down to emulate the sensor's failure.
+To test the behavior of the sensor failure fallbacks, the respective sensor driver, mentioned in section \ref{subsec:gazebo_driver}, was shut down to emulate the sensor's failure.
 
@@
 The scenarios are derived from the functional requirements and can be used to test multiple requirements as listed in the tables \ref{tab:sensor_scenarios} and \ref{tab:behavior_scenarios}.
-The acceptance criteria often contain multiple binary conditions that can only be met or failed. All acceptance criteria for a given scenario must be met for the test to be counted as a success.
-
-For the Lidar, IMU, and Odometry scenarios, the sensor driver was shut down when reaching the desired speed listed in the scenario. 
-Scenario NoPathFound1 is a scenario in which the lifecycle state gets transitioned into "inactive" and can not execute the planning action anymore when triggered.
-The BT has to recover from the inability to plan in this scenario.
-NoPathFound$\_$2 is simulating the scenario when the planner is active but fails to find a collision-free path to the desired goal.
-This scenario tests the ability to find alternative goals and navigate to a nearby goal.
-In the Collision$\_$1 scenario, an obstacle gets moved so close to the robot that a collision is registered.
-The forced collision tests the ability to escape a collision and resume navigation afterward.
-A new navigation goal is published after the behavior tree exits the Collision Fallback Routine.
-Collision$\_$2 scenario is testing the same behavior but during active navigation.
-A flat obstacle is spawned in the path of the robot.
-The obstacle is too low to be detected by the laserscanner, and a collision occurs.
-The robot must be able to recover from the crash and keep on navigating towards the goal with an updated global path in this scenario.
-The Battery$\_$1 scenario tests if the robot tries to navigate toward a distant goal with a low battery charge.
-The simulated battery is emptied to a low level, and a new navigation goal is set.
-If the battery charge runs below zero percent during the navigation, the test fails.
+The acceptance criteria often contain multiple binary conditions that can only be met or failed.
+All acceptance criteria for a given scenario must be met for the test to be counted as a success.
 
 \begin{table}[ht]
+	\centering
 	\caption{Sensor Scenarios}
 	\label{tab:sensor_scenarios}
-	\begin{tabular}{| m{0.1\textwidth} | m{0.17\textwidth}| m{0.3\textwidth} | m{0.3\textwidth}|} 
+	\renewcommand{\arraystretch}{1.5}
+	\resizebox{0.9\textwidth}{!}{%
+		\begin{tabular}{| m{0.1\textwidth} | m{0.2\textwidth}| m{0.3\textwidth} | m{0.3\textwidth}|} 
 			\hline
-  	\textbf{Name} & \textbf{Related Requirements }& \textbf{Description} & \textbf{Success Criteria}\\ 
+			\textbf{Name} & \textbf{Requirements }& \textbf{Description} & \textbf{Success Criteria}\\ 
 			\hline
@@
 		\end{tabular}
+	}
 \end{table}
 
+The scenarios execution is done as follows:
+
+\begin{itemize}
+	\item For the Lidar, IMU, and Odometry scenarios, the sensor driver was shut down when reaching the desired speed listed in the scenario.
+
+	\item Scenario NoPathFound1 is a scenario in which the lifecycle state gets transitioned into "inactive" and can not execute the planning action anymore when triggered.
+The BT has to recover from the inability to plan in this scenario.
+
+	\item NoPathFound$\_$2 is simulating the scenario when the planner is active but fails to find a collision-free path to the desired goal.
+This scenario tests the ability to find alternative goals and navigate to a nearby goal.
+
+	\item In the Collision$\_$1 scenario, an obstacle gets moved so close to the robot that a collision is registered.
+The forced collision tests the ability to escape a collision and resume navigation afterward.
+A new navigation goal is published after the behavior tree exits the Collision Fallback Routine.
+
+	\item Collision$\_$2 scenario is testing the same behavior but during active navigation.
+A flat obstacle is spawned in the path of the robot.
+The obstacle is too low to be detected by the laserscanner, and a collision occurs.
+The robot must be able to recover from the crash and keep on navigating towards the goal with an updated global path in this scenario.
+
+	\item The Battery$\_$1 scenario tests if the robot tries to navigate toward a distant goal with a low battery charge.
+The simulated battery is emptied to a low level, and a new navigation goal is set.
+If the battery charge runs below zero percent during the navigation, the test fails.
+\end{itemize}
+
 \begin{table}[ht]
+	\centering
 	\caption{Behavior Scenarios}
 	\label{tab:behavior_scenarios}
-	\begin{tabular}{| m{0.1\textwidth} | m{0.13\textwidth}| m{0.33\textwidth} | m{0.34\textwidth}|} 
+	\renewcommand{\arraystretch}{1.5}
+	\resizebox{0.9\textwidth}{!}{%
+		\begin{tabular}{| m{0.1\textwidth} | m{0.2\textwidth}| m{0.4\textwidth} | m{0.3\textwidth}|} 
 		\hline
-	\textbf{Name} & \textbf{Related Requirements} & \textbf{Description} & \textbf{Success Criteria}\\ 
+		\textbf{Name} & \textbf{Requirements} & \textbf{Description} & \textbf{Success Criteria}\\ 
 	 	\hline	
@@
 		\hline	
-
 		\end{tabular}
+	}
 \end{table}
@@
 The collision behavior during driving worked well, and the map updates are an essential feature for the success of the scenarios.
+
 Nevertheless, there is a lack of intelligence and autonomy when the robot stands still.
@@
 \begin{table}[ht]
-
+	\centering
 	\caption{Results}
 	\label{tab:results}
-	\begin{tabular}{| m{0.11\textwidth} | m{0.13\textwidth}| m{0.33\textwidth} | m{0.33\textwidth}|} 
+	\renewcommand{\arraystretch}{1.5}
+	\resizebox{0.9\textwidth}{!}{%
+		\begin{tabular}{| m{0.11\textwidth} | m{0.1\textwidth}| m{0.35\textwidth} | m{0.33\textwidth}|} 
 	 	\hline
@@
 		\end{tabular}
+	}
 \end{table}
@@
 Implementing the whole behavior planning and system supervision system eliminates a single point of failure for the system.
-Even when the Navigation2 system collapses, is the robot able to move to a degree and, more importantly, stop completely (compare table \ref{tab:tab:nonfn_req}, non$\_$fn$\_$req).
+Even when the Navigation2 system collapses, is the robot able to move to a degree and, more importantly, stop completely (compare table \ref{tab:nonfn_req}, non$\_$fn$\_$req).
 The behavior tree and algorithms implemented in the behavior tree are all deterministic (non$\_$fn$\_$req3).
 The robots' behavior goes beyond pure reactivity when navigating through environments, as the advanced behaviors use a dedicated planning phase before the actions are carried out (non$\_$fn$\_$req4).
+
 Finally, the system's performance when checking the conditions of the BT is below the desired threshold of 10ms, as the whole system runs in multiple threads, which allows the behavior tree to complete one complete execution cycle in about 5ms with the current system.

### 2022-09-26 Thesis/text/implementierung.tex
update to last chapter
ahmed's work of god

@@
 \begin{figure}[ht]
-	\includegraphics[width=1.00\textwidth]{images/component_diagram_bt.png}
+	\centering
+	\includegraphics[width=0.9\textwidth]{images/component_diagram_bt.png}
 	\caption{Component Diagram of System Architecture}
@@
 \section{Data Backup}
+
 The data backup component functions as a way to efficiently gather data from all system components and provide necessary data to the behavior tree when the planning process requires past data points.
@@
 This way, the computational load, and the network load are decreased.
-\paragraph*{}
 
@@
 Work, in this context, means that if one of the subscribed topics receives a message, the node has to execute one of its callback functions associated with that subscription.
+
 This spinning has to be done in regular intervals.
@@
 Having a constant spinning node would completely block the behavior tree's execution.
-If one tries to spin a node once when a node gets ticked in the BT, the problem of missing messages will become relevant because the BT might need longer than expected to return to a node to tick it.
+If one tries to spin a node once when a node gets ticked in the BT, the problem of missing messages becomes relevant because the BT might need longer than expected to return to a node to tick it.
 That way, the BT node responsible for saving all the relevant data can not guarantee that it has picked up on all the messages between spins.
@@
 
-\begin{center}
 \begin{table}[ht]
+	\centering
 	\caption{Types of saved data}
 	\label{tab:data_backup_types}
-	\begin{tabular}{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
+	\renewcommand{\arraystretch}{1.5}
+	%\resizebox{0.9\textwidth}{!}{%
+		\begin{tabular}{ | l | l | l | }
+			%{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
 	 	\hline
@@
 		\end{tabular}
+	%}
 \end{table}
-\end{center}
 
@@
 This component publishes the received messages on the actual topic "/cmd$\_$vel" which is subscribed by the robot controller.
-The BT can modify the Decision Gate parameters to either disregard incoming messages of Nav2 entirely or to modify them. The modification is helpful when the situation demands the robot to slow down. 
+The BT can modify the Decision Gate parameters to either disregard incoming messages of Nav2 entirely or to modify them.
+The modification is helpful when the situation demands the robot to slow down.
+
 If the Decision Gate receives messages from Nav2 and the BT, it will prioritize the BT.
@@
 \begin{figure}[ht]
-	\includegraphics[width=1.0\textwidth]{images/simplified_bt.png}
+	\centering
+	\includegraphics[width=0.9\textwidth]{images/simplified_bt.png}
 	\caption{Top Level Behavior Tree Structure}
@@
 \subsection{BT System Supervision}
+\label{subsec:BT System Supervision}
 
@@
 \begin{figure}[ht]
-	\includegraphics[width=0.4\textwidth]{images/watchdog_pattern.png}
+	\centering
+	\includegraphics[width=0.5\textwidth]{images/watchdog_pattern.png}
 	\caption{Examiner pattern for the system supervisor component \cite{konrad2003defining}}
@@
 If a node stops sending messages or responding completely, the external execution checker can inform the BT via a provided service of a problem with the node.
+
 ROS2 lifecycle nodes provide information about their health via an FSM implementation inside the node.
@@
 However, the node's output needs to be checked, too, because it might be that the FSM inside of the node did not pick up on an error and failed to change the internal lifecycle state.
-The system supervision is constantly checking the system's health. The BT is using service calls to get the health information about the components checked inside Condition Nodes. 
+
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=0.9\textwidth]{images/sensor_fallback.png}
+	\caption{Lidar Fallback}
+	\label{fig:lidar_fallback}
+\end{figure}
+
+The system supervision is constantly checking the system's health.
+The BT is using service calls to get the health information about the components checked inside Condition Nodes.
 The components get checked by the BT inside of a Fallback.
 If the condition for the execution is met, the Condition Node returns "Success", and the fallback would exit to move on to the next fallback and execution condition check (see fig.
-\ref{lidar_fallback}).
+\ref{fig:lidar_fallback}).
 If the execution checker detects a node failure, a sequence to react to the sensor failure is executed in which the robot slows down to a minimum and tries to restart the sensor.
@@
 This behavior pattern is realized for the lidar, the IMU, and the wheel odometry.
-\begin{center}
-\begin{figure}[ht]
-	\includegraphics[width=1.0\textwidth]{images/sensor_fallback.png}
-	\caption{Lidar Fallback}
-	\label{fig:lidar_fallback}
-\end{figure}
-\end{center}
 
@@
 \subsection{BT Advanced Behaviors}
+
 After the BT has checked for sensor failures and can guarantee that the environment representation is accurate, the system can check for more complicated scenarios to improve autonomy.
@@
 
-One of the biggest problems mentioned in chapter 3.1 is the inability to react to collisions.
+One of the biggest problems mentioned in section \ref{sec:Current State} is the inability to react to collisions.
 The collision can happen due to inadequate sensory coverage for detecting obstacles or general sensor inadequacy, which can not detect obstacles because of their shape or surface material.
@@
 However, even if the robot could detect all obstacles, it would still be vulnerable to outside perturbations, e.g., other robots driving into it or humans accidentally touching the robot.
+
 In order to combat this weakness, the robot must be able to detect that a collision occurred, get out of the collision state, update the map and reset all costmaps to generate a new path to the original goal.
@@
 \begin{figure}[ht]
-	\includegraphics[width=1.0\textwidth]{images/collision_fallback.png}
+	\centering
+	\includegraphics[width=0.9\textwidth]{images/collision_fallback.png}
 	\caption{Collision Fallback Behavior}
@@
 
-\subparagraph*{}
 The collision checker uses a simulated sensor to detect collisions.
@@
 The collision checker is integrated into the system supervisor, and the BT makes service calls to the execution checker to get the collision state of the robot periodically.
+
 The child nodes of the collision sequence use the sensor data backup.
@@
 %\textit{Formula for determining execution time}
-\subparagraph*{}
+
 Regarding the map reset, it has to be determined if a new map is being created with a SLAM method or if the map server provided a previously saved map.
@@
 When the map server component from Nav2 provides a static map, the map needs to be updated with a new obstacle at the point of collision and republished by the map server.
+
 The activity of the "MapUpdateAfterCollision" action node is depicted in figure \ref{fig:activity_diagram_collision}.
@@
 Hence the obstacle position is simplified to always be in front of the robot.
+A $25$ by $25$ $cm$ square with $100$ percent occupation at the collision point is added to the current map and then saved to be reloaded by the map server.
+
 %The collision position is calculated with 
@@
 
-A 25 by 25 cm square with 100 percent occupation at the collision point is added to the current map and then saved to be reloaded by the map server. 
 %as depicted in figure xx. 
@@
 
-\subparagraph*{}
 The original goal is then republished to replan the global path on the updated map. A more detailed view of the whole behavior is depicted in figure \ref{fig:activity_diagram_collision}.
 
-\begin{figure}[ht]
-	\includegraphics[width=0.8\textwidth]{images/activity_diagram_collision.png}
+\begin{figure*}[ht!]
+	\centering
+	\includegraphics[width=0.9\textwidth]{images/activity_diagram_collision.png}
 	\caption{Activity Diagram for the Collision Behavior}
 	\label{fig:activity_diagram_collision}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{center}
 \begin{figure}[ht]
-	\includegraphics[width=0.5\textwidth]{images/battery_fallback.png}
+	\centering
+	\includegraphics[width=0.6\textwidth]{images/battery_fallback.png}
 	\caption{Battery Fallback Behavior}
@@
 \end{figure}
-\end{center}
 
@@
 %\textit{Formel here}
-\subparagraph*{}
 
@@
 The consumption rates are not close to reality but are chosen to allow the rapid testing of the behavior.
+
 The battery package provides a service to get the battery charge and the values for the assumed consumption rates while idle and driving (compare \ref{tab:custom_interfaces}.
@@
 By shifting the goal to a nearby alternative goal, the robot tries to navigate, although the original goal is not reachable.
+
 If the location is still unreachable, the goal gets moved further away from the original.
@@
 
-\begin{center}
 \begin{figure}[ht]
-	\includegraphics[width=0.7\textwidth]{images/global_planning_fallback.png}
+	\centering
+	\includegraphics[width=0.6\textwidth]{images/global_planning_fallback.png}
 	\caption{Path Planning Fallback Behavior}
@@
 \subsection{Gazebo Sensor Drivers}
-%\label{sec:gazebo_driver}
+\label{subsec:gazebo_driver}
+
 For testing purposes, the simulated sensor data is not published directly on the actual topic by Gazebo.
@@
 
-\begin{center}
 \begin{figure}[ht]
-	\includegraphics[width=0.75\textwidth]{images/rqt_graph_sensor_drivers.png}
+	\centering
+	\includegraphics[width=0.7\textwidth]{images/rqt_graph_sensor_drivers.png}
 	\caption{Graph for the Sensor Driver nodes and topics}
@@
 
-Because of the necessity to have ROS nodes spinning outside of the BT thread, as discussed in 4.4.1, additional interfaces needed to be created in order to receive information through service calls.
+Because of the necessity to have ROS nodes spinning outside of the BT thread, as discussed in section \ref{subsec:BT System Supervision}, additional interfaces needed to be created in order to receive information through service calls.
 The created service types are listed in table \ref{tab:custom_interfaces}.
@@
 \begin{table}[ht]
+	\centering
 	\caption{Implemented Custom Services}
 	\label{tab:custom_interfaces}
+	\renewcommand{\arraystretch}{1.5}
+	\resizebox{0.9\textwidth}{!}{%
 		\begin{tabular}{ | m{0.2\textwidth} | m{0.33\textwidth}| m{0.33\textwidth} | m{0.2\textwidth} |} 

### 2022-09-26 Thesis/text/methoden.tex
ahmed's work of god

@@
 \section{Current State}
+\label{sec:Current State}
+
 \subsection{ROS}
-The Robot Operating System (ROS) is an open-source middleware for creating robot applications. It is not an operating system (OS) since it needs an underlying OS, mainly Linux-based, to work. 
+The Robot Operating System (ROS) is an open-source middleware for creating robot applications.
+It is not an operating system (OS) since it needs an underlying OS, mainly Linux-based, to work.
 ROS can handle the communication between different programs (called nodes) with standardized interfaces and network protocols.
@@
 \begin{table}[ht]
+	\centering
 	\caption{ROS Communication Options}
 	\label{tab:ros_interfaces}
-	\begin{tabular}{ | m{0.15\textwidth} | m{0.2\textwidth}| m{0.35\textwidth} | m{0.2\textwidth}|} 
+	\renewcommand{\arraystretch}{1.5}
+	\resizebox{0.9\textwidth}{!}{%
+		\begin{tabular}{ | m{0.15\textwidth} | m{0.3\textwidth}| m{0.3\textwidth} | m{0.2\textwidth}|} 
 	 	\hline
@@
 	 	\hline
-  	Topics & Publisher/ Subscriber & Continuous stream of data e.g. sensor data &  n:m \\ 
+	 	Topics & Publisher / Subscriber & Continuous stream of data e.g. sensor data & n:m \\ 
 	 	\hline
-  	Services & Request/ Response & Get specific data only once & One service, many clients \\ 
+	 	Services & Request / Response & Get specific data only once & One service, many clients \\ 
 	 	\hline
-  	Actions & Request/ Response and Publish/ Subscribe & Trigger executing of asynchronous goal-driven processes and receive updates & One action, many clients \\
+	 	Actions & Request / Response and Publish/ Subscribe & Trigger executing of asynchronous goal-driven processes and receive updates & One action, many clients \\
 	 	\hline
 		\end{tabular}
+	}
 \end{table}
@@
 \subsection{Navigation2}
+
 The ROS Navigation2 Stack (Nav2) combines different packages that allow mobile robots to navigate from point A to point B.
@@
 \begin{figure}[ht]
-	\includegraphics[width=1.0\textwidth]{images/nav2_architecture.png}
+	\centering
+	\includegraphics[width=0.9\textwidth]{images/nav2_architecture.png}
 	\caption{Navigation2 Architecture \cite{macenski2020}}
@@
 \end{figure}
-\subparagraph*{}
+
 These packages are often integrating a SLAM method to build or enlarge maps.
-Path planning and execution capabilities, which correspond to the global and local planner described in section 2.2, are further supported by ready-to-use planning plugins that use A* and Dijkstra algorithms for path planning and a dynamic window approach for path execution.
+Path planning and execution capabilities, which correspond to the global and local planner described in section \ref{sec:Autonomous Driving Navigation Architectures}, are further supported by ready-to-use planning plugins that use A* and Dijkstra algorithms for path planning and a dynamic window approach for path execution.
 The system sequencing is done in a hierarchical structure as a central behavior tree calls asynchronous actions from the respective planners after another, as seen in figure \ref{fig:nav_architecture}.
@@
 \end{figure}
-\subparagraph*{}
+
 The behavior tree depicted in figure \ref{fig:bt_nav} is the one that is the default Nav2 behavior and has a set of recovery behaviors already implemented (spin, wait, back up, clear map).
@@
 Nav2 uses a node lifecycle management system to start, activate, deactivate, and shut down nodes in a controlled way.
-The lifecycle approach allows the system degree to monitor the system node's startups, executions, and failures.
-\subparagraph*{}
+The lifecycle approach allows the system to monitor the system node's startups, executions, and failures.
+
 With this quantity of functions and hierarchical architecture, the ROS Navigation2 stack is an excellent basis for achieving high levels of robot autonomy.
 The software can be modified and expanded with plugins to fit the users' needs.
-When comparing the functionalities of the software stack to the proposed autonomous system architecture in figure 2.2, the things that Navigation2 is missing are a system-wide supervision layer and a more deliberative approach to behavior planning, as mentioned in chapter 2.3.
-\subparagraph*{}
+When comparing the functionalities of the software stack to the proposed autonomous system architecture in figure 	\ref{fig:autonomous_driving_architecture}, the functionalities that Navigation2 is missing are a system-wide supervision layer and a more deliberative approach to behavior planning, as mentioned in section \ref{sec:Behavior Types}.
+
 The absence of these components leads to a limited set of circumstances in which the robot can perform navigation reliably.
@@
 Unforeseen events can severely limit the reliability of the environment representation without the robot being aware of it.
+
 Examples of such events currently not handled appropriately are slipping wheels, orientation changes by external influences, or undetected, more specifically, undetectable obstacles.
-%\textit{Add statistics about random sensor failures every ~10000h} 
 This can lead to erratic movement of the robot, as reactive behaviors and more deliberative behaviors compete for authority over the robot's movement.
@@
 However, the deliberative behavior's planning is based upon a false representation of reality, leading to unsafe commands.
+
 The unsafe planning triggers reactive-type behaviors again to counteract the commands from the deliberative behavior.
@@
 \subsection{Turtlebot3}
+\label{subsec:Turtlebot3}
+
 The Turtlebot3 is a standard mobile research platform with an available ROS interface to control the robot.
 The robot is well integrated into the ROS ecosystem and is established in the literature as a system to develop and integrate new methods.
-The robot has two motors with attached wheel encoders to drive and steer, classifying it as a differential drive robot. A third omnidirectional wheel stabilizes the robot. 
-\subparagraph*{}
+The robot has two motors with attached wheel encoders to drive and steer, classifying it as a differential drive robot.
+A third omnidirectional wheel stabilizes the robot.
+
 The manufacturer provides an open-source model for simulating the robot in physics-based simulators like Gazebo, as depicted in figure \ref{fig:turtlebot}, which enables faster development and testing for the robot.
-\begin{center}
+
 \begin{figure}[ht]
+	\centering
 	\includegraphics[width=0.5\textwidth]{images/turtlebot_sim.png}
@@
 \end{figure}
-\end{center}
+
 The equipped sensors and easy simulation capabilities make the robot well suited for using the Nav2 stack and further development for behavior planning.
-
 The robot specifications are specified in table \ref{tab:turtlebot_spec}.
@@
 \begin{table}[ht]
+	\centering
 	\caption{Turtlebot3 Specifications}
 	\label{tab:turtlebot_spec}
-	\begin{tabular}{| m{0.3\textwidth} | m{0.3\textwidth} |} 
+	\renewcommand{\arraystretch}{1.5}
+	\resizebox{0.9\textwidth}{!}{%
+		\begin{tabular} {| l | l |} 
+			%{| m{0.3\textwidth} | m{0.3\textwidth} |} 
 	 	\hline
@@
 	 	Maximum rotational velocity & 1.82 rad/s (104.27 deg/s) \\
-  	\hline
-  	Maximum payload & 30kg \\
+%	 	\hline
+%	 	Maximum payload & 30kg \\
 	 	\hline
@@
 \section{Requirements}
+\label{sec:Requirements}
 
-The software requirements that improve the robot's behavior are derived from the current limitations of the standard ROS and Navigation2 setup for mobile robots described in the previous chapter. 
-The functional requirements and their acceptance criteria are listed in \ref{tab:fn_req}. The non-functional requirements are listed seperately in \ref{tab:nonfn_req}. 
+The software requirements that improve the robot's behavior are derived from the current limitations of the standard ROS and Navigation2 setup for mobile robots described in the previous section.
+The functional requirements and their acceptance criteria are listed in \ref{tab:fn_req}.
+The non-functional requirements are listed seperately in \ref{tab:nonfn_req}.
 To better contextualize the context of the systems requirements, figure \ref{fig:use_case} depicts a Use-Case diagram with the needed functions to improve the system's autonomy.
@@
 \begin{table}[ht]
+	\centering
+	\caption{Non-Functional Requirements}
+	\label{tab:nonfn_req}
+	\renewcommand{\arraystretch}{1.5}
+	\resizebox{0.9\textwidth}{!}{%
+		\begin{tabular}{ | m{0.13\textwidth} | m{0.13\textwidth}| m{0.1\textwidth} | m{0.55\textwidth} |} 
+			\hline
+			\textbf{Nr.} & \textbf{Name} & \textbf{Priority} & \textbf{Description} \\ 
+			\hline
+			non$\_$freq1 & Single Point of Failure & High & The system does not have a single point of failure	When parts of the system fail, the system maintains operability to a degree \\ 
+			\hline
+			non$\_$freq2 & Performance & High & The system's control loop guarantees fast reactions. The average frequency with which the system checks the sensors/navigation2 is higher than 100Hz (10ms) \\ 
+			\hline
+			non$\_$freq3 & Determinism & High & The outcome for a given set of inputs must be deterministic, meaning that the behavior is always executed similarly. \\
+			\hline
+			non$\_$freq4 & Deliberate & High & The robot can execute deliberative behaviors, meaning that behaviors new implemented behaviors go beyond reacting to sensor input and have a planning aspect. \\
+			\hline
+		\end{tabular}
+	}
+\end{table}
+
+\begin{table}[ht]
+	\centering
 	\caption{Functional Requirements}
 	\label{tab:fn_req}
+	\renewcommand{\arraystretch}{1.5}
+	\resizebox{0.9\textwidth}{!}{%
 		\begin{tabular}{| m{0.08\textwidth} | m{0.13\textwidth}| m{0.09\textwidth} | m{0.60\textwidth}|} 
@@
 		\end{tabular}
-\end{table}
-
-\begin{table}[ht]
-	\caption{Non-Functional Requirements}
-	\label{tab:nonfn_req}
-	\begin{tabular}{ | m{0.13\textwidth} | m{0.13\textwidth}| m{0.1\textwidth} | m{0.55\textwidth} |} 
-  	\hline
-  	\textbf{Nr.} & \textbf{Name} & \textbf{Priority} & \textbf{Description} \\ 
-  	\hline
-  	non$\_$freq1 & Single Point of Failure & High &  The system does not have a single point of failure	When parts of the system fail, the system maintains operability to a degree \\ 
-  	\hline
-  	non$\_$freq2 & Performance & High & The system's control loop guarantees fast reactions. The average frequency with which the system checks the sensors/navigation2 is higher than 100Hz (10ms) \\ 
-  	\hline
-  	non$\_$freq3 & Determinism & High & The outcome for a given set of inputs must be deterministic, meaning that the behavior is always executed similarly. \\
-  	\hline
-  	non$\_$freq4 & Deliberate & High & The robot can execute deliberative behaviors, meaning that behaviors new implemented behaviors go beyond reacting to sensor input and have a planning aspect. \\
-  	\hline
-	\end{tabular}
+	}
 \end{table}
@@
 \begin{figure}[ht]
+	\centering
 	\includegraphics[width=0.7\textwidth]{images/use_case.png}
@@
 Also, an external system can more reliably monitor other systems when its execution is detached from the performance and execution of other systems.
+
 A suitable name for the second system is the "autonomy layer", as it aims to increase the robustness and autonomy of the whole system.
@@
 The constant execution monitoring equips the system with the ability to deal with the event of node failures and component malfunctions.
-\subparagraph*{}
+
 Another required addition to the autonomy layer is a data storage of relevant system information.
@@
 Using past data points allows behaviors to become more deliberative in their approach and opens up many possibilities for intelligent behaviors like movement predictions of obstacles.
-\subparagraph*{}
+
 A third component to the autonomy layer is a behavior planner largely independent of Nav2 execution and planning.
@@
 Based on this information, the behavior planner can make a deliberative, strategic decision to increase robustness and autonomy, thus decreasing reliance on human supervision.
-\subparagraph*{}
+
 Additionally, a component to switch intelligently between speed commands from Nav2 and the autonomy layer is needed.
@@
 The proposed additional components, discussed in this chapter are colored in blue.
-Existing components, discussed in chapter 3.1, are colored in gray.
+Existing components, discussed in section \ref{sec:Current State}, are colored in gray.
 
-\begin{center}
 \begin{figure}[ht]
-	\includegraphics[width=1.0\textwidth]{images/block_diagram.png}
+	\centering 
+	\includegraphics[width=0.9\textwidth]{images/block_diagram.png}
 	\caption{System Block Diagram}
@@
 \end{figure}
-\end{center}
 
 \section{Software and Simulation}
-A comparison and assessment of the behavior planning approach were made in chapter 2.4.
+
+A comparison and assessment of the behavior planning approach were made in section \ref{sec:Behavior Planning Approaches}.
 The decision on which approach to take fell on the behavior tree method.
@@
 \subsection{Behaviortree.CPP}
+
 The Navigation2 stack uses a behavior tree to coordinate the planning layer.
@@
 This library offers a fast way to create, execute, monitor, and edit behavior trees.
-Additionally to the types of control nodes that are mentioned in chapter 2.4.2, the library adds the concept of reactivity into the catalog of control nodes.
-Reactive sequences and reactive fallbacks differ from normal ones in handling nodes that return the state "running".
-Instead of ticking the node again, the whole sequence restarts, which is very useful for continuously checking a condition and executing an asynchronous action node that gets halted when a condition is not met anymore (compare table \ref{tab:control_nodes_bt}).
+
+Additionally to the types of control nodes that are mentioned in section \ref{subsec:Behavior Trees}, the library adds the concept of reactivity into the catalog of control nodes.
+Reactive sequences and reactive fallbacks differ from normal ones in handling nodes that return the state "Running".
+Instead of ticking the node again, the whole sequence restarts, which is very useful for continuously checking a condition and executing an asynchronous action node that gets halted when a condition is not met anymore (see table \ref{tab:control_nodes_bt}).
 
 \begin{table}[ht]
+	\centering
 	\caption{Control Nodes in BT.CPP}
 	\label{tab:control_nodes_bt}
-	\begin{tabular}{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
+	\renewcommand{\arraystretch}{1.5}
+	\resizebox{0.9\textwidth}{!}{%
+		\begin{tabular}{ | l | l | l |} 
+%			{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
 	 	\hline
@@
 
-\begin{center}
 \begin{table}[ht]
+	\centering
 	\caption{Decorator Nodes in BT.CPP}
 	\label{tab:decorators_bt}
+	\renewcommand{\arraystretch}{1.5}
+	\resizebox{0.9\textwidth}{!}{%
 		\begin{tabular}{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
@@
 		\end{tabular}
+	}
 \end{table}
-\end{center}
 
-\subparagraph*{}
 To allow the tree nodes to communicate, the library provides the developer with two possibilities.
@@
 Groot is a software to create, edit, monitor, and debug behavior trees with a graphical user interface.
-The software allows the creation of behavior trees in XML files which can be directly loaded into and used by the BT.CPP library.
-Groot can monitor the live execution of behavior trees and allows the introspection of how the tree reacts to different scenarios (compare fig.
+The software allows the creation of behavior trees in XML files, which can be directly loaded into and used by the BT.CPP library.
+Groot can monitor the live execution of behavior trees and allows the introspection of how the tree reacts to different scenarios (see fig.
 \ref{fig:groot}).
 
-\begin{center}
 \begin{figure}[ht]
-		\includegraphics[width=0.7\textwidth]{images/groot.png}
+	\centering 
+	\includegraphics[width=0.9\textwidth]{images/groot.png}
 	\caption{Live monitoring of behavior trees with Groot}
@@
 \subsection{Gazebo Simulation Environment}
-The development and testing of behaviors will be done with the Gazebo simulator and the Turtlebot3 model.
+
+The development and testing of behaviors are done with the Gazebo simulator and the Turtlebot3 model.
 Gazebo is well-integrated into ROS and ROS2 and offers many ROS interfaces to control the simulation.
 Gazebo accurately models the physics of the robot and can simulate and publish the wheel odometry directly on a ROS topic.
-Furthermore, the simulator can simulate the readings for the laser scanner and IMU data via plugins.
+Furthermore, the Gazebo has an integrated ODE physics engine and can simulate the readings for the laser scanner and IMU data via plugins.
 The use of the simulator allows easier repeatability for test cases as various obstacles can be spawned at any time, and sensor failures can be easily induced, resulting in quicker overall development.
+
 Using a simulator allows slowing down or speeding up the environment time.
 The time control allows more intricate observation of fast reactive type behaviors in slow motion while also enabling the observation of the long-time robustness of the system during extended tests.
-The ROS community offers many different simulation environments to test the robot in. The environments range from simple geometric structurs, to office spaces up to complete race tracks.
+The ROS community offers many different simulation environments to test the robot in.
+The environments range from simple geometric structures, to office spaces up to complete race tracks.
 
 \subsection{OS and Software Versions}
-The software versions used to implment and test the system on are listed below in table \ref{tab:versions}.
+
+The software versions used to implement and test the system on are listed below in table \ref{tab:versions}.
 Many of the software versions are determined by the choice of the ROS distribution (distro).
 ROS2 Foxy is not the newest ROS2 distribution but is well supported and developed, so the decision was made to use this distribution.
-The decision why the turtlebot is used is laid out in chapter 3.1.3.
+The decision why the turtlebot is used is laid out in section \ref{subsec:Turtlebot3}.
 Although Python and C++ can be used to develop applications with ROS2, the behavior tree library relies on C++, which is why C++ was used for developing the whole system.
 
-\begin{center}
 \begin{table}[ht]
+	\centering
 	\caption{Used Software Versions}
 	\label{tab:versions}
-	\begin{tabular}{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
+	\renewcommand{\arraystretch}{1.5}
+	\resizebox{0.9\textwidth}{!}{%
+		\begin{tabular}{ | l | l | l |} 
+%			{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
 	 	\hline
@@
 		\end{tabular}
+	}
 \end{table}
-\end{center}

### 2022-09-26 Thesis/text/verwandt.tex
ahmed's work of god

@@
 The definition specifies the needed requirements to make a robot fully autonomous.
+
 \begin{figure}[ht]
-	\includegraphics[width=1.0\textwidth]{images/j3016graphic_2021.png}
+	\centering
+	\includegraphics[width=0.9\textwidth]{images/j3016graphic_2021.png}
 	\caption{SAE Levels Of Driving Automation \cite{J3016_202104}}
@@
 \end{figure}
+
 The automotive industry defines vehicle autonomy based on the human driver's need for supervision and possible intervention when executing complex driving tasks \cite{J3016_202104}.
 Figure \ref{fig:sae_levels} depicts the different levels of autonomy of passenger cars on a scale between no automation to completely autonomous.
+
 Both definitions rely on human supervision and guidance as the central part of what hinders a system or vehicle from becoming fully autonomous.
@@
 \section{Autonomous Driving Navigation Architectures}
+\label{sec:Autonomous Driving Navigation Architectures}
 
@@
 Most of the current autonomous driving architectures implement a hierarchical structure that follows the "Sense - Think - Act" paradigm \cite{murphy2000}.
+
 \begin{figure}[ht]
-	\includegraphics[width=1.0\textwidth]{images/autonomous_driving_architecture.png}
-	\caption{Common Autonomous Driving Architecture \cite{brooks1986} \cite{velasco2020}}
+	\centering
+	\includegraphics[width=0.9\textwidth]{images/autonomous_driving_architecture.png}
+	\caption{Common Autonomous Driving Architecture \cite{brooks1986,velasco2020}}
 	\label{fig:autonomous_driving_architecture}
 \end{figure}
+
 As shown in figure \ref{fig:autonomous_driving_architecture}, the sensory inputs, usually from multiple sensors, get processed to create an environment representation.
@@
 
-\subparagraph*{}
 The planning sequence can be further divided into global, behavioral, and local planning.
@@
 In this analogy, the behavior planner is responsible for following the traffic rules on the trip to the destination, e.g., stopping at red lights, giving the right of way to other vehicles, and following speed limits.
-The local planning, also named motion planning, takes the input from the behavior layer and is tasked to let the vehicle drive according to the executed behavior so that it stays in its lane and stops at a red light at the correct position\cite{reke2020}.
+The local planning, also named motion planning, takes the input from the behavior layer and is tasked to let the vehicle drive according to the executed behavior so that it stays in its lane and stops at a red light at the correct position \cite{reke2020}.
 
-\subparagraph*{}
 Due to safety considerations, many autonomous driving architectures implement an additional system supervision layer that monitors the execution of the other system components.
@@
 The supervisor triggers measures to restore the standard functionality or, if that is not possible, bring the robot into a safe state \cite{zimmermann2020adaptive}.
-Considering that with these architectures, vehicles can achieve an SAE level of autonomy up to level 4 \cite{bacha2008odin},  a good starting point to increase autonomy in a robotic system would be to ensure that all of the functional modules in figure xx are implemented and available.
+
+Considering that with these architectures, vehicles can achieve an SAE level of autonomy up to level 4 \cite{bacha2008odin}, a good starting point to increase autonomy in a robotic system would be to ensure that all of the functional modules in figure \ref{fig:autonomous_driving_architecture} are implemented and available.
 On this basis, higher autonomy levels are achievable by expanding the behavior planning module inside the planning layer.
@@
 \section{Behavior Types}
+\label{sec:Behavior Types}
 
@@
 \subsection{Reactive}
+
 During the emergence of behavior-based robots in the 1960s, the first type of behaviors implemented were simple reactive behaviors.
@@
 Reactive behaviors can be chained together to achieve more advanced and goal-driven behaviors.
-\subparagraph*{}
+
 Despite the possibility of more complex robot behaviors, sequential behaviors remain on the level of reactive behaviors due to the lack of a planning and decision-making cycle during their execution.
@@
 This property makes this behavior useful for system supervisors, where immediate reactions with low latency are sometimes required to ensure system safety.
+
 However, a system with a reactive behavior planning approach will always fail to meet the requirements for higher levels of autonomy due to the reasons discussed in previous chapters.
@@
 \subsection{Deliberative}
+
 Reactive behaviors suffer the drawback that they are not suitable to enable complex behaviors.
@@
 Using deliberative behaviors in the planning module enables a robot to meet the definition of autonomy, e.g., making decisions and reacting to unforeseen circumstances.
-\subparagraph*{}
+
 Concluding, the focus should be on creating deliberative behaviors to improve autonomy.
@@
 Using older data points allows the planner to proactively change the motion planning commands to better adapt to the environment.
+
 For example, a motion prediction of obstacles in highly dynamic environments would improve motion planning considerably.
@@
 A purely reactive planner could never determine if an obstacle is destined to intersect the robot's path in the future and would constantly reroute to accomplish the given goal.
+
 On the other hand, a deliberative planner can determine if the best course of action is to stay on the current path as the dynamic obstacle has already moved away by the time the robot reaches the intersection point.
@@
 \subsection{Hybrid}
+
 One can combine reactive and deliberative behaviors in a hybrid model to create reliable and intelligent systems.
@@
 \begin{figure}[ht]
-	\includegraphics[width=1.0\textwidth]{images/Deliberative_hierarchical_planning.png} 
+	\centering
+	\includegraphics[width=0.9\textwidth]{images/Deliberative_hierarchical_planning.png} 
 	\caption{Hybrid Planning Architecture \cite{arkin1998}}
@@
 When the system is not addressing these points, deliberative robots are very focused on a small problem domain and are not robust \cite{arkin1998}.
+
 The incorporation of reactive behaviors into behavior planning combats this problem.
@@
 \section{Behavior Planning Approaches}
+\label{sec:Behavior Planning Approaches}
 
@@
 \begin{figure}[ht]
-	\includegraphics[width=1.0\textwidth]{images/fsm.png} 
+	\centering
+	\includegraphics[width=0.9\textwidth]{images/fsm.png} 
 	\caption{Example of a finite state machine transition diagram \cite{wagner2006}}
@@
 
-\subparagraph*{}
 Figure \ref{fig:fsm} depicts an example state machine with different states and exemplary transitions between them.
@@
 A state transition diagram is an excellent way to understand the system behavior but can not show the details of the modeled behavior, like the actions \cite{wagner2006}.
-\subparagraph*{}
+
 Creating small state machines can easily and quickly be accomplished using various feature-rich, and performant libraries \cite{fourakis2014}. 
@@
 The nested state machines reduce the complexity of the state machine as the transitions do not need to be modeled because a nested state machine's start and final state are treated as just one state inside the bigger state machine.
+
 Despite this possibility to simplify state machines, as a modeled behavior system grows larger to accommodate more complex behaviors into the state machine, the effort to expand and maintain the state machine multiplies rapidly.
 The higher effort is due to the modeling of the transitions, transition conditions, and transition events growing with each new state that gets introduced into the state machine as existing states need to be checked and updated accordingly \cite{conner2017}.
-\subparagraph*{}
+
 The execution time of finite state machines can be fast enough that they are used to control the motors of a bipedal robot to enable the robot to balance and walk despite unexpected height variations between steps \cite{park2013}.
@@
 \subsection{Behavior Trees}
+\label{subsec:Behavior Trees}
+
 Behavior Trees (BT) are another way to model and control the behavior of autonomous systems.
@@
 Leaf nodes are also called execution nodes, while non-leaf nodes are control or control flow nodes.
-\subparagraph*{}
+
 \begin{table}[ht]
+	\centering
 	\caption{The five types of nodes \cite{iovino2022}}
 	\label{tab:node_types}
+	\renewcommand{\arraystretch}{1.5}
+	\resizebox{0.9\textwidth}{!}{%
 		\begin{tabular}{ | m{0.13\textwidth} | m{0.13\textwidth}| m{0.2\textwidth} | m{0.2\textwidth} | m{0.2\textwidth} |} 
@@
 	 	\hline
-  	Parallel & -$>$ -$>$ & If $>$= M children succeed & If $>$ N-M children fail & else \\
+	 	Parallel & -$>$ -$>$ & If $>=$ M children succeed & If $>$ N-M children fail & else \\
 	 	\hline
@@
 		\end{tabular}
+	}
 \end{table}
+
 The execution of a behavior tree is done by ticking the tree's root node.
@@
 Nodes return either "Success", "Running" or "Failure" which influences how the tick signal gets processed by the rest of the behavior tree.
+
 Control nodes can be of the type "Sequence", "Fallback" or "Condition".
@@
 If any children nodes return "Failure", the sequence node will stop ticking the remaining children and return "Failure".
+
 On the other hand, the fallback node is an equivalent of a logical "or" condition.
@@
 The primary node types and their control flow modification are listed in table \ref{tab:node_types}.
+
 Figure \ref{fig:bt_example} depicts an example of a BT with multiple levels and action nodes.
-The nodes with the arrow (-$>$) symbol are sequence nodes, and the question mark (?) symbol denotes fallback nodes.
+The nodes with the arrow (-$>$) symbol are sequence nodes, and the question mark ($?$) symbol denotes fallback nodes.
 The condition nodes are depicted as white ovals, and the action nodes are represented as gray rectangles in the behavior tree example.
 
-\begin{figure}[h!]
-	\includegraphics[width=1.0\textwidth]{images/bt_example.jpg} 
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=0.9\textwidth]{images/bt_example.jpg} 
 	\caption{Example of a behavior tree \cite{iovino2022}}
@@
 \subsection{Partially Observable Markov Decision Processes}
-Another approach to behavior planning is the Markov Decision Process(MDP) and the further advanced Partially Observable Markov Decision Process (POMDPs).
+
+Another approach to behavior planning is the Markov Decision Process (MDP) and the further advanced Partially Observable Markov Decision Process (POMDPs).
 MDPs share structural features with Finite State Machines, but the behavior planning and state transitions are stochastic and not deterministic in contrast to FSMs.
-An MDP can be described as a tuple <S, D, A, T, R>.
+An MDP can be described as a tuple $<S, D, A, T, R>$.
 Like a finite state machine, the MDP has a set of possible states \textit{S} and a set of possible transitions \textit{T} between them.
@@
 \begin{figure}[ht]
-	\includegraphics[width=1.0\textwidth]{images/mdp.png} 
+	\centering
+	\includegraphics[width=0.9\textwidth]{images/mdp.png} 
 	\caption{Example of a fully observable Markov Decision Process \cite{andrew1999reinforcement}}
@@
 The difference between an MDP and a POMDP is that the system's current state can be unknown, and the system can make observations and estimate the state to a degree.
-The POMDP is defined as a tuple <S, D, A, T, R, Z, O> in which the MDP tuple gets extended by \textit{Z} and \textit{O}.
-\textit{Z} is the observation, and \textit{O} is the probability function of seeing Z in a given state S.
-\subparagraph*{}
+The POMDP is defined as a tuple $<S, D, A, T, R, Z, O>$ in which the MDP tuple gets extended by \textit{Z} and \textit{O}.
+\textit{Z} is the observation, and \textit{O} is the probability function of seeing \textit{Z} in a given state \textit{S}.
+
 POMDPs are well suited for applications in unknown environments because the system can make decisions when the perception information is imperfect, or the environment is not fully observable.

### 2022-09-27 Thesis/text/implementierung.tex
modification to line spacing, changed bt pictures

@@
 
-The system architecture, depicted in figure \ref{fig:component_diagram}, integrates the behavior tree into the existing architecture.
+The system architecture, depicted in figure \ref{fig:component_diagram}, integrates the autonomy layer discussed in section \ref{sec:System Solution} into the existing architecture.
 The already existing components are colored in gray, and the newly added components are colored in blue.
@@
 	\centering
-	\includegraphics[width=0.9\textwidth]{images/component_diagram_bt.png}
+	\includegraphics[width=0.92\textwidth]{images/component_diagram_bt.png}
 	\caption{Component Diagram of System Architecture}
@@
 	\centering
-	\includegraphics[width=0.9\textwidth]{images/simplified_bt.png}
+	\includegraphics[width=0.95\textwidth]{images/simplified_bt_inverted.png}
 	\caption{Top Level Behavior Tree Structure}
@@
 
-The design pattern of the system supervisor is an "Watchdog" pattern.
+The design pattern of the system supervisor is a "Watchdog" pattern.
 This pattern is depicted in figure \ref{fig:watchdog_pattern}.
@@
 	\centering
-	\includegraphics[width=0.9\textwidth]{images/sensor_fallback.png}
+	\includegraphics[width=0.9\textwidth]{images/sensor_fallback_inverted.png}
 	\caption{Lidar Fallback}
@@
 	\centering
-	\includegraphics[width=0.9\textwidth]{images/collision_fallback.png}
+	\includegraphics[width=0.95\textwidth]{images/collision_fallback_inverted.png}
 	\caption{Collision Fallback Behavior}
@@
 	\centering
-	\includegraphics[width=0.6\textwidth]{images/battery_fallback.png}
+	\includegraphics[width=0.6\textwidth]{images/battery_fallback_inverted.png}
 	\caption{Battery Fallback Behavior}
@@
 	\centering
-	\includegraphics[width=0.6\textwidth]{images/global_planning_fallback.png}
+	\includegraphics[width=0.6\textwidth]{images/global_planning_fallback_inverted.png}
 	\caption{Path Planning Fallback Behavior}
@@
 	\renewcommand{\arraystretch}{1.5}
-	\resizebox{0.9\textwidth}{!}{%
-		\begin{tabular}{ | m{0.2\textwidth} | m{0.33\textwidth}| m{0.33\textwidth} | m{0.2\textwidth} |} 
+	\resizebox{0.95\textwidth}{!}{%
+		\begin{tabular}{ | m{0.2\textwidth} | m{0.42\textwidth}| m{0.45\textwidth} |} 
 	 	\hline

### 2022-09-27 Thesis/text/methoden.tex
modification to line spacing, changed bt pictures

@@
 	\resizebox{0.9\textwidth}{!}{%
-		\begin{tabular}{ | m{0.15\textwidth} | m{0.3\textwidth}| m{0.3\textwidth} | m{0.2\textwidth}|} 
+		\begin{tabular}{ | m{0.12\textwidth} | m{0.37\textwidth}| m{0.3\textwidth} | m{0.139\textwidth}|} 
 	 	\hline
@@
 	 	\hline
-	 	Actions & Request / Response and Publish/ Subscribe & Trigger executing of asynchronous goal-driven processes and receive updates & One action, many clients \\
+	 	Actions & Request / Response and Publish / Subscribe & Trigger executing of asynchronous goal-driven processes and receive updates & One action, many clients \\
 	 	\hline
@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{images/turtlebot_sim.png}
+	\includegraphics[width=0.3\textwidth]{images/turtlebot_sim.png}
 	\caption{Turtlebot3 model in Gazebo simulator}
@@
 	\renewcommand{\arraystretch}{1.5}
-	\resizebox{0.9\textwidth}{!}{%
+	\resizebox{0.8\textwidth}{!}{%
 		\begin{tabular} {| l | l |} 
@@
 	 	Size (L x W x H) & 281mm x 306mm x 141mm \\
-	 	\hline
-	 	Weight & 1.8 kg \\
 	 	\hline
@@
 	\resizebox{0.9\textwidth}{!}{%
-		\begin{tabular}{ | m{0.13\textwidth} | m{0.13\textwidth}| m{0.1\textwidth} | m{0.55\textwidth} |} 
+		\begin{tabular}{ | m{0.11\textwidth} | m{0.14\textwidth}| m{0.1\textwidth} | m{0.7\textwidth} |} 
 			\hline
@@
 			\hline
-			non$\_$freq2 & Performance & High & The system's control loop guarantees fast reactions. The average frequency with which the system checks the sensors/navigation2 is higher than 100Hz (10ms) \\ 
+			non$\_$freq2 & Performance & High & The system's control loop guarantees fast reactions. The average frequency with which the system operates is higher than 100Hz (10ms) \\ 
 			\hline
@@
 \section{System Solution}
+\label{sec:System Solution}
 

### 2022-09-27 Thesis/text/verwandt.tex
modification to line spacing, changed bt pictures

@@
 
-This thesis looks for a way to improve the autonomy of mobile robots.
+This thesis aims to increase the autonomy of mobile robots.
 In order to define an improvement in autonomous behavior, one needs to explore the definitions of the terms "automated" and "autonomy".
@@
 	\includegraphics[width=0.9\textwidth]{images/j3016graphic_2021.png}
-	\caption{SAE Levels Of Driving Automation \cite{J3016_202104}}
+	\caption{Levels Of Driving Automation \cite{J3016_202104}}
 	\label{fig:sae_levels}
@@
 
-The automotive industry defines vehicle autonomy based on the human driver's need for supervision and possible intervention when executing complex driving tasks \cite{J3016_202104}.
+The Society of Automotive Engineers (SAE) defines vehicle autonomy based on the human driver's need for supervision and possible intervention when executing complex driving tasks \cite{J3016_202104}.
 Figure \ref{fig:sae_levels} depicts the different levels of autonomy of passenger cars on a scale between no automation to completely autonomous.
+The most significant distinction between the levels is between level two and three, because a system can take the complete control over the vehicle in certain conditions.
+The SAE levels zero to two use driver support functions with no real autonomy, while levels three to five employ automated driving features.
+Another major level difference is located on the borderline between levels three and four as the vehicle can drive autonomously without a human operator who could control the vehicle as a safety fallback when the system demands it.
+A level four vehicle can theoretically be built without a steering wheel and pedals, because it is able to handle all situations it is deployed in without the need for manual driving.
+%The most significant distinction between the levels is between level three and four, because a vehicle can drive without a human driver as a safety fallback option.
 
@@
 To better understand how behavior planning influences a robot's autonomy, one can look at the latest and most used software architectures for autonomous driving on a functional level.
-Most of the current autonomous driving architectures implement a hierarchical structure that follows the "Sense - Think - Act" paradigm \cite{murphy2000}.
+The majority of the current autonomous driving architectures implement a hierarchical structure that follows the "\textit{Sense - Think - Act}" paradigm \cite{murphy2000}.
 
@@
 
-As shown in figure \ref{fig:autonomous_driving_architecture}, the sensory inputs, usually from multiple sensors, get processed to create an environment representation.
-Then the system calculates how to get from its current position to the destination and creates a path.
-A second planning calculation is triggered to compute the motion commands with the system's constraints (e.g., size, turning radius).
-These motion commands then get converted to control the motors.
+As shown in figure \ref{fig:autonomous_driving_architecture}, the sensory inputs, usually from multiple sensors, are processed to create an environment representation.
+Based on that representation, the system calculates how to get from its current position to the destination and creates a path to the destination.
+A second planning calculation is triggered to compute the motion commands while taking the system's constraints (e.g., size, turning radius) into account.
+In a final step, these motion commands then get converted to control the motors.
 
-The planning sequence can be further divided into global, behavioral, and local planning.
-Global planning, also named route planning, is responsible for outputting a path from start to goal.
+The planning sequence can be further divided into "$global$", "$behavioral$", and "$local$" planning.
+The global planning module, also named "$route$ $planning$", is responsible for outputting a path from start to goal.
 Global planning is comparable to a person using Google Maps to find the shortest or fastest path to a faraway city.
@@
 If one or more components fail to pass the health check, the supervisor is responsible for ensuring that the system does not continue the regular driving routine.
-The supervisor triggers measures to restore the standard functionality or, if that is not possible, bring the robot into a safe state \cite{zimmermann2020adaptive}.
+The supervision component triggers measures to restore the standard functionality or, if that is not possible, bring the robot into a safe state \cite{zimmermann2020adaptive}.
 
-Considering that with these architectures, vehicles can achieve an SAE level of autonomy up to level 4 \cite{bacha2008odin}, a good starting point to increase autonomy in a robotic system would be to ensure that all of the functional modules in figure \ref{fig:autonomous_driving_architecture} are implemented and available.
-On this basis, higher autonomy levels are achievable by expanding the behavior planning module inside the planning layer.
+Considering that with these architectures, vehicles can achieve an SAE level of autonomy up to level four and even five \cite{bacha2008odin}, a good starting point to increase autonomy in a robotic system would be to ensure that all of the functional modules in figure \ref{fig:autonomous_driving_architecture} are implemented and available.
+Based on this, higher autonomy levels are achievable by expanding the behavior planning module inside the planning layer.
 
@@
 
-The behavior planning module contains a set of behaviors to operate in an environment.
-The module's task is then to choose the best behavior to execute.
+The behavior planning module contains a set of tailored behaviors to operate in a specific environment.
+The module's task is to always choose the best behavior to execute to ensure the best performance for a given task.
 In this sense, behavior is defined as a mapping of sensory inputs to a pattern of actions that carry out a specific task \cite{murphy2000}.
 A good behavior planning approach provides the robot with acceptable behaviors for every possible scenario the robot is operating in.
-Different scenarios pose different challenges for the behavior planning module, which has to decide on the best behavior option to achieve a higher level goal.
+Different scenarios pose different challenges for the behavior planning module, which has to decide on the best behavioral option to achieve a higher-level goal.
 
@@
 In human behavior, this behavior resembles reflexes, like tapping on the kneecap, which unwillingly results in motion in the knee joint.
-This behavior pattern is not following the typical "Sense, Think, Act" loop but shortcuts directly from sensing to acting \cite{desilva2008}.
+This behavior pattern is not following the typical "Sense, Think, Act" loop, but shortcuts directly from sensing to acting \cite{desilva2008}.
 Reactive behaviors can be chained together to achieve more advanced and goal-driven behaviors.
 
-Despite the possibility of more complex robot behaviors, sequential behaviors remain on the level of reactive behaviors due to the lack of a planning and decision-making cycle during their execution.
-The computational load of reactive behaviors is low and thereby fast as no decision and planning process is taking place, which makes them suited in scenarios where real-time safety is of concern.
-This property makes this behavior useful for system supervisors, where immediate reactions with low latency are sometimes required to ensure system safety.
+Despite the possibility to carry out more complex robot behaviors, sequential behaviors remain largely on the level of reactive behaviors due to the lack of a planning and decision-making cycle during their execution.
+The computational load of reactive behaviors is low and therefore fast as no decision and planning process is taking place, which makes them suited in scenarios where real-time safety is of concern.
+This property makes this behavior type useful for system supervisors, where immediate reactions with low latency are sometimes required to ensure system safety.
 
-However, a system with a reactive behavior planning approach will always fail to meet the requirements for higher levels of autonomy due to the reasons discussed in previous chapters.
+However, a system with a solely reactive behavior planning approach will always fail to meet the requirements for higher levels of autonomy due to the reasons discussed in previous chapters.
 This systematic lack of autonomy does not mean reactive behaviors can not be part of highly autonomous systems.
-Their quick response time to sensor inputs makes them valuable in improving vehicle safety and reliability.
+The quick response time to sensor inputs makes them valuable in improving vehicle safety and reliability.
 
@@
 
-Reactive behaviors suffer the drawback that they are not suitable to enable complex behaviors.
+Systems purely based on reactive behaviors suffer the drawback that they do not allow for complex behaviors.
 In addition, creating sequences of reactive behaviors that produce the target behavior is a complicated task \cite{murphy2000}.
 Behaviors that involve planning and decision-making aspects are defined as deliberative.
-Unlike reactive behaviors, deliberative behaviors follow the Sense, Think, Act paradigm, and they are not mapping sensory inputs directly into motor commands.
+Unlike reactive behaviors, deliberative behaviors follow the "\textit{Sense - Think - Act}" paradigm, and they are not mapping sensory inputs directly into motor commands.
 Instead, deliberative-type behaviors can decide the best course of action before they act.
-Using deliberative behaviors in the planning module enables a robot to meet the definition of autonomy, e.g., making decisions and reacting to unforeseen circumstances.
+Using deliberative behaviors inside of the planning module enables a robot to meet the definition of autonomy, especially in regards to the making of decisions and reacting to unforeseen circumstances.
 
-Concluding, the focus should be on creating deliberative behaviors to improve autonomy.
-A behavior planner with a deliberative approach decides based on current sensor data and previously processed data.
-By incorporating older sensor information into the decision-making process, the deliberative behavior planner can predict the changes in the environment.
-Using older data points allows the planner to proactively change the motion planning commands to better adapt to the environment.
+A behavior planner with a deliberative approach can make decisions based on current sensor data and previously processed data.
+By incorporating older sensor information into the decision-making process, the deliberative behavior planner can predict changes in the environment.
+Using older data points allows the planner to proactively adapt the motion planning commands to better achieve goals in dynamic and uncertain environments.
 
 For example, a motion prediction of obstacles in highly dynamic environments would improve motion planning considerably.
-The behavior planner can adjust the motion planning with improved information on how dynamic obstacles interact with the robot's planned path in the future.
+The behavior planner can adjust the motion planning with more precise information on how dynamic obstacles interact with the robot's planned path in the future.
 A purely reactive planner could never determine if an obstacle is destined to intersect the robot's path in the future and would constantly reroute to accomplish the given goal.
@@
 Other actions in the scenario are possibly slowing down or speeding up briefly to avoid an obstacle and staying on the current path, which is calculated to be quicker than rerouting and taking a longer path.
-Generally, these cognitively more complex behaviors mimic human skills and thus make a system more autonomous.
+
+Concluding, the focus should be on creating deliberative behaviors to improve autonomy.
+Generally, these cognitively more complex behaviors mimic human skills and thus make a system more autonomous, because the need for a human operator can be decreased greatly.
 
@@
 Deliberative planning does possess limitations during the execution of generated plans as the time horizon is long-term and not able to react to fast changes in the environment.
-When the system is not addressing these points, deliberative robots are very focused on a small problem domain and are not robust \cite{arkin1998}.
+When the system is not addressing these points, deliberative robots are very focused on a narrow problem domain and are not robust when the defined system boundaries are crossed\cite{arkin1998}.
 
 The incorporation of reactive behaviors into behavior planning combats this problem.
-In this multi-layer approach, the higher level, more deliberative planners can override the reactive planners to allow the system to be more flexible and adaptable while maintaining fast reaction times.
+In this multi-layer approach, the higher level, more deliberative planners can override the reactive planners to allow the system to be more flexible in uncertain environments and adaptable while maintaining fast reaction times.
 This architecture leads to more robust robots deployed in a wider variety of environments, thus leading to higher levels of autonomy.
@@
 
-Robots that use a hybrid, hierarchical behavior planning approach need a system that decides on a high level which behaviors are to be executed.
+Robots that use a hybrid, hierarchical behavior planning approach need a system that decides on a high level which behaviors are executed.
 That system allows overriding simple reactive behaviors with more complex, deliberative behaviors when the whole system benefits from them.
 
-The different solutions for behavior execution systems all share that they use states a robot can be in.
-Different strategies and behaviors are executed during the system's runtime based on a robot's state and additional information about the environment.
+The solutions for behavior execution systems, presented in the following sections, all share the usage of states that a robot can be in.
+Different strategies and behaviors are utilized during the system's runtime based on a robot's state and additional information about the environment.
 
@@
 They describe behavior in the form of states, which trigger the execution of actions.
-A finite state machine is defined by a list of possible states S, a starting state s0, a set of state transitions delta, a set of final states F, and the input alphabet sigma.
+A finite state machine is defined by a list of possible states \textit{}, a starting state $s_{0}$, a set of state transitions $\delta$, a set of final states $F$, and the input alphabet $\sigma$.
 At any given time, only a single state is selected, and its containing actions are executed.
+
+Figure \ref{fig:fsm} depicts an example state machine with different states and exemplary transitions between them.
+Inside of the states the letters indicate the existence of actions.
+$E$ stands for entry, $I$ for Input and $X$ for Exit.
+The arrows between the states illustrate the possible transition and transition conditions for every state.
+A state transition diagram is an excellent way to understand the system behavior, but can not show the details of the modeled behavior, like the actions \cite{wagner2006}.
 
@@
 
-Figure \ref{fig:fsm} depicts an example state machine with different states and exemplary transitions between them.
-Inside of the states the letters indicate the existence of actions.
-E stands for entry, I for Input and X for Exit.
-The arrows between the states illustrate the possible transition and transition conditions for every state.
-A state transition diagram is an excellent way to understand the system behavior but can not show the details of the modeled behavior, like the actions \cite{wagner2006}.
+The execution time of finite state machines can be fast enough to control the motors of a bipedal robot to enable the robot to balance and walk despite unexpected height variations between steps \cite{park2013}.
+The high performance makes state machines very well suited for fast, reactive type behaviors while still allowing a deliberative and hierarchical approach to behavior planning.
 
-Creating small state machines can easily and quickly be accomplished using various feature-rich, and performant libraries \cite{fourakis2014}. 
-Finite state machines are often used to model reactive behavior comprehensively, and sequential behaviors in autonomous driving functions \cite{ziegler2014}.
+Creating small state machines can be accomplished easily and quickly using various feature-rich and performant libraries \cite{fourakis2014}.
+Finite state machines are often used to model reactive and sequential behaviors comprehensively in autonomous driving functions \cite{ziegler2014}.
 These finite state machines use nested state machines inside another state machine.
@@
 
-Despite this possibility to simplify state machines, as a modeled behavior system grows larger to accommodate more complex behaviors into the state machine, the effort to expand and maintain the state machine multiplies rapidly.
-The higher effort is due to the modeling of the transitions, transition conditions, and transition events growing with each new state that gets introduced into the state machine as existing states need to be checked and updated accordingly \cite{conner2017}.
-
-The execution time of finite state machines can be fast enough that they are used to control the motors of a bipedal robot to enable the robot to balance and walk despite unexpected height variations between steps \cite{park2013}.
-This property makes state machines very well suited for fast, reactive type behaviors and still allows a hierarchical approach to behavior planning.
-
+Despite the possibility of nesting state machines, the effort to expand and maintain a FSM multiplies rapidly, when the system becomes increasingly larger to integrate more complex behaviors.
+The higher effort is due to the required creation of the transitions, transition conditions, and transition events.
+The FSM needs to be updated with each new state that gets introduced into the state machine as existing states need to be checked and corrected accordingly \cite{conner2017}.
 %
@@
 Every tree has one root and many children, parent, and leaf nodes.
-Leaf nodes are also called execution nodes, while non-leaf nodes are control or control flow nodes.
+Leaf nodes are also called execution or action nodes, while non-leaf nodes are control or control flow nodes.
 
@@
 	\centering
-	\caption{The five types of nodes \cite{iovino2022}}
+	\caption{The five types of behavior tree nodes \cite{iovino2022}}
 	\label{tab:node_types}
@@
 	 	\hline
-	 	Action & shaded box & Upon completion & When impossible to complete & During completion \\
+	 	Action & shaded box & Upon completion & If impossible to complete & If node is still completing tasks \\
 	 	\hline
@@
 
-Control nodes can be of the type "Sequence", "Fallback" or "Condition".
+The relevant control nodes are of the type "Sequence", "Fallback" or "Condition".
 The condition node can not have child nodes and can only return success or failure.
 A sequence node can be compared to a logical "and" condition.
-This property of the sequence node means that once a sequence gets ticked from the parent node, it will send the tick signal to every child node as long as they return "Success" of "Running" and only return "Success" itself when every child node was successfully ticked.
+This property of the sequence node means that once a sequence gets ticked from the parent node, it will send the tick signal to every child node as long as they return "Success" or "Running" and only return "Success" itself when every child node was successfully ticked.
 If any children nodes return "Failure", the sequence node will stop ticking the remaining children and return "Failure".
 
-On the other hand, the fallback node is an equivalent of a logical "or" condition.
-The fallback node ticks its children nodes as long as they return "Failure" or "Running" and will return "Success" if one of the ticked nodes returned "success".
-It will only return "Failure" if all children nodes returned "Success".
+On the other hand, the fallback node is an equivalent of a logical "Or" condition.
+The fallback node ticks its children nodes as long as they return "Failure" or "Running" and will return "Success" if one of the ticked nodes returned "Success".
+It will only return "Failure" if all children nodes returned "Failure".
 The primary node types and their control flow modification are listed in table \ref{tab:node_types}.
+Parallel control nodes are listed for completeness, but are not discussed further in this chapter due to their limited use in robotic applications.
 
@@
 The guiding principle of MDPs is to maximize the expected reward by deciding on the best action.
-This policy Pi is what guides the behavior of the MDP \cite{feyzabadi2014riskaware}.
+This policy $\pi$ is what guides the behavior of the MDP \cite{feyzabadi2014riskaware}.
 
@@
 
-The difference between an MDP and a POMDP is that the system's current state can be unknown, and the system can make observations and estimate the state to a degree.
+The difference between an MDP and a POMDP is that in a POMDP the system's current state can be unknown, and the system can make observations and estimate the state to a degree.
 The POMDP is defined as a tuple $<S, D, A, T, R, Z, O>$ in which the MDP tuple gets extended by \textit{Z} and \textit{O}.
@@
 
-POMDPs are well suited for applications in unknown environments because the system can make decisions when the perception information is imperfect, or the environment is not fully observable.
+POMDPs are well suited for applications in unknown environments because the system can make decisions if the perception information is imperfect, or the environment is not fully observable.
 The decision process is probabilistic and is more adaptable when confronted with unknown scenarios.

### 2022-09-28 Thesis/text/einleitung.tex
modification to line spacing, changed bt pictures
fertig?!

@@
 
-Mobile robots are becoming more widespread in many domains.
+Mobile robots are becoming more widespread in many domains, like warehouse logistics, last-mile delivery, and agriculture.
 They are marketed as autonomous mobile robots, but their autonomy is often limited to particular environments and requires frequent human help to function correctly.
@@
 ROS offers many tools and functionalities to create and program robots, but a standard robot needs human supervision during the operation to ensure proper function and safety.
-There are ROS packages to safely navigate from point A to point B, but currently, typical robots cannot react to unforeseen events.
 The robot needs to be able to navigate in an uncertain environment to guarantee the safety of the robot itself and all other actors in its environment.
-These can be either other robots or humans.
+There are ROS packages to enable safe navigation capabilities, but currently, typical ROS-based robots cannot react to unforeseen events.
 
 In theory, the robot can perform fully autonomous navigation, including mapping and localization.
-However, when something unexpected happens, the robot's autonomy can not be guaranteed any more.
+However, when something unexpected happens, the robot's autonomy can not be guaranteed.
 A default robot cannot represent all possible fail states and does not detect failures on a systematic level but only inside its navigation-related subsystems.
@@
 
-	\item Chapter \ref{cha:methoden} compares the current systems and derives requirements from the key problem areas
+	\item Chapter \ref{cha:methoden} analyzes the current robotic systems and derives requirements from the key problem areas
 

### 2022-09-28 Thesis/text/diskussion.tex
fertig?!

@@
 The goal was to achieve higher levels of autonomy and robustness by reducing the number of scenarios in which a human operator is needed.
-Implementing the system supervisor component improves the system's robustness and provides a safe basis for executing more intelligent behaviors.
-The implemented behaviors to react to sensor failures and robot collisions are valuable improvements to the system and improve current ROS2 mobile robots significantly.
-The layout of the behavior tree is structured to allow more complex behaviors to be safely executed because possible scenarios that interfere with a correct environment representation are checked and dealt with before complex behaviors get executed based on that representation.
-The implemented layout choice for the tree is not mentioned in the literature. However, it has proven to be a sensible approach to model robot behavior safely and robustly. 
+Implementing the system supervisor component improved the system's robustness and provided a safe basis for executing more intelligent behaviors.
+The implemented behaviors to react to sensor failures and robot collisions were valuable improvements to the system and improved current ROS2 mobile robots significantly.
+The layout of the behavior tree was structured to allow more complex behaviors to be safely executed because possible scenarios that interfere with a correct environment representation were checked and dealt with before complex behaviors get executed based on that representation.
+The implemented layout choice for the tree was not mentioned in previous literature. However, it has proven to be a sensible approach to model robot behavior safely and robustly. 
 %The implemented behaviors and the system supervision component have improved the behavior in the simulated scenarios.
@@
 Furthermore, the system architecture is a good base for further development without facing problems or limitations in the later stages of developing behaviors.
-The architecture shows a novel way that works with the asynchronous ROS2 executor properties and allows fast execution of both the behavior planner and the system components.
+The architecture showed a novel way that works with the ROS2 executor properties and allowed fast execution of both the behavior planner and the system components.
 
-Nevertheless, the behavior tree design could be modified to use blackboard values more effectively to increase the information flow between action nodes in the tree.
-Also, the tree does not utilize the reactive control flow nodes, and no action node in the tree is designed to allow multiple threads at the current state.
-Incorporating so-called stateful nodes into the tree's structure to allow co-routines would increase the possibility of creating better and more deliberative behaviors.
+Nevertheless, the behavior tree design could have been modified to use blackboard values more effectively to increase the information flow between action nodes in the tree.
+Also, the tree did not utilize the reactive control flow nodes, and no action node in the tree was designed to allow multiple threads at the current state.
+Incorporating so-called stateful nodes into the tree's structure to allow co-routines would have increased the possibility of creating better and more deliberative behaviors.
 
@@
 
+The results must be reproduced on an actual mobile robot. The layout of the behavior tree will stay the same, but further effort is needed to adapt the action nodes of the tree for a real system.
+
 The behaviors must be more deliberative and robust to achieve more independence from human operators.
@@
 %Further research is needed on how behavior trees can manage a system with autonomous machine learning behaviors inside the tree.
-
-Lastly, the results must be reproduced on an actual mobile robot.
-The layout of the behavior tree will stay the same, but further effort is needed to adapt the action nodes of the tree for a real system.

### 2022-09-28 Thesis/text/ergebnisse.tex
modification to line spacing, changed bt pictures
fertig?!

@@
 
-The robot is put into specifically designed scenarios that should trigger behavior responses from the behavior tree component.
+The robot is put into specifically designed scenarios to trigger behavior responses from the behavior tree component.
 The scenario testing is done to evaluate the efficacy of the implemented behaviors in relation to the functional requirements listed in section \ref{sec:Requirements}.
@@
 The scenarios are executed in a simulated apartment environment as seen in figure \ref{fig:house_gazebo} and a simulated world with small round obstacles in an enclosed space as depicted in figure \ref{fig:world_gazebo}.
-The environments were mapped beforehand with the ROS2 "Slam Toolbox" package and provided via the Nav2 map server.
+The environments were mapped beforehand with the ROS2 "Slam Toolbox" package and provided by the Nav2 map server.
 
@@
 
-The robot location and navigation goals are defined in another ROS Node, which programmatically spawns the robot into the environment and sets the goal this way.
+The robot location and navigation goals are defined in another ROS node, which programmatically spawns the robot into the environment and sets the goal this way.
 This way, the same scenario can be run with and without behavior planning.
@@
 	\renewcommand{\arraystretch}{1.5}
-	\resizebox{0.9\textwidth}{!}{%
-		\begin{tabular}{| m{0.1\textwidth} | m{0.2\textwidth}| m{0.3\textwidth} | m{0.3\textwidth}|} 
+	\resizebox{0.99\textwidth}{!}{%
+		\begin{tabular}{| m{0.11\textwidth}| m{0.1\textwidth} | m{0.183\textwidth}| m{0.3\textwidth} | m{0.3\textwidth}|} 
 			\hline
-			\textbf{Name} & \textbf{Requirements }& \textbf{Description} & \textbf{Success Criteria}\\ 
+			\textbf{Sensor Tested} & \textbf{Name} & \textbf{Requirements }& \textbf{Description} & \textbf{Success Criteria}\\ 
 			\hline
-			Lidar$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Lidar node crashes & Reset the system\\ 
+			 & Lidar$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Lidar node crashes & Reset the system\\ 
+			\cline{2-5}
+			Lidar &Lidar$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
+			\cline{2-5}
+			& Lidar$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
 			\hline
-			Lidar$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
+			 & IMU$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, IMU node crashes & Reset system \\
+			\cline{2-5}
+			IMU &IMU$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
+			\cline{2-5}
+			&IMU$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
 			\hline
-			Lidar$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
-			\hline
-			IMU$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, IMU node crashes & Reset system \\
-			\hline
-			IMU$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
-			\hline
-			IMU$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
-			\hline
-			Odom$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Odom node crashes & Reset the system\\ 
-			\hline
-			Odom$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
-			\hline
-			Odom$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
+			 & Odom$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Odom node crashes & Reset the system\\ 
+			\cline{2-5}
+			Odometry &Odom$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
+			\cline{2-5}
+			&Odom$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
 			\hline
@@
 \end{table}
-
-The scenarios execution is done as follows:
-
-\begin{itemize}
-	\item For the Lidar, IMU, and Odometry scenarios, the sensor driver was shut down when reaching the desired speed listed in the scenario.
-
-	\item Scenario NoPathFound1 is a scenario in which the lifecycle state gets transitioned into "inactive" and can not execute the planning action anymore when triggered.
-The BT has to recover from the inability to plan in this scenario.
-
-	\item NoPathFound$\_$2 is simulating the scenario when the planner is active but fails to find a collision-free path to the desired goal.
-This scenario tests the ability to find alternative goals and navigate to a nearby goal.
-
-	\item In the Collision$\_$1 scenario, an obstacle gets moved so close to the robot that a collision is registered.
-The forced collision tests the ability to escape a collision and resume navigation afterward.
-A new navigation goal is published after the behavior tree exits the Collision Fallback Routine.
-
-	\item Collision$\_$2 scenario is testing the same behavior but during active navigation.
-A flat obstacle is spawned in the path of the robot.
-The obstacle is too low to be detected by the laserscanner, and a collision occurs.
-The robot must be able to recover from the crash and keep on navigating towards the goal with an updated global path in this scenario.
-
-	\item The Battery$\_$1 scenario tests if the robot tries to navigate toward a distant goal with a low battery charge.
-The simulated battery is emptied to a low level, and a new navigation goal is set.
-If the battery charge runs below zero percent during the navigation, the test fails.
-\end{itemize}
 
@@
 	\resizebox{0.9\textwidth}{!}{%
-		\begin{tabular}{| m{0.1\textwidth} | m{0.2\textwidth}| m{0.4\textwidth} | m{0.3\textwidth}|} 
+		\begin{tabular}{| m{0.1\textwidth} | m{0.19\textwidth}| m{0.4\textwidth} | m{0.3\textwidth}|} 
 			\hline
@@
 			\hline	
-		Motor$\_$1 & fn$\_$req2, fn$\_$req5 & Hardware failure on the motors & n.a. \\
-		\hline	
 		\end{tabular}
@@
 \end{table}
+
+The scenarios execution is done as follows:
+
+For the Lidar, IMU, and Odometry scenarios, the sensor driver was shut down when reaching the desired speed listed in the scenario.
+
+Scenario NoPathFound1 is a scenario in which the lifecycle state gets transitioned into "inactive" and can not execute the planning action anymore when triggered.
+The BT has to recover from the inability to plan in this scenario.
+
+NoPathFound$\_$2 is simulating the scenario when the planner is active but fails to find a collision-free path to the desired goal.
+This scenario tests the ability to find alternative goals and navigate to a nearby goal.
+
+In the Collision$\_$1 scenario, an obstacle gets moved so close to the robot that a collision is registered.
+The forced collision tests the ability to escape a collision and resume navigation afterward.
+A new navigation goal is published after the behavior tree exits the Collision Fallback Routine.
+
+Collision$\_$2 scenario is testing the same behavior but during active navigation.
+A flat obstacle is spawned in the path of the robot.
+The obstacle is too low to be detected by the laserscanner, and a collision occurs.
+The robot must be able to recover from the crash and keep on navigating towards the goal with an updated global path in this scenario.
+
+The Battery$\_$1 scenario tests if the robot tries to navigate toward a distant goal with a low battery charge.
+The simulated battery is emptied to a low level, and a new navigation goal is set.
+If the battery charge runs below zero percent during the navigation, the test fails.
 

### 2022-09-28 Thesis/text/implementierung.tex
fertig?!

@@
 
-The system architecture, depicted in figure \ref{fig:component_diagram}, integrates the autonomy layer discussed in section \ref{sec:System Solution} into the existing architecture.
-The already existing components are colored in gray, and the newly added components are colored in blue.
-The new components and their functions are described in the following sections of this chapter.
+The system architecture, depicted in figure \ref{fig:component_diagram}, integrates the autonomy layer discussed in section \ref{sec:System Solution} into the current architecture.
 
@@
 	\centering
-	\includegraphics[width=0.92\textwidth]{images/component_diagram_bt.png}
+	\includegraphics[width=0.95\textwidth]{images/component_diagram_bt.png}
 	\caption{Component Diagram of System Architecture}
@@
 \end{figure}
+
+Already existing components are colored in gray, while the components that were implemented in this thesis (i.e., Data Backup, Decision Gate, Behavior Tree and System Supervisor) are depicted in blue.
+The new components and their functions are described in the following sections of this chapter.
+The System Supervisor is explained in the BT section as a significant part of the component's functionality is realized inside of the behavior tree.
 
@@
 This component subscribes to all sensor data and saves the messages in a queue data structure.
-The most recent data gets pushed into the array, and if the array exceeds a specified length, the oldest data this data gets deleted from the array.
+The most recent data gets pushed into the array, and if the array exceeds a specified length, the oldest data this data gets deleted.
 There are two main reasons to limit the array size in this way.
-On the one side, the goal is to decrease the size of the message that gets sent to the BT to limit the network load, and on the other side, the BT often does not need to look back further than a few seconds to make a decision.
-This way, the computational load, and the network load are decreased.
-
-The BT can access the saved data by sending a service call to one of the provided services by this component.
-Service calls in ROS2 are executed with high priority and are done quickly.
-However, it adds steps to the algorithm instead of just saving the data directly in the behavior tree.
-The reason for outsourcing the data backup to another component is that the ROS node needs an executor which constantly spins the node.
-The spinning action checks the network to see if there is any work for the node.
-Work, in this context, means that if one of the subscribed topics receives a message, the node has to execute one of its callback functions associated with that subscription.
-
-This spinning has to be done in regular intervals.
-Otherwise, the node might miss incoming messages from subscribed topics.
-The way to do the spinning is to have the node constantly spun in one thread.
-Having a constant spinning node would completely block the behavior tree's execution.
-If one tries to spin a node once when a node gets ticked in the BT, the problem of missing messages becomes relevant because the BT might need longer than expected to return to a node to tick it.
-That way, the BT node responsible for saving all the relevant data can not guarantee that it has picked up on all the messages between spins.
-The synchronous nature of the ROS executor spinning the nodes is why the data backup is not stored within the BT.
-The saved data inside this component with the corresponding lookback time is listed in table \ref{tab:data_backup_types}.
+Firstly, the goal is to decrease the size of the message that gets sent to the BT to limit the network load.
+Secondly, the BT often does not need to look back further than a few seconds to make a decision.
+Subsequently, the computational and network load are decreased significantly.
+The saved data sources and corresponding lookback times are listed in table \ref{tab:data_backup_types}.
 
@@
 
+The BT can access the saved data by sending a service call to one of the provided services by this component.
+Service calls in ROS2 are executed with high priority and are done quickly.
+However, it adds steps to the algorithm instead of just saving the data directly in the behavior tree.
+The reason for outsourcing the data backup to another component is that the ROS node needs an executor which constantly spins the node.
+The spinning action checks the network to see if there is any work for the node.
+Work, in this context, means that if one of the subscribed topics receives a message, the node has to execute one of its callback functions associated with its subscription.
+
+This spinning has to be done in regular intervals.
+Otherwise, the node might miss incoming messages from subscribed topics.
+Having this action inside of the behavior tree thread would completely block the execution of the entire autonomy system.
+If one tries to spin a node only once when a node gets ticked in the BT, losing entire messages is relevant concern because the BT might need longer than expected to return to a node to tick it.
+That way, the BT node responsible for saving all the relevant data cannot guarantee that it has received all the messages between spins as multiple messages might have been received between the two ticks.
+The synchronous nature of the ROS executor is why the data backup is not stored within the BT, but in in an external ROS package to seperate the threads.
+
 \section{Command Velocity Decision Gate}
@@
 The BT can modify the Decision Gate parameters to either disregard incoming messages of Nav2 entirely or to modify them.
-The modification is helpful when the situation demands the robot to slow down.
+The modification is helpful if the situation demands the robot to slow down.
 
@@
 This situation should theoretically never occur as the BT will set the parameters for the Decision Gate accordingly and cancel the current Navigation Goal before publishing commands on its own.
-It is implemented as an additional safety measure in case the cancellation of the goal is not possible if Navigation2 is not responsive to the cancellation command.
+It is implemented as an additional safety measure in case the canceling of the goal is not possible if Navigation2 is not responsive to the cancelation command.
 
@@
 The root control node is a sequence node.
-The root node is a simple sequence because the whole system must function with increasing complexity in the tree's later parts (on the right side).
+The root node is a simple sequence because the whole system must function with increasing complexity in the tree's subsequent parts (located further on the right side of the tree).
 The complex behaviors require that all previous components and conditions work as they should.
@@
 
-\subsection{BT System Supervision}
+\subsection{System Supervision}
 \label{subsec:BT System Supervision}
 
-The design pattern of the system supervisor is a "Watchdog" pattern.
-This pattern is depicted in figure \ref{fig:watchdog_pattern}.
-The pattern is used to monitor a system from the outside and directly take corrective measures when problems are detected.
+The system supervisor is using a "Watchdog" design pattern depicted in figure \ref{fig:watchdog_pattern}.
+A watchdog is used to monitor a system from the outside and take corrective measures when problems are detected.
 
@@
 The synchronous executor would stop the execution of the whole behavior tree, which is why the component needs to be moved inside of a new thread.
-The system supervisor, also called the execution checker, is receiving messages from nodes not implemented as ROS2 lifecycle nodes.
+The system supervisor, also called the execution checker, is receiving messages from nodes that are not implemented as ROS2 lifecycle nodes.
 If a node stops sending messages or responding completely, the external execution checker can inform the BT via a provided service of a problem with the node.
@@
 The current node state can be requested via a service call.
-However, the node's output needs to be checked, too, because it might be that the FSM inside of the node did not pick up on an error and failed to change the internal lifecycle state.
+However, the node's output needs to be checked, too, because it might be that the FSM inside of the node does not pick up on an error and fails to change the internal lifecycle state.
 
@@
 The BT is using service calls to get the health information about the components checked inside Condition Nodes.
-The components get checked by the BT inside of a Fallback.
-If the condition for the execution is met, the Condition Node returns "Success", and the fallback would exit to move on to the next fallback and execution condition check (see fig.
-\ref{fig:lidar_fallback}).
+The components get checked by the BT inside of a fallback node.
+If the condition for the execution is met, the condition node returns "Success", and the fallback would exit to move on to the next fallback and execution condition check as in the lidar fallback depicted in figure  \ref{fig:lidar_fallback} .
 If the execution checker detects a node failure, a sequence to react to the sensor failure is executed in which the robot slows down to a minimum and tries to restart the sensor.
@@
 
-The timeout period for the execution checker to declare a node failed after receiving no message is set at one second.
-After that period, the goal is to enable the robot to continue progressing towards the goal.
-However, if the system can not restart the sensor in a short time after failure, the robot has to come to a stop as a sensor failure is a severe function loss for the robot.
+The timeout period for the execution checker to declare a node failed after receiving no message is set to one second.
+After a sensor failure is declared, the goal of the sensor fallback depicted in figure \ref{fig:lidar_fallback} is to enable the robot to continue progressing towards the goal.
+However, if the system cannot restart the sensor in a short time after the failure, the robot has to come to a stop as a sensor failure is a severe function loss for the system.
 
@@
 One of the biggest problems mentioned in section \ref{sec:Current State} is the inability to react to collisions.
-The collision can happen due to inadequate sensory coverage for detecting obstacles or general sensor inadequacy, which can not detect obstacles because of their shape or surface material.
+The collision can happen due to inadequate sensory coverage for detecting obstacles or general sensor inadequacy, which cannot detect obstacles because of their shape or surface material.
 Collisions would be improbable with many sensors covering all angles and heights and computer vision capabilities to detect obstacles.
-However, even if the robot could detect all obstacles, it would still be vulnerable to outside perturbations, e.g., other robots driving into it or humans accidentally touching the robot.
+However, even if the robot could detect all obstacles, it would still be vulnerable to outside perturbations (e.g., other robots driving into it or humans accidentally touching the robot).
 
@@
 The child nodes of the collision sequence use the sensor data backup.
-To get out of the collision state, the BT requests the last saved command velocities from the data backup component and modifies them to reverse with minimum speed the same way the robot got into the collision.
-The reversing action takes longer than the two seconds as the commands get published with more extended time intervals between them to adjust the driven distance to be the same as the navigation commands.
-An additional check is implemented before the reversing to ensure the robot only reverses once.
-Otherwise, the robot would drive back and forth as the data backup service would provide the BT with the commands from the last reversing action.
+To get out of the collision state, the BT requests the most recent command velocities from the data backup component and modifies them to reverse with minimum speed out of the collision.
+The reversing action takes longer than the two seconds of saved commands as they get published with extended time intervals between them to adjust the driven distance to be the same as the original navigation commands.
+An additional check is implemented before reversing to ensure the robot only reverses once.
+Otherwise, the robot would drive back and forth as the data backup service would provide the BT with commands from the last reversing action.
 
@@
 
-Regarding the map reset, it has to be determined if a new map is being created with a SLAM method or if the map server provided a previously saved map.
-In the case of the running SLAM, the recent updates to the map must be discarded, and a previous map version before the collision has to be re-utilized.
-When the map server component from Nav2 provides a static map, the map needs to be updated with a new obstacle at the point of collision and republished by the map server.
+Regarding the map reset, it has to be checked if a new map is being created with a SLAM method or if the map server provided a previously saved map.
+In the case of the running SLAM, the recent updates to the map must be discarded and a previous map version before the collision has to be re-utilized.
+If the map server component from Nav2 provides a static map, the map needs to be updated with a new obstacle at the point of collision and republished by the map server.
 
@@
 It is assumed that the collision point was right in front of the robot because the local planner favors forward drive kinematics and probably is driving forwards when a collision occurs.
-The simulated bumper sensor can provide more nuanced information about the point of contact, but this functionality can not be achieved on a real robot with IMU-based collision detection.
-Hence the obstacle position is simplified to always be in front of the robot.
+The simulated bumper sensor can provide more nuanced information about the point of contact, but this functionality can only be achieved with high developmental effort on a real robot with an IMU-based collision detection.
+In order to limit the effort of transfering and running the implemented software on a real robot, the obstacle position is simplified to always be in front of the robot.
 A $25$ by $25$ $cm$ square with $100$ percent occupation at the collision point is added to the current map and then saved to be reloaded by the map server.
+The original goal is then republished to replan the global path on the updated map.
+A more detailed view of the whole behavior is depicted in figure \ref{fig:activity_diagram_collision}.
 
@@
 % add the 
-
-The original goal is then republished to replan the global path on the updated map. A more detailed view of the whole behavior is depicted in figure \ref{fig:activity_diagram_collision}.
-
 \begin{figure*}[ht!]
 	\centering
-	\includegraphics[width=0.9\textwidth]{images/activity_diagram_collision.png}
+	\includegraphics[width=0.90\textwidth]{images/activity_diagram_collision.png}
 	\caption{Activity Diagram for the Collision Behavior}
@@
 With this behavior, the robot should be able to eventually navigate around transparent obstacles that do not reflect the laser.
-Glass doors are known to cause issues for robots based only on lidars.
+Glass doors are known to cause issues for lidar-based robots, because of their reflective properties (see section)
 
@@
 
-The battery behavior has the goal of monitoring the robot's battery and has to decide if the robot should navigate to a new goal location or not.
+The battery behavior aims to monitor the robot's battery and has to decide if the robot should navigate to a new goal location or not.
 The decision is based on the current battery charge and the estimated consumption of energy to reach the new goal.
-The robot should not begin the navigation process if it is probable that the remaining charge will not suffice for the driving.
+The robot should not begin the navigation process if it is probable that the remaining charge will not suffice to arrive at the goal.
 Figure \ref{fig:battery_fallback} depicts the behavior tree fallback.
@@
 
-The robot has to consider the idle consumption and the energy to drive the motors.
+The behavior has to consider the idle consumption and the energy to drive the motors.
 With this information, a linear function for energy expenditure was created to predict battery usage depending on the path length.
@@
 
+For the simulation, a battery package was created, which updates the battery charge every second.
+The battery package subscribes to the command velocity topic and decreases the battery charge for every message received.
+The battery package was created because the behavior needs to be tested and Gazebo does not offer this functionality.
+The consumption rates are not close to reality but are chosen to allow for a rapid testing of the behavior.
 The average values for idle and driving consumption on a real robot could be experimentally determined.
-For the simulated turtlebot, a simulated battery package was created, which updates the battery charge every second.
-The battery package subscribes to the command velocity topic and decreases the battery charge for every message received.
-The battery package was created because the behavior for this scenario needs to be tested, and Gazebo does not offer this function.
-The consumption rates are not close to reality but are chosen to allow the rapid testing of the behavior.
 
@@
 The BT node "IsBatterySufficientCheck" calls the service to compare the current battery to the calculated consumption.
-The battery behavior fallback will cancel the navigation goal if the robot can not reach the given goal with the remaining charge.
+The battery behavior fallback will cancel the navigation goal if the robot cannot reach the given goal with the remaining charge.
 
@@
 The execution checker component monitors the health of the Nav2 global planner.
-Additionally, the behavior tree analyses the global planner's output.
+Additionally, the behavior tree analyzes the global planner's output.
 If the latter condition fails, the given navigation goal is not reachable by the planning algorithm.
-By shifting the goal to a nearby alternative goal, the robot tries to navigate, although the original goal is not reachable.
+By shifting the goal to a nearby alternative goal, the robot tries to navigate to close location, although the original goal cannot be reached.
 
@@
 The goal gets moved on a straight vector towards the robot.
-Other strategies for finding alternative goals were not implemented in this behavior.
+Other strategies for finding alternative goals were not implemented in this behavior due to time-constraints.
 Figure \ref{fig:global_planning_fallback} depicts the behavior tree fallback for the behavior.
@@
 The failure of a sensor can be simulated this way by stopping the execution of the intermediary node.
-Figure \ref{fig:rqt_graph} shows the RQT graph, which depicts the ROS nodes and their subscribed/published topics.
+Figure \ref{fig:rqt_graph} depicts the ROS nodes and their subscribed and published topics.
+Topics are depicted as rectangles and nodes are shown as circles.
 
@@
 	\includegraphics[width=0.7\textwidth]{images/rqt_graph_sensor_drivers.png}
-	\caption{Graph for the Sensor Driver nodes and topics}
+	\caption{Graph for the Sensor Driver Nodes and Topics}
 	\label{fig:rqt_graph}
@@
 
-Because of the necessity to have ROS nodes spinning outside of the BT thread, as discussed in section \ref{subsec:BT System Supervision}, additional interfaces needed to be created in order to receive information through service calls.
+Because of the necessity to have ROS nodes spinning outside of the BT thread, as discussed in section \ref{subsec:BT System Supervision}, additional interfaces are created in order to receive information through service calls.
 The created service types are listed in table \ref{tab:custom_interfaces}.

### 2022-09-28 Thesis/text/methoden.tex
fertig?!

@@
 
-This chapter briefly presents the current state of mobile navigation with ROS and its limitations.
+This chapter presents the current state of mobile navigation with ROS and its limitations.
 This chapter aims to identify the critical areas in which standard mobile robots running ROS lack robustness and autonomy during autonomous mobile navigation.
-According to these critical areas, requirements are derived from them and prioritized with a risk analysis.
-This chapter proposes how behavior planning can improve the current default systems to meet these requirements.
+According to these critical areas, requirements are derived and prioritized with a risk analysis.
+This chapter proposes how behavior planning can improve the current default systems to meet the respective requirements.
 
@@
 The Robot Operating System (ROS) is an open-source middleware for creating robot applications.
-It is not an operating system (OS) since it needs an underlying OS, mainly Linux-based, to work.
+It is not an operating system (OS) since it needs an underlying OS, most commonly Linux, to work.
 ROS can handle the communication between different programs (called nodes) with standardized interfaces and network protocols.
@@
 	\renewcommand{\arraystretch}{1.5}
-	\resizebox{0.9\textwidth}{!}{%
-		\begin{tabular}{ | m{0.12\textwidth} | m{0.37\textwidth}| m{0.3\textwidth} | m{0.139\textwidth}|} 
+	\resizebox{0.95\textwidth}{!}{%
+		\begin{tabular}{ | m{0.09\textwidth} | m{0.37\textwidth}| m{0.45\textwidth} | m{0.139\textwidth}|} 
 	 	\hline
@@
 	 	\hline
-	 	Topics & Publisher / Subscriber & Continuous stream of data e.g. sensor data & n:m \\ 
+	 	Topics & Publisher / Subscriber & Continuous stream of data (e.g. sensor data) & n:m \\ 
 	 	\hline
@@
 
-With ROS2, the framework received fundamental changes and updates to its architecture and design to gain more acceptance and use in the industry.
+With ROS2, the framework received fundamental changes and updates to its architecture and design to gain more acceptance and increased usage in the industry.
 These changes allow for real-time safety, simpler certification, and security when building industrial applications and products using ROS2 \cite{ros2022}.
-
-\subsection{Navigation2}
-
-The ROS Navigation2 Stack (Nav2) combines different packages that allow mobile robots to navigate from point A to point B.
-Navigation2 is the de-facto standard for mobile navigation with a wide range of supported robots.
-Supported robot types are holonomic, differential-drive, legged, and ackerman (car-like).
-For the mobile robot to make use of the Nav2 stack, it has to be set up in a certain way to be able to generate plans and execute commands in the right way.
-Nav2 requires the mobile robot to possess a laser scan or point cloud sensor, an odometry source (such as an IMU or wheel encoders), and a set of transformations for planning and navigation.
-The stack contains tools that save and load occupancy maps, localize the robot, plan paths, execute the path, provide cost maps, build behaviors and execute recoveries \cite{macenski2020}.
-
-\begin{figure}[ht]
-	\centering
-	\includegraphics[width=0.9\textwidth]{images/nav2_architecture.png}
-	\caption{Navigation2 Architecture \cite{macenski2020}}
-	\label{fig:nav_architecture}
-\end{figure}
-
-These packages are often integrating a SLAM method to build or enlarge maps.
-Path planning and execution capabilities, which correspond to the global and local planner described in section \ref{sec:Autonomous Driving Navigation Architectures}, are further supported by ready-to-use planning plugins that use A* and Dijkstra algorithms for path planning and a dynamic window approach for path execution.
-The system sequencing is done in a hierarchical structure as a central behavior tree calls asynchronous actions from the respective planners after another, as seen in figure \ref{fig:nav_architecture}.
-
-\begin{figure}[ht]
-	\includegraphics[width=1.0\textwidth]{images/nav_bt-modified.png}
-	\caption{Navigation2 Behavior Tree}
-	\label{fig:bt_nav}
-\end{figure}
-
-The behavior tree depicted in figure \ref{fig:bt_nav} is the one that is the default Nav2 behavior and has a set of recovery behaviors already implemented (spin, wait, back up, clear map).
-These behaviors are executed if the robot can not make progress towards the set goal.
-Nav2 uses a node lifecycle management system to start, activate, deactivate, and shut down nodes in a controlled way.
-The lifecycle approach allows the system to monitor the system node's startups, executions, and failures.
-
-With this quantity of functions and hierarchical architecture, the ROS Navigation2 stack is an excellent basis for achieving high levels of robot autonomy.
-The software can be modified and expanded with plugins to fit the users' needs.
-When comparing the functionalities of the software stack to the proposed autonomous system architecture in figure 	\ref{fig:autonomous_driving_architecture}, the functionalities that Navigation2 is missing are a system-wide supervision layer and a more deliberative approach to behavior planning, as mentioned in section \ref{sec:Behavior Types}.
-
-The absence of these components leads to a limited set of circumstances in which the robot can perform navigation reliably.
-One of the core problems behavior-based robots face is generating a dependable environment representation.
-An accurate representation is the basis for the last steps in the planning part of the "Sense - Think - Act" cycle.
-Unforeseen events can severely limit the reliability of the environment representation without the robot being aware of it.
-
-Examples of such events currently not handled appropriately are slipping wheels, orientation changes by external influences, or undetected, more specifically, undetectable obstacles.
-This can lead to erratic movement of the robot, as reactive behaviors and more deliberative behaviors compete for authority over the robot's movement.
-By design, to allow more intelligent robot behaviors, the deliberative behavior can override commands from reactive behaviors.
-However, the deliberative behavior's planning is based upon a false representation of reality, leading to unsafe commands.
-
-The unsafe planning triggers reactive-type behaviors again to counteract the commands from the deliberative behavior.
-Even worse would be that the reactive behaviors may not get activated, and the robot continues to drive despite being unaware of the surroundings.
-The unsafe behavior requires a human operator would need to step in and restore the robot to full functionality by hand.
-
-%\textit{Add a closing statement about that suitability of the Nav stack for achieving autonomy}
 
@@
 The Turtlebot3 is a standard mobile research platform with an available ROS interface to control the robot.
-The robot is well integrated into the ROS ecosystem and is established in the literature as a system to develop and integrate new methods.
+The robot is well-integrated into the ROS ecosystem and is established in the literature as a system to develop and integrate new methods.
 The robot has two motors with attached wheel encoders to drive and steer, classifying it as a differential drive robot.
@@
 
-The manufacturer provides an open-source model for simulating the robot in physics-based simulators like Gazebo, as depicted in figure \ref{fig:turtlebot}, which enables faster development and testing for the robot.
+The manufacturer provides an open-source model for simulating the robot in physics-based simulators like Gazebo, as depicted in figure \ref{fig:turtlebot}, which enables faster development and testing of the created robotic applications.
 
@@
 
-The equipped sensors and easy simulation capabilities make the robot well suited for using the Nav2 stack and further development for behavior planning.
+The equipped sensors and easy simulation capabilities make the robot well-suited for using the Navigation2 stack and further development for behavior planning.
 The robot specifications are specified in table \ref{tab:turtlebot_spec}.
+The turtlebot will be used for the analysis of the current state of mobile navigation.
+Furthermore, the simulated turtlebot will be the platform where the behavior planning will be implemented and tested.
 
@@
 			\hline
-	 	Laser Distance Sensor & 360 Degree Laser Distance Sensor LDS-01 \\
+			Laser Distance Sensor (Lidar) & 360 Degree Laser Distance Sensor LDS-01 \\
 			\hline
-	 	IMU & Gyroscope 3 Axis, Accelerometer 3 Axis \\
+			Inertial Measurement Unit (IMU) & Gyroscope 3 Axis, Accelerometer 3 Axis \\
 			\hline
@@
 
-The turtlebot will be used for the analysis of the current state of mobile navigation.
-Furthermore, the simulated turtlebot will be the platform on which the behavior planning will be implemented and tested.
+\subsection{Navigation2}
+
+The ROS Navigation2 Stack (Nav2) combines different packages that allow mobile robots to navigate from point A to point B.
+Navigation2 is the de-facto standard for mobile navigation with a wide range of supported robots.
+Supported robot types are holonomic, differential-drive, legged, and ackerman (car-like).
+For the mobile robot to make use of the Nav2 stack, it has to be set up in a certain way to be able to generate plans and execute commands in the right way.
+The Nav2 setup requires the mobile robot to possess a laser scan or point cloud sensor, an odometry source (such as an Inertial Measurement Unit (IMU) or wheel encoders), a map with information about free and occupied spaces (Occupancy Grid Map (OGM)),and a set of transformations for planning and navigation (see \ref{fig:nav_architecture}).
+
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=0.9\textwidth]{images/nav2_architecture.png}
+	\caption{Navigation2 Architecture \cite{macenski2020}}
+	\label{fig:nav_architecture}
+\end{figure}
+
+The stack contains tools that save and load occupancy maps, localize the robot, plan paths, execute the path, provide cost maps, build behaviors, and execute recoveries \cite{macenski2020}.
+These packages are often integrating a Simultaneous Localization and Mapping (SLAM) method to build or enlarge maps.
+Path planning and execution capabilities, which correspond to the global and local planner described in section \ref{sec:Autonomous Driving Navigation Architectures}, are further supported by ready-to-use planning plugins that use A* and Dijkstra algorithms for global path planning and a dynamic window approach for path execution (local planning).
+The system sequencing is done in a hierarchical structure as a central behavior tree calls asynchronous actions from the respective planners after another, as seen in figure \ref{fig:nav_architecture}.
+
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=0.9\textwidth]{images/nav_bt-modified.png}
+	\caption{Navigation2 Behavior Tree}
+	\label{fig:bt_nav}
+\end{figure}
+
+The behavior tree depicted in figure \ref{fig:bt_nav} is the default Nav2 behavior and has a set of recovery behaviors already implemented (spin, wait, back up, clear map).
+These behaviors are executed if the robot can not make progress towards the set goal.
+Nav2 uses a node lifecycle management system to start, activate, deactivate, and shut down nodes in a controlled way.
+The lifecycle approach allows the system to monitor the system nodes' startups, executions, and failures.
+
+With this quantity of functions and hierarchical architecture, the ROS Navigation2 stack is an excellent basis for achieving high levels of robot autonomy.
+The software can be modified and expanded with plugins to fit the users' needs.
+When comparing the functionalities of the software stack to the proposed autonomous system architecture in figure 	\ref{fig:autonomous_driving_architecture}, the functionalities that Navigation2 is missing are a system-wide supervision layer and a more deliberative approach to behavior planning (as mentioned in section \ref{sec:Behavior Types}).
+
+The absence of these components leads to a limited set of circumstances in which the robot can perform navigation reliably.
+One of the core problems behavior-based robots face is generating a dependable environment representation.
+An accurate representation is the basis for the last steps in the planning part of the \textit{"Sense - Think - Act"} cycle.
+Unforeseen events can severely limit the reliability of the environment representation without the robot detecting the unreliability.
+
+Examples of such events currently not handled appropriately are slipping wheels, orientation changes by external influences, or undetected, more specifically, undetectable obstacles.
+This can lead to erratic movement of the robot, as reactive behaviors and more deliberative behaviors compete for authority over the robot's movement.
+By design, to allow more intelligent robot behaviors, the deliberative behavior can override commands from reactive behaviors.
+However, the deliberative behavior's planning is based upon a false representation of reality, leading to incorrect commands.
+
+The unsafe planning triggers reactive-type behaviors again to counteract the commands from the deliberative behavior.
+In a more severe scenario the reactive behaviors may not get activated at all, and the robot continues to drive despite the incorrect interpretation of the surroundings.
+This highly unsafe behavior requires a human operator to step in and restore the robot manually to full functionality.
+%\textit{Add a closing statement about that suitability of the Nav stack for achieving autonomy}
 
@@
 The software requirements that improve the robot's behavior are derived from the current limitations of the standard ROS and Navigation2 setup for mobile robots described in the previous section.
-The functional requirements and their acceptance criteria are listed in \ref{tab:fn_req}.
-The non-functional requirements are listed seperately in \ref{tab:nonfn_req}.
+The functional requirements and their acceptance criteria are listed in table \ref{tab:fn_req}.
+The non-functional requirements are listed seperately in table \ref{tab:nonfn_req}.
 To better contextualize the context of the systems requirements, figure \ref{fig:use_case} depicts a Use-Case diagram with the needed functions to improve the system's autonomy.
@@
 			\hline
-			non$\_$freq1 & Single Point of Failure & High & The system does not have a single point of failure	When parts of the system fail, the system maintains operability to a degree \\ 
+			non$\_$freq1 & Single Point of Failure & High & The system does not have a single point of failure	When parts of the system fail, the system maintains operability to a degree. \\ 
 			\hline
-			non$\_$freq2 & Performance & High & The system's control loop guarantees fast reactions. The average frequency with which the system operates is higher than 100Hz (10ms) \\ 
+			non$\_$freq2 & Performance & High & The system's control loop guarantees fast reactions. The average frequency with which the system operates is higher than 100Hz (10ms). \\ 
 			\hline
@@
 	 	\hline
-	 	fn$\_$req1 & Sensor Failure & High & The system detects sensor failure. Ensure that the system can restart sensors and decrease the speed during the time the sensor delivers limited information \\ 
+	 	fn$\_$req1 & Sensor Failure & High & The system detects sensor failure.
+Ensure that the system can restart sensors and decrease the speed during the time the sensor delivers limited information.
+\\
 	 	\hline
 	 	fn$\_$req2 & Emergency Detection & High & The system detects emergency.
-Ensure that the system can detect when the continuation on the calculated path is no longer safe (sensor failures, blockage) \\
+Ensure that the system can detect when the continuation on the calculated path is no longer safe (sensor failures, blockage).
+\\
 	 	\hline
-	 	fn$\_$req3 & Emergency Stop & High & The system can initiate emergency stops. Ensure that the system can override all commands and stop in case an emergency is detected \\
+	 	fn$\_$req3 & Emergency Stop & High & The system can initiate emergency stops. Ensure that the system can override all commands and stop in case an emergency is detected. \\
 	 	\hline 	
@@
 	 	\hline
-	 	fn$\_$req6 & Recovery & High & The system can recover from crashes. Ensure that the system can successfully reach goals despite previous crashes\\ 
+	 	fn$\_$req6 & Recovery & High & The system can recover from crashes. Ensure that the system can successfully reach goals despite previous crashes.\\ 
 	 	\hline 
@@
 To eliminate a single point of failure for the system, one needs to create an independent second system outside the existing one.
-This is to create a robust fallback behavior that does not rely on Nav2 to move.
+This aims to create a robust fallback behavior that does not rely on Nav2 to move.
 Also, an external system can more reliably monitor other systems when its execution is detached from the performance and execution of other systems.
 
-A suitable name for the second system is the "autonomy layer", as it aims to increase the robustness and autonomy of the whole system.
+A suitable name for the second system is the "autonomy layer", as its goal is to increase the robustness and autonomy of the whole system.
 The autonomy layer has to provide a comprehensive system supervisor that monitors all components' health (sensors, navigation, robot controller).
@@
 
-Another required addition to the autonomy layer is a data storage of relevant system information.
+Another required addition to the autonomy layer is a data storage of relevant system information and data.
 The stored data includes sensor data and generated maps, cost maps, positions, and speed commands.
@@
 This component could be realized as a decision gate that receives speed commands from Nav2 and the new behavior planning component and can forward, block or modify commands sent to the controller.
-This modification aspect allows high-quality local planning capabilities but enables more deliberative behaviors to happen upon this planning.
+This modification aspect allows high-quality local planning capabilities, but enables more deliberative behaviors to happen upon this planning.
 A simplified system block diagram is depicted in figure \ref{fig:block_diagram}.
@@
 
-A comparison and assessment of the behavior planning approach were made in section \ref{sec:Behavior Planning Approaches}.
-The decision on which approach to take fell on the behavior tree method.
-The reason for this is mainly the increased flexibility when creating and expanding the behavior planning compared to FSMs.
-Also, behavior trees allow complete determinism and introspection which is why POMPDs are not suitable as the behavior planning approach for the specified requirements.
+A comparison and assessment of the behavior planning approaches were made in section \ref{sec:Behavior Planning Approaches}.
+The decision which approach to use was made in favor of the behavior tree method.
+The main reason for this decision is the increased flexibility when creating and expanding the behavior planning compared to FSMs.
+Also, behavior trees allow full determinism and introspection during execution which is why the POMPD approach is not suitable as the main behavior planning approach for the specified requirements.
 
@@
 The behavior tree that is used is an expansion of the Behaviortree.CPP (BT.CPP) library.
-This library offers a fast way to create, execute, monitor, and edit behavior trees.
+This library offers a way to create, execute, monitor, and edit behavior trees.
 
@@
 
-Also, it offers Decorator Nodes which allow more control over the child node and its output.
-These Decorator Nodes can only have one child node each.
+Also, the library includes another class of Control Flow Nodes, called Decorator Nodes, which allow more control over the child node and its output.
+An important distinction to other Control Flow Nodes is that Decorator Nodes can only have one child node as compared to multiple children for sequence and fallback nodes.
 A comprehensive list of the available Decorators and descriptions is presented in table \ref{tab:decorators_bt}.
@@
 	 	\hline 	
-	 	Retry Node & If the child returns success & Ticks child as long as it returns failure. The maximum number of ticks can be defined & If the child returns running. \\
+	 	Retry Node & If the child returns success & Ticks child as long as it returns failure. The maximum number of ticks can be defined & If the child returns running \\
 	 	\hline
@@
 The software allows the creation of behavior trees in XML files, which can be directly loaded into and used by the BT.CPP library.
-Groot can monitor the live execution of behavior trees and allows the introspection of how the tree reacts to different scenarios (see fig.
-\ref{fig:groot}).
+Groot can monitor the live execution of behavior trees and allows the introspection of how the tree reacts to different scenarios.
+Figure \ref{fig:groot} shows a small segment of an active behavior tree.
 
@@
 
-%\textit{Screenshot of live monitoring}
+In this graphical user interface, the condition nodes are labeled with a "C" before the node's name to distinguish them from action nodes which are tagged with an "A".
+The colors indicate the control flow and returned results of the respective nodes.
+A green box around a node shows that the tick signal got processed and returned "Success".
+A red box signals that the node returned "Failure".
+In the depicted example the "CollisionCheck" node returned "Failure" and the fallback node ticked a sequence node called "CollisionSequence".
+The control flow has not returned to the sequence yet, which is why the sequence's border is colored orange.
+The depicted tree is currently executing an action node at the bottom of the figure named "ReverseCmdVel".
 
@@
 Gazebo accurately models the physics of the robot and can simulate and publish the wheel odometry directly on a ROS topic.
-Furthermore, the Gazebo has an integrated ODE physics engine and can simulate the readings for the laser scanner and IMU data via plugins.
-The use of the simulator allows easier repeatability for test cases as various obstacles can be spawned at any time, and sensor failures can be easily induced, resulting in quicker overall development.
+Furthermore, Gazebo has an integrated ODE physics engine and can simulate the readings for the laser scanner and IMU data via plugins.
+The use of the simulator allows easier repeatability for test cases as various obstacles can be spawned at any time, and sensor failures can be induced, resulting in quicker overall development.
 
-Using a simulator allows slowing down or speeding up the environment time.
+Using a simulator facilitates to slow down or speed up the environment time.
 The time control allows more intricate observation of fast reactive type behaviors in slow motion while also enabling the observation of the long-time robustness of the system during extended tests.
-The ROS community offers many different simulation environments to test the robot in.
+The ROS community offers many different simulation environments to test the robot.
 The environments range from simple geometric structures, to office spaces up to complete race tracks.

### 2022-09-28 Thesis/text/diskussion.tex
..

@@
 
-The implemented behaviors are a first step to increasing autonomy, but the behavior planner needs to create many more behaviors and fallbacks.
-Some behaviors not implemented in this thesis incorporate the Intel RSS rules for autonomous driving, like keeping appropriate distances, right-of-way rules, and adaptable speed based on occlusions.
+The implemented behaviors were a first step to increasing autonomy, but the behavior planner needs more behaviors and fallbacks.
+Some behaviors that are not implemented in this thesis incorporate the Intel RSS rules for autonomous driving, like keeping appropriate distances, right-of-way rules, and adaptable speed based on occlusions.
 Another area of future work is the creation of dedicated behaviors for multi-robot systems.
-E.g., sharing sensor data with other robots to overcome complete sensor failures or intelligent path planning to avoid collisions when robots can access the planned paths of other robots.
+For example the sharing sensor data with other robots to overcome complete sensor failures or intelligent path planning to avoid collisions when robots can access the planned paths of other robots.
 Also, more work is needed to define a robot's safe state or position after triggering an emergency stop. Reaching these minimum risk states after emergencies is then the duty of the behavior planner.

### 2022-09-28 Thesis/text/ergebnisse.tex
..

@@
 The missing differentiation led to a test case where the robot reversed into a wall because it was turned 90 degrees by the collision.
-That test failed because a human operator would be needed to correct the robot.
+Therefore, the test failed because a human operator would be needed to correct the robots position.
 This lack of robustness was not apparent in the driving collision scenario because the robot's force was insufficient to cause a significant shift in orientation when colliding with obstacles.
@@
 The architectural design choices meet all the non-functional requirements.
-Implementing the whole behavior planning and system supervision system eliminates a single point of failure for the system.
-Even when the Navigation2 system collapses, is the robot able to move to a degree and, more importantly, stop completely (compare table \ref{tab:nonfn_req}, non$\_$fn$\_$req).
+Implementing the behavior planning and system supervision system eliminates a single point of failure for the system.
+Even when the Navigation2 system collapses, the robot is able to move to a degree and, more importantly, stop completely (compare table \ref{tab:nonfn_req}, non$\_$fn$\_$req).
 The behavior tree and algorithms implemented in the behavior tree are all deterministic (non$\_$fn$\_$req3).

### 2022-09-28 Thesis/text/einleitung.tex
.

@@
 
-This thesis explores possibilities for improving current systems by adding a dedicated behavior planning component.
-Furthermore, the thesis provides an implementation of exemplary behaviors and provide a system architecture to incorporate behavior planning in other robots, too.
+This thesis will explore possibilities for improving current systems by adding a dedicated behavior planning component.
+Furthermore, the thesis seeks to provide an implementation of exemplary behaviors and a system architecture to incorporate behavior planning in other robots, too.
 
-The desired outcomes of this implemented software is an increase in robustness and a decrease in required human actions during several scenarios.
-In these scenarios, failures and problems are artificially induced to test the robot's abilities to behave autonomously.
+The desired outcomes of this implemented software will be an increase in robustness and a decrease in required human actions during several scenarios.
+In these scenarios, failures and problems will be artificially induced to test the robot's abilities to behave autonomously.
 

### 2022-09-29 Thesis/text/einleitung.tex
final v2

@@
 In theory, the robot can perform fully autonomous navigation, including mapping and localization.
-However, when something unexpected happens, the robot's autonomy can not be guaranteed.
+However, the robot's autonomy can not be guaranteed when something unexpected happens.
 A default robot cannot represent all possible fail states and does not detect failures on a systematic level but only inside its navigation-related subsystems.
@@
 This thesis will explore possibilities for improving current systems by adding a dedicated behavior planning component.
-Furthermore, the thesis seeks to provide an implementation of exemplary behaviors and a system architecture to incorporate behavior planning in other robots, too.
+Furthermore, the thesis seeks to provide an implementation of exemplary behaviors and a system architecture to incorporate behavior planning in other robots.
 
@@
 
-	\item Chapter \ref{cha:methoden} analyzes the current robotic systems and derives requirements from the key problem areas
+	\item Chapter \ref{cha:methoden} analyzes the current robotic systems and derives requirements from the critical problem areas
 
-	\item Chapter \ref{cha:implementierung} describes the implemented software approach including the overall system architecture
+	\item Chapter \ref{cha:implementierung} describes the implemented software approach, including the overall system architecture
 

### 2022-09-29 Thesis/text/verwandt.tex
fertig?!
final v2

@@
 The Society of Automotive Engineers (SAE) defines vehicle autonomy based on the human driver's need for supervision and possible intervention when executing complex driving tasks \cite{J3016_202104}.
-Figure \ref{fig:sae_levels} depicts the different levels of autonomy of passenger cars on a scale between no automation to completely autonomous.
-The most significant distinction between the levels is between level two and three, because a system can take the complete control over the vehicle in certain conditions.
-The SAE levels zero to two use driver support functions with no real autonomy, while levels three to five employ automated driving features.
+Figure \ref{fig:sae_levels} on page \pageref{fig:sae_levels} depicts the different levels of autonomy of passenger cars on a scale between no automation to completely autonomous.
+The most significant distinction between the levels is between levels two and three because a system can take complete control over the vehicle in certain conditions.
+The SAE levels zero to two use driver support functions with no autonomy, while levels three to five employ automated driving features.
 Another major level difference is located on the borderline between levels three and four as the vehicle can drive autonomously without a human operator who could control the vehicle as a safety fallback when the system demands it.
-A level four vehicle can theoretically be built without a steering wheel and pedals, because it is able to handle all situations it is deployed in without the need for manual driving.
-%The most significant distinction between the levels is between level three and four, because a vehicle can drive without a human driver as a safety fallback option.
+A level four vehicle can theoretically be built without a steering wheel and pedals because it can handle all situations it is deployed in without requiring manual driving.
+%The most significant distinction between the levels is between levels three and four because a vehicle can drive without a human driver as a safety fallback option.
 
@@
 
-As shown in figure \ref{fig:autonomous_driving_architecture}, the sensory inputs, usually from multiple sensors, are processed to create an environment representation.
+As shown in figure \ref{fig:autonomous_driving_architecture} (p.
+\pageref{fig:autonomous_driving_architecture}), the sensory inputs, usually from multiple sensors, are processed to create an environment representation.
 Based on that representation, the system calculates how to get from its current position to the destination and creates a path to the destination.
-A second planning calculation is triggered to compute the motion commands while taking the system's constraints (e.g., size, turning radius) into account.
+A second planning calculation is triggered to compute the motion commands while considering the system's constraints (e.g., size, turning radius).
 In a final step, these motion commands then get converted to control the motors.
@@
 Considering that with these architectures, vehicles can achieve an SAE level of autonomy up to level four and even five \cite{bacha2008odin}, a good starting point to increase autonomy in a robotic system would be to ensure that all of the functional modules in figure \ref{fig:autonomous_driving_architecture} are implemented and available.
-Based on this, higher autonomy levels are achievable by expanding the behavior planning module inside the planning layer.
+Higher autonomy levels are achievable by expanding the behavior planning module inside the planning layer.
 
@@
 The behavior planning module contains a set of tailored behaviors to operate in a specific environment.
-The module's task is to always choose the best behavior to execute to ensure the best performance for a given task.
+The module's task is always to choose the best behavior to execute to ensure the best performance for a given task.
 In this sense, behavior is defined as a mapping of sensory inputs to a pattern of actions that carry out a specific task \cite{murphy2000}.
 A good behavior planning approach provides the robot with acceptable behaviors for every possible scenario the robot is operating in.
-Different scenarios pose different challenges for the behavior planning module, which has to decide on the best behavioral option to achieve a higher-level goal.
+Different scenarios pose challenges for the behavior planning module, which has to decide on the best behavioral option to achieve a higher-level goal.
 
@@
 In human behavior, this behavior resembles reflexes, like tapping on the kneecap, which unwillingly results in motion in the knee joint.
-This behavior pattern is not following the typical "Sense, Think, Act" loop, but shortcuts directly from sensing to acting \cite{desilva2008}.
+This behavior pattern is not following the typical "Sense, Think, Act" loop but shortcuts directly from sensing to acting \cite{desilva2008}.
 Reactive behaviors can be chained together to achieve more advanced and goal-driven behaviors.
 
-Despite the possibility to carry out more complex robot behaviors, sequential behaviors remain largely on the level of reactive behaviors due to the lack of a planning and decision-making cycle during their execution.
+Despite the possibility of carrying out more complex robot behaviors, sequential behaviors remain primarily on the level of reactive behaviors due to the lack of a planning and decision-making cycle during their execution.
 The computational load of reactive behaviors is low and therefore fast as no decision and planning process is taking place, which makes them suited in scenarios where real-time safety is of concern.
@@
 Instead, deliberative-type behaviors can decide the best course of action before they act.
-Using deliberative behaviors inside of the planning module enables a robot to meet the definition of autonomy, especially in regards to the making of decisions and reacting to unforeseen circumstances.
+Using deliberative behaviors inside the planning module enables a robot to meet the definition of autonomy, especially in making decisions and reacting to unforeseen circumstances.
 
 A behavior planner with a deliberative approach can make decisions based on current sensor data and previously processed data.
-By incorporating older sensor information into the decision-making process, the deliberative behavior planner can predict changes in the environment.
+The deliberative behavior planner can predict environmental changes by incorporating older sensor information into the decision-making process.
 Using older data points allows the planner to proactively adapt the motion planning commands to better achieve goals in dynamic and uncertain environments.
@@
 Concluding, the focus should be on creating deliberative behaviors to improve autonomy.
-Generally, these cognitively more complex behaviors mimic human skills and thus make a system more autonomous, because the need for a human operator can be decreased greatly.
+Generally, these cognitively more complex behaviors mimic human skills and thus make a system more autonomous because the need for a human operator can be significantly decreased.
 
-%\textit{Talk about machine learning as deliberative behavior -> true autonomy bc learning and unknown scenarios, programmers can not cover every scenario, so a well designed/trained ai is needed to make a system really autonomous, but -> missing determinism and unaccountability of machines, hence we need behaviors that are still hard coded and deterministic, these can be highly automated.
-traditional behaviour and ai generated behaviours can and must coexist in the system (adler2019)}
+%(adler2019)
 %
@@
 
-Figure \ref{fig:hybrid_planning} shows the areas where reactive and deliberative behaviors have advantages and limitations.
+Figure \ref{fig:hybrid_planning} (p.
+\pageref{fig:hybrid_planning}) shows the areas where reactive and deliberative behaviors have advantages and limitations.
 Deliberative planning does possess limitations during the execution of generated plans as the time horizon is long-term and not able to react to fast changes in the environment.
-When the system is not addressing these points, deliberative robots are very focused on a narrow problem domain and are not robust when the defined system boundaries are crossed\cite{arkin1998}.
+When the system is not addressing these points, deliberative robots focus on a narrow problem domain and are not robust when the defined system boundaries are crossed\cite{arkin1998}.
 
@@
 The arrows between the states illustrate the possible transition and transition conditions for every state.
-A state transition diagram is an excellent way to understand the system behavior, but can not show the details of the modeled behavior, like the actions \cite{wagner2006}.
+A state transition diagram is an excellent way to understand the system behavior, but it can not show the details of the modeled behavior, like the actions \cite{wagner2006}.
 
@@
 The execution time of finite state machines can be fast enough to control the motors of a bipedal robot to enable the robot to balance and walk despite unexpected height variations between steps \cite{park2013}.
-The high performance makes state machines very well suited for fast, reactive type behaviors while still allowing a deliberative and hierarchical approach to behavior planning.
+The high performance makes state machines very well suited for fast, reactive type behaviors while allowing a deliberative and hierarchical approach to behavior planning.
 
-Creating small state machines can be accomplished easily and quickly using various feature-rich and performant libraries \cite{fourakis2014}.
-Finite state machines are often used to model reactive and sequential behaviors comprehensively in autonomous driving functions \cite{ziegler2014}.
+Creating small state machines can be accomplished easily and quickly using various feature-rich, and performant libraries \cite{fourakis2014}.
+Finite state machines are often used to comprehensively model reactive and sequential behaviors in autonomous driving functions \cite{ziegler2014}.
 These finite state machines use nested state machines inside another state machine.
@@
 
-Despite the possibility of nesting state machines, the effort to expand and maintain a FSM multiplies rapidly, when the system becomes increasingly larger to integrate more complex behaviors.
+Despite the possibility of nesting state machines, the effort to expand and maintain a FSM multiplies rapidly when the system becomes increasingly larger to integrate more complex behaviors.
 The higher effort is due to the required creation of the transitions, transition conditions, and transition events.
-The FSM needs to be updated with each new state that gets introduced into the state machine as existing states need to be checked and corrected accordingly \cite{conner2017}.
+The FSM needs to be updated with each new state introduced into the state machine as existing states need to be checked and corrected accordingly \cite{conner2017}.
 %
@@
 	\renewcommand{\arraystretch}{1.5}
-	\resizebox{0.9\textwidth}{!}{%
+	\resizebox{0.99\textwidth}{!}{%
 		\begin{tabular}{ | m{0.13\textwidth} | m{0.13\textwidth}| m{0.2\textwidth} | m{0.2\textwidth} | m{0.2\textwidth} |} 
@@
 It will only return "Failure" if all children nodes returned "Failure".
-The primary node types and their control flow modification are listed in table \ref{tab:node_types}.
-Parallel control nodes are listed for completeness, but are not discussed further in this chapter due to their limited use in robotic applications.
+The primary node types and their control flow modification are listed in table \ref{tab:node_types} (p.
+\pageref{tab:node_types}).
+Parallel control nodes are listed for completeness but are not discussed further in this chapter due to their limited use in robotic applications.
 
@@
 	\centering
-	\includegraphics[width=0.9\textwidth]{images/bt_example.jpg} 
+	\includegraphics[width=0.99\textwidth]{images/bt_example.jpg} 
 	\caption{Example of a behavior tree \cite{iovino2022}}
@@
 	\centering
-	\includegraphics[width=0.9\textwidth]{images/mdp.png} 
-	\caption{Example of a fully observable Markov Decision Process \cite{andrew1999reinforcement}}
+	\includegraphics[width=0.99\textwidth]{images/mdp.png} 
+	\caption{Example of a fully observable Markov Decision Process \cite{andrew1999reinforcement, banerjee2021real}}
 	\label{fig:mdp}
@@
 
-The difference between an MDP and a POMDP is that in a POMDP the system's current state can be unknown, and the system can make observations and estimate the state to a degree.
+The difference between an MDP and a POMDP is that in a POMDP, the system's current state can be unknown, and the system can make observations and estimate the state to a degree.
 The POMDP is defined as a tuple $<S, D, A, T, R, Z, O>$ in which the MDP tuple gets extended by \textit{Z} and \textit{O}.
@@
 
-POMDPs are well suited for applications in unknown environments because the system can make decisions if the perception information is imperfect, or the environment is not fully observable.
+POMDPs are well suited for applications in unknown environments because the system can make decisions if the perception information is imperfect or the environment is not fully observable.
 The decision process is probabilistic and is more adaptable when confronted with unknown scenarios.

### 2022-09-29 Thesis/text/diskussion.tex
final v2

@@
 
+Firstly, the results must be reproduced on an actual mobile robot.
+The layout of the behavior tree will stay the same, but further effort is needed to adapt the action nodes of the tree for a real system.
+
 The implemented behaviors were a first step to increasing autonomy, but the behavior planner needs more behaviors and fallbacks.
-Some behaviors that are not implemented in this thesis incorporate the Intel RSS rules for autonomous driving, like keeping appropriate distances, right-of-way rules, and adaptable speed based on occlusions.
+Some crucial behaviors not implemented in this thesis incorporate the Intel RSS rules for autonomous driving, like keeping appropriate distances, right-of-way rules, and adaptable speed based on occlusions.
+Further, the implemented collision detection needs to be refined to allow collision point detection to improve the whole recovery behavior's robustness.
+Also, the strategy for finding nearby alternative goals in the path planning behavior needs to be modified or changed to sample more points in all directions around the given goal.
+
 Another area of future work is the creation of dedicated behaviors for multi-robot systems.
-For example the sharing sensor data with other robots to overcome complete sensor failures or intelligent path planning to avoid collisions when robots can access the planned paths of other robots.
+For example, robots could share sensor data to overcome complete sensor failures or intelligent path planning to avoid collisions when robots can access the planned paths of other robots.
 Also, more work is needed to define a robot's safe state or position after triggering an emergency stop. Reaching these minimum risk states after emergencies is then the duty of the behavior planner.
 
-The results must be reproduced on an actual mobile robot. The layout of the behavior tree will stay the same, but further effort is needed to adapt the action nodes of the tree for a real system.
-
-The behaviors must be more deliberative and robust to achieve more independence from human operators.
-One key question in autonomous driving is the responsibility of a vehicle in a crash because, with higher levels of automation, the human is not needed anymore.
-However, a machine can not be responsible for its action because it was programmed to do so.
-Nevertheless, intelligent, robust, and independent decision-making almost always involves machine learning which makes decisions based on training with dedicated datasets and not explicitly written lines of code.
-Because a system in which the outcome for a high-risk scenario is unclear, machine learning behaviors may not be suited to be used independently.
-Therefore, an important work area is incorporating machine learning algorithms into a deterministic system, like a behavior tree.
-A behavior tree could execute and monitor the behaviors to safeguard the indeterministic behaviors and override them if necessary.
+Intelligent, robust, and independent decision-making almost always involves machine learning (ML) which makes decisions based on training data and not written lines of code.
+This property makes ML nondeterministic and, therefore, unsuitable for many high-risk and safety-relevant applications.
+Therefore, an important area of future work is how ML algorithms can be incorporated into a deterministic system like a behavior tree.
+A behavior tree could execute and monitor the ML behaviors to safeguard and override them if necessary with explicitly programmed but still highly automated behaviors.
 %Further research is needed on how behavior trees can manage a system with autonomous machine learning behaviors inside the tree.

### 2022-09-29 Thesis/text/ergebnisse.tex
final v2

@@
 
-The robot is put into specifically designed scenarios to trigger behavior responses from the behavior tree component.
+The robot is put into specific scenarios to trigger behavior responses from the behavior tree component.
 The scenario testing is done to evaluate the efficacy of the implemented behaviors in relation to the functional requirements listed in section \ref{sec:Requirements}.
@@
 	\centering
-	\begin{subfigure}[b]{0.45\textwidth}
+	\begin{subfigure}[b]{0.495\textwidth}
 		\centering
@@
 	\hfill
-	\begin{subfigure}[b]{0.45\textwidth}
+	\begin{subfigure}[b]{0.495\textwidth}
 		\centering
@@
 	\resizebox{0.9\textwidth}{!}{%
-		\begin{tabular}{| m{0.1\textwidth} | m{0.19\textwidth}| m{0.4\textwidth} | m{0.3\textwidth}|} 
+		\begin{tabular}{| m{0.12\textwidth} | m{0.19\textwidth}| m{0.4\textwidth} | m{0.35\textwidth}|} 
 			\hline
@@
 
-The scenarios execution is done as follows:
+The scenarios were conducted in different ways described in the following paragraphs.
 
-For the Lidar, IMU, and Odometry scenarios, the sensor driver was shut down when reaching the desired speed listed in the scenario.
+For the lidar, IMU, and odometry scenarios, the sensor driver was shut down when reaching the desired speed listed in the scenario.
+The location and goal are given as described in the previous chapter.
+Each sensor was tested multiple times in three different driving situations.
 
-Scenario NoPathFound1 is a scenario in which the lifecycle state gets transitioned into "inactive" and can not execute the planning action anymore when triggered.
+NoPathFound$\_$1 is a scenario in which the lifecycle state of the global planner gets transitioned into "inactive" and cannot execute the planning action anymore when triggered.
 The BT has to recover from the inability to plan in this scenario.
-
-NoPathFound$\_$2 is simulating the scenario when the planner is active but fails to find a collision-free path to the desired goal.
+NoPathFound$\_$2 simulates that the planner is active but fails to find a collision-free path to the desired goal.
 This scenario tests the ability to find alternative goals and navigate to a nearby goal.
@@
 In the Collision$\_$1 scenario, an obstacle gets moved so close to the robot that a collision is registered.
-The forced collision tests the ability to escape a collision and resume navigation afterward.
+The forced collision tests the ability to maneuver out of a collision and resume navigation afterward.
 A new navigation goal is published after the behavior tree exits the Collision Fallback Routine.
-
-Collision$\_$2 scenario is testing the same behavior but during active navigation.
+Collision$\_$2 tests the same behavior but during active navigation.
 A flat obstacle is spawned in the path of the robot.
 The obstacle is too low to be detected by the laserscanner, and a collision occurs.
-The robot must be able to recover from the crash and keep on navigating towards the goal with an updated global path in this scenario.
+The robot must be able to recover from the crash and keep on navigating towards the goal with an updated global path.
 
@@
 The simulated battery is emptied to a low level, and a new navigation goal is set.
-If the battery charge runs below zero percent during the navigation, the test fails.
+The test fails if the battery charge runs below zero percent during the navigation.
 
@@
 The system supervision component is very reliable and enables the system to recover from an unexpected sensor failure.
-The more advanced behaviors implemented can improve the autonomous handling of the scenarios.
+The more advanced behaviors can improve the autonomous handling of the scenarios.
 The Battery Scenario was successful every time it was induced.
 However, the path planning behavior for finding alternative goals was unsuccessful about 50 percent of the time.
-The test showed that alternative methods for finding new goals would probably have resulted in better performance.
-The collision behavior during driving worked well, and the map updates are an essential feature for the success of the scenarios.
 
-Nevertheless, there is a lack of intelligence and autonomy when the robot stands still.
-A collision is forced, but the implemented behavior relies on the same mechanism as during the driving scenario.
+The collision behavior during driving worked well, and the map updates are an essential feature for the success of the test scenarios.
+Nevertheless, there is a lack of intelligence and autonomy when the robot stands still, and a collision is forced.
+The implemented behavior relies on the same mechanism as during the driving scenario.
 The missing differentiation led to a test case where the robot reversed into a wall because it was turned 90 degrees by the collision.
-Therefore, the test failed because a human operator would be needed to correct the robots position.
+Therefore, the test failed because a human operator would be needed to correct the robot's position.
 This lack of robustness was not apparent in the driving collision scenario because the robot's force was insufficient to cause a significant shift in orientation when colliding with obstacles.
@@
 	\resizebox{0.9\textwidth}{!}{%
-		\begin{tabular}{| m{0.11\textwidth} | m{0.1\textwidth}| m{0.35\textwidth} | m{0.33\textwidth}|} 
+		\begin{tabular}{| m{0.11\textwidth} | m{0.11\textwidth}| m{0.235\textwidth} | m{0.19\textwidth}|} 
 			\hline
-	 	\textbf{Name} & \textbf{Number of Runs} & \textbf{Percentage Succesful Without Behavior Planning} & \textbf{Percentage Succesful With Behavior Planning}\\ 
+			\multirow{2}{*}{\textbf{Name}} & \multirow{2}{0.11\textwidth}{\textbf{Number of Runs}} & \multicolumn{2}{|c|}{\textbf{Percentage Successful}} \\
+			\cline{3-4}
+			& & \textbf{Without Behavior Planning} & \textbf{With Behavior Planning} \\
 			\hline
@@
 			\hline
-	 	Motor$\_$1 & 10 & 0 & 0 \\ 
-	 	\hline
 		\end{tabular}
@@
 
+The results show that the functional requirements (as defined in \ref{sec:Requirements}) are met in almost all cases.
+Fn$\_$req1 to fn$\_$req5 are all completely fulfilled.
+Fn$\_$req6 is only partly fulfilled as the test showed potential shortcomings of the implemented collision behavior.
+Fn$\_$req7 satisfies the acceptance criteria, but fn$\_$req8 is only partly met by the implemented behavior to find alternative goals.
+Finally, the battery behavior meets all acceptance criteria, and fn$\_$req9 is fulfilled.
+
 The architectural design choices meet all the non-functional requirements.
 Implementing the behavior planning and system supervision system eliminates a single point of failure for the system.
-Even when the Navigation2 system collapses, the robot is able to move to a degree and, more importantly, stop completely (compare table \ref{tab:nonfn_req}, non$\_$fn$\_$req).
+Even when the Navigation2 system collapses, the robot can move to a degree and, more importantly, stop completely (compare table \ref{tab:nonfn_req}, non$\_$fn$\_$req).
 The behavior tree and algorithms implemented in the behavior tree are all deterministic (non$\_$fn$\_$req3).

### 2022-09-29 Thesis/text/implementierung.tex
final v2

@@
 The behavior tree's commands have a higher priority and disable the forwarding of the Navigation2 commands.
-%The behavior tree can access the sensor data when needed through the Data Backup component.
-The System Supervision component controls the health of all components, but the connections are not depicted in the component diagram for reasons of clarity and comprehensibility.
+%The behavior tree can access the sensor data through the Data Backup component.
+The System Supervision component controls the health of all components, but the connections are not depicted in the component diagram for clarity and comprehensibility.
 
@@
 
-Already existing components are colored in gray, while the components that were implemented in this thesis (i.e., Data Backup, Decision Gate, Behavior Tree and System Supervisor) are depicted in blue.
+Already existing components are colored in gray, while the components implemented in this thesis (i.e., Data Backup, Decision Gate, Behavior Tree, and System Supervisor) are depicted in blue.
 The new components and their functions are described in the following sections of this chapter.
-The System Supervisor is explained in the BT section as a significant part of the component's functionality is realized inside of the behavior tree.
+The System Supervisor is explained in the BT section as a significant part of the component's functionality is realized inside the behavior tree.
 
@@
 
-The data backup component functions as a way to efficiently gather data from all system components and provide necessary data to the behavior tree when the planning process requires past data points.
+The data backup component functions as a method to efficiently gather data from all system components and provide it to the behavior tree when the planning process requires past data points.
 This component subscribes to all sensor data and saves the messages in a queue data structure.
@@
 Otherwise, the node might miss incoming messages from subscribed topics.
-Having this action inside of the behavior tree thread would completely block the execution of the entire autonomy system.
-If one tries to spin a node only once when a node gets ticked in the BT, losing entire messages is relevant concern because the BT might need longer than expected to return to a node to tick it.
+This action inside the behavior tree thread would completely block the execution of the entire autonomy system.
+If one tries to spin a node only once when a node gets ticked in the BT, losing entire messages is a relevant concern because the BT might need longer than expected to return to a node to tick it.
 That way, the BT node responsible for saving all the relevant data cannot guarantee that it has received all the messages between spins as multiple messages might have been received between the two ticks.
-The synchronous nature of the ROS executor is why the data backup is not stored within the BT, but in in an external ROS package to seperate the threads.
+The synchronous nature of the ROS executor is why the data backup is not stored within the BT, but in an external ROS package to separate the threads.
 
@@
 
-A simplified behavior tree only depicting the top-level behavior control fallback nodes is shown in figure \ref{fig:top_level_bt}.
+A simplified behavior tree only depicting the top-level behavior control fallback nodes is shown in figure \ref{fig:top_level_bt} on page \pageref{fig:top_level_bt}.
 The tree receives the initial tick signal from the root in an infinite loop.
@@
 
-The system supervisor is using a "Watchdog" design pattern depicted in figure \ref{fig:watchdog_pattern}.
+The system supervisor uses a "Watchdog" design pattern depicted in figure \ref{fig:watchdog_pattern}.
 A watchdog is used to monitor a system from the outside and take corrective measures when problems are detected.
@@
 The synchronous executor would stop the execution of the whole behavior tree, which is why the component needs to be moved inside of a new thread.
-The system supervisor, also called the execution checker, is receiving messages from nodes that are not implemented as ROS2 lifecycle nodes.
+The system supervisor, also called the execution checker, receives messages from nodes not implemented as ROS2 lifecycle nodes.
 If a node stops sending messages or responding completely, the external execution checker can inform the BT via a provided service of a problem with the node.
@@
 The current node state can be requested via a service call.
-However, the node's output needs to be checked, too, because it might be that the FSM inside of the node does not pick up on an error and fails to change the internal lifecycle state.
+However, the node's output needs to be checked, too, because it might be that the FSM inside the node does not pick up on an error and fails to change the internal lifecycle state.
 
@@
 The components get checked by the BT inside of a fallback node.
-If the condition for the execution is met, the condition node returns "Success", and the fallback would exit to move on to the next fallback and execution condition check as in the lidar fallback depicted in figure  \ref{fig:lidar_fallback} .
+If the condition for the execution is met, the condition node returns "Success", and the fallback would exit to move on to the next fallback and execution condition check as in the lidar fallback depicted in figure  \ref{fig:lidar_fallback}.
 If the execution checker detects a node failure, a sequence to react to the sensor failure is executed in which the robot slows down to a minimum and tries to restart the sensor.
@@
 In order to combat this weakness, the robot must be able to detect that a collision occurred, get out of the collision state, update the map and reset all costmaps to generate a new path to the original goal.
+The described behavior is depicted in figure \ref{fig:collision_fallback}.
 
@@
 
-The activity of the "MapUpdateAfterCollision" action node is depicted in figure \ref{fig:activity_diagram_collision}.
+The activity of the "MapUpdateAfterCollision" action node is depicted in figure \ref{fig:activity_diagram_collision} on page \pageref{fig:activity_diagram_collision}.
 The collision point is estimated by the robot's pose, which refers to the robot base link and the robot footprint.
 It is assumed that the collision point was right in front of the robot because the local planner favors forward drive kinematics and probably is driving forwards when a collision occurs.
-The simulated bumper sensor can provide more nuanced information about the point of contact, but this functionality can only be achieved with high developmental effort on a real robot with an IMU-based collision detection.
-In order to limit the effort of transfering and running the implemented software on a real robot, the obstacle position is simplified to always be in front of the robot.
+The simulated bumper sensor can provide more nuanced information about the point of contact.
+However, this functionality can only be achieved with high developmental effort on a real robot with IMU-based collision detection.
+In order to limit the effort of transferring and running the implemented software on a real robot, the obstacle position is simplified to always be directly in front of the robot.
 A $25$ by $25$ $cm$ square with $100$ percent occupation at the collision point is added to the current map and then saved to be reloaded by the map server.
-The original goal is then republished to replan the global path on the updated map.
-A more detailed view of the whole behavior is depicted in figure \ref{fig:activity_diagram_collision}.
+The original goal is then republished to replan the global path on the updated map..
 
@@
 %\textit{Formula here}
-
 %as depicted in figure xx. 
@@
 With this behavior, the robot should be able to eventually navigate around transparent obstacles that do not reflect the laser.
-Glass doors are known to cause issues for lidar-based robots, because of their reflective properties (see section)
+Glass doors are known to cause issues for lidar-based robots because of their reflective properties (see section \ref{subsec:nav2}).
 
@@
 The robot should not begin the navigation process if it is probable that the remaining charge will not suffice to arrive at the goal.
-Figure \ref{fig:battery_fallback} depicts the behavior tree fallback.
+Figure \ref{fig:battery_fallback} on page \pageref{fig:battery_fallback} depicts the battery fallback.
 
@@
 
-For the simulation, a battery package was created, which updates the battery charge every second.
+A battery package was created for the simulation, which updates the battery charge every second.
 The battery package subscribes to the command velocity topic and decreases the battery charge for every message received.
-The battery package was created because the behavior needs to be tested and Gazebo does not offer this functionality.
-The consumption rates are not close to reality but are chosen to allow for a rapid testing of the behavior.
+The battery package was created because the behavior needs to be tested, but Gazebo does not offer this functionality.
+The consumption rates are not close to reality but are chosen to allow for rapid testing of the behavior.
 The average values for idle and driving consumption on a real robot could be experimentally determined.
@@
 If the latter condition fails, the given navigation goal is not reachable by the planning algorithm.
-By shifting the goal to a nearby alternative goal, the robot tries to navigate to close location, although the original goal cannot be reached.
+By shifting the goal to a nearby alternative goal, the robot tries to navigate to a close location, although the original goal cannot be reached.
 
@@
 The goal gets moved on a straight vector towards the robot.
-Other strategies for finding alternative goals were not implemented in this behavior due to time-constraints.
+This behavior did not implement other strategies for finding alternative goals due to time constraints of the thesis.
 Figure \ref{fig:global_planning_fallback} depicts the behavior tree fallback for the behavior.
@@
 The failure of a sensor can be simulated this way by stopping the execution of the intermediary node.
-Figure \ref{fig:rqt_graph} depicts the ROS nodes and their subscribed and published topics.
-Topics are depicted as rectangles and nodes are shown as circles.
-
-\begin{figure}[ht]
-	\centering
-	\includegraphics[width=0.7\textwidth]{images/rqt_graph_sensor_drivers.png}
-	\caption{Graph for the Sensor Driver Nodes and Topics}
-	\label{fig:rqt_graph}
-\end{figure}
+%Figure \ref{fig:rqt_graph} depicts the ROS sensor nodes and their subscribed and published topics.
+Topics are illustrated as rectangles and nodes are shown as circles.
+The "/clock" and "/rosout" topic are unrelated to the described functionality and are required to exist to work with the simulation and ROS.
+%
+%\begin{figure}[ht]
+%	\centering
+%	\includegraphics[width=0.7\textwidth]{images/rqt_graph_sensor_drivers.png}
+%	\caption{Graph for the Sensor Driver Nodes and Topics}
+%	\label{fig:rqt_graph}
+%\end{figure}
 
@@
 
-Because of the necessity to have ROS nodes spinning outside of the BT thread, as discussed in section \ref{subsec:BT System Supervision}, additional interfaces are created in order to receive information through service calls.
+Because of the necessity to have ROS nodes spinning outside of the BT thread, as discussed in section \ref{subsec:BT System Supervision}, additional interfaces are created to receive information through service calls.
 The created service types are listed in table \ref{tab:custom_interfaces}.

### 2022-09-29 Thesis/text/methoden.tex
final v2

@@
 The equipped sensors and easy simulation capabilities make the robot well-suited for using the Navigation2 stack and further development for behavior planning.
-The robot specifications are specified in table \ref{tab:turtlebot_spec}.
-The turtlebot will be used for the analysis of the current state of mobile navigation.
+The robot specifications are specified in table \ref{tab:turtlebot_spec} (p.\pageref{tab:turtlebot_spec}).
+The turtlebot will be used to analyze the current state of mobile navigation.
 Furthermore, the simulated turtlebot will be the platform where the behavior planning will be implemented and tested.
@@
 \subsection{Navigation2}
+\label{subsec:nav2}
 
@@
 For the mobile robot to make use of the Nav2 stack, it has to be set up in a certain way to be able to generate plans and execute commands in the right way.
-The Nav2 setup requires the mobile robot to possess a laser scan or point cloud sensor, an odometry source (such as an Inertial Measurement Unit (IMU) or wheel encoders), a map with information about free and occupied spaces (Occupancy Grid Map (OGM)),and a set of transformations for planning and navigation (see \ref{fig:nav_architecture}).
+The Nav2 setup requires the mobile robot to possess a laser scan or point cloud sensor, an odometry source (such as an Inertial Measurement Unit (IMU) or wheel encoders), a map with information about free and occupied spaces (Occupancy Grid Map (OGM)), and a set of transformations for planning and navigation (see figure \ref{fig:nav_architecture}).
 
@@
 Path planning and execution capabilities, which correspond to the global and local planner described in section \ref{sec:Autonomous Driving Navigation Architectures}, are further supported by ready-to-use planning plugins that use A* and Dijkstra algorithms for global path planning and a dynamic window approach for path execution (local planning).
-The system sequencing is done in a hierarchical structure as a central behavior tree calls asynchronous actions from the respective planners after another, as seen in figure \ref{fig:nav_architecture}.
+The system sequencing is done in a hierarchical structure as a central behavior tree calls asynchronous actions from the respective planners after another, as seen in figure \ref{fig:nav_architecture} on page \pageref{fig:nav_architecture}.
 
@@
 The behavior tree depicted in figure \ref{fig:bt_nav} is the default Nav2 behavior and has a set of recovery behaviors already implemented (spin, wait, back up, clear map).
-These behaviors are executed if the robot can not make progress towards the set goal.
+These behaviors are executed if the robot can not progress towards the set goal.
 Nav2 uses a node lifecycle management system to start, activate, deactivate, and shut down nodes in a controlled way.
@@
 Examples of such events currently not handled appropriately are slipping wheels, orientation changes by external influences, or undetected, more specifically, undetectable obstacles.
-This can lead to erratic movement of the robot, as reactive behaviors and more deliberative behaviors compete for authority over the robot's movement.
+These events can lead to erratic movement of the robot, as reactive behaviors and more deliberative behaviors compete for authority over the robot's movement.
 By design, to allow more intelligent robot behaviors, the deliberative behavior can override commands from reactive behaviors.
@@
 The unsafe planning triggers reactive-type behaviors again to counteract the commands from the deliberative behavior.
-In a more severe scenario the reactive behaviors may not get activated at all, and the robot continues to drive despite the incorrect interpretation of the surroundings.
+In a more severe scenario, the reactive behaviors may not get activated, and the robot continues to drive despite the incorrect interpretation of the surroundings.
 This highly unsafe behavior requires a human operator to step in and restore the robot manually to full functionality.
@@
 The software requirements that improve the robot's behavior are derived from the current limitations of the standard ROS and Navigation2 setup for mobile robots described in the previous section.
-The functional requirements and their acceptance criteria are listed in table \ref{tab:fn_req}.
-The non-functional requirements are listed seperately in table \ref{tab:nonfn_req}.
-To better contextualize the context of the systems requirements, figure \ref{fig:use_case} depicts a Use-Case diagram with the needed functions to improve the system's autonomy.
+The non-functional requirements are listed separately in table \ref{tab:nonfn_req}.
+The functional requirements and their acceptance criteria are listed in table \ref{tab:fn_req} on page \pageref{tab:fn_req}.
+Figure \ref{fig:use_case} on page \pageref{fig:use_case} depicts a Use-Case diagram with the functions needed to improve the system's autonomy and better contextualize the system's requirements.
 
@@
 	\renewcommand{\arraystretch}{1.5}
-	\resizebox{0.9\textwidth}{!}{%
-		\begin{tabular}{| m{0.08\textwidth} | m{0.13\textwidth}| m{0.09\textwidth} | m{0.60\textwidth}|} 
+	\resizebox{0.95\textwidth}{!}{%
+		\begin{tabular}{| m{0.08\textwidth} | m{0.15\textwidth}| m{0.09\textwidth} | m{0.60\textwidth}|} 
 			\hline
@@
 
+The prioritization is based on the severity of the consequences for the system.
+The functional requirements one to six focus on safety and robustness of the robot and are therefore prioritized high.
+Functional requirements seven to nine focus on extending the autonomy of the robot, but a failure in these requirements do not pose as high of a risk the other requirements.
+
 \begin{figure}[ht]
@@
 This component could be realized as a decision gate that receives speed commands from Nav2 and the new behavior planning component and can forward, block or modify commands sent to the controller.
-This modification aspect allows high-quality local planning capabilities, but enables more deliberative behaviors to happen upon this planning.
+This modification aspect allows high-quality local planning capabilities but enables more deliberative behaviors to happen upon this planning.
 A simplified system block diagram is depicted in figure \ref{fig:block_diagram}.
-The proposed additional components, discussed in this chapter are colored in blue.
+The proposed additional components discussed in this chapter are colored in blue.
 Existing components, discussed in section \ref{sec:Current State}, are colored in gray.
@@
 A comparison and assessment of the behavior planning approaches were made in section \ref{sec:Behavior Planning Approaches}.
-The decision which approach to use was made in favor of the behavior tree method.
+The decision on which approach to use was made in favor of the behavior tree method.
 The main reason for this decision is the increased flexibility when creating and expanding the behavior planning compared to FSMs.
-Also, behavior trees allow full determinism and introspection during execution which is why the POMPD approach is not suitable as the main behavior planning approach for the specified requirements.
+Also, behavior trees allow complete determinism and introspection during execution which is why the POMPD approach is unsuitable as the primary behavior planning approach for the specified requirements.
 
@@
 
-Additionally to the types of control nodes that are mentioned in section \ref{subsec:Behavior Trees}, the library adds the concept of reactivity into the catalog of control nodes.
+Additionally to the types of control nodes mentioned in section \ref{subsec:Behavior Trees}, the library adds the concept of reactivity into the catalog of control nodes.
 Reactive sequences and reactive fallbacks differ from normal ones in handling nodes that return the state "Running".
@@
 	\renewcommand{\arraystretch}{1.5}
-	\resizebox{0.9\textwidth}{!}{%
+	\resizebox{0.8\textwidth}{!}{%
 		\begin{tabular}{ | l | l | l |} 
@@
 Also, the library includes another class of Control Flow Nodes, called Decorator Nodes, which allow more control over the child node and its output.
-An important distinction to other Control Flow Nodes is that Decorator Nodes can only have one child node as compared to multiple children for sequence and fallback nodes.
+An essential distinction to other Control Flow Nodes is that Decorator Nodes can only have one child node compared to multiple children for sequence and fallback nodes.
 A comprehensive list of the available Decorators and descriptions is presented in table \ref{tab:decorators_bt}.
@@
 	\renewcommand{\arraystretch}{1.5}
-	\resizebox{0.9\textwidth}{!}{%
-		\begin{tabular}{ | m{0.2\textwidth} | m{0.25\textwidth}| m{0.25\textwidth} | m{0.2\textwidth} |} 
+	\resizebox{0.95\textwidth}{!}{%
+		\begin{tabular}{ | m{0.15\textwidth} | m{0.4\textwidth}| m{0.4\textwidth} | m{0.2\textwidth} |} 
 			\hline
@@
 			\hline
-	 	Repeat Node & Ticks child as long as it returns success. The maximum number of repeats can be defined & If the child returns failure & If the child returns running \\
+			Repeat Node & Ticks child as long as it returns success. Number of repeats can be defined & If the child returns failure & If the child returns running \\
 			\hline 	
-	 	Retry Node & If the child returns success & Ticks child as long as it returns failure. The maximum number of ticks can be defined & If the child returns running \\
+			Retry Node & If the child returns success & Ticks child as long as it returns failure. Number of ticks can be defined & If the child returns running \\
 			\hline
@@
 To allow the tree nodes to communicate, the library provides the developer with two possibilities.
-Either node can use the blackboard, a dictionary (key/value) that all nodes of the tree can read and write.
+Either node can use the blackboard, a dictionary (key/value) that all tree nodes can read and write.
 Alternatively, two nodes can be connected through ports which allows direct communication between two nodes via a key/value.
@@
 
-Groot is a software to create, edit, monitor, and debug behavior trees with a graphical user interface.
+Groot is a program to create, edit, monitor, and debug behavior trees with a graphical user interface.
 The software allows the creation of behavior trees in XML files, which can be directly loaded into and used by the BT.CPP library.
@@
 The colors indicate the control flow and returned results of the respective nodes.
-A green box around a node shows that the tick signal got processed and returned "Success".
-A red box signals that the node returned "Failure".
-In the depicted example the "CollisionCheck" node returned "Failure" and the fallback node ticked a sequence node called "CollisionSequence".
+A green box around a node shows that the tick signal got processed and returned "Success", and a red box signals that the node returned "Failure".
+In the depicted example, the "CollisionCheck" node returned "Failure," and the fallback node ticked a sequence node called "CollisionSequence".
 The control flow has not returned to the sequence yet, which is why the sequence's border is colored orange.
-The depicted tree is currently executing an action node at the bottom of the figure named "ReverseCmdVel".
+The depicted tree is executing an action node at the bottom of the figure named "ReverseCmdVel".
 
@@
 Furthermore, Gazebo has an integrated ODE physics engine and can simulate the readings for the laser scanner and IMU data via plugins.
-The use of the simulator allows easier repeatability for test cases as various obstacles can be spawned at any time, and sensor failures can be induced, resulting in quicker overall development.
+The simulator allows easier repeatability for test cases as various obstacles can be spawned at any time, and sensor failures can be induced, resulting in quicker overall development.
 
-Using a simulator facilitates to slow down or speed up the environment time.
+Using a simulator facilitates slowing down or speeding up the environment time.
 The time control allows more intricate observation of fast reactive type behaviors in slow motion while also enabling the observation of the long-time robustness of the system during extended tests.
 The ROS community offers many different simulation environments to test the robot.
-The environments range from simple geometric structures, to office spaces up to complete race tracks.
+The environments range from simple geometric structures to office spaces up to complete race tracks.
 
@@
 
-The software versions used to implement and test the system on are listed below in table \ref{tab:versions}.
+The software versions used to implement and test the system are listed below in table \ref{tab:versions}.
 Many of the software versions are determined by the choice of the ROS distribution (distro).

### 2022-09-29 Thesis/text/ergebnisse.tex
pagerefs

@@
 Fn$\_$req6 is only partly fulfilled as the test showed potential shortcomings of the implemented collision behavior.
-Fn$\_$req7 satisfies the acceptance criteria, but fn$\_$req8 is only partly met by the implemented behavior to find alternative goals.
-Finally, the battery behavior meets all acceptance criteria, and fn$\_$req9 is fulfilled.
+Fn$\_$req7 satisfies the acceptance criteria, but fn$\_$req8 is only partly met by the implemented solution to find alternative goals.
+Finally, the battery behavior meets all acceptance criteria, and thus fn$\_$req9 is fulfilled.
 

### 2022-11-11 Code/bt_workspace/src/bt/include/update_map_after_collision.hpp
Update to Collision Behavior

@@ -133,15 +133,15 @@ class UpdateMapAfterCollision : public BT::SyncActionNode
     {
         if(debug)
             RCLCPP_INFO(node_->get_logger(), "Converting to vector 2d.");
-        for (size_t i = 0; i < map.info.width-1; i++)
+        for (size_t i = 0; i < map.info.height - 1; i++) // columns are the outer loop
         {   
             // if(debug)
                 // RCLCPP_INFO(node_->get_logger(), "i: %d", i);
-            for (size_t j = 0; j < map.info.height -1; j++)
+            for (size_t j = 0; j < map.info.width -1; j++) // rows are the inner loop
             {   
                 // if(debug)
                     // RCLCPP_INFO(node_->get_logger(), "j: %d", j);
-                vec_2d.at(i).at(j) = map.data.at(i * map.info.height + j);
+                vec_2d.at(i).at(j) = map.data.at(i * map.info.width + j);
             }            
         }       
     }
@@ -210,9 +210,9 @@ class UpdateMapAfterCollision : public BT::SyncActionNode
         // then it should modify the costmap by adding a 5x5 square at the location of the crash  
         if(debug)
             RCLCPP_INFO(node_->get_logger(), "Adding occupation to costmap vector."); 
-        for (size_t i = x_px -2; i < x_px +2; i++)
+        for (size_t i = y_px -2; i < y_px +2; i++)
         {
-            for (size_t j = y_px -2; j < y_px +2; j++)
+            for (size_t j = x_px -2; j < x_px +2; j++)
             {
                 costmap_2d.at(i).at(j) = 100; 
             }
@@ -230,11 +230,11 @@ class UpdateMapAfterCollision : public BT::SyncActionNode
             RCLCPP_INFO(node_->get_logger(), "Converting to one dimensional data array and putting it into the map msg.");
         std::vector<int8_t> data((map.info.height * map.info.width), -1);
 
-        for (size_t i = 0; i < costmap_vector_2d.size(); i++) //Rows
+        for (size_t i = 0; i < costmap_vector_2d.size(); i++) //Columns outer loop
         {   
             // if(debug)
                 // RCLCPP_INFO(node_->get_logger(), "i: %d", i);
-            for (size_t j = 0; j < costmap_vector_2d.at(i).size(); j++) // Columns
+            for (size_t j = 0; j < costmap_vector_2d.at(i).size(); j++) // Rows inner loop
             {
[…]

### 2022-11-11 Thesis/text/methoden.tex
Update to Collision Behavior

@@
 The software requirements that improve the robot's behavior are derived from the current limitations of the standard ROS and Navigation2 setup for mobile robots described in the previous section.
-The non-functional requirements are listed separately in table \ref{tab:nonfn_req}.
-The functional requirements and their acceptance criteria are listed in table \ref{tab:fn_req} on page \pageref{tab:fn_req}.
+The non-functional requirements and descriptions are listed in table \ref{tab:nonfn_req}.
+The functional requirements and their acceptance criteria are listed afterwards in table \ref{tab:fn_req} on page \pageref{tab:fn_req}.
 Figure \ref{fig:use_case} on page \pageref{fig:use_case} depicts a Use-Case diagram with the functions needed to improve the system's autonomy and better contextualize the system's requirements.
@@
 			\hline
-			\textbf{Nr.} & \textbf{Name} & \textbf{Priority} & \textbf{Description} \\ 
+			\textbf{Nr.} & \textbf{Name} & \textbf{Priority} & \textbf{Description/Acceptance criteria} \\ 
 			\hline
@@
 	\renewcommand{\arraystretch}{1.5}
-	\resizebox{0.95\textwidth}{!}{%
+	\resizebox{0.9\textwidth}{!}{%
 		\begin{tabular}{| m{0.08\textwidth} | m{0.15\textwidth}| m{0.09\textwidth} | m{0.60\textwidth}|} 

### 2022-11-11 Code/bt_workspace/src/bt_tests/src/scenario1.cpp

@@ -54,7 +54,7 @@ void BehaviorTester::spawn_entity()
 
 
 
-  auto result = spawn_client_handle_->async_send_request()
+  // auto result = spawn_client_handle_->async_send_request();
 }
 
 int main(int argc, char ** argv)

### 2023-01-29 Paper/paper.tex
Paper template skeleton

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[letterpaper, 10pt, conference]{ieeeconf}
\overrideIEEEmargins			% to meet printer requirements
\IEEEoverridecommandlockouts	% to override locked commands

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%REQUIRED PACKAGES%
\usepackage{multirow}
\usepackage{rotating}
\usepackage{float}
\usepackage{caption}
\usepackage{lipsum}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{euscript}
\usepackage{ctable}
\graphicspath{{./Figures/}}
\usepackage[nonumberlist,acronym]{glossaries}
\usepackage[hidelinks]{hyperref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% correct hyphenation
\hyphenation{temp-orary}

% glossaries
\newacronym{ROS}{ROS}{Robot Operating System}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\title{Behavioral Planning Approaches To Improve Autonomous Behaviour During
	Navigation for Mobile Robots Running ROS}

\author{Lukas Evers$^{*,1}$, Umut Uzunoglu$^{1}$, and Ahmed Hussein$^{1}$ \textit{Senior Member, IEEE}%
    \thanks{$^{*}$ Corresponding author }%
    \thanks{$^{1}$ IAV GmbH, Berlin, Germany \newline
		{\tt\small lukas.evers@iav.de, umut.uzunoglu@iav.de, ahmed.hussein@ieee.org}}%
}

\maketitle
\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}

This document is a model and instructions for \LaTeX.
This file define the components of your paper [title, text, heads, etc.]. 
*CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, or Math in Paper Title or Abstract.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:Introduction}

 Lorem ipsum dolor sit amet, consectetur adipiscing elit. Nam in turpis laoreet magna elementum commodo. Interdum et malesuada fames ac ante ipsum primis in faucibus. Mauris non massa accumsan, molestie velit nec, efficitur leo. Aenean nunc arcu, molestie vitae lectus sit amet, malesuada sagittis lacus. Curabitur fringilla massa ac tellus vestibulum pretium. Sed neque metus, aliquet fringilla tincidunt placerat, vehicula quis justo. Nulla vulputate luctus risus, sed porttitor mauris imperdiet nec. Vestibulum nunc ligula, vestibulum eget nibh sed, volutpat rutrum nisl. Curabitur ultrices nulla urna, et molestie dui rutrum id. Duis faucibus lacinia porttitor. Ut vel commodo nunc~\gls{ROS}~\cite{quigley2009}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Related Work}
\label{sec:RelatedWork}

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.9\linewidth]{fig1.png}
    \caption{Figure Caption}
    \label{fig:figureLabel}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Proposed Approach}
\label{sec:ProposedApproach}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental Work}
\label{sec:ExperimentalWork}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Results and Discussion}
\label{sec:ResultsAndDiscussion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion and Future Recommendations}
\label{sec:Conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\vfill
%\section*{ACKNOWLEDGMENT}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\addtolength{\textheight}{-12cm}
%\vspace{10mm}
\bibliographystyle{IEEEtran}
\bibliography{paper}
\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2023-01-29 Paper/paper.tex
Updated abstract and title

@@
 
-\title{Behavioral Planning Approaches To Improve Autonomous Behaviour During
-	Navigation for Mobile Robots Running ROS}
+\title{Enhancing Autonomy of Mobile Robots with Behavioral Tree using ROS2}
 
@@
 
-This document is a model and instructions for \LaTeX.
-This file define the components of your paper [title, text, heads, etc.]. 
-*CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, or Math in Paper Title or Abstract.
+This paper proposes a behavioral planning approach to improve the autonomy of mobile robots using the Robot Operating System (ROS2).
+The study aims to address the issue of human intervention that is often required during the operation of autonomous mobile robots.
+The system includes a monitoring system, sensor data storage, and behavior tree design and implementation.
+The effectiveness of the approach was evaluated using the Gazebo simulator, with the behavior tree handling various scenarios such as sensor failures, collisions, unreachable goals, and low battery state.
+The simulation results showed a significant improvement in the robot's autonomy and resilience to failures.
+Additional validation is being performed through real-world experiments to confirm the results.
 

### 2023-01-29 Paper/paper.tex
Updated state of the art

@@
 
-This paper proposes a behavioral planning approach to improve the autonomy of mobile robots using the Robot Operating System (ROS2).
-The study aims to address the issue of human intervention that is often required during the operation of autonomous mobile robots.
+This paper proposes a state of the art review and a behavioral planning approach to improve the autonomy of mobile robots using the Robot Operating System (ROS2).
+The study aims to address the issue of human intervention that is often required during the operation of autonomous mobile robots and improve the efficiency and performance of the robots.
 The system includes a monitoring system, sensor data storage, and behavior tree design and implementation.
 The effectiveness of the approach was evaluated using the Gazebo simulator, with the behavior tree handling various scenarios such as sensor failures, collisions, unreachable goals, and low battery state.
-The simulation results showed a significant improvement in the robot's autonomy and resilience to failures.
-Additional validation is being performed through real-world experiments to confirm the results.
+The simulation results showed a significant improvement in the robot's autonomy and resilience to failures, providing evidence for the feasibility of the proposed approach.
+To further validate the results, additional experiments are being performed in real-world settings.
 
@@
 
- Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Nam in turpis laoreet magna elementum commodo.
-Interdum et malesuada fames ac ante ipsum primis in faucibus.
-Mauris non massa accumsan, molestie velit nec, efficitur leo.
-Aenean nunc arcu, molestie vitae lectus sit amet, malesuada sagittis lacus.
-Curabitur fringilla massa ac tellus vestibulum pretium.
-Sed neque metus, aliquet fringilla tincidunt placerat, vehicula quis justo.
-Nulla vulputate luctus risus, sed porttitor mauris imperdiet nec.
-Vestibulum nunc ligula, vestibulum eget nibh sed, volutpat rutrum nisl.
-Curabitur ultrices nulla urna, et molestie dui rutrum id.
-Duis faucibus lacinia porttitor.
-Ut vel commodo nunc~\gls{ROS}~\cite{quigley2009}.
+Mobile robots are increasingly being used in a variety of fields, including warehouse logistics, last-mile delivery, and agriculture.
+Despite being marketed as autonomous, they often require human supervision to operate correctly and are limited in their autonomy to specific environments.
+The Robot Operating System (ROS) is a popular platform for building and programming robots, but the standard robot still requires human intervention to ensure proper functioning and safety in uncertain environments.
+
+This paper focuses on how behavior planning can increase the robustness and autonomy of robots running ROS2.
+Behavior planning allows the robot to react to failures and problems and decide on alternative courses of action to mitigate risks and achieve its goals.
+The paper will explore the possibility of improving current systems by adding a dedicated behavior planning component and provide an implementation of exemplary behaviors and a system architecture for incorporating behavior planning in other robots.
+
+The study aims to demonstrate that the implementation of behavior planning can increase the robot's robustness and decrease the need for human intervention in various scenarios.
+The paper will present the results obtained from selected scenarios, where failures and problems will be artificially induced to test the robot's ability to behave autonomously.
+The paper will summarize the results and discuss the acceptance criteria for the requirements.
+The conclusion will provide an outlook for future work and recommendations.
 
@@
 
-\section{Related Work}
-\label{sec:RelatedWork}
+\section{State of the Art}
+\label{sec:StateOfTheArt}
+
+\subsection{Autonomous Driving Navigation Architectures}
+
+Autonomous driving software architectures follow the "\textit{Sense - Think - Act}" paradigm \cite{murphy2000}, as shown in Figure \ref{fig:autonomous_driving_architecture}.
+Most architectures implement a hierarchical structure in which sensory inputs are processed to create an environmental representation, leading to a path from the current position to the destination computed by global planning.
+This is followed by behavioral planning, which ensures that the vehicle obeys traffic rules, and local planning, which translates motion commands while taking into account system constraints.
+An additional system monitoring layer monitors the execution of other components and triggers actions in case of failures \cite{zimmermann2020adaptive}.
+Vehicles can achieve (Society of Automotive Engineers) SAE levels of autonomy up to levels four and five \cite{bacha2008odin}.
+Expanding the behavior planning module can lead to higher levels of autonomy \cite{reke2020}.
 
@@
 	\centering
-    \includegraphics[width=0.9\linewidth]{fig1.png}
-    \caption{Figure Caption}
-    \label{fig:figureLabel}
+	\includegraphics[width=0.9\linewidth]{Figures/autonomous_driving_architecture.png}
+	\caption{Common Autonomous Driving Architecture \cite{brooks1986,velasco2020}}
+	\label{fig:autonomous_driving_architecture}
 \end{figure}
+
+\subsection{Behavior Types}
+
+The behavior planning module is responsible for choosing the best behavior to execute in a specific environment to ensure the best performance for a given task.
+A behavior planning approach provides the robot with acceptable behaviors for every possible scenario the robot is operating in.
+
+\subsubsection{Reactive Approach}
+
+The first type of behaviors implemented in behavior-based robots were reactive behaviors.
+This maps sensory input directly to motor commands, similar to reflexes in humans.
+Reactive behaviors are fast due to their low computational load, and are suitable in scenarios where real-time safety is a concern.
+However, a system with a solely reactive behavior planning approach will not meet requirements for higher levels of autonomy.
+Reactive behaviors can still be valuable in improving vehicle safety and reliability \cite{desilva2008}.
+
+\subsubsection{Deliberative Approach}
+
+Deliberative behaviors are defined as behaviors that involve planning and decision-making.
+Unlike reactive behaviors, deliberative behaviors follow the "Sense, Think, Act" paradigm and can decide the best course of action before acting \cite{murphy2000}.
+A behavior planner with a deliberative approach can make decisions based on current and previously processed data, and can predict environmental changes.
+Deliberative behaviors improve autonomy and mimic human skills, reducing the need for a human operator.
+
+\subsubsection{Hybrid Approach}
+
+A hybrid model combining reactive and deliberative behaviors can create reliable and intelligent systems.
+The hybrid approach combines the quick reaction times of reactive behaviors with the decision-making ability of deliberative behaviors, leading to more robust robots and higher levels of autonomy.
+The higher-level deliberative planners can override reactive planners, allowing the system to be more flexible in uncertain environments while still maintaining fast reaction times \cite{arkin1998}.
+
+\subsection{Behavior Planning Approaches}
+
+Robots that use a hierarchical behavior planning approach need a system to determine which behaviors to execute.
+Different approaches for behavior planning use states to determine which strategies and behaviors are utilized.
+This article presents and compares Finite State Machines (FSMs), Behavior Trees (BTs), and Partially Observable Markov Decision Processes (POMDPs).
+
+\subsubsection{Finite State Machines}
+
+FSMs are a popular choice for behavior planning in robotics.
+The fundamental idea behind FSMs is to describe behavior in terms of states that trigger actions.
+FSMs are defined by a set of states, a starting state, state transitions, final states, and an input alphabet.
+At any given time, a specific state machine is selected and its actions are executed.
+The benefits of FSMs include their speed and ease of implementation, especially when using libraries, which can make the development process much simpler and faster.
+However, as the complexity of the problem increases, so does the effort required to implement the FSM.
+Despite these limitations, FSMs are still a widely used approach in robotics due to their simplicity and effectiveness in controlling motors \cite{wagner2006}.
+
+\subsubsection{Behavior Trees}
+
+BTs are a popular approach for modeling AI for non-player characters in computer games.
+Unlike FSMs, BTs have a hierarchical structure, with nodes that represent actions or sequences of actions.
+The tree consists of a root node, child nodes, parent nodes, and leaf nodes (execution or action nodes).
+The behavior tree is executed by ticking the root node, which will then be recursively ticked down the tree until the execution of an action node is triggered.
+Control nodes, such as "Sequence," "Fallback," and "Condition," determine the order and conditions of execution for the action nodes.
+BTs provide a flexible and intuitive way to represent complex AI behaviors, and are widely used in the game development industry \cite{iovino2022}.
+
+\subsubsection{Partially Observable Markov Decision Processes}
+
+POMDPs are a more sophisticated approach for decision-making in uncertain environments.
+The basic idea behind POMDPs is to model decision-making as a process of choosing actions based on the probability distribution of the system's state.
+A POMDP consists of a set of states, actions, observations, rewards, and a transition model.
+POMDPs use belief states, which represent the probability distribution of the system's state, to make decisions.
+The solution to a POMDP is a policy, which is a mapping of belief states to actions.
+POMDPs are more complex than FSMs and BTs, but provide a more robust and flexible way to handle uncertainty in robotic decision-making \cite{feyzabadi2014riskaware}.
+
+In conclusion, the behavior planning module of an autonomous driving system plays a crucial role in determining the best behavior to execute in a given environment.
+The behavior planning approach can range from reactive to deliberative to a hybrid of both.
+FSMs and BTs are two popular approaches for behavior planning in robotics, while POMDPs are a more sophisticated approach for decision-making in uncertain environments.
+FSMs are fast and easy to implement, while BTs provide a flexible and intuitive representation of complex AI behaviors.
+POMDPs are more complex but provide a robust and flexible way to handle uncertainty.
+The proposed approach for this study will use a BT approach for the behavior planning module, offering a hierarchical structure for modeling AI and a flexible representation of complex behaviors.
 

### 2023-01-29 Paper/paper.tex
Updated proposed approach

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Proposed Approach}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Experimental Work}
-\label{sec:ExperimentalWork}
+In this work, the Robot Operating System 2.0 (ROS2) \cite{ros2022} framework was selected as the platform to test the hypothesis.
+ROS2 is a newer version of the open-source middleware, Robot Operating System (ROS) \cite{quigley2009}, that has received updates to its architecture and design for improved real-time safety, certification, and security for industrial applications.
+ROS2's Nav2 package, which is used for navigation and localization, has some limitations.
+For example, it may not handle events such as slipping wheels or orientation changes by external influences appropriately.
+This can lead to erratic movement of the robot, as well as incorrect planning commands.
+In severe cases, the robot may continue to drive despite incorrect interpretation of its surroundings, which is highly unsafe and requires manual intervention.
+Therefore, to achieve higher levels of autonomy, it is necessary to define requirements that address these limitations and elevate the level of safety and reliability.
+
+\subsection{Requirements}
+
+The limitations of the standard ROS and Navigation2 setup for mobile robots were identified and used to derive software requirements for improving the robot's behavior.
+The non-functional requirements and their descriptions are listed in Table \ref{tab:fn_req}.
+The functional requirements and their acceptance criteria are also listed in the same table, with prioritization based on the severity of consequences for the system.
+The first six functional requirements prioritize safety and robustness, while functional requirements seven to nine focus on increasing robot autonomy with lower risk potential.
+
+\begin{table*}[ht]
+	\centering
+	\caption{Non-Functional and Functional Requirements}
+	\label{tab:fn_req}
+	\renewcommand{\arraystretch}{1.25}
+		\begin{tabular}{|m{0.06\textwidth}|m{0.15\textwidth}|m{0.05\textwidth}|m{0.64\textwidth}|} 
+			\hline
+			\textbf{Nr.} & \textbf{Name} & \textbf{Priority} & \textbf{Description/Acceptance criteria} \\ 
+			\hline
+			non$\_$freq1 & Single Point of Failure & High & The system does not have a single point of failure. If parts of the system fail, the system maintains some level of functionality. \\ 
+			\hline
+			non$\_$freq2 & Performance & High & The control loop of the system guarantees fast reactions. The average frequency at which the system operates is higher than 100Hz (10ms). \\ 
+			\hline
+			non$\_$freq3 & Determinism & High & The outcome for a given set of inputs must be deterministic, which means that the behavior will always be executed in a similar way. \\
+			\hline
+			non$\_$freq4 & Deliberate & High & The robot is able to perform deliberative behaviors, which means that the newly implemented behaviors go beyond the reaction to the sensor input and have a planning aspect.
+\\
+			\hline
+			fn$\_$req1 & Sensor Failure & High & The system detects sensor failure.
+Ensure that the system can restart sensors and decrease the speed during the time the sensor delivers limited information.
+\\
+			\hline
+			fn$\_$req2 & Emergency Detection & High & The system detects emergency.
+Ensure that the system can detect when the continuation on the calculated path is no longer safe (sensor failures, blockage).
+\\
+			\hline
+			fn$\_$req3 & Emergency Stop & High & The system can initiate emergency stops. Ensure that the system can override all commands and stop in case an emergency is detected. \\
+			\hline 	
+			fn$\_$req4 & Override Navigation2 & High & The system can override navigation2. Ensure that the system's commands can always override the commands coming from navigation2. \\
+			\hline
+			fn$\_$req5 & Maintain operability & High & The robot executes commands as long as it is safe.
+Ensure that the robot keeps driving if it is safe even when system functions are not working correctly.
+\\
+			\hline
+			fn$\_$req6 & Recovery & High & The system can recover from crashes. Ensure that the system can successfully reach goals despite previous crashes.\\ 
+			\hline 
+			fn$\_$req7 & Control Path Planning & Medium & The system controls and rates the quality of the planned paths. \\
+			\hline	
+			fn$\_$req8 & Reset Goals & Medium & The system can reset and override goals set by the user so that the goal is reachable by planners. \\
+			\hline
+			fn$\_$req9 & Robot Range & Medium & Ensure that the robot will not run out of battery during navigation to a goal. \\	
+			\hline
+		\end{tabular}
+\end{table*}
+
+\subsection{System Solution}
+
+This work focuses on extending the capabilities of a mobile robot system by adding an autonomous layer to the existing navigation setup.
+The requirements for improving the robot's behavior were derived from the limitations of the current setup and are prioritized based on their severity.
+The autonomy layer is designed to increase the robustness and autonomy of the whole system and includes three main components: a system supervisor, a data storage system, and a behavior planner.
+The system supervisor monitors the health of all components, including sensors and navigation, and provides constant execution monitoring.
+The data storage system stores relevant system information and data, allowing for more deliberative behaviors.
+The behavior planner processes information from the execution checker, the navigation system, and sensor data to make decisions on how to increase autonomy.
+Finally, a decision gate is needed to intelligently switch between speed commands from the navigation system and the autonomy layer.
+A block diagram of the proposed system is shown in Figure \ref{fig:block_diagram}.
+
+\begin{figure}[ht]
+	\centering 
+	\includegraphics[width=0.9\linewidth]{Figures/block_diagram.png}
+	\caption{Proposed System Block Diagram: The additional components, discussed in this work, are colored blue while the existing components are colored gray}
+	\label{fig:block_diagram}
+\end{figure}
+
+\subsection{Software and Simulation}
+
+The paper compares and assesses behavior planning approaches, eventually choosing the behavior tree method for its increased flexibility and determinism.
+The Nav2 stack implements behavior trees using the Behaviortree.CPP (BT.CPP) library, which provides a way to create, execute, monitor, and edit behavior trees and allows communication between nodes.
+The behavior trees can be created using the Groot software, which is a graphical user interface for creating, editing, and debugging behavior trees.
+The behavior trees are developed and tested using the Gazebo simulator and the Turtlebot3 model.
+The simulator allows for easier testing and offers control over the simulation time.
+The software versions used are with the choice of using ROS2 Foxy and C++ for development.
+
+\subsection{System Architecture}
+
+The system architecture in Figure \ref{fig:component_diagram} shows the integration of the autonomy layer into current architecture, with existing components in gray and new components (Data Backup, Decision Gate, Behavior Tree, and System Supervisor) in blue.
+The System Supervisor is explained in the BT section, as it's part of the behavior tree's functionality.
+
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/component_diagram_bt.png}
+	\caption{Component Diagram of System Architecture}
+	\label{fig:component_diagram}
+\end{figure}
+
+The data backup component gathers data from system components and provides it to the behavior tree for past data points.
+It subscribes to sensor data and saves the messages in a queue data structure.
+The array size is limited to decrease network load and computational requirements.
+The behavior tree accesses the saved data through a service call to one of the component's services.
+The data backup component operates in a separate ROS node to avoid blocking the execution of the autonomy system.
+
+The decision gate component modifies or blocks velocity commands from Nav2.
+It receives messages from Nav2 and the behavior tree, and prioritizes messages from the behavior tree.
+The component is an added safety measure in case Nav2 is not responsive to a cancellation command.
+
+The system supervisor component uses a Watchdog design pattern to monitor the system and take corrective measures when problems are detected.
+The component operates partly outside the behavior tree's thread to avoid blocking its execution.
+The component receives messages from nodes and informs the behavior tree of problems with a node through a provided service.
+The behavior tree uses service calls to check component health in Condition Nodes.
+If a node fails, the robot slows down and tries to restart the sensor.
+If the restart fails twice, the robot stops.
+The timeout for declaring a node failed is set to one second.
+
+\subsection{Behavior Tree Structure}
+
+The top-level behavior control fallback nodes of a simplified behavior tree are shown in Figure \ref{fig:top_level_bt}.
+The tree receives the tick signal from a sequence root node.
+The root node functions to increase the system's robustness by ensuring all fundamental components are functioning properly before executing higher-level behaviors.
+Details for each behavior will follow.
+
+\begin{figure}[ht]
+	\centering
+	\includegraphics[width=0.9\linewidth]{Figures/simplified_bt_inverted.png}
+	\caption{Top Level Behavior Tree Structure}
+	\label{fig:top_level_bt}
+\end{figure}
+
+The robot's collision behavior is meant to handle collisions and react to them.
+Collisions occur due to inadequate sensor coverage or a lack of detection capabilities.
+The collision behavior involves detecting a collision, getting out of the collision state, updating the map, and resetting all costmaps to generate a new path to the goal.
+The collision checker uses a simulated sensor and is integrated into the system supervisor.
+The behavior tree (BT) requests the most recent command velocities and modifies them to reverse out of the collision.
+The map needs to be updated with a new obstacle at the collision point and republished.
+The estimated collision point is simplified to be directly in front of the robot, and a square with 100\% occupancy is added to the map.
+The original goal is then republished to replan the global path.
+
+The battery behavior monitors the robot's battery and decides if the robot should navigate to a new goal location.
+The decision is based on the current battery charge and the estimated energy consumption to reach the new goal.
+The energy expenditure was predicted using a linear function and considering the idle consumption and energy to drive the motors.
+A safety factor was added to the energy estimation function to consider scenarios where the robot may need to reroute or stop.
+A battery package was created for simulation purposes and updates the battery charge every second.
+The battery package subscribes to the command velocity topic and decreases the battery charge for every message received.
+The consumption rates are not close to reality but were chosen for rapid testing.
+
+The path planning behavior involves generating a safe path for the robot to navigate from its current location to a goal location.
+The path is generated based on the robot's current location, goal location, and the environment.
+The path planning behavior makes use of the map data and costmaps to generate a path that avoids obstacles.
+The costmap is updated with the new obstacle information to ensure the path is safe.
+The global and local planners generate a path, and the velocity commands are sent to the robot.
+The robot's velocity is monitored, and if it deviates from the planned path, the path is replanned to ensure the robot stays on course.
+
+\subsection{ROSBot Adaption}
+
+The ROSBot robot requires several crucial modules to be developed for successful real-world experiments.
+One such module is the battery monitoring module, which is responsible for monitoring the battery level and providing updates to the rest of the system.
+This is crucial as it ensures the robot operates safely within its battery capacity and prevents unexpected power-off during operation.
+Another important module is the collision detection module, which helps the robot detect obstacles in its environment and respond accordingly.
+This module is essential for the robot's safety as it prevents damage to itself or other actors in the environment.
+The development of these modules is essential for ensuring the ROSBot's robustness and safe operation in real-world scenarios.
 

### 2023-01-29 Paper/paper.tex
Updated the results and conclusion

@@
 
-\begin{figure}[ht]
+\begin{figure*}[ht]
 	\centering
@@
 	\label{fig:top_level_bt}
-\end{figure}
+\end{figure*}
 
@@
 
+\subsection{Setup}
+
+The efficacy of the behavior tree component in the robot was evaluated through scenario testing in a simulated apartment environment (as shown in Figure \ref{fig:house_gazebo}) and a simulated world with small round obstacles in an enclosed space (as shown in Figure \ref{fig:world_gazebo}).
+These scenarios were designed to trigger behavior responses from the component and to assess its ability to meet the functional requirements specified earlier.
+The environments were pre-mapped using the ROS2 "SLAM Toolbox" package and provided by the Nav2 map server.
+The robot location and navigation goals were defined in a separate ROS node, which programmatically positioned the robot within the environment and set its goals.
+This approach allowed the same scenario to be run with and without behavior planning.
+To test the behavior of the sensor failure fallbacks, the relevant sensor driver was shut down to emulate sensor failure.
+
+\begin{figure}[ht]
+	\centering
+	\begin{subfigure}[b]{0.49\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{Figures/house_env.png}
+		\caption{Apartment Environment}
+		\label{fig:house_gazebo}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[b]{0.49\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{Figures/world_env.png}
+		\caption{Enclosed Environment}
+		\label{fig:world_gazebo}
+	\end{subfigure}
+	\caption{Gazebo World Environment}
+	\label{fig:three graphs}
+\end{figure}
+
+\subsection{Scenarios}
+
+The scenarios, derived from functional requirements, test multiple requirements as listed in tables \ref{tab:behavior_scenarios} and \ref{tab:sensor_scenarios}.
+To be counted as a success, all acceptance criteria for a scenario must be met.
+The scenarios are conducted by shutting down sensor drivers, publishing navigation goals, and simulating obstacles and battery conditions.
+Examples include: NoPathFound$_1$ tests recovery from inability to plan, NoPathFound$_2$ tests alternative goal navigation, Collision$_1$ tests maneuvering out of collisions, Collision$_2$ tests recovering from crashes, and Battery$_1$ tests low battery navigation.
+
+\begin{table}[ht]
+	\centering
+	\caption{Behavior Scenarios}
+	\label{tab:behavior_scenarios}
+	\renewcommand{\arraystretch}{1.25}
+	\resizebox{0.95\linewidth}{!}{%
+		\begin{tabular}{| m{0.11\linewidth} | m{0.19\linewidth}| m{0.4\linewidth} | m{0.3\linewidth}|} 
+			\hline
+			\textbf{Name} & \textbf{Requirements} & \textbf{Description} & \textbf{Success Criteria}\\ 
+			\hline	
+			NoPath Found$\_$1 & fn$\_$req7 & Path to goal can not be calculated, robot is standing still & 
+			Reset system, calculate path to goal, reach goal \\
+			\hline
+			NoPath Found$\_$2 & fn$\_$req7, fn$\_$req8 & Goal is unreachable, robot is standing still & Alternative goals, close to the original are tested to be reachable \\
+			\hline
+			Collision$\_$1 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & Robot is standing still and a collision with the robot is caused & Robot can get out of collision state, navigation to goals still working \\
+			\hline
+			Collision$\_$2 & fn$\_$req2, fn$\_$req4, fn$\_$req6 & The robot is driving and collides with an undetected obstacle (0.25m/s) & 
+			Robot can get out of collision state, navigation to goals still working, undetected obstacle gets added to map \\
+			\hline
+			Battery$\_$1 & fn$\_$req9 & The robot battery runs low & The robot will not drive to a goal that is outside of its reachable range \\
+			\hline	
+		\end{tabular}
+	}
+\end{table}
+
+\begin{table}[ht]
+	\centering
+	\caption{Sensor Scenarios}
+	\label{tab:sensor_scenarios}
+	\renewcommand{\arraystretch}{1.25}
+	\resizebox{0.95\linewidth}{!}{%
+		\begin{tabular}{| m{0.1\linewidth}| m{0.1\linewidth} | m{0.2\linewidth}| m{0.3\linewidth} | m{0.3\textwidth}|} 
+			\hline
+			\textbf{Sensor Tested} & \textbf{Name} & \textbf{Requirements }& \textbf{Description} & \textbf{Success Criteria}\\ 
+			\hline
+			& Lidar$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Lidar node crashes & Reset the system\\ 
+			\cline{2-5}
+			Lidar &Lidar$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
+			\cline{2-5}
+			& Lidar$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
+			\hline
+			& IMU$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, IMU node crashes & Reset system \\
+			\cline{2-5}
+			IMU &IMU$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
+			\cline{2-5}
+			&IMU$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
+			\hline
+			& Odom$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Odom node crashes & Reset the system\\ 
+			\cline{2-5}
+			Odometry &Odom$\_$2 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s straight towards goal (1m away) & Reset system, reach goal \\ 
+			\cline{2-5}
+			&Odom$\_$3 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Navigating with 0.25 m/s and 0.5 rad/s towards goal (1m away) & Reset system, reach goal\\
+			\hline
+		\end{tabular}
+	}
+\end{table}
+
+\subsection{Results}
+
+The results in table \ref{tab:results} show an increase in the successful handling of the test scenarios for the majority of test cases.
+The system supervision component is very reliable and enables the system to recover from an unexpected sensor failure.
+The more advanced behaviors can improve the autonomous handling of the scenarios.
+The battery scenario was successful every time it was induced.
+
+\begin{table}[ht]
+	\centering
+	\caption{Results}
+	\label{tab:results}
+	\renewcommand{\arraystretch}{1.5}
+	\resizebox{0.95\linewidth}{!}{%
+		\begin{tabular}{| m{0.25\linewidth} | m{0.15\linewidth}| m{0.3\linewidth} | m{0.3\linewidth}|} 
+			\hline
+			\multirow{2}{*}{\textbf{Name}} & \multirow{2}{0.11\textwidth}{\textbf{Number of Runs}} & \multicolumn{2}{|c|}{\textbf{Percentage Successful}} \\
+			\cline{3-4}
+			& & \textbf{Without BT} & \textbf{With BT} \\
+			\hline
+			Lidar$\_$1 & 5 & 0 & 100 \\ 
+			\hline
+			Lidar$\_$2 & 5 & 0 & 100 \\ 
+			\hline
+			Lidar$\_$3 & 5 & 0 & 100 \\ 
+			\hline
+			IMU$\_$1 & 5 & 0 & 100 \\ 
+			\hline
+			IMU$\_$2 & 5 & 0 & 100 \\ 
+			\hline
+			IMU$\_$3 & 5 & 0 & 100 \\ 
+			\hline
+			Odom$\_$1 & 5 & 0 & 100 \\ 
+			\hline
+			Odom$\_$2 & 5 & 0 & 100 \\ 
+			\hline
+			Odom$\_$3 & 5 & 0 & 100 \\ 
+			\hline
+			NoPath Found$\_$1 & 20 & 0 & 100 \\ 
+			\hline
+			NoPath Found$\_$2 & 20 & 0 & 55 \\ 
+			\hline
+			Collision$\_$1 & 20 & 0 & 95 \\ 
+			\hline
+			Collision$\_$2 & 20 & 0 & 100 \\ 
+			\hline
+			Battery$\_$1 & 10 & 0 & 100 \\ 
+			\hline
+		\end{tabular}
+	}
+\end{table}
+
+However, there is a lack of intelligence and autonomy when the robot stands still, and a collision is forced.
+The test showed potential shortcomings of the implemented collision behavior.
+The results show that the functional requirements (as defined in \ref{tab:fn_req}) are met in almost all cases, with the exception of fn$_req6$ and fn$_req8$.
+The architectural design choices meet all the non-functional requirements.
+The system's performance when checking the conditions of the BT is below the desired threshold of $10ms$, allowing for efficient execution of behavior tree cycles.
+
+The system also performed well in terms of performance, with the whole system running in multiple threads, allowing the behavior tree to complete one complete execution cycle in about 5ms.
+Overall, the results suggest that the implemented behavior planning and system supervision system is a promising solution for autonomous navigation in complex environments.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{sec:Conclusion}
+
+This paper aimed to improve the autonomy and robustness of ROS2 mobile robots by incorporating advanced behavior planning.
+The implementation of a system supervisor component improved the system's robustness and provided a safe basis for executing more intelligent behaviors.
+The behavior tree was structured to allow for safe execution of complex behaviors and to check for possible scenarios that could interfere with the correct representation of the environment.
+The system architecture provides a good base for future development and showed a novel way of working with the ROS2 executor properties for fast execution of the behavior planner and system components.
+
+The behavior tree design could have been improved by using blackboard values more effectively to increase information flow between action nodes.
+Additionally, incorporating stateful nodes into the tree structure would have increased the possibility of creating better and more deliberative behaviors.
+The tree did not utilize reactive control flow nodes, and no action node was designed to allow multiple threads at the current state.
+These modifications would have further increased the level of autonomy and robustness of the system.
+
+Further efforts are needed to improve the autonomy of the behavior planner by adding more behaviors and fallbacks.
+To increase robustness, the implementation of Intel RSS rules for autonomous driving \cite{gassmann2019towards}, like maintaining safe distances, right-of-way rules, and adaptable speed based on occlusions is crucial.
+The collision detection system also needs to be refined to allow for collision point detection.
+The strategy for finding alternative goals in the path planning behavior needs to be modified to sample more points in all directions around the given goal.
+Additionally, there is a need for dedicated behaviors for multi-robot systems, such as sharing sensor data and intelligent path planning to avoid collisions.
+Future work should also focus on defining a safe state for the robot after triggering an emergency stop.
+The integration of machine learning algorithms into the behavior tree is another important area of future work, as this will allow for intelligent, robust, and independent decision-making.
 

### 2023-01-31 Paper/paper.tex
changed figure in paper

@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/simplified_bt_inverted.png}
+	\includegraphics[width=0.9\linewidth]{Figures/top_level_bt.png}
 	\caption{Top Level Behavior Tree Structure}
@@
 
-The ROSBot robot requires several crucial modules to be developed for successful real-world experiments.
+The ROSBot 2.0 by Husarion has been selected as the test platform for real world scenarios.
+It is a four-wheeled robot equipped with a suite of sensors, including LiDAR, IMU, wheel encoders and range sensors.
+To advance the ROSBot to a higher level of autonomy the proposed system architecture has been implemented.
+The battery charge module and the collision detection module were required to be adapated from the simulation to function on the real robot.
+The collision detection module is based on the IMU data instead of a simulated bumper sensor.
+The battery module estimates the remaining charge by taking the battery voltage, the power consumption of every component and the movement of the robot into account thus assuring that the robot has enough charge left to navigate to a goal.
+
+% The ROSBot robot requires several crucial modules to be developed for successful real-world experiments.
 One such module is the battery monitoring module, which is responsible for monitoring the battery level and providing updates to the rest of the system.

### 2023-01-31 Code/bt_workspace/src/sensor_data_backup/include/sensor_data_backup/sensor_data_backup.hpp
changes to pose subscription from PoseWithCovarianceStamped to PoseStamped

@@ -10,7 +10,7 @@
 // #include "bt_msgs/srv/pub_cmd_vel.hpp"
 #include "bt_msgs/srv/get_twist_array.hpp"
 #include "std_srvs/srv/empty.hpp"
-#include "geometry_msgs/msg/pose_with_covariance_stamped.hpp"
+// #include "geometry_msgs/msg/pose_stamped.hpp"
 #include "bt_msgs/srv/get_last_goal.hpp"
 #include "bt_msgs/srv/get_pose.hpp"
 #include "bt_msgs/srv/get_last_map.hpp"
@@ -65,7 +65,7 @@ class SensorDataBackup : public rclcpp::Node
 
         void map_callback(const nav_msgs::msg::OccupancyGrid::SharedPtr msg);
 
-        void pose_update_callback(const geometry_msgs::msg::PoseWithCovarianceStamped::SharedPtr msg);
+        void pose_update_callback(const geometry_msgs::msg::PoseStamped::SharedPtr msg);
 
         void global_costmap_callback(const nav_msgs::msg::OccupancyGrid::SharedPtr msg);
 
@@ -74,7 +74,7 @@ class SensorDataBackup : public rclcpp::Node
         // rclcpp::Subscription<sensor_msgs::msg::LaserScan>::SharedPtr sub_lidar;
         // rclcpp::Subscription<nav_msgs::msg::Odometry>::SharedPtr sub_odom;
         // rclcpp::Subscription<sensor_msgs::msg::Imu>::SharedPtr sub_imu;
-        rclcpp::Subscription<geometry_msgs::msg::PoseWithCovarianceStamped>::SharedPtr sub_pose;
+        rclcpp::Subscription<geometry_msgs::msg::PoseStamped>::SharedPtr sub_pose;
         rclcpp::Subscription<nav_msgs::msg::OccupancyGrid>::SharedPtr sub_map_;
         rclcpp::Publisher<nav_msgs::msg::OccupancyGrid>::SharedPtr pub_map_;
         rclcpp::Subscription<geometry_msgs::msg::Twist>::SharedPtr sub_cmd_vel_;
@@ -100,7 +100,7 @@ class SensorDataBackup : public rclcpp::Node
         std::vector<nav_msgs::msg::OccupancyGrid> vector_map;
         geometry_msgs::msg::PoseStamped last_goal;
         std::vector<geometry_msgs::msg::Twist> vector_cmd_vel_;
[…]

### 2023-01-31 Code/bt_workspace/src/sensor_data_backup/src/sensor_data_backup.cpp

@@ -11,7 +11,7 @@ SensorDataBackup::SensorDataBackup(const std::string & node_name) : Node(node_na
     service_pub_last_goal_ = this->create_service<std_srvs::srv::Empty>("pub_last_goal_service", std::bind(&SensorDataBackup::pub_last_goal_service_callback, this, std::placeholders::_1, std::placeholders::_2));
     
     sub_map_ = this->create_subscription<nav_msgs::msg::OccupancyGrid>("map", rclcpp::QoS(rclcpp::KeepLast(1)).transient_local().reliable(), std::bind(&SensorDataBackup::map_callback, this, std::placeholders::_1));
-    sub_pose = this->create_subscription<geometry_msgs::msg::PoseWithCovarianceStamped>("amcl_pose", 1, std::bind(&SensorDataBackup::pose_update_callback, this, std::placeholders::_1));
+    sub_pose = this->create_subscription<geometry_msgs::msg::PoseStamped>("robot_pose", 1, std::bind(&SensorDataBackup::pose_update_callback, this, std::placeholders::_1));
 
     sub_global_costmap = this->create_subscription<nav_msgs::msg::OccupancyGrid>("global_costmap/costmap", 1, std::bind(&SensorDataBackup::global_costmap_callback, this, std::placeholders::_1));
     service_get_collision_pose_ = this->create_service<bt_msgs::srv::GetPose>("get_collision_pose_service", std::bind(&SensorDataBackup::get_collision_pose_service_callback, this, std::placeholders::_1, std::placeholders::_2));
@@ -37,7 +37,7 @@ void SensorDataBackup::send_updated_map_callback(
     pub_map_->publish(last_map);
 
 }
-void SensorDataBackup::pose_update_callback(const geometry_msgs::msg::PoseWithCovarianceStamped::SharedPtr msg)
+void SensorDataBackup::pose_update_callback(const geometry_msgs::msg::PoseStamped::SharedPtr msg)
 {
     if (debug)
         RCLCPP_INFO(this->get_logger(), "Received AMCL Pose");
@@ -116,7 +116,7 @@ void SensorDataBackup::get_last_pose_service_callback(
     const bt_msgs::srv::GetPose_Request::SharedPtr request,
     const bt_msgs::srv::GetPose_Response::SharedPtr response)
 {
-    response->last_robot_pose = this->vector_poses_.back().pose.pose;
+    response->last_robot_pose = this->vector_poses_.back().pose;
 }
 
[…]

### 2023-01-31 Paper/paper.tex
battery charge calculation added

@@
 The energy expenditure was predicted using a linear function and considering the idle consumption and energy to drive the motors.
-A safety factor was added to the energy estimation function to consider scenarios where the robot may need to reroute or stop.
+A safety factor was added to the energy estimation function to consider scenarios where the robot may need to stop navigation..
 A battery package was created for simulation purposes and updates the battery charge every second.
@@
 The collision detection module is based on the IMU data instead of a simulated bumper sensor.
-The battery module estimates the remaining charge by taking the battery voltage, the power consumption of every component and the movement of the robot into account thus assuring that the robot has enough charge left to navigate to a goal.
 
-% The ROSBot robot requires several crucial modules to be developed for successful real-world experiments.
-One such module is the battery monitoring module, which is responsible for monitoring the battery level and providing updates to the rest of the system.
-This is crucial as it ensures the robot operates safely within its battery capacity and prevents unexpected power-off during operation.
-Another important module is the collision detection module, which helps the robot detect obstacles in its environment and respond accordingly.
-This module is essential for the robot's safety as it prevents damage to itself or other actors in the environment.
-The development of these modules is essential for ensuring the ROSBot's robustness and safe operation in real-world scenarios.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The battery module estimates the remaining charge by taking the battery voltage, the power consumption of every component and the movement of the robot into account thus assuring that the robot has enough charge left to navigate to a given goal.
+Following function is used to determine the current state of charge of the battery: \[Batt_{SoC}(t) = Batt_{SoC}(t-1)-\frac{\frac{P_{components}}{V_{measured}}}{Batt_{capacity}}* t_{timestep} *100\] where $Batt_{SoC}$ is the charge in percentage, $t$ is the time of calculation in hours, $t_{timestep}$ the time step in seconds and $P_{components}$ the combined power consumption of every component connected to the robot in Watt.
+$P_{components}$ changes depending on the speed of the robot.
+The calculation assumes a fully charged battery.
 

### 2023-02-01 Paper/paper.tex
updated paper

@@
 An additional system monitoring layer monitors the execution of other components and triggers actions in case of failures \cite{zimmermann2020adaptive}.
-Vehicles can achieve (Society of Automotive Engineers) SAE levels of autonomy up to levels four and five \cite{bacha2008odin}.
+Vehicles can achieve Society of Automotive Engineers (SAE) levels of autonomy up to levels four and five \cite{bacha2008odin}.
 Expanding the behavior planning module can lead to higher levels of autonomy \cite{reke2020}.
@@
 This can lead to erratic movement of the robot, as well as incorrect planning commands.
-In severe cases, the robot may continue to drive despite incorrect interpretation of its surroundings, which is highly unsafe and requires manual intervention.
+In severe cases, the robot may continue to drive despite incorrect interpretation of its surroundings, which is highly unsafe and requires human intervention.
 Therefore, to achieve higher levels of autonomy, it is necessary to define requirements that address these limitations and elevate the level of safety and reliability.
@@
 
-The limitations of the standard ROS and Navigation2 setup for mobile robots were identified and used to derive software requirements for improving the robot's behavior.
+The limitations of the standard ROS2 and Navigation2 setup for mobile robots were identified and used to derive software requirements for improving the robot's behavior.
 The non-functional requirements and their descriptions are listed in Table \ref{tab:fn_req}.
@@
 The behavior planner processes information from the execution checker, the navigation system, and sensor data to make decisions on how to increase autonomy.
-Finally, a decision gate is needed to intelligently switch between speed commands from the navigation system and the autonomy layer.
+Finally, a decision gate is needed to intelligently switch between velocity commands from the navigation system and the autonomy layer.
 A block diagram of the proposed system is shown in Figure \ref{fig:block_diagram}.
@@
 The behavior tree uses service calls to check component health in Condition Nodes.
-If a node fails, the robot slows down and tries to restart the sensor.
+If a node fails, the robot slows down and tries to restart the node.
 If the restart fails twice, the robot stops.
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{Figures/top_level_bt.png}
+	\includegraphics[width=0.5\linewidth]{Figures/top_level_bt.png}
 	\caption{Top Level Behavior Tree Structure}
@@
 
-The battery behavior monitors the robot's battery and decides if the robot should navigate to a new goal location.
+The battery behavior monitors the robot's battery and decides if the robot should navigate to a given goal.
 The decision is based on the current battery charge and the estimated energy consumption to reach the new goal.
@@
 The battery module estimates the remaining charge by taking the battery voltage, the power consumption of every component and the movement of the robot into account thus assuring that the robot has enough charge left to navigate to a given goal.
-Following function is used to determine the current state of charge of the battery: \[Batt_{SoC}(t) = Batt_{SoC}(t-1)-\frac{\frac{P_{components}}{V_{measured}}}{Batt_{capacity}}* t_{timestep} *100\] where $Batt_{SoC}$ is the charge in percentage, $t$ is the time of calculation in hours, $t_{timestep}$ the time step in seconds and $P_{components}$ the combined power consumption of every component connected to the robot in Watt.
+Following function is used to determine the current state of charge of the battery: \[Batt_{SoC}(t) = Batt_{SoC}(t-1)-\frac{\frac{P_{components}}{V_{measured}}}{Batt_{capacity}}* t_{timestep} *100\] where $Batt_{SoC}$ is the charge in percentage, $t$ the time of calculation in hours, $V_{measured}$ the voltage measurements of the ROSBot, $Batt_{capacity}$ the capacity of the battery pack, $t_{timestep}$ the time step in seconds and $P_{components}$ the combined power consumption of every component connected to the robot in Watt.
 $P_{components}$ changes depending on the speed of the robot.
@@
 The scenarios are conducted by shutting down sensor drivers, publishing navigation goals, and simulating obstacles and battery conditions.
-Examples include: NoPathFound$_1$ tests recovery from inability to plan, NoPathFound$_2$ tests alternative goal navigation, Collision$_1$ tests maneuvering out of collisions, Collision$_2$ tests recovering from crashes, and Battery$_1$ tests low battery navigation.
+Examples include: NoPathFound$\_1$ tests recovery from inability to plan, NoPathFound$\_2$ tests alternative goal navigation, Collision$\_1$ tests maneuvering out of collisions, Collision$\_2$ tests recovering from crashes, and Battery$\_1$ tests low battery navigation.
 
@@
 	\renewcommand{\arraystretch}{1.25}
-	\resizebox{0.95\linewidth}{!}{%
-		\begin{tabular}{| m{0.11\linewidth} | m{0.19\linewidth}| m{0.4\linewidth} | m{0.3\linewidth}|} 
+	\resizebox{0.99\linewidth}{!}{%
+		\begin{tabular}{| m{0.15\linewidth} | m{0.19\linewidth}| m{0.3\linewidth} | m{0.4\linewidth}|} 
 			\hline
@@
 	\renewcommand{\arraystretch}{1.25}
-	\resizebox{0.95\linewidth}{!}{%
-		\begin{tabular}{| m{0.1\linewidth}| m{0.1\linewidth} | m{0.2\linewidth}| m{0.3\linewidth} | m{0.3\textwidth}|} 
+	\resizebox{0.99\linewidth}{!}{%
+		\begin{tabular}{| m{0.12\linewidth}| m{0.1\linewidth} | m{0.22\linewidth}| m{0.5\linewidth} | m{0.08\textwidth}|} 
 			\hline
@@
 			\hline
-			& Lidar$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Lidar node crashes & Reset the system\\ 
+			& Lidar$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Lidar node crashes & Reset system\\ 
 			\cline{2-5}
@@
 			\hline
-			& Odom$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Odom node crashes & Reset the system\\ 
+			& Odom$\_$1 & fn$\_$req1, fn$\_$req2, fn$\_$req3, fn$\_$req4, fn$\_$req5 & Robot is standing still, Odom node crashes & Reset system\\ 
 			\cline{2-5}
@@
 	\renewcommand{\arraystretch}{1.5}
-	\resizebox{0.95\linewidth}{!}{%
-		\begin{tabular}{| m{0.25\linewidth} | m{0.15\linewidth}| m{0.3\linewidth} | m{0.3\linewidth}|} 
+	\resizebox{0.99\linewidth}{!}{%
+		\begin{tabular}{| m{0.25\linewidth} | m{0.15\linewidth}| m{0.15\linewidth} | m{0.15\linewidth}| m{0.15\linewidth}| m{0.15\linewidth}|} 
 			\hline
-			\multirow{2}{*}{\textbf{Name}} & \multirow{2}{0.11\textwidth}{\textbf{Number of Runs}} & \multicolumn{2}{|c|}{\textbf{Percentage Successful}} \\
-			\cline{3-4}
-			& & \textbf{Without BT} & \textbf{With BT} \\
+			\multirow{2}{*}{\textbf{Name}} & \multirow{2}{0.11\textwidth}{\textbf{Number of Runs}} & \multicolumn{4}{|c|}{\textbf{Percentage Successful}} \\
+			\cline{3-6}
+			& & \multicolumn{2}{|c|}{Without BT} & \multicolumn{2}{|c|}{With BT} \\
+			\cline{3-6}
+			& & \textbf{Simulation} & \textbf{Real World} & \textbf{Simulation} & \textbf{Real World} \\		
 			\hline
-			Lidar$\_$1 & 5 & 0 & 100 \\ 
+			Lidar$\_$1 & 5 & 0 & 0 & 100 &100\\ 
 			\hline
-			Lidar$\_$2 & 5 & 0 & 100 \\ 
+			Lidar$\_$2 & 5 & 0 & 0 & 100 &100\\ 
 			\hline
-			Lidar$\_$3 & 5 & 0 & 100 \\ 
+			Lidar$\_$3 & 5 & 0 & 0 & 100 &100\\ 
 			\hline
-			IMU$\_$1 & 5 & 0 & 100 \\ 
+			IMU$\_$1 & 5 & 0 &  0& 100 &100\\ 
 			\hline
-			IMU$\_$2 & 5 & 0 & 100 \\ 
+			IMU$\_$2 & 5 & 0 & 0 & 100 &100\\ 
 			\hline
-			IMU$\_$3 & 5 & 0 & 100 \\ 
+			IMU$\_$3 & 5 & 0 & 0 & 100 &100\\ 
 			\hline
-			Odom$\_$1 & 5 & 0 & 100 \\ 
+			Odom$\_$1 & 5 & 0 & 0 & 100 &100\\ 
 			\hline
-			Odom$\_$2 & 5 & 0 & 100 \\ 
+			Odom$\_$2 & 5 & 0 & 0 & 100 &100\\ 
 			\hline
-			Odom$\_$3 & 5 & 0 & 100 \\ 
+			Odom$\_$3 & 5 & 0 & 0 & 100 &100\\ 
 			\hline
-			NoPath Found$\_$1 & 20 & 0 & 100 \\ 
+			NoPath Found$\_$1 & 20 & 0 & 0 & 100 &100\\ 
 			\hline
-			NoPath Found$\_$2 & 20 & 0 & 55 \\ 
+			NoPath Found$\_$2 & 20 & 0 & 0 & 55 &30\\ 
 			\hline
-			Collision$\_$1 & 20 & 0 & 95 \\ 
+			Collision$\_$1 & 20 & 0 & 0 & 95 &60\\ 
 			\hline
-			Collision$\_$2 & 20 & 0 & 100 \\ 
+			Collision$\_$2 & 20 & 0 & 0 & 100 &75\\ 
 			\hline
-			Battery$\_$1 & 10 & 0 & 100 \\ 
+			Battery$\_$1 & 10 & 0 & 0 & 100 &100\\ 
 			\hline
@@
 The test showed potential shortcomings of the implemented collision behavior.
-The results show that the functional requirements (as defined in \ref{tab:fn_req}) are met in almost all cases, with the exception of fn$_req6$ and fn$_req8$.
+The results show that the functional requirements (as defined in \ref{tab:fn_req}) are met in almost all cases, with the exception of fn$\_$req6 and fn$\_$req8.
 The architectural design choices meet all the non-functional requirements.
@@
 
-The system also performed well in terms of performance, with the whole system running in multiple threads, allowing the behavior tree to complete one complete execution cycle in about 5ms.
+The system also performed well in terms of performance, with the whole system running in multiple threads, allowing the behavior tree to complete one complete execution cycle in about $5ms$.
 Overall, the results suggest that the implemented behavior planning and system supervision system is a promising solution for autonomous navigation in complex environments.

### 2023-02-01 Paper/paper.tex
updated email adresses
Minor title modification before submission

@@
 
-\title{Enhancing Autonomy of Mobile Robots with Behavioral Tree using ROS2}
+\title{Enhancing Autonomy of Mobile Robots with Behavior Tree using ROS2}
 
@@
     \thanks{$^{*}$ Corresponding author }%
-    \thanks{$^{1}$ IAV GmbH, Berlin, Germany \newline
-		{\tt\small lukas.evers@iav.de, umut.uzunoglu@iav.de, ahmed.hussein@ieee.org}}%
+    \thanks{$^{1}$ Digital Lab, IAV GmbH, Berlin, Germany \newline
+		{\tt\small lukas.e@web.de, umutuzunoglu@hotmail.de, ahmed.hussein@ieee.org}}%
 }
